### 2012-03-14 cpomdp/camdp-xadd/src/doc/draft6-after first review3.tex

\documentclass{article}
\usepackage{proceed}

\def\argmax{\operatornamewithlimits{arg\max}}

\title{Exact Symbolic Dynamic Programming for Continuous POMDPs}

\author{}

% The author names and affiliations should appear only in the accepted paper.
%
%\author{ {\bf Harry Q.~Bovik\thanks{Footnote for author to give an 
%alternate address.}} \\  
%Computer Science Dept. \\  
%Cranberry University\\ 
%Pittsburgh, PA 15213 \\ 
%\And 
%{\bf Coauthor}  \\ 
%Affiliation          \\ 
%Address \\              
%\And 
%{\bf Coauthor}   \\ 
%Affiliation \\          
%Address    \\           
%(if needed)\\ 
%}

\begin{document}

\maketitle

\begin{abstract} 
Decision-theoretic planning problems try to find the optimal sequence of actions for a given environment. Using continuous models in such problems is naturally closer to the real-world. Partially Observable MDPs model uncertain environments where the underlying state is not be fully observable.

While previous work have provided solutions for discrete states and/or observations, here we define the first exact solution to both continuous state and observations. This symbolic dynamic programming approach avoids enumerating the state and observation space providing solutions for our general discrete and continuous partially observable MDP. The main contribution defines the relevant observation partitions associated with each state partition. This allows us to perform the Bellman backup operation efficiently. We address the problems in POMDP continuous planning and solutions to solve them symbolically. %To compactly represent the exact value functions Affine algebraic decision diagrams are used. 
\end{abstract}

\section{Introduction}

The following example is used throughout the paper to help better understand the definitions.

%define example* here, just the model
\textbf{Example} \textsc{(Power Plant)}
\emph{The steam generation system of a power plant aims to provide hot steam to the steam turbine which in turn produces electricity. Feed-water is exposed to heat in the tubes leading to the water tank . Evaporating the water is performed under specific pressure and temperature inside the attached water tubes. 
Mixing water and steam pressures makes reading tank levels and temperatures very hard and uncertain. For this reason, the process is modeled using POMDP in the literature. The temperature is assumed to be a continuous variable that is not observed. Noisy observations of the temperature are used to decide on the action of closing or opening the valve.}
Most work in the literature have assumed discretization of the state and observation space ~\cite{steam2}. Here we take a continuous approach using the DC-POMDP framework.

\section{DC-POMDP Model} 
%%%% define DC-MDP
We assume familiarity with MDPs and introduce Discrete and Continuous Partially Observable MDPs as an extension to Discrete and Continuous State MDPs ~\cite{sanner_uai11}. A Discrete and Continuous partially observable MDP (DC-POMDP) is a tuple $\langle
\mathcal{S},\mathcal{A},\mathcal{O},\mathcal{T},\mathcal{R},\mathcal{Z},\gamma,h \rangle$. 
States are represented by vectors of $(\vec{d_s},\vec{x_s}) = ( d_{s_1},\ldots,d_{s_n},x_{s_1},\ldots,x_{s_m} )$ where each $d_{s_i} \in \{ 0,1 \}$ ($1 \leq i \leq n$) is a discrete boolean and each $x_{s_j} \in \mathbb{R}$ ($1 \leq j \leq m$) is a continuous variable. Actions are presented by a finite set of actions $\{ a_1, \ldots, a_p \}$. We define discrete and continuous observations by the vector $(\vec{d_o},\vec{x_o}) = ( d_{o_1},\ldots,d_{o_n},x_{o_1},\ldots,x_{o_m} )$ where each $d_{o_i} \in \{ 0,1 \}$ ($1 \leq i \leq n$) is boolean and each $x_{o_j} \in \mathcal{R}$ ($1 \leq j \leq m$) is a continuous variable.

Three functions are required for modeling DC-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state given the action and previous state%$P(\vec{d}',\vec{x}'|\cdots,a)$
; (2)  $\mathcal{R}:\mathcal{S}\times\mathcal{A}\rightarrow\mathbb{R}$ a reward function which returns the immediate reward of taking an action in some state and (3) an observation function defined as $\mathcal{Z} : \mathcal{S} \times \mathcal{A} \times \mathcal{O} \rightarrow [ 0, 1 ]$  which gives the probability of an observation given the outcome of a state after executing an action.  A discount factor $\gamma, \; 0 \leq \gamma \leq 1$ is included in the model to account for future rewards in higher horizons.

%%%% forms of transition, observation and reward function
To define the transition function, we use the fact that the state variables of this POMDP can be factored into a Dynamic Bayesian Network (DBN). Not allowing synchronic arcs between the discrete and continuous variables defines their joint transition to be independent:

{\footnotesize
\begin{equation}
P(\vec{d_s}',\vec{x_s}'|\vec{d_s},\vec{x_s},a) = 
\prod_{i=1}^n P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \prod_{j=1}^m P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a). \nonumber 
\end{equation}}
The joint observation function can also be defined considering the direct correspondence between the state and observation variables:
{\footnotesize
\begin{equation}
P(\vec{d_o},\vec{x_o}|\vec{d_s},\vec{x_s},a) = 
\prod_{i=1}^n P(d_{o_i}|d_{s_i},a) \prod_{j=1}^m P(x_{o_j}|x_{s_j},a). \nonumber 
\end{equation}}

The \emph{binary} variables are represented using conditional probability functions (CPF) in the form of $P(d_i'|\vec{d},\vec{x},a)$. For \emph{continuous} variables
the CPF $P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ is represented by \emph{piecewise
linear equations} (PLEs) that are first-order Markov and deterministic, i.e. the probabilities are encoded using the Dirac $\delta[\cdot]$ function.

The reward function $R(\vec{d},\vec{x},a)$ is set to be any general piecewise linear  function that could facilitate pruning for space efficiency via bilinear solvers.

For the \textsc{Power Plant} example, the temperature $t_s \in \mathbb{R}$ is modelled as the continuous state variable. Noisy observation of the temperature $t_o \in \mathbb{R}$ is used to help decide the optimal action to either open or close the pressure valve $a \in \{open,close\}$. 
For the transition probability, we use the Dirac function to model the deterministic equations. We can also define stochasticity in the transition model using boolean random variables that are sampled stochastically.  
\begin{align}
P(t_s'|\vec{t_s},a)= \delta\left[ t_s' - 
\begin{cases}
 (a=open) &: t_s + 50 \\ 
(a \neq open) &: t_s - 35 \\
\end{cases}
\right]\nonumber
\end{align}
%contribution? 
For observations we can consider discrete noise, modeled using CSEs. The probability that the observed temperature is equal to the actual temperature is $0.9$  while it indicates a higher temperature with probability of $0.1$. For the open action, observations are more disposed to noise:

\begin{align}
P(t_o|t_s',close) = 
\begin{cases}
 \delta\left[ t_o=t_s \right] &: 0.9 \\
 \delta\left[ t_o=1.5*t_s \right] &: 0.1 \\
\end{cases}
\\
P(t_o|t_s',open) = 
\begin{cases}
 \delta\left[ t_o=1.1*t_s \right] &: 0.8 \\
 \delta\left[ t_o=1.4*t_s \right] &: 0.2 \\
\end{cases}
\end{align}
%We can extend this to consider continuous noise such as a uniform noise distribution which is more realistic in many domains. For a certain interval, the probability is a fixed number, while it is zero elsewhere. The integral of the uniform distribution should sum to one. 
%\begin{align}
%P(t_o|t_s',close) = 
%\begin{cases}
% (t_o>t_s-5) \wedge (t_o<t_s+5) &: 0.1 \\
% \neg((t_o>t_s-5) \wedge (t_o<t_s+5)) &: 0 \\
%\end{cases}
%\end{align}

The reward function can be any linear function of the state or action. Going above a threshold temperature (e.g. $T=200$) will cause an explosion in the plant when trying to close the valve which results in the negative reward of $-1000$. Staying below this temperature is safe and will produce electricity an gain the reward of $100$. The reward of an open valve is $-1$.
\begin{align}
R(t_s,a) = 
\begin{cases}
 (a=open) &: -1 \\
(a \neq open)\wedge (t_s>T) &: -1000 \\
(a \neq open)\wedge \neg(t_s>T) &: 100 \\
\end{cases}\nonumber
\end{align}
This example will try to maximize its reward using observations on the temperature. The observations recieved before closing the plant to a safe temperature are therefore critical.

%write about policy in POMDPs
\section{DC-POMDP solution}
The DC-POMDP will choose actions in order to compute an optimal plan over time. This plan is the policy $\pi$ of the agent which depends on the history of actions and observations over a Markovian state. In the partially observable case where there is uncertainty on the states, the policy is a function over a continuous set of probability distributions over $|\mathcal{S}|$. Observations don't provide a unique state therefore a probability of the state is stored as the belief state $b$.

 At any of the horizon steps, a DC-POMDP policy $\pi(b)$ is defined as the iterative plan on executing an action and observing all probable outcomes of the set of observations from the $(h-1)$-step. The belief state is updated after this according to the Bayes rule: $\sum_{d_s}\int_{x_s} b(xd_s)T(xd'_s,xd_s,a)Z(xd'_s,xd_o,a)$  where for simiplicity we use ($xd_s = x_s,d_s$) and ($xd_o=x_o,d_o$).
The optimal policy is defined over the belief state as the sum of the expected discounted rewards over horizon $h$ starting with belief $b_0$.
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = E_{\pi^*} \left[ \sum_{h=0}^{H} \gamma^t \cdot r_h \Big| \vec{b}_0 = \vec{b} \right]\nonumber
\end{equation}
The next section provides the mathematics required to solve this equation in DC-POMDPs.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 1 - policy tree
\begin{figure}[t!]
%\vspace{-1mm}
\begin{center}
\includegraphics[height=4cm]{cond_plan2}
\end{center}
\vspace{-2mm}
\caption{\footnotesize Example conditional plan $\beta^h$ for discrete observations.}
\label{fig:cond_plan}
\vspace{-1mm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Value Iteration solution} 
%solutions with belief
Solving the value function for the optimal policy is performed using Dynamic programming algorithms such as value iteration. 
The decision at each time step is to pick an action based on past actions and observations. The possible set of conditional plans $\beta^h$ correspond to a decision tree. Figure 1 shows the decision tree for the simple case of three horizons with discrete observations. It represents a conditional plan consisting of an action and two possible observation strategies. For continuous observations, conditional plans are decision trees with infinitely many branches. In our solution we show that instead of a policy tree with infinite branches, we can create partitions based on relevant observation values, resulting in a policy tree similar to that of figure 1. We prove that a finer grain partitioning of the observation space has the same properties of the minimal partition set. From here on we assume this finite set of observation partitioning as $\mathcal{O}$.

The value iteration algorithm finds the optimal value for each horizon starting with an initial belief. Although the number of belief states is infinite, the optimal value function for $h$ is proven to be a piecewise linear and convex function of the belief state \cite{smallwoodSondik}. The optimal value function is the maximum value over a finite set of alpha vectors  $\alpha^h_i$:%that define the hyperplanes:
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = \max_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot \vec{b}
\end{equation}
where the belief states $\vec{b}$ and each $\alpha^h_i \in \Gamma^h$ are of dimension $|\mathcal{S}|$. At each of the $h$-steps the value function is parameterized by the set of  $\alpha$-vectors that partition the belief state. 
At a particular belief point, we can compute the next value function: 
\begin{align*}
V^{h+1}(b) = \max_a \left( b \cdot r_a + \gamma \sum_o p(xd_o|a,b)V^h(b|xd_o,a)\right)\nonumber
\end{align*}
Replacing the previous iteration value with (3) and also updating the belief after executing action $a$ and observing $xd_o$ results in the following equation: 
\begin{align*}
V^{h+1}(b) = \max_a \left( b \cdot r_a + 
\gamma \sum_o \max_{\langle g^h_{a,o,j} \rangle_j}  b \cdot g^h_{a,o,j} \right)\nonumber
\end{align*}
where $r_a = r(x_s,d_s,a)$ and  
\begin{align}
g^h_{a,o,j} =  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s)
\end{align}

Calculating all possible ways of building the value function independent of a certain belief state can also define the optimal value function. To derive the $\Gamma$-set of all possible $\alpha$-vectors for each horizon, we use the value iteration algorithm from (Monahan)~\cite{monahan82}. 
Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (4):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as 
$\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
\textrm{\large $\boxplus$} Q = \{ \vec{p} + \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}
\begin{align}
\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \\
\Gamma^h  &= \bigcup_a \Gamma^h_a 
\end{align}
Each action-dependent $\Gamma$ sums over the possible set of $g^h_{a,o,j}(\cdot)$ given a relevant observation partition $(xd_o)$ and for all $\alpha^{h-1}_{j} \in \Gamma^{h-1} $. The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.  
For each horizon, the optimal $\alpha$-vector is obtained from the maximum over all $\alpha$-vectors: 
\begin{align}
\alpha^* = \argmax_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot b \nonumber
\end{align}

For discrete observations, the cross-sum operation takes the set of $\alpha$ -vectors for each observation and produces a sum over vectors that depends on the number of previous iteration vectors. Thus the $\alpha$-vector set grows exponentially in the number of observations, i.e., $|\Gamma^{h}| = |\mathcal{A}|
|\Gamma^{h-1}|^{|\mathcal{O}|}$ during the cross-sum operation. For this reason even for a small number of discrete observations, tractable solutions may be hard to define. Here we do not restrict our observations to discrete ones, but extend it to a fully continuous approach where continuous observations and states are allowed in the DC-POMDP. We use the number of  observation partitions $|\mathcal{O}|$ as the branching factor in the policy tree. We present the symbolic value iteration algorithm and approaches to solving this problem in the following section.

\section{Symbolic Dynamic Programming} 
Dynamic programming provides exact solutions to POMDP problems for small domains. As the domain grows, fully enumerating the state, action and observation space leads to intractable solutions even for low horizons. Avoiding this enumeration often requires approximating the solutions. We claim an exact solution using symbolic dynamic programming. The SDP solution for fully observable domains has been presented in \cite{sanner_uai11} %using piecewise constant functions
Our symbolic algorithm deals with continuous observations and states by generating the relevant observation partition at every time step and performing the backup operation using $\alpha$-vectors. We define the requirements to implement the value iteration for DC-POMDP.

\subsection{Case Representation and Extended ADDs}
% operations, max, restrict, substitute
%overview + example plant
The \textsc{Power Plant} example represented all functions using the case structure which can be generally defined as:
{%\footnotesize 
\begin{align}
f = 
\begin{cases}
  \phi_1: & f_1 \\ 
 \vdots&\vdots\\ 
  \phi_k: & f_k \\ 
\end{cases} \nonumber
\end{align}
}
Where $\phi_i$ are disjoint logical formulae defined over the state $(\vec{d},\vec{x})$ with logical ($\land,\lor,\neg$) combinations of boolean variables and inequalities ($\geq,>,\leq,<$) over continuous variables.  
The $f_i$ are function definitions either linear or quadratic in the continuous parameters.

For \emph{unary operations} such as scalar multiplication $c\cdot f$ (for some constant $c \in \mathbb{R}$) or negation $-f$ on case statements is simply to apply the operation on each case partition $f_i$ ($1 \leq i \leq k$).

A \emph{binary operation} on two case statements, takes the cross-product of the logical partitions of each case statement and performs the corresponding operation on the resulting paired partitions.  The cross-sum $\oplus$ of two cases is defined as the following:

{\footnotesize 
\begin{center}
\begin{tabular}{r c c c l}
&
\hspace{-6mm} 
  $\begin{cases}
    \phi_1: & f_1 \\ 
    \phi_2: & f_2 \\ 
  \end{cases}$
$\oplus$
&
\hspace{-4mm}
  $\begin{cases}
    \psi_1: & g_1 \\ 
    \psi_2: & g_2 \\ 
  \end{cases}$
&
\hspace{-2mm} 
$ = $
&
\hspace{-2mm}
  $\begin{cases}
  \phi_1 \wedge \psi_1: & f_1 + g_1 \\ 
  \phi_1 \wedge \psi_2: & f_1 + g_2 \\ 
  \phi_2 \wedge \psi_1: & f_2 + g_1 \\ 
  \phi_2 \wedge \psi_2: & f_2 + g_2 \\ 
  \end{cases}$
\end{tabular}
\end{center}
}
\normalsize

Likewise $\ominus$ and $\otimes$ are defined by subtracting or multiplying partition values.  Inconsistent partitions can be discarded when they are irrelevant to the function value.
% Do I need maximization for VI? 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Define V^2 here?
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.4\textwidth]{v1.pdf}
\end{center}
\vspace{-3mm}
\caption{%\footnotesize 
The optimal value function for \textsc{Power Plant}
as a decision diagram: 
the \emph{true} branch is solid, the \emph{false}
branch is dashed.} 
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%xadd representation
The data structure of \emph{extended ADD} (XADD)~\cite{sanner_uai11} is used to support
case statements and the required operations.  Here we use continuous SDP for both states and observations in DC-POMDPs. The value functions and $\alpha$-vectors partition into states of equal value, both of which can be represented using XADDs. For increasing efficiency linear constraint feasibility checkers such as LP solvers allow pruning unreachable paths in the XADD.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{substitute}{Substitute}

\Begin
{
	$\textsc{O}:=\emptyset$ \;	
	\For {all $\alpha_j^{'h}$ in  $\Gamma^h$}    
	{
    	\emph{Relevant state partitions $\phi_j^{'h}$ of $\alpha_j^{'h}$}\\   
   	 	\For {all $xd'_{s_i}$ in $\phi_j^{'h}$}
   		 {
    		$\phi_{xd_s} := $ $\phi\{xd_{s_i}\setminus xd_{o_i}$\}\ \;
    		$\alpha_{j,o_i}^{h} := P(xd_{o_i}|xd'_{s_i},a) * \phi_{xd_s}$;\
    	}
    	$\textsc{O}^h := \oplus \alpha_{j,o}^{'h}$\;
    }
    \Return{$\textsc{O}^{h}$} \;
}
\caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a$) $\longrightarrow$ $\textsc{O}^{h}$ }
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\Begin{

    \emph{// Continuous regression, marginal integration}\\
    \For {all $x_{s_j}'$ in $\alpha_{j,o}^{'h}$}
    {
         $G_a^h := \int \textsc{O}^{h} \otimes P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a) \, d_{x_{s_j}'}$\;
    }
    \emph{// Discrete regression, marginal summation}\\
    \For {all $d_{s_i}'$ in $\alpha_{j,o}^{'h}$}
    {
         $G_a^h := \left[ G_a^h \otimes P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \right]|_{d_{s_i}' = 1}$\\
         \hspace{8mm} $\oplus \left[ G_a^h \otimes P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \right]|_{d_{s_i}'= 0}$\;
    }
    \Return{$G_a^h$} \;
}
\caption{\footnotesize \texttt{Backup}($\alpha_{j,o}^{'h},a,$) $\longrightarrow$ $G_a^h$ \label{alg:regress}}
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\incmargin{.5em}
%\linesnumbered
%\begin{algorithm}[t!]
%\vspace{-.5mm}
%\dontprintsemicolon
%\Begin{
%	\emph{any function f has $i$ partitions like $\phi_i:f_i$}\;	
%	\emph{compute UB and LB from $\phi_i$ using constraints on $var$}\;
%    $I=$ \emph{ any $var$-independent $\phi_i$} \;
%    $F = $ \emph{differentiate $f_i$}\;
%    $F = I \otimes [F(UB) - F(LB)]$\;
%    \Return{$F$} \;
%}
%\caption{\footnotesize \texttt{VE}($var,f$) $\longrightarrow$ $F$ }
%\vspace{-1mm}
%\end{algorithm}
%\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Continuous Observation Space} 
%progression definition
The main operations required for value iteration in DC-POMDP is generating relevant observations and perform the bellman backup operation. 
%Progression is the inverse of regression. The \emph{regression} of a formula $\psi$ through an action $a$ is a formula $\psi'$ that holds prior to $a$ being performed iff $\psi$ holds after $a$. \emph{Progression} takes a formula $\psi$ and an action $o$ and represents the formula $\psi'$ that would hold after action $o$ is performed in a situation satisfying $\psi$.

The \texttt{GenerateRelObs} algorithm defines the operation of generating relevant observation partitions. Starting with an empty set of relevant observation partitions $\textsc{O}^{h}$, the algorithm iterates through all possible $\alpha$-vectors of the previous horizon. For each of the vectors, we use the state partitions of that $\alpha$-vector where the logical formulas $\phi_j^{'h}$ of the cases are defined as the state partitions.

Lines 5--7 iterate on all the state variables (both discrete and continuous) of a certain state partition $\phi_j^{'h}$. In line 6 the state variables are substituted with the observation variables based on the observation model provided. In the next line the probability attached to the observation given this state and the input action is multiplied in the relevant observation partition to produce a single case partition with a logical formula depending on observations and a probability based on the observation model.
In other words, lines 5-7 take a state partition and return an observation partition that is only relevant to the states according to $\mathcal{Z} : \mathcal{S} \times \mathcal{A} \times \mathcal{O} \rightarrow [ 0, 1 ]$. 
Finally for every specific $\alpha$-vector, the relevant observation partitions are added to the set of all relevant observations $\textsc{O}^{h}$ regardless of the vector they came from. The operator $\oplus$ only performs the operation on the logical formulas $\phi_i$ of the case partitions,without actually adding the function values.

To illustrate the algorithm with an example, we assume a discrete %and continuous 
noise distribution model such as (1). We explain each step of the algorithm using the \texttt{Power Plant} example. Assume that after $h=1$ the set of $\alpha$-vectors are defined as:
\begin{align}
\alpha_1^1(s) &= 
\begin{cases}
 (t_s>200) &: -1000 \\
\neg(t_s>200) &: 100 \\
\end{cases}
\alpha_2^1(s) &= \top &: -1
\end{align}
The relevant state partition only considers the logical formulas: $\{t_s>200, t_s<200\}$. For horizon $h=2$ and specific action $close$ we derive the relevant set of observations for both $\alpha_1^1$ and $\alpha_2^1$. The discrete noise observation model (1) is used for the substitution step where $t_o$ is substituted with $t_s$ most of the times(i.e, with probability 0.9) and $2/3t_o=t_s$ with probability of 0.1. By replacing these inside the state partitions, the relevant observation partitions are formed: 
\begin{align*}
\begin{tabular}{r c c c l}
&
\hspace{-6mm} 
$\alpha_1^1(t_s'>200) $&= 
$\begin{cases}
  t_o>200 \\
   t_o>300  \\
 \end{cases}$
 \\
 &
\hspace{-6mm} 
$\alpha_1^1(t_s'<200) $ &= 
$\begin{cases}
 t_o<200 \\
 t_o<300 \\
\end{cases}$
&
\hspace{-4mm} 
$\alpha_2^1(t_s'>200) = \top$
\end{tabular}
\end{align*}
The probabilities of each partition is then multiplied according to the noise model and the two vectors are added to the relevant observation set $\textsc{O}^{h}$. In our example $\alpha_2^1$ will leave no effect in the final set proposing a partitioning on the observation space depending on the relevant state space partition. The probability of receiving observations between ($200,300$),given that the next state has high temperatures (higher than $200$) is $0.9$. We name the three partitions as the observation partitions of $O_1,O_2,O_3$:

\begin{align*}
O_1:P(200<t_o<300 |t_s',close) = 
\begin{cases}
 % 2*t_o>200 &: 0.7 \\
 %2/3*t_o>200 &: 0.3 \\
% 2*t_o<200 &: 0.7 \\
 %2/3*t_o<200 &: 0.3 \\
  t_s'>200 &: 0.9 \\
\neg(t_s'>200) &: 0.1 \\
 \end{cases}
 \\
 \hspace{-4mm} 
O_2:P(t_o>300|t_s',close) = 
\begin{cases}
 t_s'>200 &: 0.1 \\
 \neg(t_s'>200) &: 0.0 \\
 \end{cases}
\\
\hspace{-4mm} 
O_3:P(t_o<200|t_s',close) = 
\begin{cases}
 t_s'<200 &: 0.9 \\
 \neg(t_s'>200) &: 0.0 \\
\end{cases}
\end{align*}
%For continuous noise models such as (2) lines 8--11 is used. To progress a continuous observation from a continuous next state according to noise, we use a symbolic variable elimination inference approach %mentioned in \cite{sve}. 
%To compute  $\int_{x_1=-\infty}^{\infty} f \, dx_1$ for some case-statement function $f$ we define the the following: 
%\vspace{-2mm}
%{\footnotesize
%\begin{equation}
%\int_{x_1=-\infty}^{\infty} \sum_i \mathbb{I}[\phi_i] \cdot f_i \, dx_1 \; = \; \sum_i \int_{x_1=-\infty}^{\infty} \mathbb{I}[\phi_i] \cdot f_i \, dx_1 \nonumber
%\vspace{-1mm}
%\end{equation}
%}
%where $\mathbb{I}[\cdot]$ is an indicator function taking value 1 when its argument is true, 0 when it is false.  We can compute the integrals separately for each case partition and $\sum$ the results using the case sum $\oplus$.
%
%For any single case partition, the integral is computed by a number of steps defined in the VE algorithm. For each case in $\alpha^h_j$ and for each state variable that is to be eliminated $x_{s_i}$, we perform a step of the VE algorithm. First the upper and lower bounds of the logical formulas of $\phi_{i,j} \in \alpha^h_j$ are computed. Any value less than ($<$ , $\leq$) of $x_{s_i}$ is considered a lower bound (LB) and any value greater is an upper bound. For the noise model of (2) and according to the first case statement of $\alpha^1_1$ the LB and UB are defined as below: $LB := max \left\langle 200,t_o - 5 \right\rangle , UB:=\left\langle t_o + 5\right\rangle$
%
%If a certain $\phi_{i,j}$ does not constrain on $x_{s_i}$, then this is added to an indicator function  $\mathbb{I}$. After evaluating the integral of the function $f_{i,j}$, the result is obtained by substituting the bounds and multiplying by the independent constraints. 
%\vspace{-1mm}
%\begin{equation*}
%\mathbb{I}[0.1] \int_{x_{s_i}=\mathit{LB}}^{\mathit{UB}} (-1000) \, dx_{s_i} = 0.1  \left[ -1000 * x_{s_i} \bigg|_{x_{s_i}=\mathit{LB}}^{x_1=\mathit{UB}} \right]
%\end{equation*}
%
%The case statement of the observation probability is defined over all partitions of $\alpha^h_j$: 
%\begin{align*}
%P(t_o|t_s'>200,close) &=
%\begin{cases}
% % 2*t_o>200 &: 0.7 \\
% %2/3*t_o>200 &: 0.3 \\
%% 2*t_o<200 &: 0.7 \\
% %2/3*t_o<200 &: 0.3 \\
%  t_o>205 &: -10,000 \\
% t_o<205 &: -1195 - 1000*t_o \\
% \end{cases}
% \\
% P(t_o|t_s'<200,close) &= 
%\begin{cases}
% t_o<195 &: 305 - 100*t_o \\
% t_o>195 &: 1000 \\
%\end{cases}
%\end{align*}

% regression in value iteration
Backup of an observation-weighted transition function where the weights are determined according to the relevant observation partitions is explained in the \texttt{Regress}. It is composed of two main parts:

\emph{Boolean restriction} is defined in lines 5--8 where $f|_{b=v}$ restricts a boolean variable $b$ occurring in $f$ to values $v \in \{ 0,1 \}$. This can be done by simply appending $b=v$ to each case partition (logical $\wedge$ of $b=v$ and the logical first-order formulas of $f$).

\emph{Continuous regression} $\int f(x_{s_j}') \otimes P(x_{s_j}'|\cdots) dx_{s_j}'$ in line 3--4 has the form $\delta[x_{s_j}' - h(\vec{z})]$ for the probability over the next state, where $h(\vec{z})$ is a case statement and $\vec{z}$ does not contain
$x_{s_j}'$.  A symbolic substitution is defined for this integration such that any occurrence of $x_{s_j}'$ in $f(x_{s_j}')$ is substituted with $h(\vec{z})$: 
$\int f(x_{s_j}') \otimes \delta[x_{s_j}' - h(\vec{z})] dx_{s_j}' = f(x'_j) \{ x_{s_j}' / h(\vec{z}) \}$

In our example the $\alpha$-vectors of the previous horizon use the transition probability to produce the substituted case partitions for all continuous state variables $x'_{s_i}$:

\begin{align}
\hspace{-4mm} 
G_{close}^2= 
\begin{cases}
(t_s'>235) &: -1000 \\
\neg(t_s'>235) &: 100 \\
\end{cases}
\end{align}
\begin{align*}
G_{open}^2=  \top &: -1
\end{align*}

\subsection{Value iteration for DC-POMDPs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{backup}{Backup}
\SetKwFunction{genObs}{GenRelObs}
\SetKwFunction{prune}{Prune}
\SetKwFunction{remapWithPrimes}{Prime}
\Begin
{
   $V^0:=0, h:=0$\;
   \While{$h < H$}
   {
       $h:=h+1, \Gamma^h :=\emptyset$\;
       \ForEach {$a \in A$}
       {
			$\Gamma_{a}^h :=\emptyset$ \;       		
       		$\textsc{O}^{h-1} \,:=\,$ \genObs{$\Gamma^{h-1},a$}\;
       		 \ForEach {$o_i \in \textsc{O}^{h-1}$}
       		 {
				\ForEach {$\alpha_j^{h-1} \in \Gamma^{h-1}$}
       			{
   	 		  		$\alpha_j^{h'}=$ \remapWithPrimes{$\alpha_j^{h}$} \;
   	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
   	 		    	$\Gamma_{a,o_i,j}^h \,:=\, o_i \otimes$ \backup{$\alpha_j^{h'},a$}\;
       	      	}
       	      	$\Gamma_{a,o}^h\,:=\,\textrm{\large $\boxplus$} \Gamma_{a,o_i}^h$\;
       	     }
           $\Gamma_a^{h} \,:=\,R_a \oplus \gamma \cdot \Gamma_{a,o}^h$\;
            $\Gamma^{h} \,:=\, \Gamma^{h} \cup \Gamma_a^{h}$\;
        }  
        	%monahan's pruning first generates all vectors, then prunes
              $V^{h} \,:=\, $\prune{$\Gamma^h$} \;
              $\pi^{*,h} \,:=\, \argmax_{a} \, \Gamma_a^{h}$\;

       \If{$V^h = V^{h-1}$}
           {break $\,$ \emph{// Terminate if early convergence}\;}
   }
     \Return{$(V^h,\pi^{*,h})$} \;
}
\caption{\footnotesize \texttt{VI}(DC-POMDP, $H$) $\longrightarrow$ $(V^h,\pi^{*,h})$ \label{alg:vi}}
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%overview + example plant
Having defined the two main algorithms to produce the relevant observation partitions and the backup operation, we now explain the value iteration algorithm as defined in \texttt{VI} for DC-POMDPs. Each step of the algorithm is explained using the \textsc{Power plant} example for the first three horizons.

For $h=0$, $V^0$ is given in line 2.  For $h=1$ we choose an action, for example $close$ to start with. The $\alpha$-vector set from $h=0$ is empty so line 7 of generating the relevant observation partitions is empty. Also priming the vectors and obtaining the product of relevant observations and the backup is an empty set. In line 15 the reward $R(s,close)$ is added to obtain $\alpha^1_{1,close}$ as defined by (8). The same operation is repeated for the $open$ action where $R(s,open)=-1$. Line 16 defines the final $\Gamma$-set: 
$\Gamma^1 = \{ \alpha^1_{1,open}, \alpha^1_{1,close} \}$.

For $h=2$, for each of the discrete actions, first we generate the relevant set of observations. This has been done for the $close$ action in the previous section. Next for each of the relevant partitions and both $\alpha$-vectors of the previous $\Gamma$ set lines 8--15 are executed.

Line 10 will substitute prime values for all state variables. Line 12 produces the backup $\alpha$-vector and multiplies it by the weight derived from the relevant observation partition. For each observation, two vectors are formed for each corresponding $\alpha$-vector. For example line 12 for observations $O_2,O_3$ and $\alpha_1^1,\alpha_2^1$ is defined as:

\begin{align*}
%\begin{cases}
%\Gamma_{close,O_1,1}^2= 
%\begin{cases}
%(200<t_o<300) \wedge (t_s'>235) &: -900 \\
%(200<t_o<300)\wedge (200<t_s'<235) &: 90 \\
%(200<t_o<300) \wedge (t_s'<200) &: -10 \\
%\end{cases}
%\\
%\Gamma_{close,O_1,2}^2= 
%\begin{cases}
%(200<t_o<300) \wedge (t_s'>200) &: -0.9 \\
%(200<t_o<300) \wedge (t_s'<200) &: -0.1 \\
%\end{cases}
%\end{cases}
%\\
\begin{cases}
\Gamma_{close,O_2,1}^2= 
\begin{cases}
(t_o>300) \wedge (t_s'>235) &: -100 \\
(t_o>300)\wedge (200<t_s'<235) &: 10 \\
(t_o>300) \wedge (t_s'<200) &: 0 \\
\end{cases}
\\
\Gamma_{close,O_2,2}^2= 
\begin{cases}
(t_o>300)\wedge (t_s'>200) &: -0.1 \\
(t_o>300) \wedge (t_s'<200) &: 0 \\
\end{cases}
\end{cases}
\\
\begin{cases}
\Gamma_{close,O_3,1}^2= 
\begin{cases}
(t_o<200) \wedge (t_s'>235) &: -900 \\
(t_o<200)\wedge (200<t_s'<235) &: 90 \\
(t_o<200) \wedge (t_s'<200) &: 0 \\
\end{cases}
\\
\Gamma_{close,O_3,2}^2= 
\begin{cases}
(t_o<200) \wedge (t_s'>200) &: -0.9 \\
(t_o<200)\wedge (t_s'<200) &: 0 \\
\end{cases}
\end{cases}
\end{align*}

Next in line 13 we have to take the cross-sum of the two sets which takes every possible combination of observations according to the $\alpha$-vectors. This means for $O_2 \textrm{\large $\boxplus$} O_3$ we have the following vectors:

\begin{align*}
\begin{cases}
\begin{cases}
(t_s'>235) &: -1000 \\
(200<t_s'<235) &: 100 \\
(t_s'<200) &: 0 \\
\end{cases}
\\
\begin{cases}
(t_s'>235) &: -100.9 \\
(<200<t_s'<235) &: 9.1 \\
(t_s'<200) &: 0 \\
\end{cases}
\\
\begin{cases}
 (t_s'>235) &: -900 \\
 (200<t_s'<235) &: 89.9 \\
 (t_s'<200) &: 0 \\
\end{cases}
\\
\begin{cases}
(t_s'>200) &: -1 \\
(t_s'<200) &: 0 \\
\end{cases}
\end{cases}
\end{align*}

Where the observation partitions ($t_o>300 , t_o<200$) are inconsistent and therefore omitted from the result. These 4 vectors are added with the two vectors from $\Gamma_{close,O_1}^2$ to get $2^3$ different vectors. All 8 vectors are multiplied by the discount $\gamma$ and added with the reward in line 14 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
% need to add this?
The same operation is performed for the $open$ action and 8 different vectors are formed here too. In line 15 the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors (16 in total) for the next horizon.

In line 16 this final set is using in the \texttt{Prune} algorithm. For all $\alpha$-vectors in the current $\Gamma$-set first an LP-solver is used to prune any inconsistency in the branches of this vector. Next a pairwise dominance test is performed where for all $j,k$: 
$\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$
The result of this pairwise test is a dominant $\alpha$-vector which is added to the set of $V^h$. As an example from the set $\Gamma^2_{open,close}$ the two $\alpha$-vectors of $\alpha_{1,open}(\vec{s}) \geq \alpha_{1,close}(\vec{s})$ are tested: 
\begin{equation*}
(t_s>200) \wedge (200<t_o<300) \geq (t_s>150) \wedge (220<t_o<280)
\end{equation*}
The second $\alpha$-vector is omitted from the overall answer since it is dominated by $\alpha_{1,open}$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\SetKwFunction{lpsolve}{LP-Solve}
\SetKwFunction{dominancy}{Dominance}
\Begin
{
   		$V^h :=0$ \;
       \ForEach {$\alpha_j^h \in \Gamma^h$}
       {
           $\alpha_j^h \,:=\,$\lpsolve{$\alpha_j^h$} \;
           \ForEach {$\alpha_k^h \in \Gamma^h$}
           {
				\If {$j \neq k$}  
				{         
           			$V^h \,:=\,\oplus$ \dominancy{$\alpha_j^h,\alpha_k^h$} \; 
           		}  
           	}
       }
     \Return{$(V^h)$} \;
}
\caption{\footnotesize \texttt{Prune}($\Gamma^h$) $\longrightarrow$ $(V^h)$ }
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Empirical Results}

\section{Related Work}

\section{Conclusion}

\subsubsection*{References} 
\bibliography{dcpomdp}
\bibliographystyle{plain}

\end{document}

### 2012-03-15 cpomdp/camdp-xadd/src/CPOMDP/ComputeGamma.java

@@ -1,14 +1,22 @@
 package CPOMDP;
 
+import java.util.ArrayList;
+import java.util.Arrays;
 import java.util.HashMap;
 import java.util.HashSet;
 import java.util.Map;
 
 import util.IntTriple;
 import xadd.XADD;
+import xadd.XADD.ArithExpr;
 import xadd.XADD.BoolDec;
 import xadd.XADD.DeltaFunctionSubstitution;
+import xadd.XADD.DoubleExpr;
+import xadd.XADD.OperExpr;
+import xadd.XADD.VarExpr;
+import xadd.XADD.XADDINode;
 import xadd.XADD.XADDNode;
+import xadd.XADD.XADDTNode;
 
 public class ComputeGamma {
 
@@ -23,6 +31,8 @@ public class ComputeGamma {
 	}
 	public int computeGamma(COAction a, HashMap<Integer, Integer> _previousgammaSet_h) {
 
+		//first generate relevant observation partitions
+		generateRelObs(a,_previousgammaSet_h);
 		// for each of the alpha-vectors in the previous gammaset
 		int crossSum = _context.getTermNode(XADD.ZERO);
 		for (Map.Entry<Integer,Integer> j : _previousgammaSet_h.entrySet())
@@ -65,7 +75,7 @@ public class ComputeGamma {
 				
 				// Cache result
 				_pomdp._logStream.println("-->: " + _context.getString(q));
-				//_pomdp._hmContRegrCache.put(new IntTriple(_contRegrKey), q);
+				_pomdp._hmContRegrCache.put(new IntTriple(_contRegrKey), q);
 			}
 			
 			// Regress boolean variables second
@@ -90,7 +100,7 @@ public class ComputeGamma {
 			}
 			
 			//finished regress, for each G_{a,j}^h add to the cross-sum result
-			crossSum = _context.apply(crossSum,q, _context.SUM);
+			crossSum = _context.apply(crossSum, q, _context.SUM);
 			
 		}
 		crossSum = _context.apply(a._reward,  
@@ -101,12 +111,107 @@ public class ComputeGamma {
 		for (Integer constraint : _pomdp._alConstraints)
 			crossSum = _context.apply(crossSum, constraint, XADD.PROD);
 		if (_pomdp._alConstraints.size() > 0)
-			//crossSum = _context.reduceLP(crossSum);
+			crossSum = _context.reduceLP(crossSum,_pomdp._alContAllVars);
 		
 		// Optional Display
[…]

### 2012-03-15 trunk/cpomdp/camdp-xadd/src/CPOMDP/CPOMDP.java

@@ -101,7 +101,7 @@ public class CPOMDP {
 		parser.buildPOMDP(input);
 		_context._hmMaxVal = parser._maxCVal;
 		_context._hmMinVal = parser._minCVal;
-		//_context._alBooleanVars = parser.getBVars();
+		_context._hsBooleanVars = parser.getBVars();
 
 		_alConstraints = parser.getConstraints();
 		_nMaxIter = parser.getIterations();
@@ -196,7 +196,7 @@ public class CPOMDP {
 				int regr = _gammaHelper.computeGamma(me.getValue(),_previousgammaSet_h);
 				//the result is Gamma^h_a, add this to the current set of Gamma^h_a
 				regr = _context.reduceLinearize(regr);
-				//regr = _context.reduceLP(regr);
+				regr = _context.reduceLP(regr,_alContAllVars);
 				
 				_currentgammaSet_h.put(counter, regr);
 				
@@ -300,7 +300,7 @@ public class CPOMDP {
 		sb.append("CVars:       " +	_alContAllVars + " = " + _hsContSVars + " + " + _hsContOVars + "\n");
 		sb.append("Min-values:  " + _context._hmMinVal + "\n");
 		sb.append("Max-values:  " + _context._hmMaxVal + "\n");
-		//sb.append("BVars:       " + _context._alBooleanVars + "/" + _hsBoolSVars + "\n");
+		sb.append("BVars:       " + _context._hsBooleanVars + "/" + _hsBoolSVars + "\n");
 		sb.append("BOVars:       " + _hsBoolOVars + "\n");
 		sb.append("OVars:       " + _hsContOVars + "\n");
 		sb.append("Order:       " + _context._alOrder + "\n");

### 2012-03-16 cpomdp/camdp-xadd/src/CPOMDP/ComputeGamma.java

@@ -6,12 +6,16 @@ import java.util.HashMap;
 import java.util.HashSet;
 import java.util.Map;
 
+import org.omg.stub.java.rmi._Remote_Stub;
+
 import util.IntTriple;
 import xadd.XADD;
 import xadd.XADD.ArithExpr;
 import xadd.XADD.BoolDec;
+import xadd.XADD.Decision;
 import xadd.XADD.DeltaFunctionSubstitution;
 import xadd.XADD.DoubleExpr;
+import xadd.XADD.ExprDec;
 import xadd.XADD.OperExpr;
 import xadd.XADD.VarExpr;
 import xadd.XADD.XADDINode;
@@ -23,7 +27,7 @@ public class ComputeGamma {
 	XADD _context = null;
 	CPOMDP _pomdp = null;
 	private IntTriple _contRegrKey = new IntTriple(-1,-1,-1);
-
+	HashMap<Integer,PartitionObsState> _obspartitionset= new HashMap<Integer,PartitionObsState>();
 	public ComputeGamma(XADD xadd,CPOMDP pomdp)
 	{
 		_context = xadd;
@@ -121,8 +125,9 @@ public class ComputeGamma {
 	
 	private void generateRelObs(COAction a, HashMap<Integer, Integer> _previousgammaSet_h) 
 	{
-
-		HashMap<String,ArithExpr> obs_subs = new HashMap<String,ArithExpr>();
+		//first find the set of observation substitutes for state variables 
+		HashMap<String,ArithExpr> obs_subs = null;
+		boolean low = true;
 		for (int i=0;i<a._contObs.size();i++)
 		{
 			//first find the obs_subs that are to be substituted in each alpha-vector. 
@@ -134,36 +139,135 @@ public class ComputeGamma {
 				//either a TNode which means no noise, or INODE where there is noise
 				if (obsnode instanceof XADDTNode)
 				{
+					obs_subs = new HashMap<String,ArithExpr>();
 					obs_var = getVarCoeff(obs_var,((XADDTNode) obsnode)._expr);
 					obs_subs.put(a._contState.get(i), obs_var);
 				}
 				else if (obsnode instanceof XADDINode)
 				{
+					obs_subs = new HashMap<String,ArithExpr>();
 					XADDTNode high = (XADDTNode) _context.getNode(((XADDINode) obsnode)._high);
 					obs_var = getVarCoeff(obs_var,high._expr);
 					obs_var.makeCanonical();
[…]

### 2012-03-16 cpomdp/camdp-xadd/src/CPOMDP/PartitionObsState.java

@@ -1,0 +1,38 @@
+package CPOMDP;
+
+import java.util.HashMap;
+
+public class PartitionObsState {
+		HashMap <Integer, StateObsVector> _relObsProb ;
+		
+		public PartitionObsState()
+		{
+			_relObsProb = new HashMap<Integer, StateObsVector>(); 
+		}
+		
+		public void setnewPartition(StateObsVector so)
+		{
+			_relObsProb.put(_relObsProb.size(),so);
+		}
+	
+		public int sizePartitionObsState()
+		{
+			return _relObsProb.size();
+		}
+
+		public HashMap<Integer, StateObsVector> get_relObsProb() {
+			return _relObsProb;
+		}
+
+		public void set_relObsProb(HashMap<Integer, StateObsVector> _relObsProb) {
+			this._relObsProb = _relObsProb;
+		}
+		
+		public StateObsVector get_relObsProb(int j) {
+			return _relObsProb.get(j);
+		}
+
+		public void set_relObsProb(StateObsVector so) {
+			this._relObsProb.put(_relObsProb.size(), so);
+		}
+}

### 2012-03-16 trunk/cpomdp/camdp-xadd/src/CPOMDP/StateObsVector.java

@@ -1,0 +1,39 @@
+package CPOMDP;
+
+import xadd.XADD.ArithExpr;
+import xadd.XADD.CompExpr;
+
+public class StateObsVector {
+		Double probability;
+		CompExpr state;
+		CompExpr obs; 
+		
+		public StateObsVector(CompExpr s,CompExpr o,Double p)
+		{
+			state = s;
+			obs = o;
+			probability = p;
+		}
+		public CompExpr getState() {
+			return state;
+		}
+
+		public void setState(CompExpr state) {
+			this.state = state;
+		}
+
+		public CompExpr getObs() {
+			return obs;
+		}
+
+		public void setObs(CompExpr obs) {
+			this.obs = obs;
+		}
+		public Double getProbability() {
+			return probability;
+		}
+
+		public void setProbability(Double probability) {
+			this.probability = probability;
+		}
+}

### 2012-05-15 cpomdp/camdp-xadd/src/CPOMDP/cpomdp.java

@@ -3,10 +3,13 @@ package CPOMDP;
 import graph.Graph;
 
 import java.io.BufferedReader;
+import java.io.BufferedWriter;
 import java.io.File;
 import java.io.FileNotFoundException;
 import java.io.FileOutputStream;
 import java.io.FileReader;
+import java.io.FileWriter;
+import java.io.IOException;
 import java.io.PrintStream;
 import java.math.BigDecimal;
 import java.text.DecimalFormat;
@@ -46,14 +49,14 @@ public class cpomdp {
 	/* For printing */
 	public static DecimalFormat _df = new DecimalFormat("#.###");
 	public static PrintStream _logStream = null;
-
+	public static int filecounter = 0;
 	/* Static variables */
 	public static long _lTime; // For timing purposes
 	public static Runtime RUNTIME = Runtime.getRuntime();
 
 	/* Local vars */
 	public boolean DISPLAY_2D = false;
-	public boolean DISPLAY_3D = false;
+	public boolean DISPLAY_3D = true;
 
 	public String _problemFile = null;
 	public String _logFileRoot = null;
@@ -177,7 +180,7 @@ public class cpomdp {
 		//define belief: 
 		//1 belief for now: 
 			ArrayList l0 =new ArrayList();
-			/*	l0.add("[-6 + t*1 <= 0]");
+				/*l0.add("[-6 + t*1 <= 0]");
 				ArrayList l0t = new ArrayList();
 				ArrayList l0f = new ArrayList();
 				l0t.add("[-2 + t*1 >= 0]");
@@ -322,9 +325,9 @@ public class cpomdp {
 				}
 
 				//pruning the _alpha vectors
-				_currentgammaSet_h = dominanceTest(_currentgammaSet_h);
+			//	_currentgammaSet_h = dominanceTest(_currentgammaSet_h);
 				
-				
+				_logStream.println("Number of Alpha vectors after pruning for belief "+ (i+1) + "is : " + _currentgammaSet_h.size());
 				//now for this belief, compute the alpha vector that will result in the maximum value
 				int alphaValue[] = new int[_currentgammaSet_h.size()];
 				double maximumAlpha = 0;
@@ -365,6 +368,8 @@ public class cpomdp {
 			{
 				_currentgammaSet_h.put(j, beliefBasedVectors[j]);
[…]

### 2012-05-15 trunk/cpomdp/camdp-xadd/src/CPOMDP/contObs/ContComputeGamma.java

@@ -1,5 +1,8 @@
 package CPOMDP.contObs;
 
+import java.io.FileNotFoundException;
+import java.io.FileOutputStream;
+import java.io.PrintStream;
 import java.util.ArrayList;
 import java.util.Arrays;
 import java.util.HashMap;
@@ -37,6 +40,9 @@ public class ContComputeGamma {
 	XADD _context = null;
 	cpomdp _pomdp = null;
 	private IntTriple _contRegrKey = new IntTriple(-1,-1,-1);
+	public String _logFileRoot;
+	public static PrintStream _logStream = null;
+
 	HashMap<Integer,ObsPartition> _obspartitionset= new HashMap<Integer,ObsPartition>();
 	ObsPartition partition = new ObsPartition();
 	int partitionNo =0;
@@ -46,6 +52,14 @@ public class ContComputeGamma {
 	{
 		_context = xadd;
 		_pomdp = pomdp;
+		_logFileRoot = _pomdp.InsertDirectory("src/CPOMDP/contObs/obsPartitions", _pomdp.RESULTS_DIR).replace("-", "_");
+		try {
+			_logStream = new PrintStream(new FileOutputStream(_logFileRoot + ".log"));
+		} catch (FileNotFoundException e) {
+			System.err.println(e);
+			e.printStackTrace();
+			System.exit(1);
+		}
 		
 	}
 	public int[] computeGamma(COAction a, HashMap<Integer, Integer> _previousgammaSet_h, int belief) {
@@ -131,13 +145,31 @@ public class ContComputeGamma {
 		//now we have the regressed alpha's and the probability of each observation partition. 
 		//for each observation, multipy in the alphas
 		//the state in the observation set are not in form of XADD, so need to convert them
-
-		int[][] regressedAlpha = new int[_obspartitionset.size()][newalphas.size()];
+		HashMap<Integer,ObsPartition> _newobspartitionset= new HashMap<Integer,ObsPartition>();
+		int counter=0;
 		for (int i = 0;i< _obspartitionset.size();i++)
 		{
-
+			
 			double d = _obspartitionset.get(i).getProbability();
-			if (d< 0.00000001) d = 0.0;
+			if (d< 0.05) 
+			{
+				d = 0.0;
+				//_obspartitionset.remove(i);
+			}
+			else 
+				{
+					_newobspartitionset.put(counter, _obspartitionset.get(i));
+					counter++;
+				}
+		}
[…]

### 2012-05-18 trunk/cpomdp/camdp-xadd/src/doc/NIPS/draft2-long version.tex

\documentclass{article} % For LaTeX2e
\usepackage{nips12submit_e,times}
%\documentstyle[nips12submit_09,times,art10]{article} % For LaTeX 2.09

\def\argmax{\operatornamewithlimits{arg\max}}

\title{Exact Symbolic Dynamic Programming for Continuous POMDPs}

\author{
%Zahra Zamani \\
%ANU & NICTA\\
%Canberra, Australia \\
%\texttt{zahra.zamani@anu.edu.au} \\
%\And
%Coauthor \\
%Affiliation \\
%Address \\
%\texttt{email} \\
%\AND
%Coauthor \\
%Affiliation \\
%Address \\
%\texttt{email} \\
%\And
%Coauthor \\
%Affiliation \\
%Address \\
%\texttt{email} \\
%\And
%Coauthor \\
%Affiliation \\
%Address \\
%\texttt{email} \\
%(if needed)\\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

%\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

\begin{abstract}
Decision-theoretic planning problems try to find the optimal sequence of actions for a given environment. Using continuous models in such problems is naturally closer to the real-world. Partially Observable MDPs model uncertain environments where the underlying state is not be fully observable.

While previous work have provided solutions for discrete states and/or observations, here we define the first exact solution to both continuous state and observations. This symbolic dynamic programming approach avoids enumerating the state and observation space providing solutions for our general discrete and continuous partially observable MDP. The main contribution of this paper is to use continuous states with discrete observations and extending it to continuous observations by defining relevant observation partitions associated with each state partition. This allows us to perform the Bellman backup operation efficiently. We address the problems in POMDP continuous planning and solutions to solve them symbolically. %To compactly represent the exact value functions Affine algebraic decision diagrams are used. 
\end{abstract}

\section{Introduction}
% Write intro here
In many complex systems complete information of the state is not available due to noisy or poor sensors and readings of them. Partially observable Markov Decision Processes (POMDPs)\cite{POMDPfirst} can be used for planning under such uncertainty in abstract problems. In many real-world problems, resources and sensor inputs are naturally modeled using continuous variables.

Unfortunately most POMDP techniques can not be applied to such settings without assuming prior knowledge about the problem domain. Most methods use some approximation to discretize the continuous state, observation or action space which increases computational complexity \cite{Thrun99h}. Exact solutions to a multi-variate continuous state and observation setting has not been approached for the curse of dimensionality problem. In assuming a continuous observation setting, the agent must search for all possible observations that are read from the environment. This can quickly become infeasible even for small problem sizes.

In this paper we provide an exact solution to a subset of POMDPs with continuous states and observations. Our symbolic dynamic programming (SDP) approach assumes discrete noise and piecewise linear dynamics and functions. We formulize our problems using a Discrete and Continuous POMDP (DC-POMDP) model and apply our SDP algorithm. The main problem that rises in any problem modeled using DC-POMDP is the definition of different observation partitions and their probability. We apply symbolic solutions such as integration and maximization to provide an optimal closed-form solution to our problem domains. 
As an example we consider the following power plant example which has mainly been solved in the literature using discrete sets of states and observations~\cite{steam2}. We use the example throughout the paper to help better understand the definitions.

\textbf{Example} \textsc{(Power Plant)}
\emph{The steam generation system of a power plant aims to provide hot steam to the steam turbine which in turn produces electricity. Feed-water is exposed to heat in the tubes leading to the water tank . Evaporating the water is performed under specific pressure and temperature inside the attached water tubes. 
Mixing water and steam pressures makes reading tank levels and temperatures very hard and uncertain. For this reason, the process is modeled using POMDP in the literature. The temperature is assumed to be a continuous variable that is not observed. Noisy observations of the temperature are used to decide on the action of closing or opening the valve.}

We need to extend the basic SDP framework from \cite{sanner_uai11} by adding continuous observations to the model. The solution to a closed-form value function is similar to the algorithm from \cite{monahan82} apart from the fact that no explicit observation is defined in the continuous case. We redefine the value-iteration algorithm using case representations that are implemented using Extended Algebric decision diagrams (XADD) and provide an efficient automatic way to partition the entire observation space into piecewise linear functions where each partition has a relative probability. We use the univariate \textsc{Power Plant} example to define the main operations and algorithms and also provide empirical results for a multi-variate version of the problem.

\section{DC-POMDP Model} 
%%%% no need to define DC-MDP
We assume familiarity with MDPs and introduce Discrete and Continuous Partially Observable MDPs as an extension to Discrete and Continuous State MDPs ~\cite{sanner_uai11}. A Discrete and Continuous partially observable MDP (DC-POMDP) is a tuple $\langle
\mathcal{S},\mathcal{A},\mathcal{O},\mathcal{T},\mathcal{R},\mathcal{Z},\gamma,h \rangle$. 
States are represented by vectors of $(\vec{d_s},\vec{x_s}) = ( d_{s_1},\ldots,d_{s_n},x_{s_1},\ldots,x_{s_m} )$ where each $d_{s_i} \in \{ 0,1 \}$ ($1 \leq i \leq n$) is a discrete boolean and each $x_{s_j} \in  \mathbb{R}$ ($1 \leq j \leq m$) is a continuous variable. Actions are presented by a finite set of actions $\{ a_1, \ldots, a_p \}$. We define discrete and continuous observations by the vector $(\vec{d_o},\vec{x_o}) = ( d_{o_1},\ldots,d_{o_n},x_{o_1},\ldots,x_{o_m} )$ where each $d_{o_i} \in \{ 0,1 \}$ ($1 \leq i \leq n$) is boolean and each $x_{o_j} \in \mathcal{R}$ ($1 \leq j \leq m$) is a continuous variable.

Three functions are required for modeling DC-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state given the action and previous state%$P(\vec{d}',\vec{x}'|\cdots,a)$
; (2)  $\mathcal{R}:\mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R}$ a reward function which returns the immediate reward of taking an action in some state and (3) an observation function defined as $\mathcal{Z} : \mathcal{S} \times \mathcal{A} \times \mathcal{O} \rightarrow [ 0, 1 ]$  which gives the probability of an observation given the outcome of a state after executing an action.  A discount factor $\gamma, \; 0 \leq \gamma \leq 1$ is included in the model to account for future rewards in higher horizons.

%%%% forms of transition, observation and reward function
To define the transition function, we use the fact that the state variables of this POMDP can be factored into a Dynamic Bayesian Network (DBN). Not allowing synchronic arcs between the discrete and continuous variables defines their joint transition to be independent:

{\footnotesize
\begin{equation}
P(\vec{d_s}',\vec{x_s}'|\vec{d_s},\vec{x_s},a) = 
\prod_{i=1}^n P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \prod_{j=1}^m P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a). \nonumber 
\end{equation}}
The joint observation function can also be defined considering the direct correspondence between the state and observation variables:
{\footnotesize
\begin{equation}
P(\vec{d_o},\vec{x_o}|\vec{d_s},\vec{x_s},a) = 
\prod_{i=1}^n P(d_{o_i}|d_{s_i},a) \prod_{j=1}^m P(x_{o_j}|x_{s_j},a). \nonumber 
\end{equation}}

The \emph{binary} variables are represented using conditional probability functions (CPF) in the form of $P(d_i'|\vec{d},\vec{x},a)$. For \emph{continuous} variables
the CPF $P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ is represented by \emph{piecewise
linear equations} (PLEs) that are first-order Markov and deterministic, i.e. the probabilities are encoded using the Dirac $\delta[\cdot]$ function.  The reward function $R(\vec{d},\vec{x},a)$ is set to be any general piecewise linear  function that could facilitate pruning for space efficiency via bilinear solvers.

For the \textsc{Power Plant} example, the temperature $t_s \in \mathbb{R}$ is modelled as the continuous state variable. Noisy observation of the temperature $t_o \in \mathbb{R}$ is used to help decide the optimal action to either open or close the pressure valve $a \in \{open,close\}$. 
For the transition probability, we use the Dirac function to model the deterministic equations. We can also define stochasticity in the transition model using boolean random variables that are sampled stochastically.  The reward function can be any linear function of the state or action. Going above a threshold temperature (e.g. $T=10$) will cause an explosion in the plant when trying to close the valve which results in the negative reward of $-1000$. Staying below this temperature is safe and will produce electricity an gain the reward of $100$. The reward of an open valve is $-1$.
\begin{align}
P(t_s'|\vec{t_s},a)= \delta\left[ t_s' - 
\begin{cases}
 (a=open) &: t_s - 5 \\ 
(a \neq open) &: t_s + 7 \\
\end{cases}
\right]\nonumber
, 
R(t_s,a) = 
\begin{cases}
 (a=open) &: -1 \\
(a \neq open)\wedge (t_s>T) &: -1000 \\
(a \neq open)\wedge \neg(t_s>T) &: 100 \\
\end{cases}\nonumber
\end{align}
%contribution? 
This example will try to maximize its reward using observations on the temperature. The observations recieved before closing the plant to a safe temperature are therefore critical.  For the observation model we begin with a discrete model for two observations $t_{o_1},t_{o_2}$ such as the following:  
\begin{align}
P(t_{o_1}|t_s',open) = 
\begin{cases}
 \delta\left[ t_s <= 15 \right] &: 0.9 \\
 \delta\left[ \neg (t_s <= 15)\right] &: 0.1 \\
\end{cases}\nonumber
, 
P(t_{o_2}|t_s',open) = 
\begin{cases}
 \delta\left[ t_s <= 15 \right] &: 0.1 \\
 \delta\left[ \neg (t_s <= 15)\right] &: 0.9 \\
\end{cases}\nonumber
\\
P(t_{o_1}|t_s',close) = 
\begin{cases}
 \delta\left[ t_s <= 15 \right] &: 0.1 \\
 \delta\left[ \neg (t_s <= 15)\right] &: 0.8 \\
\end{cases}\nonumber
,
P(t_{o_2}|t_s',close) = 
\begin{cases}
 \delta\left[ t_s <= 15 \right] &: 0.9 \\
 \delta\left[ \neg (t_s <= 15)\right] &: 0.2 \\
\end{cases} \nonumber
\end{align}

After defining the solution to the discrete observation with a continuous state DC-POMDP, we can extend this to consider continuous observations using a continuous uniform noise distribution which is more realistic in many domains. For a certain interval, the probability is a fixed number, while it is zero elsewhere. The integral of the uniform distribution should sum to one. 
\begin{align}
P(t_o|t_s',open) = 
\begin{cases}
 (t_o>t_s-5) \wedge (t_o<t_s+5) &: 0.1 \\
 \neg((t_o>t_s-5) \wedge (t_o<t_s+5)) &: 0 \\
\end{cases}\nonumber
\end{align}

Next we define the general solution to solving DC-POMDPs.

\section{DC-POMDP solution}
The DC-POMDP will choose actions in order to compute an optimal plan over time. This plan is the policy $\pi$ of the agent  depending on the history of actions and observations over a Markovian state. In the partially observable case where there is uncertainty on the states, the policy is a function over a continuous set of probability distributions over $|\mathcal{S}|$. Observations don't provide a unique state therefore the probability of the state is stored as the belief state $b$.

 At any of the horizon steps, a DC-POMDP policy $\pi(b)$ is defined as the iterative plan on executing an action and observing all probable outcomes of the set of observations from the $(h-1)$-step. The belief state is updated after this according to the Bayes rule: $\sum_{d_s}\int_{x_s} b(xd_s)T(xd'_s,xd_s,a)Z(xd'_s,xd_o,a)$  where for simiplicity we use ($xd_s = x_s,d_s$) and ($xd_o=x_o,d_o$).
The optimal policy is defined over the belief state as the sum of the expected discounted rewards over horizon $h$ starting with belief $b_0$.
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = E_{\pi^*} \left[ \sum_{h=0}^{H} \gamma^t \cdot r_h \Big| \vec{b}_0 = \vec{b} \right]\nonumber
\end{equation}
%The next section provides the mathematics required to solve this equation in DC-POMDPs.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 1 - policy tree
\begin{figure}[t!]
%\vspace{-1mm}
\begin{center}
\includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
\includegraphics[width=0.45\textwidth]{pics/b1-4.pdf}
\end{center}
\vspace{-2mm}
\caption{\footnotesize Left: Example conditional plan $\beta^h$ for discrete observations. Right: The optimal value function for \textsc{Power Plant}
as a decision diagram: 
the \emph{true} branch is solid, the \emph{false}
branch is dashed.}
\label{fig:cond_plan}
\vspace{-1mm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}[t]
%\begin{center}
%\includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
%\end{center}
%\vspace{-3mm}
%\caption{%\footnotesize 
%The optimal value function for \textsc{Power Plant}
%as a decision diagram: 
%the \emph{true} branch is solid, the \emph{false}
%branch is dashed.} 
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Dynamic Programming solution} 
Solving the value function for the optimal policy is performed using Dynamic programming algorithms such as value iteration. 
The decision at each time step is to pick an action based on past actions and observations. The possible set of conditional plans $\beta^h$ correspond to a decision tree. Figure 1 left, shows the decision tree for the simple case of three horizons with discrete observations. It represents a conditional plan consisting of an action and two possible observation strategies. For continuous observations, conditional plans are decision trees with infinitely many branches. In our solution we show that instead of a policy tree with infinite branches, we can create partitions based on relevant observation values, resulting in a policy tree similar to that of figure 1. %From here on we assume this finite set of observation partitioning as $\mathcal{O}$.

We define the value iteration algorithm for discrete observations and discrete states that finds the optimal value for each horizon starting with an initial belief. Although the number of belief states is infinite, the optimal value function for $h$ is proven to be a piecewise linear and convex function of the belief state \cite{smallwoodSondik}. The optimal value function is the maximum value over a finite set of alpha vectors  $\alpha^h_i$:%that define the hyperplanes:
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = \max_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot \vec{b}
\end{equation}
where the belief states $\vec{b}$ and each $\alpha^h_i \in \Gamma^h$ are of dimension $|\mathcal{S}|$. At each of the $h$-steps the value function is parameterized by the set of  $\alpha$-vectors that partition the belief state. 
At a particular belief point, we can compute the next value function: 
\begin{align*}
V^{h+1}(b) = \max_a \left( b \cdot r_a + \gamma \sum_o p(xd_o|a,b)V^h(b|xd_o,a)\right)\nonumber
\end{align*}
Replacing the previous iteration value with (3) and also updating the belief after executing action $a$ and observing $xd_o$ results in the following equation: 
\begin{align*}
V^{h+1}(b) = \max_a \left( b \cdot r_a + 
\gamma \sum_o \max_{\langle g^h_{a,o,j} \rangle_j}  b \cdot g^h_{a,o,j} \right)\nonumber
\end{align*}
where $r_a = r(x_s,d_s,a)$ and  
$g^h_{a,o,j} =  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s)$

Calculating all possible ways of building the value function independent of a certain belief state can define the optimal value function. To derive the $\Gamma$-set of all possible $\alpha$-vectors for each horizon, we use the value iteration algorithm from ~\cite{monahan82}. 
Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (4):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as 
$\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
\textrm{\large $\boxplus$} Q = \{ \vec{p} + \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}
\begin{align}
\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \\
\Gamma^h  &= \bigcup_a \Gamma^h_a 
\end{align}
Each action-dependent $\Gamma$ sums over the possible set of $g^h_{a,o,j}(\cdot)$ given an observation model to derive discrete observations $xd_o$ and for all $\alpha^{h-1}_{j} \in \Gamma^{h-1} $. The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.  
For each horizon, the optimal $\alpha$-vector is obtained from the maximum over all $\alpha$-vectors: 
$\alpha^* = \argmax_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot b $

Here the cross-sum operation takes the set of $\alpha$ -vectors for each observation and produces a sum over vectors that depends on the number of previous iteration vectors. Thus the $\alpha$-vector set grows exponentially in the number of observations, i.e., $|\Gamma^{h}| = |\mathcal{A}||\Gamma^{h-1}|^{|\mathcal{O}|}$ during the cross-sum operation. For this reason even for a small number of discrete observations, tractable solutions may be hard to define. Here we do not restrict our observations to discrete ones, but extend it to a fully continuous approach where continuous observations and states are allowed in the DC-POMDP. We use the number of  observation partitions $|\mathcal{O}|$ as the branching factor in the policy tree. We present the symbolic value iteration algorithm in the case of continuous observations using relevant observation partitioning of this continuous space.

\section{Symbolic Dynamic Programming} 
Dynamic programming provides exact solutions to POMDP problems for small domains. As the domain grows, fully enumerating the state, action and observation space leads to intractable solutions even for low horizons. Avoiding this enumeration often requires approximating the solutions. We claim an exact solution using symbolic dynamic programming. The SDP solution for fully observable domains has been presented in \cite{sanner_uai11} %using piecewise constant functions
Our symbolic algorithm deals with continuous observations and states by generating the relevant observation partition at every time step and performing the backup operation using $\alpha$-vectors. We define the requirements to implement the value iteration algorithm for DC-POMDP.

\subsection{Case Representation and Extended ADDs}
% operations, max, restrict, substitute
%overview + example plant
The \textsc{Power Plant} example represented all functions using the case structure which can be generally defined as:
{%\footnotesize 
\begin{align}
f = 
\begin{cases}
  \phi_1: & f_1 \\ 
 \vdots&\vdots\\ 
  \phi_k: & f_k \\ 
\end{cases} \nonumber
\end{align}
}
Where $\phi_i$ are disjoint logical formulae defined over the state $(\vec{d},\vec{x})$ with logical ($\land,\lor,\neg$) combinations of boolean variables and inequalities ($\geq,>,\leq,<$) over continuous variables.  
The $f_i$ are function definitions either linear or quadratic in the continuous parameters.

For \emph{unary operations} such as scalar multiplication $c\cdot f$ (for some constant $c \in \mathbb{R}$) or negation $-f$ on case statements is simply to apply the operation on each case partition $f_i$ ($1 \leq i \leq k$).

A \emph{binary operation} on two case statements, takes the cross-product of the logical partitions of each case statement and performs the corresponding operation on the resulting paired partitions.  The cross-sum $\oplus$ of two cases is defined as the following:
{\footnotesize 
\begin{center}
\begin{tabular}{r c c c l}
&
\hspace{-6mm} 
  $\begin{cases}
    \phi_1: & f_1 \\ 
    \phi_2: & f_2 \\ 
  \end{cases}$
$\oplus$
&
\hspace{-4mm}
  $\begin{cases}
    \psi_1: & g_1 \\ 
    \psi_2: & g_2 \\ 
  \end{cases}$
&
\hspace{-2mm} 
$ = $
&
\hspace{-2mm}
  $\begin{cases}
  \phi_1 \wedge \psi_1: & f_1 + g_1 \\ 
  \phi_1 \wedge \psi_2: & f_1 + g_2 \\ 
  \phi_2 \wedge \psi_1: & f_2 + g_1 \\ 
  \phi_2 \wedge \psi_2: & f_2 + g_2 \\ 
  \end{cases}$
\end{tabular}
\end{center}
}
\normalsize
Likewise $\ominus$ and $\otimes$ are defined by subtracting or multiplying partition values.  Inconsistent partitions can be discarded when they are irrelevant to the function value.
The next three operations are required in defining the value iteration algorithm. A \emph{symbolic case maximization} is defined as below:
\vspace{-4mm}

{\footnotesize
\begin{center}
\begin{tabular}{r c c c l}
&
\hspace{-7mm} $\mathrm{max} \Bigg(
  \begin{cases}
    \phi_1: \hspace{-2mm} & \hspace{-2mm} f_1 \\ 
    \phi_2: \hspace{-2mm} & \hspace{-2mm} f_2 \\ 
  \end{cases}$
$,$
&
\hspace{-4mm}
  $\begin{cases}
    \psi_1: \hspace{-2mm} & \hspace{-2mm} g_1 \\ 
    \psi_2: \hspace{-2mm} & \hspace{-2mm} g_2 \\ 
  \end{cases} \Bigg)$
&
\hspace{-4mm} 
$ = $
&
\hspace{-4mm}
  $\begin{cases}
  \phi_1 \wedge \psi_1 \wedge f_1 > g_1    : & \hspace{-2mm} f_1 \\ 
  \phi_1 \wedge \psi_1 \wedge f_1 \leq g_1 : & \hspace{-2mm} g_1 \\ 
  \phi_1 \wedge \psi_2 \wedge f_1 > g_2    : & \hspace{-2mm}f_1 \\ 
  \phi_1 \wedge \psi_2 \wedge f_1 \leq g_2 : & \hspace{-2mm} g_2 \\ 
  %\vdots & \vdots
  \phi_2 \wedge \psi_1 \wedge f_2 > g_1    : & \hspace{-2mm} f_2 \\ 
  \phi_2 \wedge \psi_1 \wedge f_2 \leq g_1 : & \hspace{-2mm} g_1 \\ 
  \phi_2 \wedge \psi_2 \wedge f_2 > g_2    : & \hspace{-2mm} f_2 \\ 
  \phi_2 \wedge \psi_2 \wedge f_2 \leq g_2 : & \hspace{-2mm} g_2 \\ 
  \end{cases}$
\end{tabular}
\end{center}
}
\emph{Restriction} takes a function $f$ to restrict only in cases
that satisfy some formula $\phi$, which we write as $f|_{\phi}$.  
This can be done by simply appending $\phi$ to each case partition
as follows:
{\footnotesize
\begin{center}
\begin{tabular}{r c c l}
&
\hspace{-6mm} 
  $f = \begin{cases}
    \phi_1: & f_1 \\ 
   \vdots&\vdots\\ 
    \phi_k: & f_k \\ 
  \end{cases}$
&

&
\hspace{-2mm}
  $f|_{\phi} = \begin{cases}
    \phi_1 \land \phi : & f_1 \\ 
   \vdots&\vdots\\ 
    \phi_k \land \phi : & f_k \\ 
  \end{cases}$
\end{tabular}
\end{center}
}

\emph{Symbolic substitution} simply takes a set $\sigma$ of variables and their substitutions, where
the LHS of the substitution operator $/$ represents the substitution variable and the
RHS of the $/$ represents the expression that should be substituted in its place.
Hence to perform a continuous regression on a more general
representation, we obtain that $\int_{x_j'} P(x_j'|\cdots) V'^{h} dx_j'$
\begin{align*}
    = \begin{cases}
    \phi_1: & V'^{h} \{ x_j' = f_1 \} \\ 
   \vdots&\vdots\\ 
    \phi_k: & V'^{h} \{ x_j' = f_k \}  \\ 
  \end{cases}
\end{align*}

%xadd representation
The data structure of \emph{extended ADD} (XADD)~\cite{sanner_uai11} is used to support
case statements and the required operations.  Here we use continuous SDP for both states and observations in DC-POMDPs. The value functions and $\alpha$-vectors partition into states of equal value, both of which can be represented using XADDs. For increasing efficiency linear constraint feasibility checkers such as LP solvers allow pruning unreachable paths in the XADD. The right picture in Figure 1 is an example of an XADD representation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{backup}{Backup}
\SetKwFunction{genObs}{GenRelObs}
\SetKwFunction{prune}{Prune}
\SetKwFunction{remapWithPrimes}{Prime}
\Begin
{
   $V^0:=0, h:=0$\;
   \While{$h < H$}
   {
       $h:=h+1, \Gamma^h :=\emptyset$\;
       \ForEach {$a \in A$}
       {
			$\Gamma_{a}^h :=\emptyset$ \;       		
       		\If {$ContObs =true$}
       			{$P(xd_{o_i}|xd_{s_i}) \,:=\,$ \genObs{$\Gamma^{h-1},a,b_i$}\;}
       		 \ForEach {$xd_{o_i} \in \textsc{O}^{h-1}$}
       		 {
				\ForEach {$\alpha_j^{h-1} \in \Gamma^{h-1}$}
       			{
   	 		  		$\alpha_j^{h'}=$ \remapWithPrimes{$\alpha_j^{h}$} \;
   	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
   	 		    	$\Gamma_{a,xd_{o_i},j}^h \,:=\, P(xd_{o_i}|xd_{s_i}) \otimes$ \backup{$\alpha_j^{h'},a$}\;
       	      	}
       	      	$\Gamma_{a,xd_o}^h\,:=\,\textrm{\large $\boxplus$} \Gamma_{a,xd_{o_i}}^h$\;
       	     }
           $\Gamma_a^{h} \,:=\,R_a \oplus \gamma \cdot \Gamma_{a,xd_o}^h$\;
            $\Gamma^{h} \,:=\, \Gamma^{h} \cup \Gamma_a^{h}$\;
        }  
        	%monahan's pruning first generates all vectors, then prunes
              $\Gamma^h \,:=\, $\prune{$\Gamma^h$} \;
              $V^h \,:=\, \mathrm{max}_{\alpha_j \in \Gamma^h} b_i \cdot \alpha_j$\;
              $\pi^{*,h} \,:=\, \argmax_{a} \, \Gamma_a^{h}$\;

       \If{$V^h = V^{h-1}$}
           {break $\,$ \emph{// Terminate if early convergence}\;}
   }
     \Return{$(V^h,\pi^{*,h})$} \;
}
\caption{\footnotesize \texttt{VI}(DC-POMDP, $H$,$ContObs$, $b_i$) $\longrightarrow$ $(V^h,\pi^{*,h})$ \label{alg:vi}}
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\incmargin{.5em}
%\linesnumbered
%\begin{algorithm}[t!]
%\vspace{-.5mm}
%\dontprintsemicolon
%\Begin{
%	\emph{any function f has $i$ partitions like $\phi_i:f_i$}\;	
%	\emph{compute UB and LB from $\phi_i$ using constraints on $var$}\;
%    $I=$ \emph{ any $var$-independent $\phi_i$} \;
%    $F = $ \emph{differentiate $f_i$}\;
%    $F = I \otimes [F(UB) - F(LB)]$\;
%    \Return{$F$} \;
%}
%\caption{\footnotesize \texttt{VE}($var,f$) $\longrightarrow$ $F$ }
%\vspace{-1mm}
%\end{algorithm}
%\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Value iteration for DC-POMDPs}

For continuous state and discrete observations, we use the symbolic version of Monahan's algorithm as the DC-POMDP solution. The same approach is required for continuous observations only based on point-based dynamic programming backups[]. Here for a given belief, the best conditional plan corresponding to an $\alpha$-vector is computed using point-based backups: 
\begin{align*}
\alpha^{h+1}(b) = R(b) + \gamma \int_{xd_o} P(xd_o|a,b)\alpha^{h+1}(b)\nonumber
\end{align*}
where $P(xd_o|a,b) = \int_{xd_s}\int_{xd'_s}b(s)P(xd'_s|xd_s,a)P(xd_o|xd'_s,a)$

Given an initial set of beliefs, we can compute the relevant observation partitions explained in the next section and then continue in a similar fashion of having discrete observations. To visualize we explain of the value iteration algorithm as defined in \texttt{VI} for DC-POMDPs. Each step of the algorithm is explained using the \textsc{Power plant} example for the first three horizons. Note that if the input to this algorithm is continuous observations $ContObs=true$ then line 8 finds the relevant observation partitions, else discrete observations are assumed and their probabilities are given in the problem specifications. Also a belief is given as the input and often several beliefs are required to cover the belief space and thus the algorithm is executed for each one of $b_i \in \mathbb{B}$.

For $h=0$, $V^0$ is given in line 2.  For $h=1$ we choose an action, for example $close$ to start with. The $\alpha$-vector set from $h=0$ is empty so line 7--8 of generating the relevant observation partitions is empty. Also priming the vectors and obtaining the product of relevant observations and the backup is an empty set. In line 15 the reward $R(s,close)$ is added to obtain $\alpha^1_{1,close}$ as defined by (8). The same operation is repeated for the $open$ action where $R(s,open)=-1$. Line 16 defines the final $\Gamma$-set: 
$\Gamma^1 = \{ \alpha^1_{1,open}, \alpha^1_{1,close} \}$.

The \texttt{Backup} algorithm is defined in a separate algorithm on the next page which takes the primed version of the variables and performs regression on the $\alpha$-vectors. 
\emph{Boolean restriction} is defined in lines 5--8 where $f|_{b=v}$ restricts a boolean variable $b$ occurring in $f$ to values $v \in \{ 0,1 \}$. This can be done by simply appending $b=v$ to each case partition (logical $\wedge$ of $b=v$ and the logical first-order formulas of $f$).

\emph{Continuous regression} $\int f(x_{s_j}') \otimes P(x_{s_j}'|\cdots) dx_{s_j}'$ in line 3--4 has the form $\delta[x_{s_j}' - h(\vec{z})]$ for the probability over the next state, where $h(\vec{z})$ is a case statement and $\vec{z}$ does not contain
$x_{s_j}'$.  A symbolic substitution is defined for this integration such that any occurrence of $x_{s_j}'$ in $f(x_{s_j}')$ is substituted with $h(\vec{z})$: 
$\int f(x_{s_j}') \otimes \delta[x_{s_j}' - h(\vec{z})] dx_{s_j}' = f(x'_j) \{ x_{s_j}' / h(\vec{z}) \}$

For $h=2$, for each of the discrete actions, first we generate the probability of the relevant set of observations in the \texttt{GenRelObs} algorithm. This has been done for the $close$ action in the next section and we use the results here. 
\begin{align}
P(z_k)=
\begin{cases}
 (2<t_o<6) &: 0.2 \\
(-4<t_o<2) &: 0.6\\
(-8<t_o<-4) &:0.1999
\end{cases} 
\nonumber
\end{align}
Next for each of the relevant partitions and both $\alpha$-vectors of the previous $\Gamma$ set lines 9--16 are executed. Line 11 will substitute prime values for all state variables. Line 13 produces the backup $\alpha$-vector and multiplies it by the weight derived from the relevant observation partition. For each observation, two vectors are formed for each corresponding $\alpha$-vector. For example line 13 for the observation partitions above and $\alpha_1^1,\alpha_2^1$ is defined as:
\begin{align*}
\Gamma_{close,z_1,1}^2= 
\begin{cases} 
(t_s'>15) &: -600 \\ 
\neg(t_s'>15) &: 60 \\
\end{cases}
\hspace{15mm} 
\Gamma_{close,z_1,2}^2= \top &:-0.6
\\
\Gamma_{close,z_2,1}^2= 
\begin{cases}
(t_s'>15) &: -199.999 \\ 
\neg(t_s'>15) &: 19.999 
\end{cases}
\hspace{12mm} 
\Gamma_{close,z_2,2}^2= \top &:-0.1999
\\
\Gamma_{close,z_3,1}^2= 
\begin{cases}
(t_s'>15) &: -200 \\ 
\neg(t_s'>15) &: 20 
\end{cases}
\hspace{15mm} 
\Gamma_{close,z_3,2}^2= \top &:-0.2
\end{align*}
Next in line 14 we have to take the cross-sum of the two sets which takes every possible combination of observations according to the $\alpha$-vectors. 
\begin{align*}
\Gamma_{close,1}^2 &= 
\begin{cases} 
(t_s'>15) &: -800.2 \\ 
\neg(t_s'>15) &: 79.8 \\
\end{cases}
\hspace{10mm} 
\Gamma_{close,2}^2= 
\begin{cases} 
(t_s'>15) &: -1000 \\ 
\neg(t_s'>15) &: 100 \\
\end{cases}
\\
\Gamma_{close,3}^2&= 
\begin{cases} 
(t_s'>15) &: -800.2 \\ 
\neg(t_s'>15) &: 79.8 \\
\end{cases}
\hspace{10mm}
\Gamma_{close,4}^2= 
\begin{cases} 
(t_s'>15) &: -600.4 \\ 
\neg(t_s'>15) &: 59.6 \\
\end{cases}
\\
\Gamma_{close,5}^2&= 
\begin{cases} 
(t_s'>15) &: -400.5999 \\ 
\neg(t_s'>15) &: 39.3999 \\
\end{cases}
\hspace{5mm} 
\Gamma_{close,6}^2= 
\begin{cases} 
(t_s'>15) &: -200.7999 \\ 
\neg(t_s'>15) &: 19.999 \\
\end{cases}
\\
\Gamma_{close,7}^2&= 
\begin{cases} 
(t_s'>15) &: -200.8 \\ 
\neg(t_s'>15) &: 19.2 \\
\end{cases}
\hspace{10mm} 
\Gamma_{close,8}^2= \top:-1
\end{align*}
All 8 vectors are multiplied by the discount $\gamma$ and added with the reward in line 15 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
% need to add this?
The same operation is performed for the $open$ action and then in line 15 the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors for the next horizon.

In line 17 this final set is using in the \texttt{Prune} algorithm. For all $\alpha$-vectors in the current $\Gamma$-set first an LP-solver is used to prune any inconsistency in the branches of this vector. Next a pairwise dominance test is performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$
The result of this pairwise test is a dominant $\alpha$-vector which is added to the set of $\Gamma^h$. 
In the last step, we find the $\alpha$-vector that maximizes the value-function with respect to this belief and save it as the new $\Gamma$-set for the next horizon. The action from which this $\alpha$-vector was chosen from is the maximum policy. 
This operation is performed for each of the belief points $b_i \in B$ and for each belief point one dominant $\alpha$-vector is stored in the $\Gamma$-set. The algorithm iterates until convergence. 
In the next section we discuss the main part of our algorithm which is generating relevant observation partitions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\Begin{

    \emph{// Continuous regression, marginal integration}\\
    \For {all $xd_{s_j}'$ in $\alpha_{j,o}^{'h}$}
    {
         $G_a^h := \int P(z) \otimes P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a) \, d_{x_{s_j}'}$\;
    }
    \emph{// Discrete regression, marginal summation}\\
    \For {all $d_{s_i}'$ in $\alpha_{j,o}^{'h}$}
    {
         $G_a^h := \left[ G_a^h \otimes P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \right]|_{d_{s_i}' = 1}$\\
         \hspace{8mm} $\oplus \left[ G_a^h \otimes P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \right]|_{d_{s_i}'= 0}$\;
    }
    \Return{$G_a^h$} \;
}
\caption{\footnotesize \texttt{Backup}($\alpha_{j,o}^{'h},a,P(z)$) $\longrightarrow$ $G_a^h$ \label{alg:regress}}
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\SetKwFunction{lpsolve}{LP-Solve}
\SetKwFunction{dominancy}{Dominance}
\Begin
{
   		$V^h :=0$ \;
       \ForEach {$\alpha_j^h \in \Gamma^h$}
       {
           $\alpha_j^h \,:=\,$\lpsolve{$\alpha_j^h$} \;
           \ForEach {$\alpha_k^h \in \Gamma^h$}
           {
				\If {$j \neq k$}  
				{         
           			$V^h \,:=\,\oplus$ \dominancy{$\alpha_j^h,\alpha_k^h$} \; 
           		}  
           	}
       }
     \Return{$(V^h)$} \;
}
\caption{\footnotesize \texttt{Prune}($\Gamma^h$) $\longrightarrow$ $(V^h)$ }
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Continuous Observation Space} 
%progression definition
The main operations required for value iteration in DC-POMDP is generating relevant observations to perform the bellman backup operation. 
%Progression is the inverse of regression. The \emph{regression} of a formula $\psi$ through an action $a$ is a formula $\psi'$ that holds prior to $a$ being performed iff $\psi$ holds after $a$. \emph{Progression} takes a formula $\psi$ and an action $o$ and represents the formula $\psi'$ that would hold after action $o$ is performed in a situation satisfying $\psi$.

The \texttt{GenRelObs} algorithm defines the operation of generating relevant observation partitions. 
We use the second iteration of our \texttt{Power Plant} example to demonstrate the steps  required to partition the entire continuous observation space. This continuous nature requires assuming some belief state and obtaining the relevant partitions according to that belief. 
Starting with an empty set of relevant observation partitions $z$, the algorithm first eliminates the next state variable $xd'_s$. It multiplies the observation model and the transition model in the set of $\alpha$-vectors of the previous horizon.
$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $
This operation is similar to that of \texttt{Regress} plus the observation regressed model. For continuous variables the continuous regression and for discrete variables boolean restriction is used as defined in the previous section. 
Assume that after $h=1$ the set of $\alpha$-vectors are defined as:
\begin{align}
\alpha_1^1(s) &= 
\begin{cases}
 (t_s>200) &: -1000 \\
\neg(t_s>200) &: 100 \\
\end{cases}
\hspace{10mm} 
\alpha_2^1(s) = \top: -1 \nonumber
\end{align}
For $h=2$ the resulting $\alpha$-vectors after this regression step are functions of the observation and current state: 
\begin{align}
\alpha_1^2(s,o) &= 
\begin{cases}
 (t_s<15)\wedge (t_s - 10 < t_o<t_s) &: 10 \\
(t_s>15)\wedge (t_s - 10 < t_o<t_s) &: -100  \\
\neg(t_s - 10 < t_o<t_s) &: 0
\end{cases}
\hspace{0mm} 
\alpha_2^2(s,o) &= \begin{cases}
(t_s - 10 < t_o<t_s) &: -0.1 \\
\neg(t_s - 10 < t_o<t_s) &: 0
\end{cases}
\nonumber
\end{align}
We need the $\alpha$-vectors to partition the observation space, thus next we eliminate the current state variable. During this integration the current state variable is factored out by multiplying the belief in each of the current $\alpha$-vectors.  The resulting $\delta$-functions only depend on the observation: $\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$.
For the belief $b_1: U[t_s;2,6]$ the two vectors above the resulting $\delta$-functions are defined below: 
\begin{align}
\delta_{1}(o) &= 
\begin{cases}
 (2<t_o<6) &: 15 - 2.5*t_o \\
(-4<t_o<2) &: 10 \\
(-8<t_o<-4) &: 20 + 2.5*t_o 
\end{cases}
\hspace{5mm} 
\delta_{2}(o) &= \begin{cases}
 (2<t_o<6) &: -15 + 0.025*t_o \\
(-4<t_o<2) &: -0.1 \\
(-8<t_o<-4) &: -0.2 - 0.025*t_o 
\end{cases}
\nonumber
\end{align}
The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_1$. According to continuous observation space defined in \cite{pascalPomdp}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space. They propose numerical solutions to finding the roots of any two observation dependant function. Here we use the symbolic power of the max-operator to find all the maximum boundary regions of two or more $\delta$-functions at the same time. For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximuum: 
\begin{align}
\mathrm{max} \Bigg(\delta_{1}(o),\delta_{1}(o)\Bigg) &= 
\begin{cases}
 (2<t_o<6) &: 15 - 2.5*t_o \\
(-4<t_o<2) &: 10 \\
(-8<t_o<-4) &: 20 + 2.5*t_o 
\end{cases}
\nonumber
\end{align}
The three partitions define 3 observation regions which we can now use in a similar fashion to a discrete observation set . Thus we only need to find the probability of each of the three distinct observation partitions to use in Monahan's algorithm. If we integrate out the state and observation from the observation model and the belief state regardless of a particular $\alpha$-vector, it integrates to one: 
$\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*P(s) = 1$

Now that we have partitions on the observation space, we only need to multiply the indicators of each case partition in this formula to obtain the probability of each partition such that it sums to one: 
$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$
For our example the following holds: 
\begin{align}
P(z_k)=
\begin{cases}
 (2<t_o<6) &: 0.2 \\
(-4<t_o<2) &: 0.6\\
(-8<t_o<-4) &:0.1999
\end{cases} 
\nonumber
\end{align}
Hence we can now use the algorithms and methods of a discrete observation setting using the probabilities of the partitioned observation space. 
%Figure ~\ref{fig:obsPartition} shows the observation partitions obtained in the 1D problem instance for different horizons and their related probabilities. In a lyx file, needed?
Next we present some results for 2-Dimensional continuous observation spaces.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{substitute}{Substitute}

\Begin
{
	$z:=0$;\	
		\For {all $\alpha_j$ in  $\Gamma^{h-1}$}    
		{
    	$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $;\
		}  
		\For {all $j$ in $\alpha_j(xd_s,xd_o)$}    
		{
		$\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$\}\ \;
		}
		\For {all $j$ in $\delta_{j}(xd_o)$}    
		{
		$z := max(\delta_j(xd_o),z)$\;
    	}
    	\emph{// $\phi_{z_k}$ logical term on each partition of $z$}\\
   		$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$ \;
    \Return{$P(z)$} \;
    %do this for each belief in B
}
\caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a,b_i$) $\longrightarrow$ $P(z)$ }
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}[tbp!]
%\vspace{-2mm}
%\centering
%\includegraphics[width=0.42\textwidth]{pics/nodes.pdf}\\
%\vspace{-2mm}
%\includegraphics[width=0.42\textwidth]{pics/time.pdf}
%\vspace{-2mm}
%\caption{\footnotesize space 
%and time vs. horizon.
%}
%\label{fig:timeSpace}
%\vspace{-4mm}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Empirical Results}
We evaluated our continuous POMDP solution using XADDs on the one-dimensional
\textsc{Power Plant} example and another variant of this problem with two variables. \footnote{
Full problem specifications and Java code to reproduce these experiments are currently 
in Google Code; anonymity restrictions prevent us from providing this link in the submitted paper version.}
We start by defining the 2-dimensional description of this problem:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\vspace{-2mm}
\centering
\includegraphics[width=0.3\textwidth]{pics/time2.pdf}
\includegraphics[width=0.3\textwidth]{pics/nodes2.pdf}
\includegraphics[width=0.3\textwidth]{pics/alpha-vectors2.pdf}
\vspace{-3mm}
\caption{\footnotesize 
{\it (left)} Space vs Horizon; 
{\it (middle)} Time vs Horizon; 
{\it (right)} Number of $\alpha$-vectors vs Horizon.
}
\label{fig:timeSpace}
\vspace{-4mm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

{\bf \textsc{2D-Power Plant}:} We consider the more complex model of the power plant similar to \cite{steam2} where the pressure inside the water tank must be controlled to avoid water mixing into the steam or explosion of the tank. We model the pressure variable $p$ as a partially observable variable from the observation readings of the pressure valve $po$. The two actions of increase and decrease are defined based on the change in both the temperature and the pressure. For the increase action the transition functions and the reward function is defined as below: 
\begin{align}
P(p_s'|\vec{p_s},a)&= \delta\left[ p_s' - 
\begin{cases}
 (p+10> 20) &: 20 \\ 
\neq (p+10> 20) &: p_s + 10 \\
\end{cases}
\right]\nonumber
\\
P(t_s'|\vec{t_s},a)&= \delta\left[ t_s' -  t_s +10 \right]\nonumber
\\
R(t_s,p_s,a) &= 
\begin{cases}
(5 \leq p_s \leq 15)\wedge (95 \leq t_s \leq 105)&:50\\
(5 \leq p_s \leq 15)\wedge (t_s \leq 95)&: -1\\
(5 \leq p_s \leq 15)\wedge (t_s \geq 95)&: -3\\				
(p_s \leq 5) &: -3\\						
(p_s \geq 15) &: -5\\ 
\end{cases}\nonumber
\end{align}
There is a high reward for staying withing the safe temperature and pressure range since it produces power, else depending on how safe it is to have values higher or lower than the safe range, penalty is defined.
As for the decrease action, the transition functions reduce the temperature by 5 units and the pressure by 10 units as long as the pressure stays above zero. For the reward function, we assume that there is always a small penalty for decreasing the values because power can not be generated.

For the observation model we consider two continuous uniform distributions such as the following:  
\begin{align}
P(t_o|t_s') &= 
\begin{cases}
 (t_o>t_s + 80) \wedge (t_o<t_s+ 105) &: 0.4 \\
 \neg( (t_o>t_s + 80) \wedge (t_o<t_s+ 105)) &: 0 \\
\end{cases}\nonumber
\end{align}
\begin{align}
P(p_o|p_s') &= 
\begin{cases}
 (p_o>p_s) \wedge (p_o<p_s+10) &: 0.1 \\
 \neg( (p_o>p_s) \wedge (p_o<p_s+10)) &: 0 \\
\end{cases}\nonumber
\end{align}

We define two rectangular uniform beliefs around the regions of rewarding, so that one needs to increase the values while the other should decrease them: 
\begin{align}
b_1 = 
\begin{cases}
(p_o>p_s) \wedge (p_o<p_s+10) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 100)&: 0.01 \\
 \neg( (p_o>p_s) \wedge (p_o<p_s+10) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 100)) &: 0 \\
\end{cases}\nonumber
\end{align}
\begin{align}
b_2 = 
\begin{cases}
(p_o>p_s +10) \wedge (p_o<p_s+30) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 130)&: 0.00125 \\
 \neg((p_o>p_s +10) \wedge (p_o<p_s+30) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 130)) &: 0 
\end{cases}\nonumber
\end{align}
% I can draw this and the 1D belief
In Figure ~\ref{fig:timeSpace}, a time and space analysis of
the two versions of the problem have been performed for up to 6 horizons. Comparing the two problem sizes demonstrates the effect of the number of state-observation variables in our algorithm. As the algorithm progresses, the time required to compute the probability of the partitions and finding the maximum value function increases for both problem sizes and significantly more for the 2D version. The stability in time is due to the fact that each stage almost takes the same time since it has converged quickly.%???
Increase in the problem size, increases the partition numbers on the observation space and this produces more $\alpha$-vectors which also effecting the space required to perform the algorithm. The number of vectors stays the same for most horizons and they drop after convergence in the 2D problem instance. The number of nodes in each iteration increases as more space is required to perform higher iterations.

In Figure ~\ref{fig:3D}  we show plots of the value function for different iterations of the 2D problem instance. Each plot demonstrates the value as a function of the pressure and temperature. Starting with the first iteration, the value is highest for the reward range ($5<p<15 \wedge 95<t<105$) and -1 or less for other places. In the fifth iteration, the value function has partitioned into more pieces showing how higher temperatures can increase the value without considering the effect of the pressure. In the last plot, horizon 6 has better tuned the value function so that higher temperatures and pressures increase the value function and also within the reward range, finer grain partitions have been formed. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% policy2, v2plot.pdf, v9plot
% annotation?
\begin{figure*}[tbp!]
\vspace{-2mm}
\centering
\includegraphics[width=0.31\textwidth]{pics/2d1.pdf}
\includegraphics[width=0.31\textwidth]{pics/2d9.pdf}
\includegraphics[width=0.31\textwidth]{pics/2d111.pdf}
\vspace{-3mm}
\caption{\footnotesize 
{\it (left)} Value function for first iteration; 
{\it (middle)} $V^5(p,t)$; 
{\it (right)} $V^6(p,t)$.
}
\label{fig:3D}
\vspace{-4mm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Related Work} 
This work has used the concept of continuous states and the SDP solution to MDPs from ~\cite{sanner_uai11} and continuous observations from the work of \cite{pascalPomdp}. It is exclusive in the sense that it brings together the continuous state and observation POMDPs using a symbolic approach. 
In many applications of the POMDP framework, the state and observation space have been considered as discrete values such as ~\cite{steam2} here we avoid this general simplification and work with the real values from sensors.
%sampling, monte carlo, particle filter, gaussians, perseus, pergeus

There has been prior work on approximate solutions to POMDPs with large or continuous state spaces ~\cite{Thrun99h}, ~\cite{Perseus} and some has been extended to the continuous observation setting using methods such as sampling ~\cite{PerseusObs}.
In most approximate continuous state POMDP work, continuous or large discrete actions has been used. Thus we can extend the current work using \cite{zamaniAAAI} to contain continuous actions and provide exact or approximate solutions.

\section{Conclusion}
We presented the first exact solution to continuous states and observations in a DC-POMDP framework. Transitions and reward functions are defined as piecewise linear functions of the continuous states. Continuous observations require definition of explicit partitions based on the observation model and the belief state of the agent. The key contribution is to define the partitions on the observation space and derive the related probabilities using symbolic integration and maximization. We extend the application to more than one state-observation setting and present the results.

%\subsubsection*{References} 
\bibliography{dcpomdp}
\bibliographystyle{plain}

\end{document}

### 2012-05-22 cpomdp/camdp-xadd/src/doc/NIPS/draft3-10page.tex
new 10 page draft

--- 2012-05-18 trunk/cpomdp/camdp-xadd/src/doc/NIPS/draft2-long version.tex
@@
 \begin{abstract}
-Decision-theoretic planning problems try to find the optimal sequence of actions for a given environment.
-Using continuous models in such problems is naturally closer to the real-world.
-Partially Observable MDPs model uncertain environments where the underlying state is not be fully observable.
-
+Decision-theoretic planning searches for the optimal sequence of actions for a given uncertain environment.
+Partially Observable MDPs (POMDP) can model such environments where the underlying state is not be fully observable.
+%While discrete models of the environment is often used for planning, continuous models are naturally closer to the real-world.
 While previous work have provided solutions for discrete states and/or observations, here we define the first exact solution to both continuous state and observations.
-This symbolic dynamic programming approach avoids enumerating the state and observation space providing solutions for our general discrete and continuous partially observable MDP.
-The main contribution of this paper is to use continuous states with discrete observations and extending it to continuous observations by defining relevant observation partitions associated with each state partition.
-This allows us to perform the Bellman backup operation efficiently.
-We address the problems in POMDP continuous planning and solutions to solve them symbolically.
-%To compactly represent the exact value functions Affine algebraic decision diagrams are used.
+The symbolic dynamic programming approach in this paper avoids enumerating the state and observation space providing solutions for general multi-variate discrete and continuous POMDPs.
+Our main contribution is that for a fixed set of reachable belief states and a fixed set of actions, only a \textit{finite} partitioning of the continuous observation space is required for optimality.
+Using function representation of Affine algebraic decision diagrams (XADDs) and symbolic case operators we derive the automatic partitioning for piecewise linear dynamics %such as the reward, transition and observation function.
+We demonstrate empirical results on an operation research domain to show the \textit{first exact solution} to continuous state and observation POMDPs.  
+%to use continuous states with discrete observations and extending it to continuous observations by defining relevant observation partitions %associated with each state partition.
+%This allows us to perform the Bellman backup operation efficiently.
+%We address the problems in POMDP continuous planning and solutions to solve them symbolically. %To compactly represent the exact value functions Affine algebraic decision diagrams are used. 
 \end{abstract}
@@
 In many complex systems complete information of the state is not available due to noisy or poor sensors and readings of them.
-Partially observable Markov Decision Processes (POMDPs)\cite{POMDPfirst} can be used for planning under such uncertainty in abstract problems.
-In many real-world problems, resources and sensor inputs are naturally modeled using continuous variables.
-
+Partially observable Markov Decision Processes (POMDPs)\cite{POMDPfirst} can be used for planning under such uncertainty in such problems.
+The solution to a POMDP is to find an answer to the fully-observable belief-MDP, whose states are probability distributions over the state space of the original POMDP \cite{kaebling}
+Real-world resources and sensor inputs are often naturally modelled using continuous variables. 
 Unfortunately most POMDP techniques can not be applied to such settings without assuming prior knowledge about the problem domain.
 Most methods use some approximation to discretize the continuous state, observation or action space which increases computational complexity \cite{Thrun99h}.
+The optimal value function of belief-MDPs is piecewiselinear and convex in these discrete cases \cite{sondik71} but has also been proven in the setting of only continuous states.
+\cite{PerseusObs}
+
 Exact solutions to a multi-variate continuous state and observation setting has not been approached for the curse of dimensionality problem.
@@
 For this reason, the process is modeled using POMDP in the literature.
-The temperature is assumed to be a continuous variable that is not observed.
-Noisy observations of the temperature are used to decide on the action of closing or opening the valve.}
+The temperature $t_s \in \mathbb{R}$ is modelled as a continuous state variable.
+Noisy observation of the temperature $t_o \in \mathbb{R}$ is used to help decide the optimal action to either open or close the pressure valve $a \in \{open,close\}$.}
 
@@
 
-Three functions are required for modeling DC-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state given the action and previous state%$P(\vec{d}',\vec{x}'|\cdots,a)$
+Three functions are required for modeling DC-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state given the action and previous state %$P(\vec{d}',\vec{x}'|\cdots,a)$
 ; (2)  $\mathcal{R}:\mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R}$ a reward function which returns the immediate reward of taking an action in some state and (3) an observation function defined as $\mathcal{Z} : \mathcal{S} \times \mathcal{A} \times \mathcal{O} \rightarrow [ 0, 1 ]$  which gives the probability of an observation given the outcome of a state after executing an action.
@@
 %%%% forms of transition, observation and reward function
-To define the transition function, we use the fact that the state variables of this POMDP can be factored into a Dynamic Bayesian Network (DBN).
-Not allowing synchronic arcs between the discrete and continuous variables defines their joint transition to be independent:
-
-{\footnotesize
-\begin{equation}
-P(\vec{d_s}',\vec{x_s}'|\vec{d_s},\vec{x_s},a) = 
-\prod_{i=1}^n P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \prod_{j=1}^m P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a). \nonumber 
-\end{equation}}
-The joint observation function can also be defined considering the direct correspondence between the state and observation variables:
+To define the transition and observation function, we use the fact that the state variables of this POMDP can be factored into a Dynamic Bayesian Network (DBN).
+Not allowing synchronic arcs between the discrete and continuous variables defines their joint transition or observation to be independent:
+%{\footnotesize
+%\begin{equation}
+%P(\vec{d_s}',\vec{x_s}'|\vec{d_s},\vec{x_s},a) = 
+%\prod_{i=1}^n P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \prod_{j=1}^m P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a). \nonumber 
+%\end{equation}}
+%The joint observation function can also be defined considering the direct correspondence between the state and observation variables:
 {\footnotesize
@@
 \end{equation}}
-
-The \emph{binary} variables are represented using conditional probability functions (CPF) in the form of $P(d_i'|\vec{d},\vec{x},a)$. For \emph{continuous} variables
-the CPF $P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ is represented by \emph{piecewise
+\emph{Binary} variables are represented using conditional probability functions (CPF) in the form of $P(d_i'|\vec{d},\vec{x},a)$. \emph{Continuous} variables
+$P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ are represented by \emph{piecewise
 linear equations} (PLEs) that are first-order Markov and deterministic, i.e.
 the probabilities are encoded using the Dirac $\delta[\cdot]$ function.
-The reward function $R(\vec{d},\vec{x},a)$ is set to be any general piecewise linear  function that could facilitate pruning for space efficiency via bilinear solvers.
-
-For the \textsc{Power Plant} example, the temperature $t_s \in \mathbb{R}$ is modelled as the continuous state variable.
-Noisy observation of the temperature $t_o \in \mathbb{R}$ is used to help decide the optimal action to either open or close the pressure valve $a \in \{open,close\}$.
-For the transition probability, we use the Dirac function to model the deterministic equations.
-We can also define stochasticity in the transition model using boolean random variables that are sampled stochastically.
-The reward function can be any linear function of the state or action.
-Going above a threshold temperature (e.g.
-$T=10$) will cause an explosion in the plant when trying to close the valve which results in the negative reward of $-1000$.
-Staying below this temperature is safe and will produce electricity an gain the reward of $100$.
-The reward of an open valve is $-1$.
-\begin{align}
-P(t_s'|\vec{t_s},a)= \delta\left[ t_s' - 
-\begin{cases}
- (a=open) &: t_s - 5 \\ 
-(a \neq open) &: t_s + 7 \\
-\end{cases}
-\right]\nonumber
-, 
-R(t_s,a) = 
-\begin{cases}
- (a=open) &: -1 \\
-(a \neq open)\wedge (t_s>T) &: -1000 \\
-(a \neq open)\wedge \neg(t_s>T) &: 100 \\
-\end{cases}\nonumber
-\end{align}
-%contribution? 
-This example will try to maximize its reward using observations on the temperature.
-The observations recieved before closing the plant to a safe temperature are therefore critical.
-For the observation model we begin with a discrete model for two observations $t_{o_1},t_{o_2}$ such as the following:
-\begin{align}
-P(t_{o_1}|t_s',open) = 
-\begin{cases}
- \delta\left[ t_s <= 15 \right] &: 0.9 \\
- \delta\left[ \neg (t_s <= 15)\right] &: 0.1 \\
-\end{cases}\nonumber
-, 
-P(t_{o_2}|t_s',open) = 
-\begin{cases}
- \delta\left[ t_s <= 15 \right] &: 0.1 \\
- \delta\left[ \neg (t_s <= 15)\right] &: 0.9 \\
-\end{cases}\nonumber
-\\
-P(t_{o_1}|t_s',close) = 
-\begin{cases}
- \delta\left[ t_s <= 15 \right] &: 0.1 \\
- \delta\left[ \neg (t_s <= 15)\right] &: 0.8 \\
-\end{cases}\nonumber
-,
-P(t_{o_2}|t_s',close) = 
-\begin{cases}
- \delta\left[ t_s <= 15 \right] &: 0.9 \\
- \delta\left[ \neg (t_s <= 15)\right] &: 0.2 \\
-\end{cases} \nonumber
-\end{align}
-
-After defining the solution to the discrete observation with a continuous state DC-POMDP, we can extend this to consider continuous observations using a continuous uniform noise distribution which is more realistic in many domains.
-For a certain interval, the probability is a fixed number, while it is zero elsewhere.
-The integral of the uniform distribution should sum to one.
-\begin{align}
-P(t_o|t_s',open) = 
-\begin{cases}
- (t_o>t_s-5) \wedge (t_o<t_s+5) &: 0.1 \\
- \neg((t_o>t_s-5) \wedge (t_o<t_s+5)) &: 0 \\
-\end{cases}\nonumber
-\end{align}
-
-Next we define the general solution to solving DC-POMDPs.
-
-\section{DC-POMDP solution}
-The DC-POMDP will choose actions in order to compute an optimal plan over time.
-This plan is the policy $\pi$ of the agent  depending on the history of actions and observations over a Markovian state.
-In the partially observable case where there is uncertainty on the states, the policy is a function over a continuous set of probability distributions over $|\mathcal{S}|$.
-Observations don't provide a unique state therefore the probability of the state is stored as the belief state $b$.
-
- At any of the horizon steps, a DC-POMDP policy $\pi(b)$ is defined as the iterative plan on executing an action and observing all probable outcomes of the set of observations from the $(h-1)$-step.
-The belief state is updated after this according to the Bayes rule: $\sum_{d_s}\int_{x_s} b(xd_s)T(xd'_s,xd_s,a)Z(xd'_s,xd_o,a)$  where for simiplicity we use ($xd_s = x_s,d_s$) and ($xd_o=x_o,d_o$).
-The optimal policy is defined over the belief state as the sum of the expected discounted rewards over horizon $h$ starting with belief $b_0$.
-\begin{equation}
-V^h_{\Pi^*}(\vec{b}) = E_{\pi^*} \left[ \sum_{h=0}^{H} \gamma^t \cdot r_h \Big| \vec{b}_0 = \vec{b} \right]\nonumber
-\end{equation}
-%The next section provides the mathematics required to solve this equation in DC-POMDPs.
+The reward function $R(\vec{d},\vec{x},a)$ is set to be any general piecewise linear  function that could facilitate pruning for space efficiency.% via bilinear solvers.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Dynamic Programming solution} 
+For the \textsc{Power Plant} example, the transition and reward functions are defined as below:
+% For the transition probability, we use the Dirac function to model the deterministic equations.
+We can also define stochasticity in the transition model using boolean random variables that are sampled stochastically.
+The reward function can be any linear function of the state or action.
+Going above a threshold temperature (e.g.
+$T=10$) will cause an explosion in the plant when trying to close the valve which results in the negative reward of $-1000$.
+Staying below this temperature is safe and will produce electricity an gain the reward of $100$.
+The reward of an open valve is $-1$.
+{\footnotesize
+\begin{align}
+P(t_s'|\vec{t_s},a)= \delta\left[ t_s' - 
+\begin{cases}
+ (a=open) &: t_s - 5 \\ 
+(a \neq open) &: t_s + 7 \\
+\end{cases}
+\right]\nonumber
+\hspace{5mm}
+R(t_s,a) = 
+\begin{cases}
+ (a=open) &: -1 \\
+(a \neq open)\wedge (t_s>T) &: -1000 \\
+(a \neq open)\wedge \neg(t_s>T) &: 100 \\
+\end{cases}\nonumber
+\end{align}
+}
+%contribution? 
+This example will try to maximize its reward using observations on the temperature.
+The observations received before closing the plant to a safe temperature are therefore critical.
+For the observation model we begin with a discrete model for two observations $t_{o_1},t_{o_2}$ such as the following for the $open$ action:
+{\footnotesize
+\begin{align}
+P(t_{o_1}|t_s',open) = 
+\begin{cases}
+ \delta\left[ t_s <= 15 \right] &: 0.9 \\
+ \delta\left[ \neg (t_s <= 15)\right] &: 0.1 \\
+\end{cases}\nonumber
+\hspace{5mm}
+P(t_{o_2}|t_s',open) = 
+\begin{cases}
+ \delta\left[ t_s <= 15 \right] &: 0.1 \\
+ \delta\left[ \neg (t_s <= 15)\right] &: 0.9 \\
+\end{cases}\nonumber
+\end{align}
+}
+%\\
+%P(t_{o_1}|t_s',close) = 
+%\begin{cases}
+% \delta\left[ t_s <= 15 \right] &: 0.1 \\
+% \delta\left[ \neg (t_s <= 15)\right] &: 0.8 \\
+%\end{cases}\nonumber
+%,
+%P(t_{o_2}|t_s',close) = 
+%\begin{cases}
+% \delta\left[ t_s <= 15 \right] &: 0.9 \\
+% \delta\left[ \neg (t_s <= 15)\right] &: 0.2 \\
+%\end{cases} \nonumber
+%\end{align}
+After defining the solution to the discrete observation with a continuous state DC-POMDP, we can extend this to consider continuous observations using a continuous uniform noise distribution which is more realistic in many domains.
+%For a certain interval, the probability is a fixed number, while it is zero elsewhere.
+The integral of the uniform distribution should sum to one.
+{\footnotesize
+\begin{align}
+P(t_o|t_s',open) = 
+\begin{cases}
+ (t_o>t_s-5) \wedge (t_o<t_s+5) &: 0.1 \\
+ \neg((t_o>t_s-5) \wedge (t_o<t_s+5)) &: 0 \\
+\end{cases}\nonumber
+\end{align}
+}
+Next we define the general solution to solving DC-POMDPs.
+
+\section{DC-POMDP solution}
+The DC-POMDP will choose actions in order to compute an optimal plan over time.
+This plan is the policy $\pi$ of the agent  depending on the history of actions and observations over a Markovian state.
+In the partially observable case where there is uncertainty on the states, the policy is a function over a continuous set of probability distributions over $|\mathcal{S}|$ known as the belief state $b = P(x_s,d_s)$.
+
+At any of the horizon steps, a DC-POMDP policy $\pi(b)$ is defined as the iterative plan on executing an action and observing all probable outcomes of the set of observations from the $(h-1)$-step.
+The belief state is updated after this according to the Bayes rule: $\sum_{d_s}\int_{x_s} b(xd_s)T(xd'_s,xd_s,a)Z(xd'_s,xd_o,a)$.
+\footnote{For simiplicity we use ($xd_s = x_s,d_s$) and ($xd_o=x_o,d_o$)}
+The optimal policy is defined over the belief state as the sum of the expected discounted rewards over horizon $h$ starting with belief $b_0$.
+{\footnotesize
+\begin{equation}
+V^h_{\Pi^*}(\vec{b}) = E_{\pi^*} \left[ \sum_{h=0}^{H} \gamma^t \cdot r_h \Big| \vec{b}_0 = \vec{b} \right]\nonumber
+\end{equation}
+}
 Solving the value function for the optimal policy is performed using Dynamic programming algorithms such as value iteration. 
-The decision at each time step is to pick an action based on past actions and observations.
+%The decision at each time step is to pick an action based on past actions and observations. 
 The possible set of conditional plans $\beta^h$ correspond to a decision tree.
@@
 %From here on we assume this finite set of observation partitioning as $\mathcal{O}$.
-
-We define the value iteration algorithm for discrete observations and discrete states that finds the optimal value for each horizon starting with an initial belief.
-Although the number of belief states is infinite, the optimal value function for $h$ is proven to be a piecewise linear and convex function of the belief state \cite{smallwoodSondik}.
-The optimal value function is the maximum value over a finite set of alpha vectors  $\alpha^h_i$:%that define the hyperplanes:
-\begin{equation}
-V^h_{\Pi^*}(\vec{b}) = \max_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot \vec{b}
-\end{equation}
-where the belief states $\vec{b}$ and each $\alpha^h_i \in \Gamma^h$ are of dimension $|\mathcal{S}|$.
-At each of the $h$-steps the value function is parameterized by the set of  $\alpha$-vectors that partition the belief state.
-At a particular belief point, we can compute the next value function: 
-\begin{align*}
-V^{h+1}(b) = \max_a \left( b \cdot r_a + \gamma \sum_o p(xd_o|a,b)V^h(b|xd_o,a)\right)\nonumber
-\end{align*}
-Replacing the previous iteration value with (3) and also updating the belief after executing action $a$ and observing $xd_o$ results in the following equation: 
-\begin{align*}
-V^{h+1}(b) = \max_a \left( b \cdot r_a + 
-\gamma \sum_o \max_{\langle g^h_{a,o,j} \rangle_j}  b \cdot g^h_{a,o,j} \right)\nonumber
-\end{align*}
-where $r_a = r(x_s,d_s,a)$ and  
-$g^h_{a,o,j} =  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s)$
-
-Calculating all possible ways of building the value function independent of a certain belief state can define the optimal value function.
-To derive the $\Gamma$-set of all possible $\alpha$-vectors for each horizon, we use the value iteration algorithm from ~\cite{monahan82}.
-Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (4):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as
-$\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
-\textrm{\large $\boxplus$} Q = \{ \vec{p} + \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}
-\begin{align}
-\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \\
-\Gamma^h  &= \bigcup_a \Gamma^h_a 
-\end{align}
-Each action-dependent $\Gamma$ sums over the possible set of $g^h_{a,o,j}(\cdot)$ given an observation model to derive discrete observations $xd_o$ and for all $\alpha^{h-1}_{j} \in \Gamma^{h-1} $.
-The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.
-For each horizon, the optimal $\alpha$-vector is obtained from the maximum over all $\alpha$-vectors: 
-$\alpha^* = \argmax_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot b $
-
-Here the cross-sum operation takes the set of $\alpha$ -vectors for each observation and produces a sum over vectors that depends on the number of previous iteration vectors.
-Thus the $\alpha$-vector set grows exponentially in the number of observations, i.e., $|\Gamma^{h}| = |\mathcal{A}||\Gamma^{h-1}|^{|\mathcal{O}|}$ during the cross-sum operation.
-For this reason even for a small number of discrete observations, tractable solutions may be hard to define.
-Here we do not restrict our observations to discrete ones, but extend it to a fully continuous approach where continuous observations and states are allowed in the DC-POMDP.
-We use the number of  observation partitions $|\mathcal{O}|$ as the branching factor in the policy tree.
-We present the symbolic value iteration algorithm in the case of continuous observations using relevant observation partitioning of this continuous space.
-
-\section{Symbolic Dynamic Programming} 
-Dynamic programming provides exact solutions to POMDP problems for small domains.
-As the domain grows, fully enumerating the state, action and observation space leads to intractable solutions even for low horizons.
-Avoiding this enumeration often requires approximating the solutions.
-We claim an exact solution using symbolic dynamic programming.
-The SDP solution for fully observable domains has been presented in \cite{sanner_uai11} %using piecewise constant functions
-Our symbolic algorithm deals with continuous observations and states by generating the relevant observation partition at every time step and performing the backup operation using $\alpha$-vectors.
-We define the requirements to implement the value iteration algorithm for DC-POMDP.
-
-\subsection{Case Representation and Extended ADDs}
-% operations, max, restrict, substitute
-%overview + example plant
-The \textsc{Power Plant} example represented all functions using the case structure which can be generally defined as:
-{%\footnotesize 
-\begin{align}
-f = 
-\begin{cases}
-  \phi_1: & f_1 \\ 
- \vdots&\vdots\\ 
-  \phi_k: & f_k \\ 
-\end{cases} \nonumber
-\end{align}
-}
-Where $\phi_i$ are disjoint logical formulae defined over the state $(\vec{d},\vec{x})$ with logical ($\land,\lor,\neg$) combinations of boolean variables and inequalities ($\geq,>,\leq,<$) over continuous variables.
-The $f_i$ are function definitions either linear or quadratic in the continuous parameters.
-
-For \emph{unary operations} such as scalar multiplication $c\cdot f$ (for some constant $c \in \mathbb{R}$) or negation $-f$ on case statements is simply to apply the operation on each case partition $f_i$ ($1 \leq i \leq k$).
-
-A \emph{binary operation} on two case statements, takes the cross-product of the logical partitions of each case statement and performs the corresponding operation on the resulting paired partitions.
-The cross-sum $\oplus$ of two cases is defined as the following:
-{\footnotesize 
-\begin{center}
-\begin{tabular}{r c c c l}
-&
-\hspace{-6mm} 
-  $\begin{cases}
-    \phi_1: & f_1 \\ 
-    \phi_2: & f_2 \\ 
-  \end{cases}$
-$\oplus$
-&
-\hspace{-4mm}
-  $\begin{cases}
-    \psi_1: & g_1 \\ 
-    \psi_2: & g_2 \\ 
-  \end{cases}$
-&
-\hspace{-2mm} 
-$ = $
-&
-\hspace{-2mm}
-  $\begin{cases}
-  \phi_1 \wedge \psi_1: & f_1 + g_1 \\ 
-  \phi_1 \wedge \psi_2: & f_1 + g_2 \\ 
-  \phi_2 \wedge \psi_1: & f_2 + g_1 \\ 
-  \phi_2 \wedge \psi_2: & f_2 + g_2 \\ 
-  \end{cases}$
-\end{tabular}
-\end{center}
-}
-\normalsize
-Likewise $\ominus$ and $\otimes$ are defined by subtracting or multiplying partition values.  Inconsistent partitions can be discarded when they are irrelevant to the function value.
-The next three operations are required in defining the value iteration algorithm. A \emph{symbolic case maximization} is defined as below:
-\vspace{-4mm}
-
-{\footnotesize
-\begin{center}
-\begin{tabular}{r c c c l}
-&
-\hspace{-7mm} $\mathrm{max} \Bigg(
-  \begin{cases}
-    \phi_1: \hspace{-2mm} & \hspace{-2mm} f_1 \\ 
-    \phi_2: \hspace{-2mm} & \hspace{-2mm} f_2 \\ 
-  \end{cases}$
-$,$
-&
-\hspace{-4mm}
-  $\begin{cases}
-    \psi_1: \hspace{-2mm} & \hspace{-2mm} g_1 \\ 
-    \psi_2: \hspace{-2mm} & \hspace{-2mm} g_2 \\ 
-  \end{cases} \Bigg)$
-&
-\hspace{-4mm} 
-$ = $
-&
-\hspace{-4mm}
-  $\begin{cases}
-  \phi_1 \wedge \psi_1 \wedge f_1 > g_1    : & \hspace{-2mm} f_1 \\ 
-  \phi_1 \wedge \psi_1 \wedge f_1 \leq g_1 : & \hspace{-2mm} g_1 \\ 
-  \phi_1 \wedge \psi_2 \wedge f_1 > g_2    : & \hspace{-2mm}f_1 \\ 
-  \phi_1 \wedge \psi_2 \wedge f_1 \leq g_2 : & \hspace{-2mm} g_2 \\ 
-  %\vdots & \vdots
-  \phi_2 \wedge \psi_1 \wedge f_2 > g_1    : & \hspace{-2mm} f_2 \\ 
-  \phi_2 \wedge \psi_1 \wedge f_2 \leq g_1 : & \hspace{-2mm} g_1 \\ 
-  \phi_2 \wedge \psi_2 \wedge f_2 > g_2    : & \hspace{-2mm} f_2 \\ 
-  \phi_2 \wedge \psi_2 \wedge f_2 \leq g_2 : & \hspace{-2mm} g_2 \\ 
-  \end{cases}$
-\end{tabular}
-\end{center}
-}
-\emph{Restriction} takes a function $f$ to restrict only in cases
-that satisfy some formula $\phi$, which we write as $f|_{\phi}$.  
-This can be done by simply appending $\phi$ to each case partition
-as follows:
-{\footnotesize
-\begin{center}
-\begin{tabular}{r c c l}
-&
-\hspace{-6mm} 
-  $f = \begin{cases}
-    \phi_1: & f_1 \\ 
-   \vdots&\vdots\\ 
-    \phi_k: & f_k \\ 
-  \end{cases}$
-&
-
-&
-\hspace{-2mm}
-  $f|_{\phi} = \begin{cases}
-    \phi_1 \land \phi : & f_1 \\ 
-   \vdots&\vdots\\ 
-    \phi_k \land \phi : & f_k \\ 
-  \end{cases}$
-\end{tabular}
-\end{center}
-}
-
-\emph{Symbolic substitution} simply takes a set $\sigma$ of variables and their substitutions, where
-the LHS of the substitution operator $/$ represents the substitution variable and the
-RHS of the $/$ represents the expression that should be substituted in its place.
-Hence to perform a continuous regression on a more general
-representation, we obtain that $\int_{x_j'} P(x_j'|\cdots) V'^{h} dx_j'$
-\begin{align*}
-    = \begin{cases}
-    \phi_1: & V'^{h} \{ x_j' = f_1 \} \\ 
-   \vdots&\vdots\\ 
-    \phi_k: & V'^{h} \{ x_j' = f_k \}  \\ 
-  \end{cases}
-\end{align*}
-
-%xadd representation
-The data structure of \emph{extended ADD} (XADD)~\cite{sanner_uai11} is used to support
-case statements and the required operations.
-Here we use continuous SDP for both states and observations in DC-POMDPs.
-The value functions and $\alpha$-vectors partition into states of equal value, both of which can be represented using XADDs.
-For increasing efficiency linear constraint feasibility checkers such as LP solvers allow pruning unreachable paths in the XADD.
-The right picture in Figure 1 is an example of an XADD representation.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+For discrete observations and discrete states the number of belief states is infinite, but the optimal value function for $h$ is proven to be a piecewise linear and convex function of the belief state \cite{smallwoodSondik}.
+The optimal value function is the maximum value over a finite set of alpha vectors  $\alpha^h_i$:%that define the hyperplanes:
+{\footnotesize
+\begin{equation}
+V^h_{\Pi^*}(\vec{b}) = \max_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot \vec{b}
+\end{equation}
+}
+where the belief states $\vec{b}$ and each $\alpha^h_i \in \Gamma^h$ are of dimension $|\mathcal{S}|$.
+
+Point-based dynamic programming backups compute the optimal value function given a belief state $b$.
+The best conditional plan for the given belief, corresponds to an $\alpha$-vector which is computed using point-based backups:
+{\footnotesize
+\begin{align*}
+\alpha^{h+1}(b) = R(b) + \gamma \int_{xd_o} P(xd_o|a,b)\alpha^{h+1}(b)\nonumber
+\end{align*}
+}
+where $P(xd_o|a,b) = \int_{xd_s}\int_{xd'_s}b(s)P(xd'_s|xd_s,a)P(xd_o|xd'_s,a)$.
+
+Calculating all possible ways of building the value function independent of a certain belief state can also define the optimal value function.
+To derive the $\Gamma$-set of all possible $\alpha$-vectors for each horizon, we use the value iteration algorithm from ~\cite{monahan82}.
+Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (2):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as
+$\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
+\textrm{\large $\boxplus$} Q = \{ \vec{p} + \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}
+{\footnotesize
+\begin{align}
+V^{h+1}(b) &= \max_a \left( b \cdot r_a + 
+\gamma \sum_o \max_{\langle g^h_{a,o,j} \rangle_j}  b \cdot g^h_{a,o,j} \right) \\
+\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \\
+\Gamma^h  &= \bigcup_a \Gamma^h_a 
+\end{align}
+}
+where $r_a = r(x_s,d_s,a)$ and  
+$g^h_{a,o,j} =  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s)$
+%Each action-dependent $\Gamma$ sums over the possible set of $g^h_{a,o,j}(\cdot)$ given an observation model to derive discrete observations $xd_o$ and for all $\alpha^{h-1}_{j} \in \Gamma^{h-1} $.
+The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.
+Lines 9--16 of the \textsc{VI} algorithm show the above in a general algorithm used for both discrete and continuous observations. 
+For all $\alpha$-vectors in the current $\Gamma$-set first an LP-solver is used to prune any inconsistency in the branches of this vector in line 17.
+A pairwise dominance test is performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$
+The result of this pairwise test is a dominant $\alpha$-vector which is added to the set of $\Gamma^h$. 
+%Here the cross-sum operation takes the set of $\alpha$ -vectors for each observation and produces a sum over vectors that depends on the number of previous iteration vectors.
+
+Note that the $\alpha$-vector set grows exponentially in the number of observations, i.e., $|\Gamma^{h}| = |\mathcal{A}||\Gamma^{h-1}|^{|\mathcal{O}|}$ during the cross-sum operation.
+For this reason even for a small number of discrete observations, tractable solutions may be hard to define.
+Here we do not restrict our observations to discrete ones, but extend it to a fully continuous approach where lines 7--8 will be executed to generate relevant observation partitions based on a belief state $b_i$.
+We use the number of  observation partitions $|\mathcal{O}|$ as the branching factor in the policy tree.
+We present the symbolic approach required to partition this continuous space.
+
+\section{Symbolic Dynamic Programming} 
+Dynamic programming provides exact solutions to POMDP problems for small domains.
+As the domain grows, fully enumerating the state, action and observation space leads to intractable solutions even for low horizons.
+Avoiding this enumeration often requires approximating the solutions.
+We claim an exact solution using symbolic dynamic programming.
+The SDP solution for fully observable domains has been presented in \cite{sanner_uai11} %using piecewise constant functions
+Our symbolic algorithm deals with continuous observations and states by generating the relevant observation partition at every time step and performing the backup operation using $\alpha$-vectors.
+We define the requirements to implement the value iteration algorithm for DC-POMDP.
+
+\subsection{Case Representation and Extended ADDs}
+% operations, max, restrict, substitute
+%overview + example plant
+The \textsc{Power Plant} example represented all functions using the case structure which can be generally defined as:
+{\footnotesize 
+\begin{align}
+f = 
+\begin{cases}
+  \phi_1: & f_1 \\ 
+ \vdots&\vdots\\ 
+  \phi_k: & f_k \\ 
+\end{cases} \nonumber
+\end{align}
+}
+Where $\phi_i$ are disjoint logical formulae defined over the state $(\vec{d},\vec{x})$ with logical ($\land,\lor,\neg$) combinations of boolean variables and inequalities ($\geq,>,\leq,<$) over continuous variables.
+The $f_i$ are function definitions either linear or quadratic in the continuous parameters.
+
+For \emph{unary operations} such as scalar multiplication $c\cdot f$ (for some constant $c \in \mathbb{R}$) or negation $-f$ on case statements is simply to apply the operation on each case partition $f_i$ ($1 \leq i \leq k$).
+A \emph{binary operation} on two case statements, takes the cross-product of the logical partitions of each case statement and performs the corresponding operation on the resulting paired partitions.
+The cross-sum $\oplus$ of two cases is defined as the following:
+{\footnotesize 
+\begin{center}
+\begin{tabular}{r c c c l}
+&
+\hspace{-6mm} 
+  $\begin{cases}
+    \phi_1: & f_1 \\ 
+    \phi_2: & f_2 \\ 
+  \end{cases}$
+$\oplus$
+&
+\hspace{-4mm}
+  $\begin{cases}
+    \psi_1: & g_1 \\ 
+    \psi_2: & g_2 \\ 
+  \end{cases}$
+&
+\hspace{-2mm} 
+$ = $
+&
+\hspace{-2mm}
+  $\begin{cases}
+  \phi_1 \wedge \psi_1: & f_1 + g_1 \\ 
+  \phi_1 \wedge \psi_2: & f_1 + g_2 \\ 
+  \phi_2 \wedge \psi_1: & f_2 + g_1 \\ 
+  \phi_2 \wedge \psi_2: & f_2 + g_2 \\ 
+  \end{cases}$
+\end{tabular}
+\end{center}
+}
+Likewise $\ominus$ and $\otimes$ are defined by subtracting or multiplying partition values.  Inconsistent partitions can be discarded when they are irrelevant to the function value.
+The next three operations are required in defining the value iteration algorithm. A \emph{symbolic case maximization} is defined as below:
+\vspace{-4mm}
+{\footnotesize
+\begin{center}
+\begin{tabular}{r c c c l}
+&
+\hspace{-7mm} $\mathrm{max} \Bigg(
+  \begin{cases}
+    \phi_1: \hspace{-2mm} & \hspace{-2mm} f_1 \\ 
+    \phi_2: \hspace{-2mm} & \hspace{-2mm} f_2 \\ 
+  \end{cases}$
+$,$
+&
+\hspace{-4mm}
+  $\begin{cases}
+    \psi_1: \hspace{-2mm} & \hspace{-2mm} g_1 \\ 
+    \psi_2: \hspace{-2mm} & \hspace{-2mm} g_2 \\ 
+  \end{cases} \Bigg)$
+&
+\hspace{-4mm} 
+$ = $
+&
+\hspace{-4mm}
+  $\begin{cases}
+  \phi_1 \wedge \psi_1 \wedge f_1 > g_1    : & \hspace{-2mm} f_1 \\ 
+  \phi_1 \wedge \psi_1 \wedge f_1 \leq g_1 : & \hspace{-2mm} g_1 \\ 
+  \phi_1 \wedge \psi_2 \wedge f_1 > g_2    : & \hspace{-2mm}f_1 \\ 
+  \phi_1 \wedge \psi_2 \wedge f_1 \leq g_2 : & \hspace{-2mm} g_2 \\ 
+  \vdots & \vdots
+%  \phi_2 \wedge \psi_1 \wedge f_2 > g_1    : & \hspace{-2mm} f_2 \\ 
+%  \phi_2 \wedge \psi_1 \wedge f_2 \leq g_1 : & \hspace{-2mm} g_1 \\ 
+%  \phi_2 \wedge \psi_2 \wedge f_2 > g_2    : & \hspace{-2mm} f_2 \\ 
+%  \phi_2 \wedge \psi_2 \wedge f_2 \leq g_2 : & \hspace{-2mm} g_2 \\ 
+  \end{cases}$
+\end{tabular}
+\end{center}
+}
+\emph{Restriction} takes a function $f$ to restrict only in cases
+that satisfy some formula $\phi$, which we write as $f|_{\phi}$.  
+This can be done by simply appending $\phi$ to each case partition
+as the left side shows:
+{\footnotesize
+\begin{center}
+\begin{tabular}{r c c l}
+%&
+%\hspace{-6mm} 
+%  $f = \begin{cases}
+%    \phi_1: & f_1 \\ 
+%   \vdots&\vdots\\ 
+%    \phi_k: & f_k \\ 
+%  \end{cases}$
+%&
+&
+\hspace{0mm}
+  $f|_{\phi} = \begin{cases}
+    \phi_1 \land \phi : & f_1 \\ 
+   \vdots&\vdots\\ 
+    \phi_k \land \phi : & f_k \\ 
+  \end{cases}$
+&
+\hspace{15mm}
+$\int_{x_j'} P(x_j'|\cdots) V'^{h} dx_j' = \begin{cases}
+    \phi_1: & V'^{h} \{ x_j' = f_1 \} \\ 
+   \vdots&\vdots\\ 
+    \phi_k: & V'^{h} \{ x_j' = f_k \}  \\ 
+  \end{cases}$
+  \end{tabular}
+\end{center}
+}
+\emph{Symbolic substitution} simply takes a set $\sigma$ of variables and their substitutions, where
+the LHS of the substitution operator $/$ represents the substitution variable and the
+RHS of the $/$ represents the expression that should be substituted in its place.
+Hence to perform a continuous regression on a more general
+representation, we obtain the right side of the above.
+
+%xadd representation
+The data structure of \emph{extended ADD} (XADD)~\cite{sanner_uai11} is used to support
+case statements and the required operations.
+Here we use continuous SDP for both states and observations in DC-POMDPs.
+The value functions and $\alpha$-vectors partition into states of equal value, both of which can be represented using XADDs.
+For increasing efficiency linear constraint feasibility checkers such as LP solvers allow pruning unreachable paths in the XADD.
+The right picture in Figure 1 is an example of an XADD representation.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Value iteration for DC-POMDPs}
+\subsection{PBVI for DC State and Discrete Observations}
 
-For continuous state and discrete observations, we use the symbolic version of Monahan's algorithm as the DC-POMDP solution.
-The same approach is required for continuous observations only based on point-based dynamic programming backups[].
-Here for a given belief, the best conditional plan corresponding to an $\alpha$-vector is computed using point-based backups:
-\begin{align*}
-\alpha^{h+1}(b) = R(b) + \gamma \int_{xd_o} P(xd_o|a,b)\alpha^{h+1}(b)\nonumber
-\end{align*}
-where $P(xd_o|a,b) = \int_{xd_s}\int_{xd'_s}b(s)P(xd'_s|xd_s,a)P(xd_o|xd'_s,a)$
+For discrete and continuous states and discrete observations, we use the symbolic version of Monahan's algorithm as the DC-POMDP solution.
+The same approach is required for continuous observations only based on point-based dynamic programming backups.
 
 Given an initial set of beliefs, we can compute the relevant observation partitions explained in the next section and then continue in a similar fashion of having discrete observations.
-To visualize we explain of the value iteration algorithm as defined in \texttt{VI} for DC-POMDPs.
+%To visualize we explain of the value iteration algorithm as defined in \texttt{VI} for DC-POMDPs.
 Each step of the algorithm is explained using the \textsc{Power plant} example for the first three horizons.
@@
 
-For $h=0$, $V^0$ is given in line 2.
+For each iteration, the $\alpha$-vector set are case functions that can be primed (line 11--12). 
+%For $h=0$, $V^0$ is given in line 2.
 For $h=1$ we choose an action, for example $close$ to start with.
@@
 Line 16 defines the final $\Gamma$-set:
-$\Gamma^1 = \{ \alpha^1_{1,open}, \alpha^1_{1,close} \}$.
-
-The \texttt{Backup} algorithm is defined in a separate algorithm on the next page which takes the primed version of the variables and performs regression on the $\alpha$-vectors. 
-\emph{Boolean restriction} is defined in lines 5--8 where $f|_{b=v}$ restricts a boolean variable $b$ occurring in $f$ to values $v \in \{ 0,1 \}$.
+%$\Gamma^1 = \{ \alpha^1_{1,open}, \alpha^1_{1,close} \}$.
+The \texttt{Backup} algorithm in line 13 takes the primed version of the variables and performs regression on the $\alpha$-vectors. 
+\emph{Boolean restriction} is defined where $f|_{b=v}$ restricts a boolean variable $b$ occurring in $f$ to values $v \in \{ 0,1 \}$.
 This can be done by simply appending $b=v$ to each case partition (logical $\wedge$ of $b=v$ and the logical first-order formulas of $f$).
 
-\emph{Continuous regression} $\int f(x_{s_j}') \otimes P(x_{s_j}'|\cdots) dx_{s_j}'$ in line 3--4 has the form $\delta[x_{s_j}' - h(\vec{z})]$ for the probability over the next state, where $h(\vec{z})$ is a case statement and $\vec{z}$ does not contain
+\emph{Continuous regression} $\int f(x_{s_j}') \otimes P(x_{s_j}'|\cdots) dx_{s_j}'$ has the form $\delta[x_{s_j}' - h(\vec{z})]$ for the probability over the next state, where $h(\vec{z})$ is a case statement and $\vec{z}$ does not contain
 $x_{s_j}'$.  A symbolic substitution is defined for this integration such that any occurrence of $x_{s_j}'$ in $f(x_{s_j}')$ is substituted with $h(\vec{z})$: 
@@
 
-For $h=2$, for each of the discrete actions, first we generate the probability of the relevant set of observations in the \texttt{GenRelObs} algorithm.
+%For $h=2$, for each of the discrete actions, first we generate the probability of the relevant set of observations in the \texttt{GenRelObs} algorithm.
 This has been done for the $close$ action in the next section and we use the results here.
-\begin{align}
-P(z_k)=
-\begin{cases}
- (2<t_o<6) &: 0.2 \\
-(-4<t_o<2) &: 0.6\\
-(-8<t_o<-4) &:0.1999
-\end{cases} 
-\nonumber
-\end{align}
-Next for each of the relevant partitions and both $\alpha$-vectors of the previous $\Gamma$ set lines 9--16 are executed.
-Line 11 will substitute prime values for all state variables.
-Line 13 produces the backup $\alpha$-vector and multiplies it by the weight derived from the relevant observation partition.
-For each observation, two vectors are formed for each corresponding $\alpha$-vector.
-For example line 13 for the observation partitions above and $\alpha_1^1,\alpha_2^1$ is defined as:
-\begin{align*}
-\Gamma_{close,z_1,1}^2= 
-\begin{cases} 
-(t_s'>15) &: -600 \\ 
-\neg(t_s'>15) &: 60 \\
-\end{cases}
-\hspace{15mm} 
-\Gamma_{close,z_1,2}^2= \top &:-0.6
-\\
-\Gamma_{close,z_2,1}^2= 
-\begin{cases}
-(t_s'>15) &: -199.999 \\ 
-\neg(t_s'>15) &: 19.999 
-\end{cases}
-\hspace{12mm} 
-\Gamma_{close,z_2,2}^2= \top &:-0.1999
-\\
-\Gamma_{close,z_3,1}^2= 
-\begin{cases}
-(t_s'>15) &: -200 \\ 
-\neg(t_s'>15) &: 20 
-\end{cases}
-\hspace{15mm} 
-\Gamma_{close,z_3,2}^2= \top &:-0.2
-\end{align*}
-Next in line 14 we have to take the cross-sum of the two sets which takes every possible combination of observations according to the $\alpha$-vectors. 
-\begin{align*}
-\Gamma_{close,1}^2 &= 
-\begin{cases} 
-(t_s'>15) &: -800.2 \\ 
-\neg(t_s'>15) &: 79.8 \\
-\end{cases}
-\hspace{10mm} 
-\Gamma_{close,2}^2= 
-\begin{cases} 
-(t_s'>15) &: -1000 \\ 
-\neg(t_s'>15) &: 100 \\
-\end{cases}
-\\
-\Gamma_{close,3}^2&= 
-\begin{cases} 
-(t_s'>15) &: -800.2 \\ 
-\neg(t_s'>15) &: 79.8 \\
-\end{cases}
-\hspace{10mm}
-\Gamma_{close,4}^2= 
-\begin{cases} 
-(t_s'>15) &: -600.4 \\ 
-\neg(t_s'>15) &: 59.6 \\
-\end{cases}
-\\
-\Gamma_{close,5}^2&= 
-\begin{cases} 
-(t_s'>15) &: -400.5999 \\ 
-\neg(t_s'>15) &: 39.3999 \\
-\end{cases}
-\hspace{5mm} 
-\Gamma_{close,6}^2= 
-\begin{cases} 
-(t_s'>15) &: -200.7999 \\ 
-\neg(t_s'>15) &: 19.999 \\
-\end{cases}
-\\
-\Gamma_{close,7}^2&= 
-\begin{cases} 
-(t_s'>15) &: -200.8 \\ 
-\neg(t_s'>15) &: 19.2 \\
-\end{cases}
-\hspace{10mm} 
-\Gamma_{close,8}^2= \top:-1
-\end{align*}
-All 8 vectors are multiplied by the discount $\gamma$ and added with the reward in line 15 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
+%\begin{align}
+%P(z_k)=
+%\begin{cases}
+% (2<t_o<6) &: 0.2 \\
+%(-4<t_o<2) &: 0.6\\
+%(-8<t_o<-4) &:0.1999
+%\end{cases} 
+%\nonumber
+%\end{align}
+%Next for each of the relevant partitions and both $\alpha$-vectors of the previous $\Gamma$ set lines 9--16 are executed. Line 11 will substitute prime values for all state variables. 
+Line 13 produces the backup $\alpha$-vector and multiplies it by the weight derived from the relevant observation partition in the continuous case or observation probabilities in the problem description.
+%For each observation, two vectors are formed for each corresponding $\alpha$-vector. For example line 13 for the observation partitions above and $\alpha_1^1,\alpha_2^1$ is defined as:
+%\begin{align*}
+%\Gamma_{close,z_1,1}^2= 
+%\begin{cases} 
+%(t_s'>15) &: -600 \\ 
+%\neg(t_s'>15) &: 60 \\
+%\end{cases}
+%\hspace{15mm} 
+%\Gamma_{close,z_1,2}^2= \top &:-0.6
+%\\
+%\Gamma_{close,z_2,1}^2= 
+%\begin{cases}
+%(t_s'>15) &: -199.999 \\ 
+%\neg(t_s'>15) &: 19.999 
+%\end{cases}
+%\hspace{12mm} 
+%\Gamma_{close,z_2,2}^2= \top &:-0.1999
+%\\
+%\Gamma_{close,z_3,1}^2= 
+%\begin{cases}
+%(t_s'>15) &: -200 \\ 
+%\neg(t_s'>15) &: 20 
+%\end{cases}
+%\hspace{15mm} 
+%\Gamma_{close,z_3,2}^2= \top &:-0.2
+%\end{align*}
+Line 14 takes the cross-sum of the two sets which takes every possible combination of observations according to the $\alpha$-vectors. 
+%\begin{align*}
+%\Gamma_{close,1}^2 &= 
+%\begin{cases} 
+%(t_s'>15) &: -800.2 \\ 
+%\neg(t_s'>15) &: 79.8 \\
+%\end{cases}
+%\hspace{10mm} 
+%\Gamma_{close,2}^2= 
+%\begin{cases} 
+%(t_s'>15) &: -1000 \\ 
+%\neg(t_s'>15) &: 100 \\
+%\end{cases}
+%\\
+%\Gamma_{close,3}^2&= 
+%\begin{cases} 
+%(t_s'>15) &: -800.2 \\ 
+%\neg(t_s'>15) &: 79.8 \\
+%\end{cases}
+%\hspace{10mm}
+%\Gamma_{close,4}^2= 
+%\begin{cases} 
+%(t_s'>15) &: -600.4 \\ 
+%\neg(t_s'>15) &: 59.6 \\
+%\end{cases}
+%\\
+%\Gamma_{close,5}^2&= 
+%\begin{cases} 
+%(t_s'>15) &: -400.5999 \\ 
+%\neg(t_s'>15) &: 39.3999 \\
+%\end{cases}
+%\hspace{5mm} 
+%\Gamma_{close,6}^2= 
+%\begin{cases} 
+%(t_s'>15) &: -200.7999 \\ 
+%\neg(t_s'>15) &: 19.999 \\
+%\end{cases}
+%\\
+%\Gamma_{close,7}^2&= 
+%\begin{cases} 
+%(t_s'>15) &: -200.8 \\ 
+%\neg(t_s'>15) &: 19.2 \\
+%\end{cases}
+%\hspace{10mm} 
+%\Gamma_{close,8}^2= \top:-1
+%\end{align*}
+All vectors are multiplied by the discount $\gamma$ and added with the reward in line 15 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
 % need to add this?
-The same operation is performed for the $open$ action and then in line 15 the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors for the next horizon.
-
-In line 17 this final set is using in the \texttt{Prune} algorithm.
+The same operation is performed for all actions and then in line 15 the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors for the next horizon. 
+%In line 17 this final set is using in the \texttt{Prune} algorithm.
 For all $\alpha$-vectors in the current $\Gamma$-set first an LP-solver is used to prune any inconsistency in the branches of this vector.
-Next a pairwise dominance test is performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$
-The result of this pairwise test is a dominant $\alpha$-vector which is added to the set of $\Gamma^h$. 
+Next a pairwise dominance test is performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$The result of this pairwise test is a dominant $\alpha$-vector which is added to the set of $\Gamma^h$.
 In the last step, we find the $\alpha$-vector that maximizes the value-function with respect to this belief and save it as the new $\Gamma$-set for the next horizon.
@@
 In the next section we discuss the main part of our algorithm which is generating relevant observation partitions.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\incmargin{.5em}
+%\linesnumbered
+%\begin{algorithm}[t!]
+%\vspace{-.5mm}
+%\dontprintsemicolon
+%\Begin{
+%   
+%    \emph{// Continuous regression, marginal integration}\\
+%    \For {all $xd_{s_j}'$ in $\alpha_{j,o}^{'h}$}
+%    {
+%         $G_a^h := \int P(z) \otimes P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a) \, d_{x_{s_j}'}$\;
+%    }
+%    \emph{// Discrete regression, marginal summation}\\
+%    \For {all $d_{s_i}'$ in $\alpha_{j,o}^{'h}$}
+%    {
+%         $G_a^h := \left[ G_a^h \otimes P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \right]|_{d_{s_i}' = 1}$\\
+%         \hspace{8mm} $\oplus \left[ G_a^h \otimes P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \right]|_{d_{s_i}'= 0}$\;
+%    }
+%    \Return{$G_a^h$} \;
+%}
+%\caption{\footnotesize \texttt{Backup}($\alpha_{j,o}^{'h},a,P(z)$) $\longrightarrow$ $G_a^h$ \label{alg:regress}}
+%\vspace{-1mm}
+%\end{algorithm}
+%\decmargin{.5em}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\incmargin{.5em}
+%\linesnumbered
+%\begin{algorithm}[t!]
+%\vspace{-.5mm}
+%\SetKwFunction{lpsolve}{LP-Solve}
+%\SetKwFunction{dominancy}{Dominance}
+%\Begin
+%{
+%   		$V^h :=0$ \;
+%       \ForEach {$\alpha_j^h \in \Gamma^h$}
+%       {
+%           $\alpha_j^h \,:=\,$\lpsolve{$\alpha_j^h$} \;
+%           \ForEach {$\alpha_k^h \in \Gamma^h$}
+%           {
+%				\If {$j \neq k$}  
+%				{         
+%           			$V^h \,:=\,\oplus$ \dominancy{$\alpha_j^h,\alpha_k^h$} \; 
+%           		}  
+%           	}
+%       }
+%     \Return{$(V^h)$} \;
+%}
+%\caption{\footnotesize \texttt{Prune}($\Gamma^h$) $\longrightarrow$ $(V^h)$ }
+%\vspace{-1mm}
+%\end{algorithm}
+%\decmargin{.5em}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\incmargin{.5em}
-\linesnumbered
-\begin{algorithm}[t!]
-\vspace{-.5mm}
-\dontprintsemicolon
-\Begin{
-
-    \emph{// Continuous regression, marginal integration}\\
-    \For {all $xd_{s_j}'$ in $\alpha_{j,o}^{'h}$}
-    {
-         $G_a^h := \int P(z) \otimes P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a) \, d_{x_{s_j}'}$\;
-    }
-    \emph{// Discrete regression, marginal summation}\\
-    \For {all $d_{s_i}'$ in $\alpha_{j,o}^{'h}$}
-    {
-         $G_a^h := \left[ G_a^h \otimes P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \right]|_{d_{s_i}' = 1}$\\
-         \hspace{8mm} $\oplus \left[ G_a^h \otimes P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \right]|_{d_{s_i}'= 0}$\;
-    }
-    \Return{$G_a^h$} \;
-}
-\caption{\footnotesize \texttt{Backup}($\alpha_{j,o}^{'h},a,P(z)$) $\longrightarrow$ $G_a^h$ \label{alg:regress}}
-\vspace{-1mm}
-\end{algorithm}
-\decmargin{.5em}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\incmargin{.5em}
-\linesnumbered
-\begin{algorithm}[t!]
-\vspace{-.5mm}
-\SetKwFunction{lpsolve}{LP-Solve}
-\SetKwFunction{dominancy}{Dominance}
-\Begin
-{
-   		$V^h :=0$ \;
-       \ForEach {$\alpha_j^h \in \Gamma^h$}
-       {
-           $\alpha_j^h \,:=\,$\lpsolve{$\alpha_j^h$} \;
-           \ForEach {$\alpha_k^h \in \Gamma^h$}
-           {
-				\If {$j \neq k$}  
-				{         
-           			$V^h \,:=\,\oplus$ \dominancy{$\alpha_j^h,\alpha_k^h$} \; 
-           		}  
-           	}
-       }
-     \Return{$(V^h)$} \;
-}
-\caption{\footnotesize \texttt{Prune}($\Gamma^h$) $\longrightarrow$ $(V^h)$ }
-\vspace{-1mm}
-\end{algorithm}
-\decmargin{.5em}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Continuous Observation Space} 
+\subsection{PBVI for DC State and DC Observations} 
 %progression definition
@@
 \emph{Progression} takes a formula $\psi$ and an action $o$ and represents the formula $\psi'$ that would hold after action $o$ is performed in a situation satisfying $\psi$.
+The \texttt{GenRelObs} algorithm defines the steps required to generate these partitions. 
+We use the second iteration of the \texttt{Power Plant} example to demonstrate partitioning of the entire continuous observation space.
+The main assumption is to consider some belief state and obtaining the relevant partitions according to that belief.
+We choose these beliefs such that they cover the entire probable state space; e.g.
+$b_1: U[t_s;2,6]$ and $b_2: U[t_s;6,13]$.
 
-The \texttt{GenRelObs} algorithm defines the operation of generating relevant observation partitions. 
-We use the second iteration of our \texttt{Power Plant} example to demonstrate the steps  required to partition the entire continuous observation space.
-This continuous nature requires assuming some belief state and obtaining the relevant partitions according to that belief.
 Starting with an empty set of relevant observation partitions $z$, the algorithm first eliminates the next state variable $xd'_s$.
-It multiplies the observation model and the transition model in the set of $\alpha$-vectors of the previous horizon.
-$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $
-This operation is similar to that of \texttt{Regress} plus the observation regressed model.
+It multiplies the observation model and the transition model in the set of $\alpha$-vectors of the previous horizon:
+$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $. 
+This operation is similar to that of \texttt{Regress} multiplied by the observation regressed model.
 For continuous variables the continuous regression and for discrete variables boolean restriction is used as defined in the previous section.
-Assume that after $h=1$ the set of $\alpha$-vectors are defined as:
+Assuming that after $h=1$ the set of $\alpha$-vectors are defined as:
+{\footnotesize
 \begin{align}
@@
 \end{align}
-For $h=2$ the resulting $\alpha$-vectors after this regression step are functions of the observation and current state: 
-\begin{align}
-\alpha_1^2(s,o) &= 
-\begin{cases}
- (t_s<15)\wedge (t_s - 10 < t_o<t_s) &: 10 \\
-(t_s>15)\wedge (t_s - 10 < t_o<t_s) &: -100  \\
-\neg(t_s - 10 < t_o<t_s) &: 0
-\end{cases}
-\hspace{0mm} 
-\alpha_2^2(s,o) &= \begin{cases}
-(t_s - 10 < t_o<t_s) &: -0.1 \\
-\neg(t_s - 10 < t_o<t_s) &: 0
-\end{cases}
-\nonumber
-\end{align}
-We need the $\alpha$-vectors to partition the observation space, thus next we eliminate the current state variable.
-During this integration the current state variable is factored out by multiplying the belief in each of the current $\alpha$-vectors.
-The resulting $\delta$-functions only depend on the observation: $\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$.
-For the belief $b_1: U[t_s;2,6]$ the two vectors above the resulting $\delta$-functions are defined below: 
-\begin{align}
-\delta_{1}(o) &= 
-\begin{cases}
- (2<t_o<6) &: 15 - 2.5*t_o \\
-(-4<t_o<2) &: 10 \\
-(-8<t_o<-4) &: 20 + 2.5*t_o 
-\end{cases}
-\hspace{5mm} 
-\delta_{2}(o) &= \begin{cases}
- (2<t_o<6) &: -15 + 0.025*t_o \\
-(-4<t_o<2) &: -0.1 \\
-(-8<t_o<-4) &: -0.2 - 0.025*t_o 
-\end{cases}
-\nonumber
-\end{align}
-The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_1$.
-According to continuous observation space defined in \cite{pascalPomdp}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space.
-They propose numerical solutions to finding the roots of any two observation dependant function.
-Here we use the symbolic power of the max-operator to find all the maximum boundary regions of two or more $\delta$-functions at the same time.
-For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximuum:
-\begin{align}
-\mathrm{max} \Bigg(\delta_{1}(o),\delta_{1}(o)\Bigg) &= 
-\begin{cases}
- (2<t_o<6) &: 15 - 2.5*t_o \\
-(-4<t_o<2) &: 10 \\
-(-8<t_o<-4) &: 20 + 2.5*t_o 
-\end{cases}
-\nonumber
-\end{align}
-The three partitions define 3 observation regions which we can now use in a similar fashion to a discrete observation set .
-Thus we only need to find the probability of each of the three distinct observation partitions to use in Monahan's algorithm.
-If we integrate out the state and observation from the observation model and the belief state regardless of a particular $\alpha$-vector, it integrates to one:
-$\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*P(s) = 1$
-
-Now that we have partitions on the observation space, we only need to multiply the indicators of each case partition in this formula to obtain the probability of each partition such that it sums to one:
-$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$
-For our example the following holds: 
-\begin{align}
-P(z_k)=
-\begin{cases}
- (2<t_o<6) &: 0.2 \\
-(-4<t_o<2) &: 0.6\\
-(-8<t_o<-4) &:0.1999
-\end{cases} 
-\nonumber
-\end{align}
-Hence we can now use the algorithms and methods of a discrete observation setting using the probabilities of the partitioned observation space. 
-%Figure ~\ref{fig:obsPartition} shows the observation partitions obtained in the 1D problem instance for different horizons and their related probabilities. In a lyx file, needed?
-Next we present some results for 2-Dimensional continuous observation spaces.
-
+}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+For $h=2$ the resulting $\alpha$-vectors after this regression step are functions of the observation and current state: 
+{\footnotesize
+\begin{align}
+\alpha_1^2(s,o) &= 
+\begin{cases}
+ (t_s<15)\wedge (t_s - 10 < t_o<t_s) &: 10 \\
+(t_s>15)\wedge (t_s - 10 < t_o<t_s) &: -100  \\
+\neg(t_s - 10 < t_o<t_s) &: 0
+\end{cases}
+\hspace{4mm} 
+\alpha_2^2(s,o) &= \begin{cases}
+(t_s - 10 < t_o<t_s) &: -0.1 \\
+\neg(t_s - 10 < t_o<t_s) &: 0
+\end{cases}
+\nonumber
+\end{align}
+}
+We need the $\alpha$-vectors to partition on the observation space, thus next we eliminate the current state variable.
+During this integration the current state variable is factored out by multiplying the belief in each of the current $\alpha$-vectors.
+The resulting $\delta$-functions only depend on the observation: $\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$.
+For the belief $b_1: U[t_s;2,6]$ the two vectors above the resulting $\delta$-functions are defined below: 
+{\footnotesize
+\begin{align}
+\delta_{1}(o) &= 
+\begin{cases}
+ (2<t_o<6) &: 15 - 2.5*t_o \\
+(-4<t_o<2) &: 10 \\
+(-8<t_o<-4) &: 20 + 2.5*t_o 
+\end{cases}
+\hspace{5mm} 
+\delta_{2}(o) &= \begin{cases}
+ (2<t_o<6) &: -15 + 0.025*t_o \\
+(-4<t_o<2) &: -0.1 \\
+(-8<t_o<-4) &: -0.2 - 0.025*t_o 
+\end{cases}
+\nonumber
+\end{align}
+}
+The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_1$.
+According to continuous observation space defined in \cite{pascalPomdp}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space.
+In this work, numerical solutions are proposed to find the roots of any two observation dependant function.
+Here we use the symbolic power of the max-operator to find all the maximum boundary regions of two or more $\delta$-functions at the same time.
+For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximuum:
+{\footnotesize
+\begin{align}
+\mathrm{max} \Bigg(\delta_{1}(o),\delta_{1}(o)\Bigg) &= 
+\begin{cases}
+ (2<t_o<6) &: 15 - 2.5*t_o \\
+(-4<t_o<2) &: 10 \\
+(-8<t_o<-4) &: 20 + 2.5*t_o 
+\end{cases}
+\nonumber
+\end{align}
+}
+The three partitions define 3 observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the three distinct observation partitions is found.
+If we integrate out the state and observation from the observation model and the belief state regardless of a particular $\alpha$-vector, it integrates to one:
+$\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*P(s) = 1$
+Thus we only need to multiply the indicators of each case partition in this formula to obtain the probability of each partition such that it sums to one: 
+$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$
+For our example the following holds: 
+{\footnotesize
+\begin{align}
+P(z_k)=
+\begin{cases}
+ (2<t_o<6) &: 0.2 \\
+(-4<t_o<2) &: 0.6\\
+(-8<t_o<-4) &:0.1999
+\end{cases} 
+\nonumber
+\end{align}
+}
+Hence we can now use the algorithms and methods of a discrete observation setting using the probabilities of the partitioned observation space in Monahan's algorithm. 
+%Figure ~\ref{fig:obsPartition} shows the observation partitions obtained in the 1D problem instance for different horizons and their related probabilities. In a lyx file, needed?
+Next we present some results for 2-Dimensional continuous observation spaces.
 
@@
 
-{\bf \textsc{2D-Power Plant}:} We consider the more complex model of the power plant similar to \cite{steam2} where the pressure inside the water tank must be controlled to avoid water mixing into the steam or explosion of the tank.
+{\bf \textsc{2D-Power Plant}:} We consider the more complex model of the power plant similar to \cite{steam2} where the pressure inside the water tank must be controlled to avoid mixing water into the steam or explosion of the tank.
 We model the pressure variable $p$ as a partially observable variable from the observation readings of the pressure valve $po$.
@@
 For the increase action the transition functions and the reward function is defined as below:
+{\footnotesize
 \begin{align}
@@
 \right]\nonumber
-\\
-P(t_s'|\vec{t_s},a)&= \delta\left[ t_s' -  t_s +10 \right]\nonumber
+\hspace{5mm} 
+P(t_s'|\vec{t_s},a)= \delta\left[ t_s' -  t_s +10 \right]\nonumber
 \\
@@
 \end{align}
+}
 There is a high reward for staying withing the safe temperature and pressure range since it produces power, else depending on how safe it is to have values higher or lower than the safe range, penalty is defined.
@@
 For the observation model we consider two continuous uniform distributions such as the following:  
+{\footnotesize
 \begin{align}
-P(t_o|t_s') &= 
+P(t_o|t_s') = 
 \begin{cases}
@@
 \end{cases}\nonumber
-\end{align}
-\begin{align}
-P(p_o|p_s') &= 
+\hspace{0mm} 
+P(p_o|p_s') = 
 \begin{cases}
@@
 \end{align}
-
-We define two rectangular uniform beliefs around the regions of rewarding, so that one needs to increase the values while the other should decrease them: 
-\begin{align}
-b_1 = 
-\begin{cases}
-(p_o>p_s) \wedge (p_o<p_s+10) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 100)&: 0.01 \\
- \neg( (p_o>p_s) \wedge (p_o<p_s+10) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 100)) &: 0 \\
-\end{cases}\nonumber
-\end{align}
-\begin{align}
-b_2 = 
-\begin{cases}
-(p_o>p_s +10) \wedge (p_o<p_s+30) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 130)&: 0.00125 \\
- \neg((p_o>p_s +10) \wedge (p_o<p_s+30) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 130)) &: 0 
-\end{cases}\nonumber
-\end{align}
+}
+We define two rectangular uniform beliefs around the regions of rewarding, so that one needs to increase the values while the other should decrease them: $b_1: U[t_s;90,100]*U[p_s;0,10]$ and $b_2: U[t_s;90,130]*U[p_s;10,30]$
+%\begin{align}
+%b_1 = 
+%\begin{cases}
+%(p_o>p_s) \wedge (p_o<p_s+10) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 100)&: 0.01 \\
+% \neg( (p_o>p_s) \wedge (p_o<p_s+10) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 100)) &: 0 \\
+%\end{cases}\nonumber
+%\end{align}
+%\begin{align}
+%b_2 = 
+%\begin{cases}
+%(p_o>p_s +10) \wedge (p_o<p_s+30) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 130)&: 0.00125 \\
+% \neg((p_o>p_s +10) \wedge (p_o<p_s+30) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 130)) &: 0 
+%\end{cases}\nonumber
+%\end{align}
 % I can draw this and the 1D belief
 In Figure ~\ref{fig:timeSpace}, a time and space analysis of
-the two versions of the problem have been performed for up to 6 horizons.
+the two versions of \textsc{Power Plant} have been performed for up to 6 horizons.
 Comparing the two problem sizes demonstrates the effect of the number of state-observation variables in our algorithm.
@@
 The stability in time is due to the fact that each stage almost takes the same time since it has converged quickly.%???
-Increase in the problem size, increases the partition numbers on the observation space and this produces more $\alpha$-vectors which also effecting the space required to perform the algorithm.
+Increase in the problem size, increases the partition numbers on the observation space and this produces more $\alpha$-vectors which also effects the space required to perform the algorithm.
 The number of vectors stays the same for most horizons and they drop after convergence in the 2D problem instance.
 The number of nodes in each iteration increases as more space is required to perform higher iterations.
+This shows that although the 2D instance takes more time and space than the 1D instance, it still converges within reasonable resources.
 
-In Figure ~\ref{fig:3D}  we show plots of the value function for different iterations of the 2D problem instance.
+In Figure ~\ref{fig:3D}  we present plots of the value function for different iterations of the 2D problem instance.
 Each plot demonstrates the value as a function of the pressure and temperature.
 Starting with the first iteration, the value is highest for the reward range ($5<p<15 \wedge 95<t<105$) and -1 or less for other places.
-In the fifth iteration, the value function has partitioned into more pieces showing how higher temperatures can increase the value without considering the effect of the pressure.
+In the fifth iteration, the value function has partitioned into more pieces, showing how higher temperatures can increase the value without considering the effect of the pressure.
 In the last plot, horizon 6 has better tuned the value function so that higher temperatures and pressures increase the value function and also within the reward range, finer grain partitions have been formed.
+Also note that the policy depends on the initial belief and for these plots they were based on $b_1$ which was lower than the reward range.
+The first action would then mean it has performed an increase while the other iterations decrease and increase respectively.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Related Work} 
+\section{Related Work and Conclusion} 
 This work has used the concept of continuous states and the SDP solution to MDPs from ~\cite{sanner_uai11} and continuous observations from the work of \cite{pascalPomdp}.
@@
 
-\section{Conclusion}
+%\section{Conclusion}
 We presented the first exact solution to continuous states and observations in a DC-POMDP framework.

### 2012-05-31 cpomdp/camdp-xadd/src/doc/NIPS/draft3-10page.tex
With additional pics for relevant partitions

@@
 \delta_{2}(o) &= \begin{cases}
- (2<t_o<6) &: -15 + 0.025*t_o \\
+ (2<t_o<6) &: -0.15 + 0.025*t_o \\
 (-4<t_o<2) &: -0.1 \\
@@
 }
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}[tbp!]
+\vspace{-6mm}
+\centering
+\hspace{-17mm}
+\includegraphics[width=0.41\textwidth]{pics/beliefs.pdf}
+\hspace{-12mm}
+\includegraphics[width=0.42\textwidth]{pics/delta1.pdf}
+\hspace{-12mm}
+\includegraphics[width=0.42\textwidth]{pics/delta2.pdf}
+\hspace{-17mm}
+\vspace{-10mm}
+\caption{\footnotesize 
+{\it (left)} Beliefs $b_1,b_2$ for 1D \texttt{Power Plant} example; 
+{\it (middle)} Observation dependent function for $b_1$ that partition the observation space into 3 regions with different probabilities $P(z_i)$ ; 
+{\it (right)} Relevant observation partitions and their probabilities for $b_2$.
+}
+\label{fig:timeSpace}
+\vspace{-4mm}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_1$.

### 2012-06-01 trunk/cpomdp/camdp-xadd/src/doc/NIPS/draft3-10page.tex
Updated main tex

--- 2012-05-31 cpomdp/camdp-xadd/src/doc/NIPS/draft3-10page.tex
@@
 
-\title{Exact Symbolic Dynamic Programming for Continuous POMDPs}
+\title{Symbolic Dynamic Programming for Continuous State and Observation POMDPs}
 
@@
 
+\newcommand{\open}{\mathit{open}}
+\newcommand{\close}{\mathit{close}}
 \newcommand{\fix}{\marginpar{FIX}}
@@
 \begin{abstract}
-Decision-theoretic planning searches for the optimal sequence of actions for a given uncertain environment.
-Partially Observable MDPs (POMDP) can model such environments where the underlying state is not be fully observable.
-%While discrete models of the environment is often used for planning, continuous models are naturally closer to the real-world.
-While previous work have provided solutions for discrete states and/or observations, here we define the first exact solution to both continuous state and observations.
-The symbolic dynamic programming approach in this paper avoids enumerating the state and observation space providing solutions for general multi-variate discrete and continuous POMDPs.
-Our main contribution is that for a fixed set of reachable belief states and a fixed set of actions, only a \textit{finite} partitioning of the continuous observation space is required for optimality.
-Using function representation of Affine algebraic decision diagrams (XADDs) and symbolic case operators we derive the automatic partitioning for piecewise linear dynamics %such as the reward, transition and observation function.
-We demonstrate empirical results on an operation research domain to show the \textit{first exact solution} to continuous state and observation POMDPs.  
-%to use continuous states with discrete observations and extending it to continuous observations by defining relevant observation partitions %associated with each state partition.
-%This allows us to perform the Bellman backup operation efficiently.
-%We address the problems in POMDP continuous planning and solutions to solve them symbolically. %To compactly represent the exact value functions Affine algebraic decision diagrams are used. 
+Partially-observable Markov decision processes (POMDPs) provide a
+powerful model for real-world sequential decision-making problems.  In
+recent years, point-based value iteration methods have proven to be
+extremely effective techniques for finding (approximately) optimal
+dynamic programming solutions to POMDPs when an initial set of belief
+states is known.  However, no point-based work has provided exact
+point-based backups for both continuous state and observation spaces,
+which we tackle in this paper.  Our key insight is that while there
+may be an infinite number of possible observations, there are only a
+finite number of observation partitionings that are relevant for
+optimal decision-making when a finite, fixed set of reachable belief
+states is known.  To this end, we make two important contributions:
+(1) we show how previous exact symbolic dynamic programming solutions
+for continuous state MDPs can be generalized to continuous state
+POMDPs with discrete observations, and (2) we show how this solution
+can be further extended via recently developed symbolic methods to
+continuous state and observations to \emph{derive} the minimal
+relevant observation partitioning for potentially correlated,
+multivariate observation spaces.
+%for domais with piecewise
+%polynomial transition and observation functions.  
+We demonstrate
+proof-of-concept results on uni- and multi-variate state and
+observation steam plant control.
 \end{abstract}
 
-\section{Introduction}
+\section{Introduction and Related Work}
 % Write intro here
-In many complex systems complete information of the state is not available due to noisy or poor sensors and readings of them.
-Partially observable Markov Decision Processes (POMDPs)\cite{POMDPfirst} can be used for planning under such uncertainty in such problems.
-The solution to a POMDP is to find an answer to the fully-observable belief-MDP, whose states are probability distributions over the state space of the original POMDP \cite{kaebling}
-Real-world resources and sensor inputs are often naturally modelled using continuous variables. 
-Unfortunately most POMDP techniques can not be applied to such settings without assuming prior knowledge about the problem domain.
-Most methods use some approximation to discretize the continuous state, observation or action space which increases computational complexity \cite{Thrun99h}.
-The optimal value function of belief-MDPs is piecewiselinear and convex in these discrete cases \cite{sondik71} but has also been proven in the setting of only continuous states.
-\cite{PerseusObs}
-
-Exact solutions to a multi-variate continuous state and observation setting has not been approached for the curse of dimensionality problem.
-In assuming a continuous observation setting, the agent must search for all possible observations that are read from the environment.
-This can quickly become infeasible even for small problem sizes.
-
-In this paper we provide an exact solution to a subset of POMDPs with continuous states and observations.
-Our symbolic dynamic programming (SDP) approach assumes discrete noise and piecewise linear dynamics and functions.
-We formulize our problems using a Discrete and Continuous POMDP (DC-POMDP) model and apply our SDP algorithm.
-The main problem that rises in any problem modeled using DC-POMDP is the definition of different observation partitions and their probability.
-We apply symbolic solutions such as integration and maximization to provide an optimal closed-form solution to our problem domains.
-As an example we consider the following power plant example which has mainly been solved in the literature using discrete sets of states and observations~\cite{steam2}.
-We use the example throughout the paper to help better understand the definitions.
-
-\textbf{Example} \textsc{(Power Plant)}
-\emph{The steam generation system of a power plant aims to provide hot steam to the steam turbine which in turn produces electricity.
-Feed-water is exposed to heat in the tubes leading to the water tank .
-Evaporating the water is performed under specific pressure and temperature inside the attached water tubes.
-Mixing water and steam pressures makes reading tank levels and temperatures very hard and uncertain.
-For this reason, the process is modeled using POMDP in the literature.
-The temperature $t_s \in \mathbb{R}$ is modelled as a continuous state variable.
-Noisy observation of the temperature $t_o \in \mathbb{R}$ is used to help decide the optimal action to either open or close the pressure valve $a \in \{open,close\}$.}
-
-We need to extend the basic SDP framework from \cite{sanner_uai11} by adding continuous observations to the model.
-The solution to a closed-form value function is similar to the algorithm from \cite{monahan82} apart from the fact that no explicit observation is defined in the continuous case.
-We redefine the value-iteration algorithm using case representations that are implemented using Extended Algebric decision diagrams (XADD) and provide an efficient automatic way to partition the entire observation space into piecewise linear functions where each partition has a relative probability.
-We use the univariate \textsc{Power Plant} example to define the main operations and algorithms and also provide empirical results for a multi-variate version of the problem.
+
+Partially-observable Markov decision processes (POMDPs) are a powerful
+modeling formalism for real-world sequential decision-making
+problems~\cite{kaebling}.  In recent years, point-based value
+iteration methods~\cite{pbvi_jair06,hsvi2,Perseus} have proved
+extremely successful at scaling (approximately) optimal POMDP
+solutions to large state spaces when a set of initial belief states is
+known.
+
+More recent work has extended point-based methods to both continuous
+state and continuous observation spaces, but no work has tackled both
+jointly without sampling.  \cite{Perseus_cont} provides exact
+point-based backups for continuous state and discrete observation
+problems (with approximate sample-based extensions to continuous
+actions and observations), while~\cite{pascal_ijcai05} provides exact
+point-based backups for discrete state and continuous observation
+problems (where multivariate observations must be independent).  While
+restricted to discrete state, \cite{pascal_ijcai05} provides an
+important insight that we exploit in this work: \emph{only a finite
+number of partitionings of the observation space are required to
+distinguish between the optimal conditional policy over a finite set
+of belief states}.
+
+Our contributions in this work are two-fold.  First, we extend
+symbolic dynamic programming for continuous state
+MDPs~\cite{sanner_uai11} to the case of continuous state and discrete
+observation POMDPs.  This provides an expressive and concrete
+instantiation of the framework in~\cite{Perseus_cont} (which only
+abstractly required that integrals were computable) in that it ensures
+the closed-form of all $\alpha$-functions for \emph{all horizons} is a
+symbolic piecewise case statement, even for POMDPs with
+\emph{arbitrary} continuous rewards and transitions with discrete
+noise (i.e., a finite mixture of deterministic transitions).  
+Second, we extend this symbolic dynamic programming algorithm
+to the case of continuous observations (while restricting the
+transition and reward functions to piecewise polynomials) and extend
+the work of~\cite{pascal_ijcai05} to \emph{derive} relevant observation
+partitions for potentially correlated multivariate continuous
+observation spaces by exploiting the piecewise polynomial integration
+operation of~\cite{sanner_aaai12} and the multivariate symbolic maximization
+technique of~\cite{sanner_uai11}.  
+We demonstrate
+proof-of-concept results on uni- and multi-variate state and
+observation steam plant control showing the tractability of our
+exact symbolic dynamic programming point-based approach.
+
+%\footnote{All code for these
+%experiments is freely available online (link suppressed) and
+%attached as anonymized supplementary material to this submission.}
+
+%We proceed as follows: after reviewing POMDPs, we discuss the need for 
+%first-order POMDPs, formalize them, and provide a lifted solution
+%via symbolic dynamic programming (SDP).  We empirically show the
+%complexity of SDP is invariant to domain size while enumerated
+%state POMDP solvers have complexity exponential in the domain size.
 
 \section{DC-POMDP Model}
-%%%% no need to define DC-MDP
-We assume familiarity with MDPs and introduce Discrete and Continuous Partially Observable MDPs as an extension to Discrete and Continuous State MDPs ~\cite{sanner_uai11}.
-A Discrete and Continuous partially observable MDP (DC-POMDP) is a tuple $\langle
-\mathcal{S},\mathcal{A},\mathcal{O},\mathcal{T},\mathcal{R},\mathcal{Z},\gamma,h \rangle$. 
-States are represented by vectors of $(\vec{d_s},\vec{x_s}) = ( d_{s_1},\ldots,d_{s_n},x_{s_1},\ldots,x_{s_m} )$ where each $d_{s_i} \in \{ 0,1 \}$ ($1 \leq i \leq n$) is a discrete boolean and each $x_{s_j} \in  \mathbb{R}$ ($1 \leq j \leq m$) is a continuous variable.
-Actions are presented by a finite set of actions $\{ a_1, \ldots, a_p \}$.
-We define discrete and continuous observations by the vector $(\vec{d_o},\vec{x_o}) = ( d_{o_1},\ldots,d_{o_n},x_{o_1},\ldots,x_{o_m} )$ where each $d_{o_i} \in \{ 0,1 \}$ ($1 \leq i \leq n$) is boolean and each $x_{o_j} \in \mathcal{R}$ ($1 \leq j \leq m$) is a continuous variable.
-
-Three functions are required for modeling DC-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state given the action and previous state %$P(\vec{d}',\vec{x}'|\cdots,a)$
-; (2)  $\mathcal{R}:\mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R}$ a reward function which returns the immediate reward of taking an action in some state and (3) an observation function defined as $\mathcal{Z} : \mathcal{S} \times \mathcal{A} \times \mathcal{O} \rightarrow [ 0, 1 ]$  which gives the probability of an observation given the outcome of a state after executing an action.
-A discount factor $\gamma, \; 0 \leq \gamma \leq 1$ is included in the model to account for future rewards in higher horizons.
-
-%%%% forms of transition, observation and reward function
-To define the transition and observation function, we use the fact that the state variables of this POMDP can be factored into a Dynamic Bayesian Network (DBN).
-Not allowing synchronic arcs between the discrete and continuous variables defines their joint transition or observation to be independent:
-%{\footnotesize
-%\begin{equation}
-%P(\vec{d_s}',\vec{x_s}'|\vec{d_s},\vec{x_s},a) = 
-%\prod_{i=1}^n P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \prod_{j=1}^m P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a). \nonumber 
-%\end{equation}}
-%The joint observation function can also be defined considering the direct correspondence between the state and observation variables:
-{\footnotesize
-\begin{equation}
+
+We assume familiarity with MDPs and introduce discrete and continuous
+partially observable MDPs (DC-POMDPs) as an extension to their DC-MDP
+subset~\cite{sanner_uai11}.  A DC-POMDP is a tuple $\langle
+\mathcal{S},\mathcal{A},\mathcal{O},\mathcal{T},\mathcal{R},\mathcal{Z},\gamma,h
+\rangle$.  States $\mathcal{S}$ are represented by vectors of
+$(\vec{d_s},\vec{x_s}) = (
+d_{s_1},\ldots,d_{s_n},x_{s_1},\ldots,x_{s_m} )$ where each $d_{s_i}
+\in \{ 0,1 \}$ ($1 \leq i \leq n$) is a discrete boolean and each
+$x_{s_j} \in \mathbb{R}$ ($1 \leq j \leq m$) is a continuous variable.
+We assume a finite, discrete action space $\mathcal{A} = \{ a_1,
+\ldots, a_p \}$. Observations
+$\mathcal{O}$ are given by the vector $(\vec{d_o},\vec{x_o}) = (
+d_{o_1},\ldots,d_{o_p},x_{o_1},\ldots,x_{o_q} )$ where each $d_{o_i}
+\in \{ 0,1 \}$ ($1 \leq i \leq p$) is boolean and each $x_{o_j} \in
+\mathbb{R}$ ($1 \leq j \leq q$) is continuous.
+
+Three functions are required for modeling DC-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state given the action and previous state%$P(\vec{d}',\vec{x}'|\cdots,a)$
+; (2)  $\mathcal{R}:\mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R}$ a reward function which returns the immediate reward of taking an action in some state; and (3) an observation function defined as $\mathcal{Z} : \mathcal{S} \times \mathcal{A} \times \mathcal{O} \rightarrow [ 0, 1 ]$  which gives the probability of an observation given the outcome of a state after executing an action.
+A discount factor $\gamma, \; 0 \leq \gamma \leq 1$ is used to discount rewards $t$ time steps into the future by $\gamma^t$.
+
+We use a dynamic Bayesian network (DBN) to compactly represent the transition model over the factored state variables and we likewise use a Bayesian network to represent the observation model:
+{\footnotesize
+\begin{align*}
+\mathcal{T}: \;\; &
+P(\vec{d_s}',\vec{x_s}'|\vec{d_s},\vec{x_s},a) = 
+\prod_{i=1}^n P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \prod_{j=1}^m P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a). \nonumber \\
+\mathcal{Z}: \;\; & 
 P(\vec{d_o},\vec{x_o}|\vec{d_s},\vec{x_s},a) = 
 \prod_{i=1}^n P(d_{o_i}|d_{s_i},a) \prod_{j=1}^m P(x_{o_j}|x_{s_j},a). \nonumber 
-\end{equation}}
-\emph{Binary} variables are represented using conditional probability functions (CPF) in the form of $P(d_i'|\vec{d},\vec{x},a)$. \emph{Continuous} variables
-$P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ are represented by \emph{piecewise
-linear equations} (PLEs) that are first-order Markov and deterministic, i.e.
-the probabilities are encoded using the Dirac $\delta[\cdot]$ function.
-The reward function $R(\vec{d},\vec{x},a)$ is set to be any general piecewise linear  function that could facilitate pruning for space efficiency.% via bilinear solvers.
+\end{align*}}
+For simplicity of algorithmic exposition, we do not allow synchronic
+arcs in the transition representation, but note that to incorporate
+such arcs only requires restrictions on the variable elimination
+ordering during dynamic programming.  In these Bayes nets, 
+probabilities over 
+\emph{discrete} variables are represented using conditional probability
+functions (CPF) in the form of
+$P(d_i'|\vec{d},\vec{x},a)$. Probabilities for \emph{continuous} variables
+$P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ are represented by
+\emph{piecewise functions} (PWFs) that are first-order Markov
+and deterministic, i.e. the probabilities are encoded using the Dirac
+$\delta[\cdot]$ function. The reward function $R(\vec{d},\vec{x},a)$
+is set to be any general piecewise function.  We next provide an intuitive
+example to make this DC-POMDP framework more concrete.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\textbf{Example} \textsc{(Power Plant)~\cite{steam2}} \emph{The steam
+generation system of a power plant aims to provide hot steam to the
+steam turbine which in turn produces electricity. Feed-water is
+exposed to heat in the tubes leading to the water tank, where water
+evaporation is performed under specific pressure and temperature.
+A reward is obtained when electricity is generated through steam
+production and the pressure and temperature are within safe ranges.
+Mixing water and steam makes the
+respective pressure and temperature observations $p_o \in \mathbb{R}$
+and $t_o \in \mathbb{R}$ on the underlying state $p_s \in \mathbb{R}$
+and $t_s \in \mathbb{R}$ highly uncertain.  Action choices are either
+to open or close a presure valve $a \in \{ \open, \close \}$.}
+
 For the \textsc{Power Plant} example, the transition and reward functions are defined as below:
@@
 \begin{cases}
- (a=open) &: t_s - 5 \\ 
-(a \neq open) &: t_s + 7 \\
+ (a=\open) &: t_s - 5 \\ 
+(a \neq \open) &: t_s + 7 \\
 \end{cases}
@@
 \begin{cases}
- (a=open) &: -1 \\
-(a \neq open)\wedge (t_s>T) &: -1000 \\
-(a \neq open)\wedge \neg(t_s>T) &: 100 \\
+ (a=\open) &: -1 \\
+(a \neq \open)\wedge (t_s>T) &: -1000 \\
+(a \neq \open)\wedge \neg(t_s>T) &: 100 \\
 \end{cases}\nonumber
@@
 The observations received before closing the plant to a safe temperature are therefore critical.
-For the observation model we begin with a discrete model for two observations $t_{o_1},t_{o_2}$ such as the following for the $open$ action:
-{\footnotesize
-\begin{align}
-P(t_{o_1}|t_s',open) = 
+For the observation model we begin with a discrete model for two observations $t_{o_1},t_{o_2}$ such as the following for the $\open$ action:
+{\footnotesize
+\begin{align}
+P(t_{o_1}|t_s',\open) = 
 \begin{cases}
@@
 \hspace{5mm}
-P(t_{o_2}|t_s',open) = 
+P(t_{o_2}|t_s',\open) = 
 \begin{cases}
@@
 %\\
-%P(t_{o_1}|t_s',close) = 
+%P(t_{o_1}|t_s',\close) = 
 %\begin{cases}
@@
 %,
-%P(t_{o_2}|t_s',close) = 
+%P(t_{o_2}|t_s',\close) = 
 %\begin{cases}
@@
 \begin{align}
-P(t_o|t_s',open) = 
+P(t_o|t_s',\open) = 
 \begin{cases}
@@
 %For $h=0$, $V^0$ is given in line 2.
-For $h=1$ we choose an action, for example $close$ to start with.
+For $h=1$ we choose an action, for example $\close$ to start with.
 The $\alpha$-vector set from $h=0$ is empty so line 7--8 of generating the relevant observation partitions is empty.
 Also priming the vectors and obtaining the product of relevant observations and the backup is an empty set.
-In line 15 the reward $R(s,close)$ is added to obtain $\alpha^1_{1,close}$ as defined by (8).
-The same operation is repeated for the $open$ action where $R(s,open)=-1$.
+In line 15 the reward $R(s,\close)$ is added to obtain $\alpha^1_{1,\close}$ as defined by (8).
+The same operation is repeated for the $\open$ action where $R(s,\open)=-1$.
 Line 16 defines the final $\Gamma$-set:
-%$\Gamma^1 = \{ \alpha^1_{1,open}, \alpha^1_{1,close} \}$.
+%$\Gamma^1 = \{ \alpha^1_{1,\open}, \alpha^1_{1,\close} \}$.
 The \texttt{Backup} algorithm in line 13 takes the primed version of the variables and performs regression on the $\alpha$-vectors. 
@@
 %For $h=2$, for each of the discrete actions, first we generate the probability of the relevant set of observations in the \texttt{GenRelObs} algorithm.
-This has been done for the $close$ action in the next section and we use the results here.
+This has been done for the $\close$ action in the next section and we use the results here.
 %\begin{align}
@@
 %\begin{align*}
-%\Gamma_{close,z_1,1}^2= 
+%\Gamma_{\close,z_1,1}^2= 
 %\begin{cases} 
@@
 %\hspace{15mm} 
-%\Gamma_{close,z_1,2}^2= \top &:-0.6
+%\Gamma_{\close,z_1,2}^2= \top &:-0.6
 %\\
-%\Gamma_{close,z_2,1}^2= 
+%\Gamma_{\close,z_2,1}^2= 
 %\begin{cases}
@@
 %\hspace{12mm} 
-%\Gamma_{close,z_2,2}^2= \top &:-0.1999
+%\Gamma_{\close,z_2,2}^2= \top &:-0.1999
 %\\
-%\Gamma_{close,z_3,1}^2= 
+%\Gamma_{\close,z_3,1}^2= 
 %\begin{cases}
@@
 %\hspace{15mm} 
-%\Gamma_{close,z_3,2}^2= \top &:-0.2
+%\Gamma_{\close,z_3,2}^2= \top &:-0.2
 %\end{align*}
@@
 %\begin{align*}
-%\Gamma_{close,1}^2 &= 
+%\Gamma_{\close,1}^2 &= 
 %\begin{cases} 
@@
 %\hspace{10mm} 
-%\Gamma_{close,2}^2= 
+%\Gamma_{\close,2}^2= 
 %\begin{cases} 
@@
 %\\
-%\Gamma_{close,3}^2&= 
+%\Gamma_{\close,3}^2&= 
 %\begin{cases} 
@@
 %\hspace{10mm}
-%\Gamma_{close,4}^2= 
+%\Gamma_{\close,4}^2= 
 %\begin{cases} 
@@
 %\\
-%\Gamma_{close,5}^2&= 
+%\Gamma_{\close,5}^2&= 
 %\begin{cases} 
@@
 %\hspace{5mm} 
-%\Gamma_{close,6}^2= 
+%\Gamma_{\close,6}^2= 
 %\begin{cases} 
@@
 %\\
-%\Gamma_{close,7}^2&= 
+%\Gamma_{\close,7}^2&= 
 %\begin{cases} 
@@
 %\hspace{10mm} 
-%\Gamma_{close,8}^2= \top:-1
+%\Gamma_{\close,8}^2= \top:-1
 %\end{align*}
@@
 The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_1$.
-According to continuous observation space defined in \cite{pascalPomdp}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space.
+According to continuous observation space defined in \cite{pascal_ijcai05}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space.
 In this work, numerical solutions are proposed to find the roots of any two observation dependant function.
@@
 \section{Related Work and Conclusion} 
-This work has used the concept of continuous states and the SDP solution to MDPs from ~\cite{sanner_uai11} and continuous observations from the work of \cite{pascalPomdp}.
+This work has used the concept of continuous states and the SDP solution to MDPs from ~\cite{sanner_uai11} and continuous observations from the work of \cite{pascal_ijcai05}.
 It is exclusive in the sense that it brings together the continuous state and observation POMDPs using a symbolic approach.
@@
 
-There has been prior work on approximate solutions to POMDPs with large or continuous state spaces ~\cite{Thrun99h}, ~\cite{Perseus} and some has been extended to the continuous observation setting using methods such as sampling ~\cite{PerseusObs}.
+There has been prior work on approximate solutions to POMDPs with large or continuous state spaces ~\cite{Thrun99h}, ~\cite{Perseus} and some has been extended to the continuous observation setting using methods such as sampling ~\cite{Perseus_cont}.
 In most approximate continuous state POMDP work, continuous or large discrete actions has been used.
-Thus we can extend the current work using \cite{zamaniAAAI} to contain continuous actions and provide exact or approximate solutions.
+Thus we can extend the current work using \cite{zamani_aaai12} to contain continuous actions and provide exact or approximate solutions.
 

### 2012-06-01 cpomdp/camdp-xadd/src/doc/NIPS/draft3-10page.tex
Edit based on previous draft

@@
 \begin{abstract}
-Decision-theoretic planning searches for the optimal sequence of actions for a given uncertain environment.
-Partially Observable MDPs (POMDP) can model such environments where the underlying state is not be fully observable.
-%While discrete models of the environment is often used for planning, continuous models are naturally closer to the real-world.
-While previous work have provided solutions for discrete states and/or observations, here we define the first exact solution to both continuous state and observations.
-The symbolic dynamic programming approach in this paper avoids enumerating the state and observation space providing solutions for general multi-variate discrete and continuous POMDPs.
-Our main contribution is that for a fixed set of reachable belief states and a fixed set of actions, only a \textit{finite} partitioning of the continuous observation space is required for optimality.
-Using function representation of Affine algebraic decision diagrams (XADDs) and symbolic case operators we derive the automatic partitioning for piecewise linear dynamics %such as the reward, transition and observation function.
-We demonstrate empirical results on an operation research domain to show the \textit{first exact solution} to continuous state and observation POMDPs.  
-%to use continuous states with discrete observations and extending it to continuous observations by defining relevant observation partitions %associated with each state partition.
-%This allows us to perform the Bellman backup operation efficiently.
-%We address the problems in POMDP continuous planning and solutions to solve them symbolically. %To compactly represent the exact value functions Affine algebraic decision diagrams are used. 
+Partially-observable Markov decision processes (POMDPs) provide a
+powerful model for real-world sequential decision-making problems.  In
+recent years, point-based value iteration methods have proven to be
+extremely effective techniques for finding (approximately) optimal
+dynamic programming solutions to POMDPs when an initial set of belief
+states is known.  However, no point-based work has provided exact
+point-based backups for both continuous state and observation spaces,
+which we tackle in this paper.  Our key insight is that while there
+may be an infinite number of possible observations, there are only a
+finite number of observation partitionings that are relevant for
+optimal decision-making when a finite, fixed set of reachable belief
+states is known.  To this end, we make two important contributions:
+(1) we show how previous exact symbolic dynamic programming solutions
+for continuous state MDPs can be generalized to continuous state
+POMDPs with discrete observations, and (2) we show how this solution
+can be further extended via recently developed symbolic methods to
+continuous state and observations to \emph{derive} the minimal
+relevant observation partitioning for potentially correlated,
+multivariate observation spaces.
+%for domais with piecewise
+%polynomial transition and observation functions.  
+We demonstrate
+proof-of-concept results on uni- and multi-variate state and
+observation steam plant control.
 \end{abstract}
 
-\section{Introduction}
+\section{Introduction and Related Work}
 % Write intro here
-In many complex systems complete information of the state is not available due to noisy or poor sensors and readings of them.
-Partially observable Markov Decision Processes (POMDPs)\cite{POMDPfirst} can be used for planning under such uncertainty in such problems.
-The solution to a POMDP is to find an answer to the fully-observable belief-MDP, whose states are probability distributions over the state space of the original POMDP \cite{kaebling}
-Real-world resources and sensor inputs are often naturally modelled using continuous variables. 
-Unfortunately most POMDP techniques can not be applied to such settings without assuming prior knowledge about the problem domain.
-Most methods use some approximation to discretize the continuous state, observation or action space which increases computational complexity \cite{Thrun99h}.
-The optimal value function of belief-MDPs is piecewiselinear and convex in these discrete cases \cite{sondik71} but has also been proven in the setting of only continuous states.
-\cite{PerseusObs}
 
-Exact solutions to a multi-variate continuous state and observation setting has not been approached for the curse of dimensionality problem.
-In assuming a continuous observation setting, the agent must search for all possible observations that are read from the environment.
-This can quickly become infeasible even for small problem sizes.
+Partially-observable Markov decision processes (POMDPs) are a powerful
+modeling formalism for real-world sequential decision-making
+problems~\cite{kaebling}.  In recent years, point-based value
+iteration methods~\cite{pbvi_jair06,hsvi2,Perseus} have proved
+extremely successful at scaling (approximately) optimal POMDP
+solutions to large state spaces when a set of initial belief states is
+known.
 
-In this paper we provide an exact solution to a subset of POMDPs with continuous states and observations.
-Our symbolic dynamic programming (SDP) approach assumes discrete noise and piecewise linear dynamics and functions.
-We formulize our problems using a Discrete and Continuous POMDP (DC-POMDP) model and apply our SDP algorithm.
-The main problem that rises in any problem modeled using DC-POMDP is the definition of different observation partitions and their probability.
-We apply symbolic solutions such as integration and maximization to provide an optimal closed-form solution to our problem domains.
-As an example we consider the following power plant example which has mainly been solved in the literature using discrete sets of states and observations~\cite{steam2}.
-We use the example throughout the paper to help better understand the definitions.
+More recent work has extended point-based methods to both continuous
+state and continuous observation spaces, but no work has tackled both
+jointly without sampling.  \cite{Perseus_cont} provides exact
+point-based backups for continuous state and discrete observation
+problems (with approximate sample-based extensions to continuous
+actions and observations), while~\cite{pascal_ijcai05} provides exact
+point-based backups for discrete state and continuous observation
+problems (where multivariate observations must be independent).  While
+restricted to discrete state, \cite{pascal_ijcai05} provides an
+important insight that we exploit in this work: \emph{only a finite
+number of partitionings of the observation space are required to
+distinguish between the optimal conditional policy over a finite set
+of belief states}.
 
-\textbf{Example} \textsc{(Power Plant)}
-\emph{The steam generation system of a power plant aims to provide hot steam to the steam turbine which in turn produces electricity.
-Feed-water is exposed to heat in the tubes leading to the water tank .
-Evaporating the water is performed under specific pressure and temperature inside the attached water tubes.
-Mixing water and steam pressures makes reading tank levels and temperatures very hard and uncertain.
-For this reason, the process is modeled using POMDP in the literature.
-The temperature $t_s \in \mathbb{R}$ is modelled as a continuous state variable.
-Noisy observation of the temperature $t_o \in \mathbb{R}$ is used to help decide the optimal action to either open or close the pressure valve $a \in \{open,close\}$.}
+Our contributions in this work are two-fold.  First, we extend
+symbolic dynamic programming for continuous state
+MDPs~\cite{sanner_uai11} to the case of continuous state and discrete
+observation POMDPs.  This provides an expressive and concrete
+instantiation of the framework in~\cite{Perseus_cont} (which only
+abstractly required that integrals were computable) in that it ensures
+the closed-form of all $\alpha$-functions for \emph{all horizons} is a
+symbolic piecewise case statement, even for POMDPs with
+\emph{arbitrary} continuous rewards and transitions with discrete
+noise (i.e., a finite mixture of deterministic transitions).  
+Second, we extend this symbolic dynamic programming algorithm
+to the case of continuous observations (while restricting the
+transition and reward functions to piecewise polynomials) and extend
+the work of~\cite{pascal_ijcai05} to \emph{derive} relevant observation
+partitions for potentially correlated multivariate continuous
+observation spaces by exploiting the piecewise polynomial integration
+operation of~\cite{sanner_aaai12} and the multivariate symbolic maximization
+technique of~\cite{sanner_uai11}.  
+We demonstrate
+proof-of-concept results on uni- and multi-variate state and
+observation steam plant control showing the tractability of our
+exact symbolic dynamic programming point-based approach.
 
-We need to extend the basic SDP framework from \cite{sanner_uai11} by adding continuous observations to the model.
-The solution to a closed-form value function is similar to the algorithm from \cite{monahan82} apart from the fact that no explicit observation is defined in the continuous case.
-We redefine the value-iteration algorithm using case representations that are implemented using Extended Algebric decision diagrams (XADD) and provide an efficient automatic way to partition the entire observation space into piecewise linear functions where each partition has a relative probability.
-We use the univariate \textsc{Power Plant} example to define the main operations and algorithms and also provide empirical results for a multi-variate version of the problem.
+%\footnote{All code for these
+%experiments is freely available online (link suppressed) and
+%attached as anonymized supplementary material to this submission.}
+
+%We proceed as follows: after reviewing POMDPs, we discuss the need for 
+%first-order POMDPs, formalize them, and provide a lifted solution
+%via symbolic dynamic programming (SDP).  We empirically show the
+%complexity of SDP is invariant to domain size while enumerated
+%state POMDP solvers have complexity exponential in the domain size.
 
 \section{DC-POMDP Model}
-%%%% no need to define DC-MDP
-We assume familiarity with MDPs and introduce Discrete and Continuous Partially Observable MDPs as an extension to Discrete and Continuous State MDPs ~\cite{sanner_uai11}.
-A Discrete and Continuous partially observable MDP (DC-POMDP) is a tuple $\langle
-\mathcal{S},\mathcal{A},\mathcal{O},\mathcal{T},\mathcal{R},\mathcal{Z},\gamma,h \rangle$. 
-States are represented by vectors of $(\vec{d_s},\vec{x_s}) = ( d_{s_1},\ldots,d_{s_n},x_{s_1},\ldots,x_{s_m} )$ where each $d_{s_i} \in \{ 0,1 \}$ ($1 \leq i \leq n$) is a discrete boolean and each $x_{s_j} \in  \mathbb{R}$ ($1 \leq j \leq m$) is a continuous variable.
-Actions are presented by a finite set of actions $\{ a_1, \ldots, a_p \}$.
-We define discrete and continuous observations by the vector $(\vec{d_o},\vec{x_o}) = ( d_{o_1},\ldots,d_{o_n},x_{o_1},\ldots,x_{o_m} )$ where each $d_{o_i} \in \{ 0,1 \}$ ($1 \leq i \leq n$) is boolean and each $x_{o_j} \in \mathcal{R}$ ($1 \leq j \leq m$) is a continuous variable.
 
-Three functions are required for modeling DC-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state given the action and previous state %$P(\vec{d}',\vec{x}'|\cdots,a)$
-; (2)  $\mathcal{R}:\mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R}$ a reward function which returns the immediate reward of taking an action in some state and (3) an observation function defined as $\mathcal{Z} : \mathcal{S} \times \mathcal{A} \times \mathcal{O} \rightarrow [ 0, 1 ]$  which gives the probability of an observation given the outcome of a state after executing an action.
-A discount factor $\gamma, \; 0 \leq \gamma \leq 1$ is included in the model to account for future rewards in higher horizons.
+We assume familiarity with MDPs and introduce discrete and continuous
+partially observable MDPs (DC-POMDPs) as an extension to their DC-MDP
+subset~\cite{sanner_uai11}.  A DC-POMDP is a tuple $\langle
+\mathcal{S},\mathcal{A},\mathcal{O},\mathcal{T},\mathcal{R},\mathcal{Z},\gamma,h
+\rangle$.  States $\mathcal{S}$ are represented by vectors of
+$(\vec{d_s},\vec{x_s}) = (
+d_{s_1},\ldots,d_{s_n},x_{s_1},\ldots,x_{s_m} )$ where each $d_{s_i}
+\in \{ 0,1 \}$ ($1 \leq i \leq n$) is a discrete boolean and each
+$x_{s_j} \in \mathbb{R}$ ($1 \leq j \leq m$) is a continuous variable.
+We assume a finite, discrete action space $\mathcal{A} = \{ a_1,
+\ldots, a_p \}$. Observations
+$\mathcal{O}$ are given by the vector $(\vec{d_o},\vec{x_o}) = (
+d_{o_1},\ldots,d_{o_p},x_{o_1},\ldots,x_{o_q} )$ where each $d_{o_i}
+\in \{ 0,1 \}$ ($1 \leq i \leq p$) is boolean and each $x_{o_j} \in
+\mathbb{R}$ ($1 \leq j \leq q$) is continuous.
 
-%%%% forms of transition, observation and reward function
-To define the transition and observation function, we use the fact that the state variables of this POMDP can be factored into a Dynamic Bayesian Network (DBN).
-Not allowing synchronic arcs between the discrete and continuous variables defines their joint transition or observation to be independent:
-%{\footnotesize
-%\begin{equation}
-%P(\vec{d_s}',\vec{x_s}'|\vec{d_s},\vec{x_s},a) = 
-%\prod_{i=1}^n P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \prod_{j=1}^m P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a). \nonumber 
-%\end{equation}}
-%The joint observation function can also be defined considering the direct correspondence between the state and observation variables:
+Three functions are required for modeling DC-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state given the action and previous state%$P(\vec{d}',\vec{x}'|\cdots,a)$
+; (2)  $\mathcal{R}:\mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R}$ a reward function which returns the immediate reward of taking an action in some state; and (3) an observation function defined as $\mathcal{Z} : \mathcal{S} \times \mathcal{A} \times \mathcal{O} \rightarrow [ 0, 1 ]$  which gives the probability of an observation given the outcome of a state after executing an action.
+A discount factor $\gamma, \; 0 \leq \gamma \leq 1$ is used to discount rewards $t$ time steps into the future by $\gamma^t$.
+
+We use a dynamic Bayesian network (DBN) to compactly represent the transition model over the factored state variables and we likewise use a Bayesian network to represent the observation model:
 {\footnotesize
-\begin{equation}
+\begin{align*}
+\mathcal{T}: \;\; &
+P(\vec{d_s}',\vec{x_s}'|\vec{d_s},\vec{x_s},a) = 
+\prod_{i=1}^n P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \prod_{j=1}^m P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a). \nonumber \\
+\mathcal{Z}: \;\; & 
 P(\vec{d_o},\vec{x_o}|\vec{d_s},\vec{x_s},a) = 
 \prod_{i=1}^n P(d_{o_i}|d_{s_i},a) \prod_{j=1}^m P(x_{o_j}|x_{s_j},a). \nonumber 
-\end{equation}}
-\emph{Binary} variables are represented using conditional probability functions (CPF) in the form of $P(d_i'|\vec{d},\vec{x},a)$. \emph{Continuous} variables
-$P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ are represented by \emph{piecewise
-linear equations} (PLEs) that are first-order Markov and deterministic, i.e.
-the probabilities are encoded using the Dirac $\delta[\cdot]$ function.
-The reward function $R(\vec{d},\vec{x},a)$ is set to be any general piecewise linear  function that could facilitate pruning for space efficiency.% via bilinear solvers.
+\end{align*}}
+For simplicity of algorithmic exposition, we do not allow synchronic
+arcs in the transition representation, but note that to incorporate
+such arcs only requires restrictions on the variable elimination
+ordering during dynamic programming.  In these Bayes nets, 
+probabilities over 
+\emph{discrete} variables are represented using conditional probability
+functions (CPF) in the form of
+$P(d_i'|\vec{d},\vec{x},a)$. Probabilities for \emph{continuous} variables
+$P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ are represented by
+\emph{piecewise functions} (PWFs) that are first-order Markov
+and deterministic, i.e. the probabilities are encoded using the Dirac
+$\delta[\cdot]$ function. The reward function $R(\vec{d},\vec{x},a)$
+is set to be any general piecewise function.  We next provide an intuitive
+example to make this DC-POMDP framework more concrete.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
+\hspace{5mm}
 \includegraphics[width=0.45\textwidth]{pics/b1-4.pdf}
@@
 \vspace{-2mm}
-\caption{\footnotesize Left: Example conditional plan $\beta^h$ for discrete observations. Right: The optimal value function for \textsc{Power Plant}
+\caption{\footnotesize Left: Example conditional plan $\beta^h$ for discrete observations. Right: The maximum $\alpha$ vector of iteration 4 for the \textsc{Power Plant}
 as a decision diagram: 
@@
 The observations received before closing the plant to a safe temperature are therefore critical.
-For the observation model we begin with a discrete model for two observations $t_{o_1},t_{o_2}$ such as the following for the $open$ action:
+For the observation model we begin with a discrete model for the observation $z$ with values $high, low$ such as the following for the $\open$ action:
 {\footnotesize
 \begin{align}
-P(t_{o_1}|t_s',open) = 
+P(z=high|t_s',\open) = 
 \begin{cases}
- \delta\left[ t_s <= 15 \right] &: 0.9 \\
- \delta\left[ \neg (t_s <= 15)\right] &: 0.1 \\
+  t_s <= 15 &: 0.9 \\
+ \neg (t_s <= 15)&: 0.1 \\
 \end{cases}\nonumber
 \hspace{5mm}
-P(t_{o_2}|t_s',open) = 
+P(z=low|t_s',\open) = 
 \begin{cases}
- \delta\left[ t_s <= 15 \right] &: 0.1 \\
- \delta\left[ \neg (t_s <= 15)\right] &: 0.9 \\
+ t_s <= 15 &: 0.1 \\
+ \neg (t_s <= 15) &: 0.9 \\
 \end{cases}\nonumber
@@
        		\If {$ContObs =true$}
-       			{$P(xd_{o_i}|xd_{s_i}) \,:=\,$ \genObs{$\Gamma^{h-1},a,b_i$}\;}
-       		 \ForEach {$xd_{o_i} \in \textsc{O}^{h-1}$}
+       			{$(O^h,P(xd_{o_i}|xd_{s_i})) \,:=\,$ \genObs{$\Gamma^{h-1},a,b_i$}\;}
+       		 \ForEach {$xd_{o_i} \in \textsc{O}^{h}$}
        		 {
@@
        			{
-   	 		  		$\alpha_j^{h'}=$ \remapWithPrimes{$\alpha_j^{h}$} \;
+   	 		  		$\alpha_j^{h'}=$ \remapWithPrimes{$\alpha_j^{h}$} 
    	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
@@
 }
-\caption{\footnotesize \texttt{VI}(DC-POMDP, $H$,$ContObs$, $b_i$) $\longrightarrow$ $(V^h,\pi^{*,h})$ \label{alg:vi}}
+\caption{\footnotesize \texttt{PBVI}(DC-POMDP, $H$,$ContObs$, $b_i$) $\longrightarrow$ $(V^h,\pi^{*,h})$ \label{alg:vi}}
 \vspace{-1mm}
@@
 \begin{align}
-V^{h+1}(b) &= \max_a \left( b \cdot r_a + 
-\gamma \sum_o \max_{\langle g^h_{a,o,j} \rangle_j}  b \cdot g^h_{a,o,j} \right) \\
+g^h_{a,o,j} &=  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \\
 \Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \\
@@
 }
-where $r_a = r(x_s,d_s,a)$ and  
-$g^h_{a,o,j} =  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s)$
+where $r_a = r(x_s,d_s,a)$.  
 %Each action-dependent $\Gamma$ sums over the possible set of $g^h_{a,o,j}(\cdot)$ given an observation model to derive discrete observations $xd_o$ and for all $\alpha^{h-1}_{j} \in \Gamma^{h-1} $.
 The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.
-Lines 9--16 of the \textsc{VI} algorithm show the above in a general algorithm used for both discrete and continuous observations. 
-For all $\alpha$-vectors in the current $\Gamma$-set first an LP-solver is used to prune any inconsistency in the branches of this vector in line 17.
-A pairwise dominance test is performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$
+Lines 7--15 of the \textsc{PBVI} algorithm show the above in a general algorithm used for both discrete and continuous observations.
+We will discuss this according to our symbolic framework in the next section.
+For all $\alpha$-vectors in the current $\Gamma$-set the \textsc{Prune} function uses an LP-solver to prune any inconsistency in the branches of this vector in line 16 similar to the approach in ~\cite{penberthy94}.
+A pairwise dominance test is also performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$
 The result of this pairwise test is a dominant $\alpha$-vector which is added to the set of $\Gamma^h$. 
@@
 }
-\emph{Restriction} takes a function $f$ to restrict only in cases
-that satisfy some formula $\phi$, which we write as $f|_{\phi}$.  
-This can be done by simply appending $\phi$ to each case partition
-as the left side shows:
-{\footnotesize
-\begin{center}
-\begin{tabular}{r c c l}
+We also require the following SDP operations on case statements already introduced in the literature: 
+\begin{itemize}
+\item Restriction $f|_{\phi}$:  Takes a function $f$ to restrict only in cases
+that satisfy some formula $\phi$ as defined in \cite{sanner_uai11}.
+\item Substitution $f\sigma$ : Takes a set $\sigma$ of variables and their substitutions, where the LHS of the substitution operator $/$ represents the substitution variable and the RHS of the $/$ represents the expression that should be substituted in its place as in \cite{sanner_uai11}.
+\item Integration $\int_{x_1} f dx_1$:  Eliminates continuous variables by taking the integral of the polynomial function $f$ with respect to linear constraints on $x_1$  \cite{sanner_aaai12}.
+\end{itemize}
+%\emph{Restriction} takes a function $f$ to restrict only in cases
+%that satisfy some formula $\phi$, which we write as $f|_{\phi}$.  
+%This can be done by simply appending $\phi$ to each case partition
+%as the left side shows:
+%{\footnotesize
+%\begin{center}
+%\begin{tabular}{r c c l}
+%%&
+%%\hspace{-6mm} 
+%%  $f = \begin{cases}
+%%    \phi_1: & f_1 \\ 
+%%   \vdots&\vdots\\ 
+%%    \phi_k: & f_k \\ 
+%%  \end{cases}$
+%%&
 %&
-%\hspace{-6mm} 
-%  $f = \begin{cases}
-%    \phi_1: & f_1 \\ 
+%\hspace{0mm}
+%  $f|_{\phi} = \begin{cases}
+%    \phi_1 \land \phi : & f_1 \\ 
 %   \vdots&\vdots\\ 
-%    \phi_k: & f_k \\ 
+%    \phi_k \land \phi : & f_k \\ 
 %  \end{cases}$
 %&
-&
-\hspace{0mm}
-  $f|_{\phi} = \begin{cases}
-    \phi_1 \land \phi : & f_1 \\ 
-   \vdots&\vdots\\ 
-    \phi_k \land \phi : & f_k \\ 
-  \end{cases}$
-&
-\hspace{15mm}
-$\int_{x_j'} P(x_j'|\cdots) V'^{h} dx_j' = \begin{cases}
-    \phi_1: & V'^{h} \{ x_j' = f_1 \} \\ 
-   \vdots&\vdots\\ 
-    \phi_k: & V'^{h} \{ x_j' = f_k \}  \\ 
-  \end{cases}$
-  \end{tabular}
-\end{center}
-}
-\emph{Symbolic substitution} simply takes a set $\sigma$ of variables and their substitutions, where
-the LHS of the substitution operator $/$ represents the substitution variable and the
-RHS of the $/$ represents the expression that should be substituted in its place.
-Hence to perform a continuous regression on a more general
-representation, we obtain the right side of the above.
+%\hspace{15mm}
+%$\int_{x_j'} P(x_j'|\cdots) V'^{h} dx_j' = \begin{cases}
+%    \phi_1: & V'^{h} \{ x_j' = f_1 \} \\ 
+%   \vdots&\vdots\\ 
+%    \phi_k: & V'^{h} \{ x_j' = f_k \}  \\ 
+%  \end{cases}$
+%  \end{tabular}
+%\end{center}
+%}
+%\emph{Symbolic substitution} simply takes a set $\sigma$ of variables and their substitutions, where
+%the LHS of the substitution operator $/$ represents the substitution variable and the
+%RHS of the $/$ represents the expression that should be substituted in its place.
+%Hence to perform a continuous regression on a more general
+%representation, we obtain the right side of the above.
 
@@
 Here we use continuous SDP for both states and observations in DC-POMDPs.
-The value functions and $\alpha$-vectors partition into states of equal value, both of which can be represented using XADDs.
+The $\alpha$-vectors partition into states of equal value, both of which can be represented using XADDs.
 For increasing efficiency linear constraint feasibility checkers such as LP solvers allow pruning unreachable paths in the XADD.
-The right picture in Figure 1 is an example of an XADD representation.
+Figure 1 (right) is an example of an XADD representation.
 
@@
 
-For discrete and continuous states and discrete observations, we use the symbolic version of Monahan's algorithm as the DC-POMDP solution.
+For discrete and continuous states and discrete observations, we use the symbolic version of Monahan's algorithm as the DC-POMDP solution represented in \textsc{PBVI}.
 The same approach is required for continuous observations only based on point-based dynamic programming backups.
@@
 Each step of the algorithm is explained using the \textsc{Power plant} example for the first three horizons.
-Note that if the input to this algorithm is continuous observations $ContObs=true$ then line 8 finds the relevant observation partitions, else discrete observations are assumed and their probabilities are given in the problem specifications.
+Note that if the input to this algorithm is continuous observations $ContObs=true$ then line 8 finds the relevant observation partitions, else discrete observations are assumed and their probabilities are given in the problem specifications such as section 2 of this paper.
 Also a belief is given as the input and often several beliefs are required to cover the belief space and thus the algorithm is executed for each one of $b_i \in \mathbb{B}$.
 
-For each iteration, the $\alpha$-vector set are case functions that can be primed (line 11--12). 
-%For $h=0$, $V^0$ is given in line 2.
-For $h=1$ we choose an action, for example $close$ to start with.
-The $\alpha$-vector set from $h=0$ is empty so line 7--8 of generating the relevant observation partitions is empty.
-Also priming the vectors and obtaining the product of relevant observations and the backup is an empty set.
-In line 15 the reward $R(s,close)$ is added to obtain $\alpha^1_{1,close}$ as defined by (8).
-The same operation is repeated for the $open$ action where $R(s,open)=-1$.
-Line 16 defines the final $\Gamma$-set:
-%$\Gamma^1 = \{ \alpha^1_{1,open}, \alpha^1_{1,close} \}$.
-The \texttt{Backup} algorithm in line 13 takes the primed version of the variables and performs regression on the $\alpha$-vectors. 
-\emph{Boolean restriction} is defined where $f|_{b=v}$ restricts a boolean variable $b$ occurring in $f$ to values $v \in \{ 0,1 \}$.
-This can be done by simply appending $b=v$ to each case partition (logical $\wedge$ of $b=v$ and the logical first-order formulas of $f$).
-
-\emph{Continuous regression} $\int f(x_{s_j}') \otimes P(x_{s_j}'|\cdots) dx_{s_j}'$ has the form $\delta[x_{s_j}' - h(\vec{z})]$ for the probability over the next state, where $h(\vec{z})$ is a case statement and $\vec{z}$ does not contain
-$x_{s_j}'$.  A symbolic substitution is defined for this integration such that any occurrence of $x_{s_j}'$ in $f(x_{s_j}')$ is substituted with $h(\vec{z})$: 
-$\int f(x_{s_j}') \otimes \delta[x_{s_j}' - h(\vec{z})] dx_{s_j}' = f(x'_j) \{ x_{s_j}' / h(\vec{z}) \}$
-
-%For $h=2$, for each of the discrete actions, first we generate the probability of the relevant set of observations in the \texttt{GenRelObs} algorithm.
-This has been done for the $close$ action in the next section and we use the results here.
-%\begin{align}
-%P(z_k)=
-%\begin{cases}
-% (2<t_o<6) &: 0.2 \\
-%(-4<t_o<2) &: 0.6\\
-%(-8<t_o<-4) &:0.1999
-%\end{cases} 
-%\nonumber
-%\end{align}
-%Next for each of the relevant partitions and both $\alpha$-vectors of the previous $\Gamma$ set lines 9--16 are executed. Line 11 will substitute prime values for all state variables. 
-Line 13 produces the backup $\alpha$-vector and multiplies it by the weight derived from the relevant observation partition in the continuous case or observation probabilities in the problem description.
-%For each observation, two vectors are formed for each corresponding $\alpha$-vector. For example line 13 for the observation partitions above and $\alpha_1^1,\alpha_2^1$ is defined as:
-%\begin{align*}
-%\Gamma_{close,z_1,1}^2= 
-%\begin{cases} 
-%(t_s'>15) &: -600 \\ 
-%\neg(t_s'>15) &: 60 \\
-%\end{cases}
-%\hspace{15mm} 
-%\Gamma_{close,z_1,2}^2= \top &:-0.6
-%\\
-%\Gamma_{close,z_2,1}^2= 
-%\begin{cases}
-%(t_s'>15) &: -199.999 \\ 
-%\neg(t_s'>15) &: 19.999 
-%\end{cases}
-%\hspace{12mm} 
-%\Gamma_{close,z_2,2}^2= \top &:-0.1999
-%\\
-%\Gamma_{close,z_3,1}^2= 
-%\begin{cases}
-%(t_s'>15) &: -200 \\ 
-%\neg(t_s'>15) &: 20 
-%\end{cases}
-%\hspace{15mm} 
-%\Gamma_{close,z_3,2}^2= \top &:-0.2
-%\end{align*}
-Line 14 takes the cross-sum of the two sets which takes every possible combination of observations according to the $\alpha$-vectors. 
-%\begin{align*}
-%\Gamma_{close,1}^2 &= 
-%\begin{cases} 
-%(t_s'>15) &: -800.2 \\ 
-%\neg(t_s'>15) &: 79.8 \\
-%\end{cases}
-%\hspace{10mm} 
-%\Gamma_{close,2}^2= 
-%\begin{cases} 
-%(t_s'>15) &: -1000 \\ 
-%\neg(t_s'>15) &: 100 \\
-%\end{cases}
-%\\
-%\Gamma_{close,3}^2&= 
-%\begin{cases} 
-%(t_s'>15) &: -800.2 \\ 
-%\neg(t_s'>15) &: 79.8 \\
-%\end{cases}
-%\hspace{10mm}
-%\Gamma_{close,4}^2= 
-%\begin{cases} 
-%(t_s'>15) &: -600.4 \\ 
-%\neg(t_s'>15) &: 59.6 \\
-%\end{cases}
-%\\
-%\Gamma_{close,5}^2&= 
-%\begin{cases} 
-%(t_s'>15) &: -400.5999 \\ 
-%\neg(t_s'>15) &: 39.3999 \\
-%\end{cases}
-%\hspace{5mm} 
-%\Gamma_{close,6}^2= 
-%\begin{cases} 
-%(t_s'>15) &: -200.7999 \\ 
-%\neg(t_s'>15) &: 19.999 \\
-%\end{cases}
-%\\
-%\Gamma_{close,7}^2&= 
-%\begin{cases} 
-%(t_s'>15) &: -200.8 \\ 
-%\neg(t_s'>15) &: 19.2 \\
-%\end{cases}
-%\hspace{10mm} 
-%\Gamma_{close,8}^2= \top:-1
-%\end{align*}
-All vectors are multiplied by the discount $\gamma$ and added with the reward in line 15 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
-% need to add this?
+For each iteration, lines 11--12 product the following equation: 
+{\footnotesize
+\begin{align}
+g^h_{a,o,j} =  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \nonumber
+\end{align}
+}
+where the $\alpha$-vector set are case functions primed by substitution in \textsc{Prime} 
+and \texttt{Backup} performs regression on the $\alpha$-vectors.
+Boolean regression is performing a symbolic restriction while a continuous regression is a symbolic substitution as defined previously.
+Line 13 takes the cross-sum of the two sets which takes every possible combination of observations according to the $\alpha$-vectors. 
+All vectors are then multiplied by the discount $\gamma$ and added with the reward in line 14 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
 The same operation is performed for all actions and then in line 15 the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors for the next horizon. 
-%In line 17 this final set is using in the \texttt{Prune} algorithm.
-For all $\alpha$-vectors in the current $\Gamma$-set first an LP-solver is used to prune any inconsistency in the branches of this vector.
-Next a pairwise dominance test is performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$The result of this pairwise test is a dominant $\alpha$-vector which is added to the set of $\Gamma^h$.
-In the last step, we find the $\alpha$-vector that maximizes the value-function with respect to this belief and save it as the new $\Gamma$-set for the next horizon.
+The $\alpha$-vector that maximizes the value-function with respect to this belief and save it as the new $\Gamma$-set for the next horizon.
 The action from which this $\alpha$-vector was chosen from is the maximum policy.
@@
 In the next section we discuss the main part of our algorithm which is generating relevant observation partitions.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\incmargin{.5em}
-%\linesnumbered
-%\begin{algorithm}[t!]
-%\vspace{-.5mm}
-%\dontprintsemicolon
-%\Begin{
-%   
-%    \emph{// Continuous regression, marginal integration}\\
-%    \For {all $xd_{s_j}'$ in $\alpha_{j,o}^{'h}$}
-%    {
-%         $G_a^h := \int P(z) \otimes P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a) \, d_{x_{s_j}'}$\;
-%    }
-%    \emph{// Discrete regression, marginal summation}\\
-%    \For {all $d_{s_i}'$ in $\alpha_{j,o}^{'h}$}
-%    {
-%         $G_a^h := \left[ G_a^h \otimes P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \right]|_{d_{s_i}' = 1}$\\
-%         \hspace{8mm} $\oplus \left[ G_a^h \otimes P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \right]|_{d_{s_i}'= 0}$\;
-%    }
-%    \Return{$G_a^h$} \;
-%}
-%\caption{\footnotesize \texttt{Backup}($\alpha_{j,o}^{'h},a,P(z)$) $\longrightarrow$ $G_a^h$ \label{alg:regress}}
-%\vspace{-1mm}
-%\end{algorithm}
-%\decmargin{.5em}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\incmargin{.5em}
-%\linesnumbered
-%\begin{algorithm}[t!]
-%\vspace{-.5mm}
-%\SetKwFunction{lpsolve}{LP-Solve}
-%\SetKwFunction{dominancy}{Dominance}
-%\Begin
-%{
-%   		$V^h :=0$ \;
-%       \ForEach {$\alpha_j^h \in \Gamma^h$}
-%       {
-%           $\alpha_j^h \,:=\,$\lpsolve{$\alpha_j^h$} \;
-%           \ForEach {$\alpha_k^h \in \Gamma^h$}
-%           {
-%				\If {$j \neq k$}  
-%				{         
-%           			$V^h \,:=\,\oplus$ \dominancy{$\alpha_j^h,\alpha_k^h$} \; 
-%           		}  
-%           	}
-%       }
-%     \Return{$(V^h)$} \;
-%}
-%\caption{\footnotesize \texttt{Prune}($\Gamma^h$) $\longrightarrow$ $(V^h)$ }
-%\vspace{-1mm}
-%\end{algorithm}
-%\decmargin{.5em}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \subsection{PBVI for DC State and DC Observations} 
-%progression definition
 The main operations required for value iteration in DC-POMDP is generating relevant observations to perform the bellman backup operation. 
-%Progression is the inverse of regression.
-The \emph{regression} of a formula $\psi$ through an action $a$ is a formula $\psi'$ that holds prior to $a$ being performed iff $\psi$ holds after $a$.
-\emph{Progression} takes a formula $\psi$ and an action $o$ and represents the formula $\psi'$ that would hold after action $o$ is performed in a situation satisfying $\psi$.
 The \texttt{GenRelObs} algorithm defines the steps required to generate these partitions. 
@@
 We choose these beliefs such that they cover the entire probable state space; e.g.
-$b_1: U[t_s;2,6]$ and $b_2: U[t_s;6,13]$.
+$b_1: U[t_s;2,6]$ and $b_2: U[t_s;6,11]$.
 
@@
 $\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $. 
-This operation is similar to that of \texttt{Regress} multiplied by the observation regressed model.
+This operation of line 3 in \textsc{GenRelObs} is similar to that of \texttt{Regress} multiplied by the observation regressed model.
 For continuous variables the continuous regression and for discrete variables boolean restriction is used as defined in the previous section.
@@
 		{
-    	$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $;\
+    	$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) \cdot P(xd'_{s_i}| xd_{s_i},a)\cdot \alpha_j $;\
 		}  
@@
 		{
-		$\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$\}\ \;
+		$\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i \cdot \alpha_j(xd_s,xd_o)$\}\ \;
 		}
@@
 }
-We need the $\alpha$-vectors to partition on the observation space, thus next we eliminate the current state variable.
+We need the $\alpha$-vectors to partition on the observation space, thus next we eliminate the current state variable in line 4--5.
 During this integration the current state variable is factored out by multiplying the belief in each of the current $\alpha$-vectors.
 The resulting $\delta$-functions only depend on the observation: $\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$.
-For the belief $b_1: U[t_s;2,6]$ the two vectors above the resulting $\delta$-functions are defined below: 
+For $b_1: U[t_s;2,6]$ the two vectors above the resulting $\delta$-functions are defined below: 
 {\footnotesize
@@
 The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_1$.
-According to continuous observation space defined in \cite{pascalPomdp}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space.
-In this work, numerical solutions are proposed to find the roots of any two observation dependant function.
+According to continuous observation space defined in \cite{pascal_ijcai05}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space.
+In this work, numerical solutions are proposed to find the roots of any two observation dependent function.
 Here we use the symbolic power of the max-operator to find all the maximum boundary regions of two or more $\delta$-functions at the same time.
-For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximuum:
+For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximum in line 6--7:
 {\footnotesize
@@
 $\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*P(s) = 1$
-Thus we only need to multiply the indicators of each case partition in this formula to obtain the probability of each partition such that it sums to one: 
+Thus we only need to multiply the indicators of each case partition as as in line 9 in this formula to obtain the probability of each partition such that it sums to one: 
 $P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$
@@
 We evaluated our continuous POMDP solution using XADDs on the one-dimensional
-\textsc{Power Plant} example and another variant of this problem with two variables. \footnote{
-Full problem specifications and Java code to reproduce these experiments are currently 
-in Google Code; anonymity restrictions prevent us from providing this link in the submitted paper version.}
+%\textsc{Power Plant} example and another variant of this problem with two variables. \footnote{
+%Full problem specifications and Java code to reproduce these experiments are currently 
+%in Google Code; anonymity restrictions prevent us from providing this link in the submitted paper version.}
 We start by defining the 2-dimensional description of this problem:
@@
 {\bf \textsc{2D-Power Plant}:} We consider the more complex model of the power plant similar to \cite{steam2} where the pressure inside the water tank must be controlled to avoid mixing water into the steam or explosion of the tank.
-We model the pressure variable $p$ as a partially observable variable from the observation readings of the pressure valve $po$.
+We model the pressure variable $p$ as a partially observable variable from the observation readings of the pressure $po$.
 The two actions of increase and decrease are defined based on the change in both the temperature and the pressure.
@@
 \begin{align}
-P(p_s'|\vec{p_s},a)&= \delta\left[ p_s' - 
+P(p_s'|\vec{p_s},inc)&= \delta\left[ p_s' - 
 \begin{cases}
  (p+10> 20) &: 20 \\ 
-\neq (p+10> 20) &: p_s + 10 \\
+\neg (p+10> 20) &: p_s + 10 \\
 \end{cases}
@@
 \hspace{5mm} 
-P(t_s'|\vec{t_s},a)= \delta\left[ t_s' -  t_s +10 \right]\nonumber
+P(t_s'|\vec{t_s},inc)= \delta\left[ t_s' - (t_s +10) \right]\nonumber
 \\
-R(t_s,p_s,a) &= 
+R(t_s,p_s,inc) &= 
 \begin{cases}
@@
 \begin{cases}
- (t_o>t_s + 80) \wedge (t_o<t_s+ 105) &: 0.4 \\
- \neg( (t_o>t_s + 80) \wedge (t_o<t_s+ 105)) &: 0 \\
+(t_s + 80<t_o<t_s+ 105) &: 0.4 \\
+ \neg (t_s + 80<t_o<t_s+ 105) &: 0 \\
 \end{cases}\nonumber
-\hspace{0mm} 
+\hspace{5mm} 
 P(p_o|p_s') = 
 \begin{cases}
- (p_o>p_s) \wedge (p_o<p_s+10) &: 0.1 \\
- \neg( (p_o>p_s) \wedge (p_o<p_s+10)) &: 0 \\
+(p_s<p_o<p_s+10) &: 0.1 \\
+ \neg(p_s<p_o<p_s+10) &: 0 \\
 \end{cases}\nonumber
@@
 Comparing the two problem sizes demonstrates the effect of the number of state-observation variables in our algorithm.
-As the algorithm progresses, the time required to compute the probability of the partitions and finding the maximum value function increases for both problem sizes and significantly more for the 2D version.
+As the algorithm progresses, the time required to compute the probability of the partitions and finding the maximum $\alpha$-vector with respect to beliefs increases for both problem sizes and significantly more for the 2D version.
 The stability in time is due to the fact that each stage almost takes the same time since it has converged quickly.%???
@@
 
-In Figure ~\ref{fig:3D}  we present plots of the value function for different iterations of the 2D problem instance.
-Each plot demonstrates the value as a function of the pressure and temperature.
+In Figure ~\ref{fig:3D}  we present plots of the maximum $\alpha$-vectors of belief $b_1$ for different iterations of the 2D problem instance.
+Each plot demonstrates the value of each vector as a function of the pressure and temperature.
 Starting with the first iteration, the value is highest for the reward range ($5<p<15 \wedge 95<t<105$) and -1 or less for other places.
 In the fifth iteration, the value function has partitioned into more pieces, showing how higher temperatures can increase the value without considering the effect of the pressure.
-In the last plot, horizon 6 has better tuned the value function so that higher temperatures and pressures increase the value function and also within the reward range, finer grain partitions have been formed.
+In the last plot, horizon 6 has better tuned the value function so that higher temperatures and pressures increase the value of the maximum $\alpha$-vector and also within the reward range, finer grain partitions have been formed.
 Also note that the policy depends on the initial belief and for these plots they were based on $b_1$ which was lower than the reward range.
@@
 \caption{\footnotesize 
-{\it (left)} Value function for first iteration; 
-{\it (middle)} $V^5(p,t)$; 
-{\it (right)} $V^6(p,t)$.
+{\it (left)}Maximum $\alpha$-vector for $b_1$ in first iteration; 
+{\it (middle)} $\alpha_{max}^5(b_1)$; 
+{\it (right)} $\alpha_{max}^5(b_1)$.%V^6(p,t)
 }

### 2012-06-01 trunk/cpomdp/camdp-xadd/src/doc/NIPS/draft4-10page.tex
updated draft4

--- 2012-06-01 cpomdp/camdp-xadd/src/doc/NIPS/draft3-10page.tex
@@
 
-\section{Introduction and Related Work}
+\section{Introduction} %and Related Work}
 % Write intro here
@@
 {\footnotesize
+\vspace{-1mm}
 \begin{align}
@@
 \end{cases}
-\right]\nonumber
-\hspace{5mm}
+\right]
+\hspace{2mm}
 R(t_s,a) = 
@@
 (a \neq \open)\wedge \neg(t_s>T) &: 100 \\
-\end{cases}\nonumber
+\end{cases}
 \end{align}
+\vspace{-4mm}
 }
+
 %contribution? 
@@
  \neg (t_s <= 15)&: 0.1 \\
-\end{cases}\nonumber
+\end{cases}
 \hspace{5mm}
@@
  \neg (t_s <= 15) &: 0.9 \\
-\end{cases}\nonumber
+\end{cases}
 \end{align}
+\vspace{-3mm}
 }
@@
 %\end{align}
+
 After defining the solution to the discrete observation with a continuous state DC-POMDP, we can extend this to consider continuous observations using a continuous uniform noise distribution which is more realistic in many domains.
@@
  \neg((t_o>t_s-5) \wedge (t_o<t_s+5)) &: 0 \\
-\end{cases}\nonumber
+\end{cases}
 \end{align}
@@
 At any of the horizon steps, a DC-POMDP policy $\pi(b)$ is defined as the iterative plan on executing an action and observing all probable outcomes of the set of observations from the $(h-1)$-step.
-The belief state is updated after this according to the Bayes rule: $\sum_{d_s}\int_{x_s} b(xd_s)T(xd'_s,xd_s,a)Z(xd'_s,xd_o,a)$.
+The belief state is updated after this according to the Bayes rule: $\sum_{d_s}\int_{x_s} b(xd_s)\mathcal{T}(xd'_s,xd_s,a)\mathcal{Z}(xd'_s,xd_o,a)$.
 \footnote{For simiplicity we use ($xd_s = x_s,d_s$) and ($xd_o=x_o,d_o$)}
@@
 For continuous observations, conditional plans are decision trees with infinitely many branches.
-In our solution we show that instead of a policy tree with infinite branches, we can create partitions based on relevant observation values, resulting in a policy tree similar to that of figure 1.
-%From here on we assume this finite set of observation partitioning as $\mathcal{O}$.
+In our solution we show that instead of a policy tree with infinite branches, we can create partitions based on relevant observation values, resulting in a policy tree similar to that of figure 1.%From here on we assume this finite set of observation partitioning as $\mathcal{O}$.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
        $h:=h+1, \Gamma^h :=\emptyset$\;
+       \ForEach {$b_i \in B$}
+       {
        	\ForEach {$a \in A$}
@@
        			{
-   	 		  		$\alpha_j^{h'}=$ \remapWithPrimes{$\alpha_j^{h}$} 
+   	 		  		$\alpha_j^{h'},:=\,$ \remapWithPrimes{$\alpha_j^{h}$} 
    	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
@@
               $\Gamma^h \,:=\, $\prune{$\Gamma^h$} \;
-              $V^h \,:=\, \mathrm{max}_{\alpha_j \in \Gamma^h} b_i \cdot \alpha_j$\;
-              $\pi^{*,h} \,:=\, \argmax_{a} \, \Gamma_a^{h}$\;
+      }
+             % $V^h \,:=\, \mathrm{max}_{\alpha_j \in \Gamma^h} b_i \cdot \alpha_j$\;
+             % $\pi^{*,h} \,:=\, \argmax_{a} \, \Gamma_a^{h}$\;
+      \ForEach {$b_i \in B$}
+      {
+     	$\alpha_{b_i}^h :=\ \argmax_{\alpha_j \in \Gamma^h} \Gamma^{h}$\;
+     	$\Gamma_{PBVI}^h :=\ \Gamma^h \cup \alpha_{b_i}^h$\;
+      }
 
@@
    }
-     \Return{$(V^h,\pi^{*,h})$} \;
+     \Return{$\Gamma_{PBVI}$} \;
 }
-\caption{\footnotesize \texttt{PBVI}(DC-POMDP, $H$,$ContObs$, $b_i$) $\longrightarrow$ $(V^h,\pi^{*,h})$ \label{alg:vi}}
+\caption{\footnotesize \texttt{PBVI}(DC-POMDP, $H$,$ContObs$, $B=\left\{b_i \right\}$) $\longrightarrow$ $(V^h,\pi^{*,h})$ \label{alg:vi}}
 \vspace{-1mm}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 For discrete observations and discrete states the number of belief states is infinite, but the optimal value function for $h$ is proven to be a piecewise linear and convex function of the belief state \cite{smallwoodSondik}.
@@
 
-Point-based dynamic programming backups compute the optimal value function given a belief state $b$.
+\textbf{Point-based dynamic programming backups} compute the optimal value function given a belief state $b$.
 The best conditional plan for the given belief, corresponds to an $\alpha$-vector which is computed using point-based backups:
@@
 The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.
-Lines 7--15 of the \textsc{PBVI} algorithm show the above in a general algorithm used for both discrete and continuous observations.
-We will discuss this according to our symbolic framework in the next section.
+Lines 7--15 of the \textsc{PBVI} algorithm show the general algorithm used for both discrete and continuous observations. We will discuss this according to our symbolic framework in the next section. 
 For all $\alpha$-vectors in the current $\Gamma$-set the \textsc{Prune} function uses an LP-solver to prune any inconsistency in the branches of this vector in line 16 similar to the approach in ~\cite{penberthy94}.
@@
 As the domain grows, fully enumerating the state, action and observation space leads to intractable solutions even for low horizons.
-Avoiding this enumeration often requires approximating the solutions.
+Avoiding this enumeration often requires approximating the solution.
 We claim an exact solution using symbolic dynamic programming.
@@
 {\footnotesize 
+\vspace{-3mm}
 \begin{align}
@@
 {\footnotesize 
+\vspace{-3mm}
 \begin{center}
@@
 \end{center}
+\vspace{-2mm}
 }
 Likewise $\ominus$ and $\otimes$ are defined by subtracting or multiplying partition values.  Inconsistent partitions can be discarded when they are irrelevant to the function value.
-The next three operations are required in defining the value iteration algorithm. A \emph{symbolic case maximization} is defined as below:
+A \emph{symbolic case maximization} is defined as below:
 \vspace{-4mm}
 {\footnotesize
+\vspace{-2mm}
 \begin{center}
@@
 \end{center}
+\vspace{-3mm}
 }
@@
 \end{itemize}
+\vspace{-2mm}
 %\emph{Restriction} takes a function $f$ to restrict only in cases
@@
 For increasing efficiency linear constraint feasibility checkers such as LP solvers allow pruning unreachable paths in the XADD.
-Figure 1 (right) is an example of an XADD representation.
+Figure 1,right is an example of an XADD representation.
 
@@
 Each step of the algorithm is explained using the \textsc{Power plant} example for the first three horizons.
-Note that if the input to this algorithm is continuous observations $ContObs=true$ then line 8 finds the relevant observation partitions, else discrete observations are assumed and their probabilities are given in the problem specifications such as section 2 of this paper.
-Also a belief is given as the input and often several beliefs are required to cover the belief space and thus the algorithm is executed for each one of $b_i \in \mathbb{B}$.
-
+Note that if the input to this algorithm is continuous observations $ContObs=true$ then line 8 finds the relevant observation partitions, else discrete observations are assumed and their probabilities are given in the problem specifications such as section 2.
+Also a set of beliefs are given as the input as often several beliefs are required to cover the belief space and thus the algorithm is executed for each one of $b_i \in B$.
 For each iteration, lines 11--12 product the following equation: 
 {\footnotesize
+\vspace{-2mm}
 \begin{align}
-g^h_{a,o,j} =  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \nonumber
+g^h_{a,o,j} =  \int_{x_{s'}} \oplus_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \nonumber
 \end{align}
 }
-where the $\alpha$-vector set are case functions primed by substitution in \textsc{Prime} 
-and \texttt{Backup} performs regression on the $\alpha$-vectors.
-Boolean regression is performing a symbolic restriction while a continuous regression is a symbolic substitution as defined previously.
-Line 13 takes the cross-sum of the two sets which takes every possible combination of observations according to the $\alpha$-vectors. 
-All vectors are then multiplied by the discount $\gamma$ and added with the reward in line 14 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
-The same operation is performed for all actions and then in line 15 the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors for the next horizon. 
-The $\alpha$-vector that maximizes the value-function with respect to this belief and save it as the new $\Gamma$-set for the next horizon.
+where the first probability is from equation (2) and the next probability is from equation (1) of the example.
+The $\alpha$-vectors are case functions primed using substitution in \textsc{Prime}.
+\texttt{Backup} performs regression on the $\alpha$-vectors where \textit{Boolean regression} is performing a symbolic restriction and \textit{Continuous regression} is a symbolic substitution, both SDP operations.
+Lines 14--16 cover closed form operations on the $\Gamma$-set such as cross-sum and union.
+Finally lines 18--20 find the maximum $\alpha$-vector within the $\Gamma^h$ set and add this to the final $\Gamma_{PBVI}^h$ if it is not already in the set.
+%Line 13 takes the cross-sum of the two sets, followed by adding the reward to the discounted result.
+For all actions the  $\Gamma$-set is added to the $\Gamma^h$ of the current horizon.
+The result is pruned as explained earlier.%which takes every possible combination of observations according to the $\alpha$-vectors.
+%All vectors are then multiplied by the discount $\gamma$ and added with the reward in line 14 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
+%The same operation is performed for all actions and then in 
+%Line 15 computes the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors for the next horizon. 
+%The $\alpha$-vector that maximizes the value-function with respect to this belief and save it as the new $\Gamma$-set for the next horizon.
 The action from which this $\alpha$-vector was chosen from is the maximum policy.
-This operation is performed for each of the belief points $b_i \in B$ and for each belief point one dominant $\alpha$-vector is stored in the $\Gamma$-set. The algorithm iterates until convergence. 
+%This operation is performed for each of the belief points $b_i \in B$ and for each belief point one dominant $\alpha$-vector is stored in the $\Gamma$-set. The algorithm iterates until convergence. 
 In the next section we discuss the main part of our algorithm which is generating relevant observation partitions.
@@
 $\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $. 
-This operation of line 3 in \textsc{GenRelObs} is similar to that of \texttt{Regress} multiplied by the observation regressed model.
+This operation of line 3 in \texttt{GenRelObs} is similar to that of \texttt{Regress} multiplied by the observation regressed model.
 For continuous variables the continuous regression and for discrete variables boolean restriction is used as defined in the previous section.
@@
 {\footnotesize
+\vspace{-3mm}
 \begin{align}
-\alpha_1^1(s) &= 
+\alpha_1^1(t_s) &= 
 \begin{cases}
@@
 \hspace{10mm} 
-\alpha_2^1(s) = \top: -1 \nonumber
+\alpha_2^1(t_s) = \top: -1 \nonumber
 \end{align}
@@
 		{
-    	$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) \cdot P(xd'_{s_i}| xd_{s_i},a)\cdot \alpha_j $;\
+    	$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) \cdot P(xd'_{s_i}| xd_{s_i},a)\cdot \alpha_j $\;
+    	\emph{// Point-based backup operation}\\
 		}  
@@
 		$\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i \cdot \alpha_j(xd_s,xd_o)$\}\ \;
+		\emph{// Generate value of $\alpha$-vector for belief $b_i$ as a function of observations}\\
 		}
@@
 		$z := max(\delta_j(xd_o),z)$\;
+		\emph{// Finding partition points on the observation space}\\
     	}
@@
    		$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$ \;
+   		\emph{// Generate probabilities and partitions of observations that distinguish optimal $\alpha$-vector (conditional plan)}\\
     \Return{$P(z)$} \;
@@
 {\footnotesize
+\vspace{-2mm}
 \begin{align}
-\alpha_1^2(s,o) &= 
+\alpha_1^2(t_s,t_o) &= 
 \begin{cases}
@@
 \end{cases}
-\hspace{4mm} 
-\alpha_2^2(s,o) &= \begin{cases}
+\hspace{2mm} 
+\alpha_2^2(t_s,t_o) &= \begin{cases}
 (t_s - 10 < t_o<t_s) &: -0.1 \\
@@
 The resulting $\delta$-functions only depend on the observation: $\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$.
-For $b_1: U[t_s;2,6]$ the two vectors above the resulting $\delta$-functions are defined below: 
+For $b_1: U[t_s;2,6]$ assume the following simple $\delta$-functions:
+%the two vectors above the resulting $\delta$-functions are defined below:
 {\footnotesize
+\vspace{-2mm}
 \begin{align}
-\delta_{1}(o) &= 
+\delta_{1}(t_o) &= 
 \begin{cases}
- (2<t_o<6) &: 15 - 2.5*t_o \\
-(-4<t_o<2) &: 10 \\
-(-8<t_o<-4) &: 20 + 2.5*t_o 
+ (-\infty < t_o< \infty) &: -0.4*t_o +6.8\\
 \end{cases}
 \hspace{5mm} 
-\delta_{2}(o) &= \begin{cases}
- (2<t_o<6) &: -0.15 + 0.025*t_o \\
-(-4<t_o<2) &: -0.1 \\
-(-8<t_o<-4) &: -0.2 - 0.025*t_o 
+\delta_{2}(t_o) &= \begin{cases}
+ (-\infty < t_o< \infty) &: 1.3*t_o +3.4\\
 \end{cases}
@@
 \end{align}
+\vspace{-5mm}
 }
+%{\footnotesize
+%\begin{align}
+%\delta_{1}(t_o) &= 
+%\begin{cases}
+% (2<t_o<6) &: 15 - 2.5*t_o \\
+%(-4<t_o<2) &: 10 \\
+%(-8<t_o<-4) &: 20 + 2.5*t_o 
+%\end{cases}
+%\hspace{5mm} 
+%\delta_{2}(t_o) &= \begin{cases}
+% (2<t_o<6) &: -0.15 + 0.025*t_o \\
+%(-4<t_o<2) &: -0.1 \\
+%(-8<t_o<-4) &: -0.2 - 0.025*t_o 
+%\end{cases}
+%\nonumber
+%\end{align}
+%}
 
@@
 {\it (left)} Beliefs $b_1,b_2$ for 1D \texttt{Power Plant} example; 
-{\it (middle)} Observation dependent function for $b_1$ that partition the observation space into 3 regions with different probabilities $P(z_i)$ ; 
-{\it (right)} Relevant observation partitions and their probabilities for $b_2$.
+{\it (middle)} Observation dependent function for $b_1$ that partition the observation space into 2 regions with different probabilities $P(z_i)$ ; 
+{\it (right)} 3 Relevant observation partitions and their probabilities for $b_2$.
 }
@@
 According to continuous observation space defined in \cite{pascal_ijcai05}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space.
-In this work, numerical solutions are proposed to find the roots of any two observation dependent function.
+In their work, numerical solutions are proposed to find the roots of any two observation dependent function.
 Here we use the symbolic power of the max-operator to find all the maximum boundary regions of two or more $\delta$-functions at the same time.
@@
 {\footnotesize
+\vspace{-2mm}
 \begin{align}
-\mathrm{max} \Bigg(\delta_{1}(o),\delta_{1}(o)\Bigg) &= 
+\mathrm{max} \Bigg(\delta_{1}(t_o),\delta_{2}(t_o)\Bigg) &= 
 \begin{cases}
- (2<t_o<6) &: 15 - 2.5*t_o \\
-(-4<t_o<2) &: 10 \\
-(-8<t_o<-4) &: 20 + 2.5*t_o 
+ (t_o<2) &: -0.4*t_o +6.8 \\
+(t_o>2) &: 1.3*t_o +3.4
 \end{cases}
@@
 }
-The three partitions define 3 observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the three distinct observation partitions is found.
+The two partitions define observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the two distinct observation partitions is found.
 If we integrate out the state and observation from the observation model and the belief state regardless of a particular $\alpha$-vector, it integrates to one:
@@
 \begin{cases}
- (2<t_o<6) &: 0.2 \\
-(-4<t_o<2) &: 0.6\\
-(-8<t_o<-4) &:0.1999
+ (t_o<2) &: 0.2 \\
+(2<t_o) &: 0.8\\
 \end{cases} 

### 2012-06-01 trunk/cpomdp/camdp-xadd/src/doc/NIPS/draft4-10page.tex

@@
    {
-       $h:=h+1, \Gamma^h :=\emptyset$\;
+       $h:=h+1, \Gamma_{PBVI}^h :=\emptyset$\;
        \ForEach {$b_i \in B$}
@@
      	$\alpha_{b_i}^h :=\ \argmax_{\alpha_j \in \Gamma^h} \Gamma^{h}$\;
-     	$\Gamma_{PBVI}^h :=\ \Gamma^h \cup \alpha_{b_i}^h$\;
+     	$\Gamma_{PBVI}^h :=\ \Gamma_{PBVI}^h \cup \alpha_{b_i}^h$\;
       }
@@
 The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.
-Lines 7--15 of the \textsc{PBVI} algorithm show the general algorithm used for both discrete and continuous observations. We will discuss this according to our symbolic framework in the next section. 
-For all $\alpha$-vectors in the current $\Gamma$-set the \textsc{Prune} function uses an LP-solver to prune any inconsistency in the branches of this vector in line 16 similar to the approach in ~\cite{penberthy94}.
+Lines 7--15 of the \texttt{PBVI} algorithm show the general algorithm used for both discrete and continuous observations. We will discuss this according to our symbolic framework in the next section. 
+For all $\alpha$-vectors in the current $\Gamma$-set the \texttt{Prune} function uses an LP-solver to prune any inconsistency in the branches of this vector in line 16 similar to the approach in ~\cite{penberthy94}.
 A pairwise dominance test is also performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$
@@
 
-For discrete and continuous states and discrete observations, we use the symbolic version of Monahan's algorithm as the DC-POMDP solution represented in \textsc{PBVI}.
+For discrete and continuous states and discrete observations, we use the symbolic version of Monahan's algorithm as the DC-POMDP solution represented in \texttt{PBVI}.
 The same approach is required for continuous observations only based on point-based dynamic programming backups.
@@
 \begin{align}
-g^h_{a,o,j} =  \int_{x_{s'}} \oplus_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \nonumber
+g^h_{a,o,j} =  \int_{x_{s'}} \large \oplus_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \nonumber
 \end{align}
@@
 where the first probability is from equation (2) and the next probability is from equation (1) of the example.
-The $\alpha$-vectors are case functions primed using substitution in \textsc{Prime}.
+The $\alpha$-vectors are case functions primed using substitution in \texttt{Prime}.
 \texttt{Backup} performs regression on the $\alpha$-vectors where \textit{Boolean regression} is performing a symbolic restriction and \textit{Continuous regression} is a symbolic substitution, both SDP operations.
@@
 In the next section we discuss the main part of our algorithm which is generating relevant observation partitions.
-
-\subsection{PBVI for DC State and DC Observations} 
-The main operations required for value iteration in DC-POMDP is generating relevant observations to perform the bellman backup operation. 
-The \texttt{GenRelObs} algorithm defines the steps required to generate these partitions. 
-We use the second iteration of the \texttt{Power Plant} example to demonstrate partitioning of the entire continuous observation space.
-The main assumption is to consider some belief state and obtaining the relevant partitions according to that belief.
-We choose these beliefs such that they cover the entire probable state space; e.g.
-$b_1: U[t_s;2,6]$ and $b_2: U[t_s;6,11]$.
-
-Starting with an empty set of relevant observation partitions $z$, the algorithm first eliminates the next state variable $xd'_s$.
-It multiplies the observation model and the transition model in the set of $\alpha$-vectors of the previous horizon:
-$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $. 
-This operation of line 3 in \texttt{GenRelObs} is similar to that of \texttt{Regress} multiplied by the observation regressed model.
-For continuous variables the continuous regression and for discrete variables boolean restriction is used as defined in the previous section.
-Assuming that after $h=1$ the set of $\alpha$-vectors are defined as:
-{\footnotesize
-\vspace{-3mm}
-\begin{align}
-\alpha_1^1(t_s) &= 
-\begin{cases}
- (t_s>200) &: -1000 \\
-\neg(t_s>200) &: 100 \\
-\end{cases}
-\hspace{10mm} 
-\alpha_2^1(t_s) = \top: -1 \nonumber
-\end{align}
-}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 		{
+		\emph{// Point-based backup operation}\\
     	$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) \cdot P(xd'_{s_i}| xd_{s_i},a)\cdot \alpha_j $\;
-    	\emph{// Point-based backup operation}\\
 		}  
@@
 		{
+		\emph{// Generate value of $\alpha$-vector for belief $b_i$ as a function of observations}\\
 		$\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i \cdot \alpha_j(xd_s,xd_o)$\}\ \;
-		\emph{// Generate value of $\alpha$-vector for belief $b_i$ as a function of observations}\\
 		}
@@
 		{
+		\emph{// Finding partition points on the observation space}\\
 		$z := max(\delta_j(xd_o),z)$\;
-		\emph{// Finding partition points on the observation space}\\
     	}
     	\emph{// $\phi_{z_k}$ logical term on each partition of $z$}\\
+    	\emph{// Generate probabilities and partitions of observations that distinguish optimal $\alpha$-vector (conditional plan)}\\
    		$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$ \;
-   		\emph{// Generate probabilities and partitions of observations that distinguish optimal $\alpha$-vector (conditional plan)}\\
     \Return{$P(z)$} \;
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-For $h=2$ the resulting $\alpha$-vectors after this regression step are functions of the observation and current state: 
+\subsection{PBVI for DC State and DC Observations} 
+The main operations required for value iteration in DC-POMDP is generating relevant observations to perform the bellman backup operation. 
+The \texttt{GenRelObs} algorithm defines the steps required to generate these partitions. 
+We use the second iteration of the \textsc{Power Plant} example to demonstrate partitioning of the entire continuous observation space.
+The main assumption is to consider some belief state and obtaining the relevant partitions according to that belief.
+We choose these beliefs such that they cover the entire probable state space; e.g.
+$b_1: U[t_s;2,6]$ and $b_2: U[t_s;6,11]$.
+
+Starting with an empty set of relevant observation partitions $z$, the algorithm first eliminates the next state variable $xd'_s$.
+It multiplies the observation model and the transition model in the set of $\alpha$-vectors of the previous horizon:
+$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $. 
+This operation of line 3 in \texttt{GenRelObs} is similar to that of \texttt{Regress} multiplied by the observation regressed model.
+For continuous variables the continuous regression and for discrete variables boolean restriction is used as defined in the previous section.
+Assuming that after $h=1$ the set of $\alpha$-vectors are defined equal to the reward function of equation (1).
+%{\footnotesize
+%\vspace{-3mm}
+%\begin{align}
+%\alpha_1^1(t_s) &= 
+%\begin{cases}
+% (t_s>200) &: -1000 \\
+%\neg(t_s>200) &: 100 \\
+%\end{cases}
+%\hspace{10mm} 
+%\alpha_2^1(t_s) = \top: -1 \nonumber
+%\end{align}
+%}
+For $h=2$ the resulting $\alpha$-vectors after the regression step are functions of the observation and current state: 
 {\footnotesize
@@
 \caption{\footnotesize 
-{\it (left)} Beliefs $b_1,b_2$ for 1D \texttt{Power Plant} example; 
+{\it (left)} Beliefs $b_1,b_2$ for 1D \textsc{Power Plant} example; 
 {\it (middle)} Observation dependent function for $b_1$ that partition the observation space into 2 regions with different probabilities $P(z_i)$ ; 
@@
 \begin{cases}
- (t_o<2) &: -0.4*t_o +6.8 \\
-(t_o>2) &: 1.3*t_o +3.4
+ z_1:(t_o<2 : -0.4*t_o +6.8) \\
+z_2:(t_o>2 : 1.3*t_o +3.4)
 \end{cases}
@@
 $P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$
-For our example the following holds: 
-{\footnotesize
-\begin{align}
-P(z_k)=
-\begin{cases}
- (t_o<2) &: 0.2 \\
-(2<t_o) &: 0.8\\
-\end{cases} 
-\nonumber
-\end{align}
-}
+For our example the probabilities are defined as $P(z_1)=0.2$ and $P(z_2)=0.8$.
+%{\footnotesize
+%\begin{align}
+%P(z_k)=
+%\begin{cases}
+% (t_o<2) &: 0.2 \\
+%(2<t_o) &: 0.8\\
+%\end{cases} 
+%\nonumber
+%\end{align}
+%}
+
 Hence we can now use the algorithms and methods of a discrete observation setting using the probabilities of the partitioned observation space in Monahan's algorithm. 
@@
 \section{Empirical Results}
-We evaluated our continuous POMDP solution using XADDs on the one-dimensional
+%We evaluated our continuous POMDP solution using XADDs on the one-dimensional
 %\textsc{Power Plant} example and another variant of this problem with two variables. \footnote{
@@
 %in Google Code; anonymity restrictions prevent us from providing this link in the submitted paper version.}
-We start by defining the 2-dimensional description of this problem:
+The 2-dimensional model description is provided below:
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The two actions of increase and decrease are defined based on the change in both the temperature and the pressure.
-For the increase action the transition functions and the reward function is defined as below:
+For the increase action we define: %the transition functions and the reward function is defined as below:
 {\footnotesize
@@
 P(t_s'|\vec{t_s},inc)= \delta\left[ t_s' - (t_s +10) \right]\nonumber
-\\
+\end{align}
+}
+There is a high reward for staying withing the safe temperature and pressure range since it produces power, else depending on how safe it is to have values higher or lower than the safe range, penalty is defined.
+{\footnotesize
+\vspace{-4mm}
+\begin{align}
 R(t_s,p_s,inc) &= 
@@
 (5 \leq p_s \leq 15)\wedge (t_s \leq 95)&: -1\\
-(5 \leq p_s \leq 15)\wedge (t_s \geq 95)&: -3\\				
-(p_s \leq 5) &: -3\\						
+%(5 \leq p_s \leq 15)\wedge (t_s \geq 95)&: -3\\				
+%(p_s \leq 5) &: -3\\						
 (p_s \geq 15) &: -5\\ 
+else &: -3
 \end{cases}\nonumber
@@
 }
-There is a high reward for staying withing the safe temperature and pressure range since it produces power, else depending on how safe it is to have values higher or lower than the safe range, penalty is defined.
 As for the decrease action, the transition functions reduce the temperature by 5 units and the pressure by 10 units as long as the pressure stays above zero.
 For the reward function, we assume that there is always a small penalty for decreasing the values because power can not be generated.
-
 For the observation model we consider two continuous uniform distributions such as the following:  
@@
 the two versions of \textsc{Power Plant} have been performed for up to 6 horizons.
-Comparing the two problem sizes demonstrates the effect of the number of state-observation variables in our algorithm.
+%Comparing the two problem sizes demonstrates the effect of the number of state-observation variables in our algorithm.
 As the algorithm progresses, the time required to compute the probability of the partitions and finding the maximum $\alpha$-vector with respect to beliefs increases for both problem sizes and significantly more for the 2D version.
-The stability in time is due to the fact that each stage almost takes the same time since it has converged quickly.%???
+%The stability in time is due to the fact that each stage almost takes the same time since it has converged quickly.%???
 Increase in the problem size, increases the partition numbers on the observation space and this produces more $\alpha$-vectors which also effects the space required to perform the algorithm.
 The number of vectors stays the same for most horizons and they drop after convergence in the 2D problem instance.
-The number of nodes in each iteration increases as more space is required to perform higher iterations.
+%The number of nodes in each iteration increases as more space is required to perform higher iterations.
 This shows that although the 2D instance takes more time and space than the 1D instance, it still converges within reasonable resources.
@@
 In Figure ~\ref{fig:3D}  we present plots of the maximum $\alpha$-vectors of belief $b_1$ for different iterations of the 2D problem instance.
-Each plot demonstrates the value of each vector as a function of the pressure and temperature.
+%Each plot demonstrates the value of each vector as a function of the pressure and temperature.
 Starting with the first iteration, the value is highest for the reward range ($5<p<15 \wedge 95<t<105$) and -1 or less for other places.
@@
 In the last plot, horizon 6 has better tuned the value function so that higher temperatures and pressures increase the value of the maximum $\alpha$-vector and also within the reward range, finer grain partitions have been formed.
-Also note that the policy depends on the initial belief and for these plots they were based on $b_1$ which was lower than the reward range.
+%Also note that the policy depends on the initial belief and for these plots they were based on $b_1$ which was lower than the reward range.
 The first action would then mean it has performed an increase while the other iterations decrease and increase respectively.
@@
 
-\section{Related Work and Conclusion} 
-This work has used the concept of continuous states and the SDP solution to MDPs from ~\cite{sanner_uai11} and continuous observations from the work of \cite{pascal_ijcai05}.
+\section{Conclusion} 
+%This work has used the concept of continuous states and the SDP solution to MDPs from ~\cite{sanner_uai11} and continuous observations from the work of \cite{pascal_ijcai05}.
 It is exclusive in the sense that it brings together the continuous state and observation POMDPs using a symbolic approach.
-In many applications of the POMDP framework, the state and observation space have been considered as discrete values such as ~\cite{steam2} here we avoid this general simplification and work with the real values from sensors.
-%sampling, monte carlo, particle filter, gaussians, perseus, pergeus
-
-There has been prior work on approximate solutions to POMDPs with large or continuous state spaces ~\cite{Thrun99h}, ~\cite{Perseus} and some has been extended to the continuous observation setting using methods such as sampling ~\cite{Perseus_cont}.
-In most approximate continuous state POMDP work, continuous or large discrete actions has been used.
+%In many applications of the POMDP framework, the state and observation space have been considered as discrete values such as ~\cite{steam2} here we avoid this general simplification and work with the real values from sensors.
+%%sampling, monte carlo, particle filter, gaussians, perseus, pergeus
+%
+%There has been prior work on approximate solutions to POMDPs with large or continuous state spaces ~\cite{Thrun99h}, ~\cite{Perseus} and some has been extended to the continuous observation setting using methods such as sampling ~\cite{Perseus_cont}.
+%In most approximate continuous state POMDP work, continuous or large discrete actions has been used.
 Thus we can extend the current work using \cite{zamani_aaai12} to contain continuous actions and provide exact or approximate solutions.
-
-%\section{Conclusion}
+%
+%%\section{Conclusion}
 We presented the first exact solution to continuous states and observations in a DC-POMDP framework.

### 2012-06-01 cpomdp/camdp-xadd/src/doc/NIPS/draft4-10page.tex
Updated paper

--- 2012-06-01 trunk/cpomdp/camdp-xadd/src/doc/NIPS/draft4-10page.tex
@@
 \newcommand{\close}{\mathit{close}}
+\newcommand{\high}{\mathit{high}}
+\newcommand{\low}{\mathit{low}}
 \newcommand{\fix}{\marginpar{FIX}}
@@
 %polynomial transition and observation functions.  
-We demonstrate
-proof-of-concept results on uni- and multi-variate state and
-observation steam plant control.
+We demonstrate our algorithm on a power plant
+control problem requiring continuous state 
+\emph{and} observations.
+%empirically evaluate on real-world inspired power 
+%proof-of-concept results on uni- and multi-variate state and
+%observation power plant control.
 \end{abstract}
@@
 problems~\cite{kaebling}.  In recent years, point-based value
-iteration methods~\cite{pbvi_jair06,hsvi2,Perseus} have proved
+iteration methods~\cite{pbvi_jair06,hsvi2,Perseus,gapmin} have proved
 extremely successful at scaling (approximately) optimal POMDP
@@
 point-based backups for discrete state and continuous observation
-problems (where multivariate observations must be independent).  While
-restricted to discrete state, \cite{pascal_ijcai05} provides an
+problems (where multivariate observations must be conditionally
+independent).  While
+restricted to discrete states, \cite{pascal_ijcai05} provides an
 important insight that we exploit in this work: \emph{only a finite
@@
 technique of~\cite{sanner_uai11}.  
-We demonstrate
-proof-of-concept results on uni- and multi-variate state and
-observation steam plant control showing the tractability of our
-exact symbolic dynamic programming point-based approach.
+We conclude by demonstrating our algorithm on a power plant
+control problem requiring bivariate continuous state and observations.
 
@@
 such arcs only requires restrictions on the variable elimination
-ordering during dynamic programming.  In these Bayes nets, 
+ordering during dynamic programming.  In these transition and
+observation Bayes nets, 
 probabilities over 
@@
 \vspace{-2mm}
-\caption{\footnotesize Left: Example conditional plan $\beta^h$ for discrete observations. Right: The maximum $\alpha$ vector of iteration 4 for the \textsc{Power Plant}
-as a decision diagram: 
+\caption{\footnotesize (left) Example conditional plan $\beta^h$ for discrete observations; (right) an example piecewise $\alpha$-vector for continuous state $t$ represented
+as a 
+%The maximum $\alpha$ vector for a belief state at iteration 4 for the \textsc{\bf Power Plant} problem in Section~\ref{}
+decision diagram: 
 the \emph{true} branch is solid, the \emph{false}
@@
 %\caption{%\footnotesize 
-%The optimal value function for \textsc{Power Plant}
+%The optimal value function for \textsc{\bf Power Plant}
 %as a decision diagram: 
@@
 
-\textbf{Example} \textsc{(Power Plant)~\cite{steam2}} \emph{The steam
+\textbf{Example} \textsc{\bf (Power Plant)~\cite{steam2}} \emph{The steam
 generation system of a power plant aims to provide hot steam to the
@@
 and $t_s \in \mathbb{R}$ highly uncertain.  Action choices are either
-to open or close a presure valve $a \in \{ \open, \close \}$.}
+to open or close a pressure valve $a \in \{ \open, \close \}$.}
 
-For the \textsc{Power Plant} example, the transition and reward functions are defined as below:
+We initially present two 
+simple examples labeled \textsc{\bf 1D-Power Plant} 
+using only the temperature state $t_s$.
+First, the transition and reward functions common to both problems
+are defined as below where simply, the temperate increments or
+decrements according to the action:
 % For the transition probability, we use the Dirac function to model the deterministic equations.
@@
 %contribution? 
-This example will try to maximize its reward using observations on the temperature.
-The observations received before closing the plant to a safe temperature are therefore critical.
-For the observation model we begin with a discrete model for the observation $z$ with values $high, low$ such as the following for the $\open$ action:
+Next we introduce the \textsc{\bf Discrete Obs. 1D-Power Plant} variant using 
+a discrete binary observation space with possible
+observations $o \in mathcal{O} = \{\high,\low\}$:
 {\footnotesize
@@
 \begin{align}
-P(z=high|t_s',\open) = 
+P(o=\high|t_s',\open) = 
 \begin{cases}
@@
 \hspace{5mm}
-P(z=low|t_s',\open) = 
+P(o=\low|t_s',\open) = 
 \begin{cases}
@@
 
-After defining the solution to the discrete observation with a continuous state DC-POMDP, we can extend this to consider continuous observations using a continuous uniform noise distribution which is more realistic in many domains.
-%For a certain interval, the probability is a fixed number, while it is zero elsewhere.
-The integral of the uniform distribution should sum to one.
+Finally we introduce the \textsc{\bf Cont. Obs. 1D-Power Plant}
+variant with continuous observation $t_o$.  
+We define the observation probability over $t_o$ as uniform on
+an interval 10 units wide, centered at $t_s$.
 {\footnotesize
@@
 }
-Next we define the general solution to solving DC-POMDPs.
 
 \section{DC-POMDP solution}
-The DC-POMDP will choose actions in order to compute an optimal plan over time.
-This plan is the policy $\pi$ of the agent  depending on the history of actions and observations over a Markovian state.
-In the partially observable case where there is uncertainty on the states, the policy is a function over a continuous set of probability distributions over $|\mathcal{S}|$ known as the belief state $b = P(x_s,d_s)$.
 
-At any of the horizon steps, a DC-POMDP policy $\pi(b)$ is defined as the iterative plan on executing an action and observing all probable outcomes of the set of observations from the $(h-1)$-step.
-The belief state is updated after this according to the Bayes rule: $\sum_{d_s}\int_{x_s} b(xd_s)\mathcal{T}(xd'_s,xd_s,a)\mathcal{Z}(xd'_s,xd_o,a)$.
-\footnote{For simiplicity we use ($xd_s = x_s,d_s$) and ($xd_o=x_o,d_o$)}
-The optimal policy is defined over the belief state as the sum of the expected discounted rewards over horizon $h$ starting with belief $b_0$.
+In a DC-POMDP, the agent does not directly observe the states and thus
+must maintain a belief state $b(xd_s) = P(xd_s)$.\footnote{For
+compactness we define $xd_s := (x_s,d_s)$ and $xd_o:=(x_o,d_o)$.}  For a
+given belief state $\vec{b} = b(\cdot)$, a POMDP policy $\pi$ can be
+represented by a tree corresponding to a conditional plan $\beta$.  An
+h-step conditional plan $\beta^h$ can be defined recursively in terms
+of $(h-1)$-step conditional plans as shown in
+Fig.~\ref{fig:cond_plan}.
+
+Our goal is to find a policy $\pi$ that maximizes the value function,
+defined as the sum of expected discounted rewards over horizon $h$
+starting from initial belief state $\vec{b}$:
+\begin{equation}
+V^h_\pi(\vec{b}) = E_{\pi} \left[ \sum\nolimits_{t=0}^{h} \gamma^t \cdot r_t \Big| \vec{b}_0 = \vec{b} \right]
+\end{equation}
+where $r_t$ is the reward obtained at time $t$ and $\vec{b}_0$ is the
+belief state at $t=0$.  For finite $h$ and belief state $\vec{b}$, the
+optimal policy $\pi$ takes the form of an $h$-step conditional plan
+$\beta^h$.  For $h = \infty$, the optimal discounted ($\gamma < 1$)
+value function can be approximated arbitrarily closely by using
+a sufficiently large, but finite $h$~\cite{kaebling}.
+
+Even when the state is continuous (but the actions and observations
+are discrete), 
+the optimal POMDP value function for finite horizon $h$ is a piecewise linear
+and convex function of the belief state $\vec{b}$~\cite{Perseus_cont} hence 
+$V^h$ is given by a maximum over a finite set of
+``$\alpha$-vectors'' $\alpha^h_i$:
+\begin{equation}
+V^h(\vec{b}) = \max\nolimits_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot \vec{b}
+\end{equation}
+Later on when we tackle continuous state \emph{and} observations,
+we note that we will dynamically derive an optimal, 
+finite partitioning of the observation
+space for a given belief state and hence reduce the continuous
+observation problem back to a discrete observation problem at every
+horizon.
+
+The $\Gamma^h$ in this optimal $h$-stage-to-go value function can be
+computed via Monahan's dynamic programming approach to \emph{value
+iteration} (VI)~\cite{monahan82}.  
+Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (2):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as
+$\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
+\textrm{\large $\boxplus$} Q = \{ \vec{p} + \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}
 {\footnotesize
-\begin{equation}
-V^h_{\Pi^*}(\vec{b}) = E_{\pi^*} \left[ \sum_{h=0}^{H} \gamma^t \cdot r_h \Big| \vec{b}_0 = \vec{b} \right]\nonumber
-\end{equation}
+\begin{align} 
+g^h_{a,o,j} &=  \int_{x_{s'}} \sum_{d_{s'}} p(o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \label{eq:backup} \\
+\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \label{eq:cross_prod} 
+\end{align}
+\begin{align}
+\Gamma^h  &= \bigcup_a \Gamma^h_a 
+\end{align}
 }
-Solving the value function for the optimal policy is performed using Dynamic programming algorithms such as value iteration. 
-%The decision at each time step is to pick an action based on past actions and observations. 
-The possible set of conditional plans $\beta^h$ correspond to a decision tree.
-Figure 1 left, shows the decision tree for the simple case of three horizons with discrete observations.
-It represents a conditional plan consisting of an action and two possible observation strategies.
-For continuous observations, conditional plans are decision trees with infinitely many branches.
-In our solution we show that instead of a policy tree with infinite branches, we can create partitions based on relevant observation values, resulting in a policy tree similar to that of figure 1.%From here on we assume this finite set of observation partitioning as $\mathcal{O}$.
+where $r_a = r(x_s,d_s,a)$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 {
-   $V^0:=0, h:=0$\;
+   $V^0:=0, h:=0, \Gamma_{PBVI}^0 = \{ \alpha_1^0 \} $\;
    \While{$h < H$}
    {
-       $h:=h+1, \Gamma_{PBVI}^h :=\emptyset$\;
+       $h:=h+1, \Gamma^h :=\emptyset, \Gamma_{PBVI}^h :=\emptyset$\;
        \ForEach {$b_i \in B$}
@@
        		\If {$ContObs =true$}
-       			{$(O^h,P(xd_{o_i}|xd_{s_i})) \,:=\,$ \genObs{$\Gamma^{h-1},a,b_i$}\;}
-       		 \ForEach {$xd_{o_i} \in \textsc{O}^{h}$}
+       			{\emph{// Continuous observation so derive observation set $O_i^h$ for belief $b_i$} \;
+			  $(O_i^h,P(O_i^h|xd_{s})) \,:=\,$ \genObs{$\Gamma_{PBVI}^{h-1},a,b_i$}\;}
+       		 \ForEach {$o \in \mathcal{O}_i^{h}$}
        		 {
-				\ForEach {$\alpha_j^{h-1} \in \Gamma^{h-1}$}
+				\ForEach {$\alpha_j^{h-1} \in \Gamma_{PBVI}^{h-1}$}
        			{
@@
    	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
-   	 		    	$\Gamma_{a,xd_{o_i},j}^h \,:=\, P(xd_{o_i}|xd_{s_i}) \otimes$ \backup{$\alpha_j^{h'},a$}\;
+%   	 		    	$\Gamma_{a,xd_{o_i},j}^h \,:=\, P(o|xd_{s}) \cdot$ \backup{$\alpha_j^{h'},a$}\;
+   	 		    	{$g_{a,o,j}^h \,:=\, $ see Eq~\eqref{eq:backup}}
        	      	}
-       	      	$\Gamma_{a,xd_o}^h\,:=\,\textrm{\large $\boxplus$} \Gamma_{a,xd_{o_i}}^h$\;
+%       	      	$\Gamma_{a,xd_o}^h\,:=\,\textrm{\large $\boxplus$} \Gamma_{a,xd_{o_i}}^h$\;
+%       	      	$\Gamma_{a,xd_o}^h\,:=\, $ see Eq~\eqref{eq:cross_prod}\;
        	     }
-           $\Gamma_a^{h} \,:=\,R_a \oplus \gamma \cdot \Gamma_{a,xd_o}^h$\;
+%           $\Gamma_a^{h} \,:=\,R_a \oplus \gamma \cdot \Gamma_{a,xd_o}^h$\;
+            $\Gamma_a^{h} \,:=\, $ see Eq~\eqref{eq:cross_prod}\;
             $\Gamma^{h} \,:=\, \Gamma^{h} \cup \Gamma_a^{h}$\;
@@
       {
-     	$\alpha_{b_i}^h :=\ \argmax_{\alpha_j \in \Gamma^h} \Gamma^{h}$\;
+     	$\alpha_{b_i}^h :=\ \argmax_{\alpha_j \in \Gamma^h} \alpha_j \cdot b_i$\;
      	$\Gamma_{PBVI}^h :=\ \Gamma_{PBVI}^h \cup \alpha_{b_i}^h$\;
@@
 
-       \If{$V^h = V^{h-1}$}
+       \If{$\Gamma_{PBVI}^h = \Gamma_{PBVI}^{h-1}$}
            {break $\,$ \emph{// Terminate if early convergence}\;}
@@
 
-For discrete observations and discrete states the number of belief states is infinite, but the optimal value function for $h$ is proven to be a piecewise linear and convex function of the belief state \cite{smallwoodSondik}.
-The optimal value function is the maximum value over a finite set of alpha vectors  $\alpha^h_i$:%that define the hyperplanes:
-{\footnotesize
-\begin{equation}
-V^h_{\Pi^*}(\vec{b}) = \max_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot \vec{b}
-\end{equation}
-}
-where the belief states $\vec{b}$ and each $\alpha^h_i \in \Gamma^h$ are of dimension $|\mathcal{S}|$.
+\textbf{Point-based value iteration (PBVI)} computes the optimal value function only for a set of belief states $\{ b_i \}$ where $b_i := P(xd_s)$.
+The idea is straightforward and the main modification needed to Monahan's algorithm in Algorithm~\ref{alg:vi} is the loop from lines 19--21 where only $\alpha$-vectors optimal at some belief state are retained for subsequent iterations.
+We note that \emph{if} all beliefs reachable in $h$ steps from some initial set can be finitely enumerated and PBVI is executed for this set of beliefs, then optimality is guaranteed for the PBVI policy over these initial beliefs up to horizon $h$.
+%In actuality in the DC-POMDP setting, the generation of continuous observations will most often lead to an infinite number of reachable belief states even in one step, so this optimality guarantee of PBVI will only hold for DC-POMDPs in very special cases when the set of reachable beliefs is finite.
+Nonetheless, PBVI has been quite successful in practice without exhaustive enumeration of all reachable beliefs~\cite{pbvi_jair06,hsvi2,Perseus,gapmin}, which motivates our use of PBVI in this work.
 
-\textbf{Point-based dynamic programming backups} compute the optimal value function given a belief state $b$.
-The best conditional plan for the given belief, corresponds to an $\alpha$-vector which is computed using point-based backups:
-{\footnotesize
-\begin{align*}
-\alpha^{h+1}(b) = R(b) + \gamma \int_{xd_o} P(xd_o|a,b)\alpha^{h+1}(b)\nonumber
-\end{align*}
-}
-where $P(xd_o|a,b) = \int_{xd_s}\int_{xd'_s}b(s)P(xd'_s|xd_s,a)P(xd_o|xd'_s,a)$.
-
-Calculating all possible ways of building the value function independent of a certain belief state can also define the optimal value function.
-To derive the $\Gamma$-set of all possible $\alpha$-vectors for each horizon, we use the value iteration algorithm from ~\cite{monahan82}.
-Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (2):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as
-$\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
-\textrm{\large $\boxplus$} Q = \{ \vec{p} + \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}
-{\footnotesize
-\begin{align}
-g^h_{a,o,j} &=  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \\
-\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \\
-\Gamma^h  &= \bigcup_a \Gamma^h_a 
-\end{align}
-}
-where $r_a = r(x_s,d_s,a)$.  
-%Each action-dependent $\Gamma$ sums over the possible set of $g^h_{a,o,j}(\cdot)$ given an observation model to derive discrete observations $xd_o$ and for all $\alpha^{h-1}_{j} \in \Gamma^{h-1} $.
-The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.
-Lines 7--15 of the \texttt{PBVI} algorithm show the general algorithm used for both discrete and continuous observations. We will discuss this according to our symbolic framework in the next section. 
-For all $\alpha$-vectors in the current $\Gamma$-set the \texttt{Prune} function uses an LP-solver to prune any inconsistency in the branches of this vector in line 16 similar to the approach in ~\cite{penberthy94}.
-A pairwise dominance test is also performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$
-The result of this pairwise test is a dominant $\alpha$-vector which is added to the set of $\Gamma^h$. 
-%Here the cross-sum operation takes the set of $\alpha$ -vectors for each observation and produces a sum over vectors that depends on the number of previous iteration vectors.
-
-Note that the $\alpha$-vector set grows exponentially in the number of observations, i.e., $|\Gamma^{h}| = |\mathcal{A}||\Gamma^{h-1}|^{|\mathcal{O}|}$ during the cross-sum operation.
-For this reason even for a small number of discrete observations, tractable solutions may be hard to define.
-Here we do not restrict our observations to discrete ones, but extend it to a fully continuous approach where lines 7--8 will be executed to generate relevant observation partitions based on a belief state $b_i$.
-We use the number of  observation partitions $|\mathcal{O}|$ as the branching factor in the policy tree.
-We present the symbolic approach required to partition this continuous space.
+We pause for a moment to make a few notes about the \texttt{PBVI}
+algorithm in the context of DC-POMDPs.  If we have a fixed discrete
+observation set, i.e., $forall h \, \mathcal{O}^h = \mathcal{O}$, then
+the algorithm proceeds as usual, but if we have continuous observation
+variables then we will need to derive a relevant set of observations
+on line 10, which we describe in Section~\ref{sec:cont_obs}.  We need
+to guarantee that the integral on line 14 can be done in closed-form
+and we guarantee this for piecewise functions in
+Section~\ref{sec:disc_obs}.  Because in general we will deal with
+$\alpha$-functions represented as piecewise functions, we only assume
+that strict dominance testing is implemented for \texttt{Prune} in
+line 18.
 
 \section{Symbolic Dynamic Programming}
-Dynamic programming provides exact solutions to POMDP problems for small domains.
-As the domain grows, fully enumerating the state, action and observation space leads to intractable solutions even for low horizons.
-Avoiding this enumeration often requires approximating the solution.
-We claim an exact solution using symbolic dynamic programming.
-The SDP solution for fully observable domains has been presented in \cite{sanner_uai11} %using piecewise constant functions
-Our symbolic algorithm deals with continuous observations and states by generating the relevant observation partition at every time step and performing the backup operation using $\alpha$-vectors.
-We define the requirements to implement the value iteration algorithm for DC-POMDP.
+
+In this section we take a symbolic dynamic programming (SDP) approach
+to performing \texttt{PBVI} in Algorithm~\ref{alg:vi}.  This requires
+us to define all of the operations used in \texttt{PBVI} on a
+closed-form representation.  For SDP, this closed-form will be a
+piecewise case statement and it will stored compactly as a decision
+diagram~\cite{sanner_uai11}.  We first review this form and then
+proceed to define SDP operations specific to the case of DC-POMDPs.
 
@@
 %overview + example plant
-The \textsc{Power Plant} example represented all functions using the case structure which can be generally defined as:
-{\footnotesize 
+The \textsc{\bf Power Plant} example represented all functions using
+the case structure which can be generally defined as {\footnotesize
 \vspace{-3mm}
@@
 }
-Where $\phi_i$ are disjoint logical formulae defined over the state $(\vec{d},\vec{x})$ with logical ($\land,\lor,\neg$) combinations of boolean variables and inequalities ($\geq,>,\leq,<$) over continuous variables.
-The $f_i$ are function definitions either linear or quadratic in the continuous parameters.
+and this is the form we use to represent all functions in a DC-POMDP.
+Here $\phi_i$ are disjoint logical formulae defined over the state $(\vec{d},\vec{x})$ with logical ($\land,\lor,\neg$) combinations of boolean variables and inequalities ($\geq,>,\leq,<$) over continuous variables.
+The $f_i$ are general functions, we will restrict them later for the
+case of continuous observation POMDPs.
 
@@
 }
-We also require the following SDP operations on case statements already introduced in the literature: 
+The following SDP operations on case statements require more detail than can be provided here, hence we refer the reader to the relevant literature:
 \begin{itemize}
@@
 \item Substitution $f\sigma$ : Takes a set $\sigma$ of variables and their substitutions, where the LHS of the substitution operator $/$ represents the substitution variable and the RHS of the $/$ represents the expression that should be substituted in its place as in \cite{sanner_uai11}.
-\item Integration $\int_{x_1} f dx_1$:  Eliminates continuous variables by taking the integral of the polynomial function $f$ with respect to linear constraints on $x_1$  \cite{sanner_aaai12}.
+\item Integration $\int_{x_1} f dx_1$:  Eliminates continuous variables by taking the integral of case statements $f$ restricted to linear constraints and polynomial values~\cite{sanner_aaai12}.
 \end{itemize}
@@
 %xadd representation
-The data structure of \emph{extended ADD} (XADD)~\cite{sanner_uai11} is used to support
-case statements and the required operations.
-Here we use continuous SDP for both states and observations in DC-POMDPs.
-The $\alpha$-vectors partition into states of equal value, both of which can be represented using XADDs.
-For increasing efficiency linear constraint feasibility checkers such as LP solvers allow pruning unreachable paths in the XADD.
-Figure 1,right is an example of an XADD representation.
+The data structure of the \emph{extended algebraic decision diagram}
+(XADD)~\cite{sanner_uai11} is used to support case statements and the
+required operations.  Figure~\ref{fig:cond_plan} (right) 
+is an example of an XADD representation.
 
@@
 \subsection{PBVI for DC State and Discrete Observations} 
+\label{sec:disc_obs}
 
@@
 %To visualize we explain of the value iteration algorithm as defined in \texttt{VI} for DC-POMDPs.
-Each step of the algorithm is explained using the \textsc{Power plant} example for the first three horizons.
+Each step of the algorithm is explained using the \textsc{\bf power plant} example for the first three horizons.
 Note that if the input to this algorithm is continuous observations $ContObs=true$ then line 8 finds the relevant observation partitions, else discrete observations are assumed and their probabilities are given in the problem specifications such as section 2.
@@
 \subsection{PBVI for DC State and DC Observations} 
+\label{sec:cont_obs}
+
 The main operations required for value iteration in DC-POMDP is generating relevant observations to perform the bellman backup operation. 
 The \texttt{GenRelObs} algorithm defines the steps required to generate these partitions. 
-We use the second iteration of the \textsc{Power Plant} example to demonstrate partitioning of the entire continuous observation space.
+We use the second iteration of the \textsc{\bf Power Plant} example to demonstrate partitioning of the entire continuous observation space.
 The main assumption is to consider some belief state and obtaining the relevant partitions according to that belief.
@@
 \caption{\footnotesize 
-{\it (left)} Beliefs $b_1,b_2$ for 1D \textsc{Power Plant} example; 
+{\it (left)} Beliefs $b_1,b_2$ for 1D \textsc{\bf Power Plant} example; 
 {\it (middle)} Observation dependent function for $b_1$ that partition the observation space into 2 regions with different probabilities $P(z_i)$ ; 
@@
 %We evaluated our continuous POMDP solution using XADDs on the one-dimensional
-%\textsc{Power Plant} example and another variant of this problem with two variables. \footnote{
+%\textsc{\bf Power Plant} example and another variant of this problem with two variables. \footnote{
 %Full problem specifications and Java code to reproduce these experiments are currently 
@@
 
-{\bf \textsc{2D-Power Plant}:} We consider the more complex model of the power plant similar to \cite{steam2} where the pressure inside the water tank must be controlled to avoid mixing water into the steam or explosion of the tank.
+{\bf \textsc{\bf 2D-Power Plant}:} We consider the more complex model of the power plant similar to \cite{steam2} where the pressure inside the water tank must be controlled to avoid mixing water into the steam or explosion of the tank.
 We model the pressure variable $p$ as a partially observable variable from the observation readings of the pressure $po$.
@@
 In Figure ~\ref{fig:timeSpace}, a time and space analysis of
-the two versions of \textsc{Power Plant} have been performed for up to 6 horizons.
+the two versions of \textsc{\bf Power Plant} have been performed for up to 6 horizons.
 %Comparing the two problem sizes demonstrates the effect of the number of state-observation variables in our algorithm.

### 2012-06-01 cpomdp/camdp-xadd/src/doc/NIPS/draft4-10page.tex
Updated paper

@@
 algorithm in the context of DC-POMDPs.  If we have a fixed discrete
-observation set, i.e., $forall h \, \mathcal{O}^h = \mathcal{O}$, then
+observation set, i.e., $\forall h \, \mathcal{O}^h = \mathcal{O}$, then
 the algorithm proceeds as usual, but if we have continuous observation
@@
 \end{itemize}
-\vspace{-2mm}
+%\vspace{-2mm}
 %\emph{Restriction} takes a function $f$ to restrict only in cases

### 2012-09-28 cpomdp/camdp-xadd/src/CPOMDP/ComputeGamma.java

@@ -150,7 +150,7 @@ public class ComputeGamma {
 			for (int j=0;j<newalphas.size();j++)
 			{
 				regressedAlpha[i][j] = _context.apply(newalphas.get(j), nodeO, _context.PROD);
-				regressedAlpha[i][j] = _context.reduceLP(regressedAlpha[i][j] , _pomdp._alContSVars);
+				regressedAlpha[i][j] = _context.reduceLP(regressedAlpha[i][j] );
 			}
 		}
 		int crossSum[] = new int[(int) Math.pow(newalphas.size(), _obspartitionset.size())];
@@ -255,9 +255,11 @@ public class ComputeGamma {
 			 int result = _context.apply(xadd1.getValue(), xadd2.getValue(), _context.PROD);
 			 
 		}
+				XADDINode n = null;
 				if (n instanceof XADDINode)
 				{
 					int sp = ((XADDINode)n)._var;
+					XADDINode o = null;
 					int op = ((XADDINode)o)._var;
 					Decision d =  _context._alOrder.get(sp);
 					Decision od=  _context._alOrder.get(op);
@@ -290,6 +292,7 @@ public class ComputeGamma {
 							 negoeC._type = _context.LT_EQ;*/
 						int oo = overlapObservation(negoeC);
 						CompExpr negoeS = new CompExpr(e._expr._type, e._expr._lhs, e._expr._rhs);
+						int branch = 0;
 						if (branch==0) 
 							negoeS._type = negoeS.flipCompOper(negoeS._type);
 						if (oo != -1)

### 2012-09-28 cpomdp/camdp-xadd/src/CPOMDP/ComputeGammaOld.java

@@ -148,7 +148,7 @@ public class ComputeGammaOld {
 			for (int j=0;j<newalphas.size();j++)
 			{
 				regressedAlpha[i][j] = _context.apply(newalphas.get(j), nodeO, _context.PROD);
-				regressedAlpha[i][j] = _context.reduceLP(regressedAlpha[i][j] , _pomdp._alContSVars);
+				regressedAlpha[i][j] = _context.reduceLP(regressedAlpha[i][j] );
 			}
 		}
 		int crossSum[] = new int[(int) Math.pow(newalphas.size(), _obspartitionset.size())];

### 2012-09-28 cpomdp/camdp-xadd/src/CPOMDP/contObs/ContComputeGamma.java

@@ -174,7 +174,7 @@ public class ContComputeGamma {
 			for (int j=0;j<newalphas.size();j++)
 			{
 				regressedAlpha[i][j] = _context.apply(newalphas.get(j), dd, _context.PROD);
-				regressedAlpha[i][j] = _context.reduceLP(regressedAlpha[i][j] , _pomdp._alContSVars);
+				regressedAlpha[i][j] = _context.reduceLP(regressedAlpha[i][j]);
 			}
 		}
 		/*int crossSum1[] = new int[(int) Math.pow(newalphas.size(), _obspartitionset.size())];
@@ -395,7 +395,7 @@ public class ContComputeGamma {
 				if (afterInt>0) 
 					afterInt =_context.computeDefiniteIntegral(afterInt,_pomdp._alContSVars.get(j));
 				else afterInt =_context.computeDefiniteIntegral(beforeInt,_pomdp._alContSVars.get(j));
-				afterInt = _context.reduceLP(afterInt, _pomdp._alContAllVars);
+				afterInt = _context.reduceLP(afterInt);
 				//or assign a new array for delta's but no need to keep them
 			}
 			newAlphas.put(i, afterInt);
@@ -417,7 +417,7 @@ public class ContComputeGamma {
 		//4- compute probability of each case partition of max
 		partition = new ObsPartition();
 		partitionNo =0;
-		max = _context.reduceLP(max, _pomdp._alContAllVars);
+		max = _context.reduceLP(max);
 		reduceProbability(max,a,belief);
 	}
 	private int reduceProbability(int node_id,COAction a,int b) {
@@ -449,7 +449,7 @@ public class ContComputeGamma {
 			//}
 				for (int i=0;i<_pomdp._alContSVars.size();i++)
 					check = _context.computeDefiniteIntegral(check, _pomdp._alContSVars.get(i));
-				check = _context.reduceLP(check, _pomdp._alContAllVars);
+				check = _context.reduceLP(check);
 				//multiply indicator
 
 				for (Map.Entry<Decision, Boolean> me : partition.get_decisions().entrySet()) 
@@ -458,12 +458,13 @@ public class ContComputeGamma {
 					double low_val = me.getValue() ? 0d : 1d;
 					check = _context.apply(check,_context.getVarNode(me.getKey(), low_val, high_val), _context.PROD);
 				}
+				check = _context.reduceLP(check);
 				//integrate only this observation
[…]

### 2012-09-28 cpomdp/camdp-xadd/src/CPOMDP/cpomdp.java

@@ -130,7 +130,7 @@ public class cpomdp {
 		_alContOVars = new ArrayList<String>(Intern(parser.getOVars())); // Retain order given in MDP file
 		_alContAllVars = new ArrayList<String>(_alContSVars);
 		_alContAllVars.addAll(_alContOVars);
-		_context._hmContinuousVars = _alContAllVars;
+		//_context._hmContinuousVars = _alContAllVars;
 
 		
 		
@@ -180,7 +180,7 @@ public class cpomdp {
 		//define belief: 
 		//1 belief for now: 
 			ArrayList l0 =new ArrayList();
-				/*l0.add("[-6 + t*1 <= 0]");
+				l0.add("[-6 + t*1 <= 0]");
 				ArrayList l0t = new ArrayList();
 				ArrayList l0f = new ArrayList();
 				l0t.add("[-2 + t*1 >= 0]");
@@ -195,23 +195,23 @@ public class cpomdp {
 				l0.add(l0f);
 				
 				ArrayList l1 =new ArrayList();
-				l1.add("[-13 + t*1 <= 0]");
+				l1.add("[-11 + t*1 <= 0]");
 				ArrayList l1t = new ArrayList();
 				ArrayList l1f = new ArrayList();
-				l1t.add("[-8 + t*1 >= 0]");
+				l1t.add("[-6 + t*1 >= 0]");
 				ArrayList l1tt = new ArrayList();
 				ArrayList l1tf = new ArrayList();
 				l1tt.add("0.2");
 				l1tf.add("0");
-				l1t.add(l0tt);
-				l1t.add(l0tf);
+				l1t.add(l1tt);
+				l1t.add(l1tf);
 				l1f.add("0");
 				l1.add(l1t);
-				l1.add(l1f);*/
+				l1.add(l1f);
 				
 				
 				//for 2d:
-				ArrayList l1 =new ArrayList();
+		/*		ArrayList l1 =new ArrayList();
 				l0.add("[-10 + p*1 <= 0]");
 				ArrayList l0t = new ArrayList();
 				ArrayList l0f = new ArrayList();
@@ -262,7 +262,7 @@ public class cpomdp {
 				l1t.add(l1tf);
 				l1f.add("0");
 				l1.add(l1t);
-				l1.add(l1f);
+				l1.add(l1f);*/
 				
 				int b0 = _context.buildCanonicalXADD(l0);
 				belief[0] = b0;
@@ -341,7 +341,11 @@ public class cpomdp {
[…]

### 2012-09-28 cpomdp/camdp-xadd/src/camdp/ComputeQFunction.java

@@ -1,0 +1,140 @@
+package camdp;
+
+import graph.Graph;
+
+import java.util.*;
+
+import util.IntTriple;
+import xadd.XADD;
+import xadd.XADD.*;
+
+public class ComputeQFunction {
+
+	public XADD _context = null;
+	public CAMDP _camdp  = null;
+
+	public ComputeQFunction(XADD context ,CAMDP camdp)
+	{
+		_context  = context;
+		_camdp = camdp;
+	}
+	
+	private IntTriple _contRegrKey = new IntTriple(-1,-1,-1);
+	
+	/**
+	 * Regress a DD through an action
+	 **/
+	public int regress(int vfun, CAction a) {
+      
+		XADDNode n = _context.getNode(vfun);
+		
+		// What state variables in vfun do we have to regress?  
+		// Note: no action params in vfun.
+		HashSet<String> state_vars_in_vfun  = n.collectVars();
+		//_camdp._logStream.println("** Regressing " + a._sName + "\n- State vars in vfun: " + state_vars_in_vfun);
+		
+		// Prime the value function
+		int q = _context.substitute(vfun, _camdp._hmPrimeSubs); 
+		//_camdp._logStream.println("- Primed value function:\n" + _context.getString(q));
+
+		// Regress continuous variables first in order given in 
+		for (String var : state_vars_in_vfun) {
+			if (!_camdp._hsContSVars.contains(var))
+				continue; // Not regressing boolean variables yet, skip
+
+			// Get cpf for continuous var'
+			String var_prime = var + "'";
+			int var_id = _context.getVarIndex( _context.new BoolDec(var_prime), false);
+			Integer dd_conditional_sub = a._hmVar2DD.get(var_prime);
+
+			//_camdp._logStream.println("- Integrating out: " + var_prime + "/" + var_id + " in\n" + _context.getString(dd_conditional_sub));
+
+			// Check cache
+			_contRegrKey.set(var_id, dd_conditional_sub, q);
+			Integer result = null;
+			if ((result = _camdp._hmContRegrCache.get(_contRegrKey)) != null) {
+				q = result;
+				continue;
+			}
+			
+			// Perform regression via delta function substitution
+			_context._hsSpecialNodes.add(dd_conditional_sub);
+			q = _context.reduceProcessXADDLeaf(dd_conditional_sub, 
+					_context.new DeltaFunctionSubstitution(var_prime, q), true);
+			
+			// Cache result
[…]

### 2012-09-28 cpomdp/camdp-xadd/src/cmdp/CMDP.java

@@ -67,7 +67,7 @@ public class CMDP {
 	public final static boolean REDUCE_LP = true;
 	
 	public final static boolean DISPLAY_Q = false;
-	public final static boolean DISPLAY_V = false;
+	public final static boolean DISPLAY_V = true;
 	public final static boolean DISPLAY_MAX = false;
 	public final static boolean DISPLAY_SUBST = false;
 	public final static boolean PRINTFINALQ = false;
@@ -195,7 +195,7 @@ public class CMDP {
 				int regr = regress(_valueDD, me.getValue());
 				
 				if (REDUCE_LP) {
-					regr = _context.reduceLP(regr, _alCVars);
+					regr = _context.reduceLP(regr);
 				}
 				
 				if (DISPLAY_Q) {
@@ -235,7 +235,7 @@ public class CMDP {
 					_context.apply(_maxDD, regr, XADD.MAX));
 			
 				if (REDUCE_LP) {
-					_maxDD = _context.reduceLP(_maxDD, _alCVars);
+					_maxDD = _context.reduceLP(_maxDD);
 				}
 				
 				if(DISPLAY_MAX){
@@ -261,7 +261,7 @@ public class CMDP {
 //				g1.addNodeColor("_temp_", "lightblue");
 //				g1.launchViewer(1300, 770);
 
-				_maxDD = _context.reduceLP(_maxDD, _alCVars);
+				_maxDD = _context.reduceLP(_maxDD);
 //				Graph g2 = _context.getGraph(_maxDD);
 //				g2.addNode("_temp_");
 //				g2.addNodeLabel("_temp_", "After [" + iter + "]");

### 2012-09-28 cpomdp/camdp-xadd/src/cmdp/Token.java

@@ -1,96 +1,0 @@
-//////////////////////////////////////////////////////////////////////
-//
-// File:     Token.java
-// Project:  MIE457F Information Retrieval System
-// Author:   Scott Sanner, University of Toronto (ssanner@cs.toronto.edu)
-// Date:     9/1/2003
-//
-//////////////////////////////////////////////////////////////////////
-
-// Package definition
-package cmdp;
-
-// Packages to import
-import java.io.*;
-import java.math.*;
-import java.util.*;
-
-/**
- * A generic token structure for TokenStream
- *
- * @version   1.0
- * @author    Scott Sanner
- * @language  Java (JDK 1.3)
- **/
-public class Token
-{
-    // Note: We could add more information to a token such as
-    // its type (e.g., symbol, number, string, etc...) but
-    // this structure suffices for an initial implementation.
-
-    /* Constants */
-    public static final int LPAREN = 1;
-    public static final int RPAREN = 2;
-    public static final int PERIOD = 3;
-    public static final int COMMA  = 4;
-    public static final int SEMI   = 5;
-    public static final int COLON  = 6;
-
-    /** Data members **/
-    public String  _sSourceDoc; // Source location for token
-    public String  _sToken;     // String associated with a token
-    public int     _nPos;       // Character position in file (absolute)
-    public int     _nLine;      // Line number in file
-    public int     _nLinePos;   // Char position on current line
-    public int     _nSymbolID;  // If token null then this is the symbol ID
-    public boolean _bInteger;   // Indicates if the String is numeric
-    public BigInteger _biIntValue;  // Integer value if an integer
-
-    /** Constructor
-     *  @param token    String associated with token
-     *  @param src      Source file/URL for token
-     *  @param pos      Absolute char position in file
-     *  @param line     Absolute line position in a file
-     *  @param line_pos Char position relative to line
-     **/
-    public Token(String token, String src, int pos, 
-		 int line, int line_pos) {
-	_sToken     = token;
-	_sSourceDoc = src;
-	_nPos       = pos;
-	_nLine      = line;
-	_nLinePos   = line_pos;
-	_biIntValue = null;
-
-	if (_sToken.length() == 1) {
-	    char c = _sToken.charAt(0);
-	    switch(c) {
[…]

### 2012-09-28 cpomdp/camdp-xadd/src/logic/kb/prop/PropKbCNF.java

@@ -1,0 +1,1191 @@
+////////////////////////////////////////////////////////////////////////////////////
+//
+// First-Order Algebraic Decision Diagram Package
+//
+// Class:  PropKb - A Kb interface for propositional theorem proving
+//                  (includes a String based interface!)
+// Author: Scott Sanner (ssanner@cs.toronto.edu)
+// Date:   7/25/03
+//
+// TODO:
+// -----
+// - Update for current ADD version
+// - Make an FOL Kb to support different types of FOL Kb creations
+//   and queries (takes both strings and FOL structs as queries
+//   and tell)
+// - NOTE: Will have to handle paramodulation manually
+// - LATER: Eventually should have LALR parser for Prop formula and
+//   build trees directly from this - then use this to interface
+//   with ADDs.  This is currently done but with a simple and
+//   potentially buggy recursive descent parser - nevertheless, a
+//   propositional structure can be retrieved from a String!
+// - Printing for prop formula
+// - ADDs for partition sets
+// - Have PropTerms and ADDs use variable name mapping - just set a HashMap?
+//
+////////////////////////////////////////////////////////////////////////////////////
+
+package logic.kb.prop;
+
+import java.io.*;
+import java.util.*;
+
+//import logic.add.*;
+import logic.kb.*;
+
+import util.MapList;
+import util.WinUNIX;
+
+public class PropKbCNF {
+
+	// Kb status constants
+	public static final int INDETERMINATE = 0;
+	public static final int TAUTOLOGY = 1;
+	public static final int INCONSISTENT = 2;
+
+	// Data members for formula storage and var->id mapping
+	public int _nLiteralIDCount; // id counter for this kb
+	public ArrayList _alLiteralList; // for id, returns PropLiteral
+	public ArrayList _alAssignment; // assignment for given var id
+	public HashMap _hmLiteralToID; // maps literal string -> id
+	public HashSet<HashSet<PropFormula.Term>> _cnfCurrentKb; // could add undo's here later :)
+
+	// Constructor (no params)
+	public PropKbCNF() {
+		_nLiteralIDCount = 1;
+		_alLiteralList = new ArrayList();
[…]

### 2012-09-28 cpomdp/camdp-xadd/src/xadd/SerialClone.java

@@ -1,0 +1,85 @@
+package xadd;
+
+import java.io.ByteArrayInputStream;
+import java.io.ByteArrayOutputStream;
+import java.io.IOException;
+import java.io.InputStream;
+import java.io.InvalidClassException;
+import java.io.ObjectInputStream;
+import java.io.ObjectOutputStream;
+import java.io.ObjectStreamClass;
+import java.io.OutputStream;
+import java.util.LinkedList;
+import java.util.Queue;
+
+public class SerialClone {
+    public static <T> T clone(T x) {
+	try {
+	    return cloneX(x);
+	} catch (IOException e) {
+	    throw new IllegalArgumentException(e);
+	} catch (ClassNotFoundException e) {
+	    throw new IllegalArgumentException(e);
+	}
+    }
+
+    private static <T> T cloneX(T x) throws IOException, ClassNotFoundException {
+	ByteArrayOutputStream bout = new ByteArrayOutputStream();
+	CloneOutput cout = new CloneOutput(bout);
+	cout.writeObject(x);
+	byte[] bytes = bout.toByteArray();
+	
+	ByteArrayInputStream bin = new ByteArrayInputStream(bytes);
+	CloneInput cin = new CloneInput(bin, cout);
+
+	@SuppressWarnings("unchecked")  // thanks to Bas de Bakker for the tip!
+	T clone = (T) cin.readObject();
+	return clone;
+    }
+
+    private static class CloneOutput extends ObjectOutputStream {
+	Queue<Class<?>> classQueue = new LinkedList<Class<?>>();
+
+	CloneOutput(OutputStream out) throws IOException {
+	    super(out);
+	}
+
+	@Override
+	protected void annotateClass(Class<?> c) {
+	    classQueue.add(c);
+	}
+
+	@Override
+	protected void annotateProxyClass(Class<?> c) {
+	    classQueue.add(c);
+	}
+    }
+
+    private static class CloneInput extends ObjectInputStream {
+	private final CloneOutput output;
+
+	CloneInput(InputStream in, CloneOutput output) throws IOException {
+	    super(in);
+	    this.output = output;
+	}
+
+    	@Override
+	protected Class<?> resolveClass(ObjectStreamClass osc)
+	throws IOException, ClassNotFoundException {
+	    Class<?> c = output.classQueue.poll();
+	    String expected = osc.getName();
+	    String found = (c == null) ? null : c.getName();
+	    if (!expected.equals(found)) {
[…]

### 2012-09-28 cpomdp/camdp-xadd/src/xadd/TestXADDDist.java

@@ -199,18 +199,22 @@ public class TestXADDDist {
 
 	public static void Plot3DXADD(XADD context, int xadd, double low_x,
 			double inc_x, double high_x, double low_y, double inc_y,
-			double high_y, String xVar, String yVar, String title) {
+			double high_y, String xVar, String yVar, String title,String filename) {
 
 		Plot3DXADD(context, xadd, low_x, inc_x, high_x, low_y, inc_y, high_y,
 				new HashMap<String, Boolean>(), new HashMap<String, Double>(),
-				xVar, yVar, title);
+				xVar, yVar, title,filename);
 	}
 
 	public static void Plot3DXADD(XADD context, int xadd, double low_x,
 			double inc_x, double high_x, double low_y, double inc_y,
 			double high_y, HashMap<String, Boolean> static_bvars,
 			HashMap<String, Double> static_dvars, String xVar, String yVar,
-			String title) {
+			String title,String filename) {
+		//print to file
+		try {
+ 		counter++;
+        BufferedWriter out = new BufferedWriter(new FileWriter(filename+counter+".txt"));
 
 		// Create a Simple 2D XY plot window.
 		ArrayList<Double> alX = new ArrayList<Double>();
@@ -237,7 +241,9 @@ public class TestXADDDist {
 				static_dvars.remove(yVar);
 
 				if (z > 0.1d)
-					System.out.println("f(" + x + "," + y + ") = " + z);
+				{
+					//System.out.println("f(" + x + "," + y + ") = " + z);
+				}
 				xArr[i][j] = x;
 				yArr[i][j] = y;
 				zArr[i][j] = z; //x + y; //z;
@@ -246,9 +252,14 @@ public class TestXADDDist {
 		for (int i=0;i<alY.size();i++)
 		{
 			for (int j=0;j<alX.size();j++)
-				System.out.print(" " + zArr[i][j]);
-			System.out.println();
+			{
+				//System.out.print(" " + zArr[i][j]);
+         		out.append(" " + zArr[i][j]);
+			}
+			//System.out.println();
+            out.newLine();
 		}
+        out.close();
 		//String title = "f(" + xVar + "," + yVar + ") @ " + static_bvars + " " + static_dvars;
 		Plot2D aPlot = new ContourPlot(xArr, yArr, zArr, 12, false, title, 
 				xVar, yVar, null, null);
@@ -281,6 +292,9 @@ public class TestXADDDist {
[…]

### 2012-09-28 cpomdp/camdp-xadd/src/xadd/XADD.java

@@ -15,6 +15,8 @@ package xadd;
 
 import graph.Graph;
 
+import java.io.PrintStream;
+import java.io.Serializable;
 import java.text.DecimalFormat;
 import java.util.ArrayList;
 import java.util.Arrays;
@@ -26,11 +28,14 @@ import java.util.List;
 import java.util.Map;
 
 import logic.kb.fol.FOPC;
+import logic.kb.prop.PropKbCNF;
 import lpsolve.LP;
 import lpsolve.LpSolve;
 import util.IntPair;
 import util.IntTriple;
+import util.MapList;
 import util.Pair;
+import xadd.XADD.Expr;
 import cmdp.HierarchicalParser;
 
 /**
@@ -40,29 +45,24 @@ public class XADD {
 
 	// Visualization
 	public final static boolean GRAPH_USE_COLOR = true;
-
 	// Flags
-	public final static boolean USE_CANONICAL_NODES = true; // Store nodes in
-															// canonical format?
-	public final static boolean CHECK_MIN_MAX = false; // Will be buggy if
-														// min/max of expr
-	// not at extrema of domain
-
-	public final static boolean USE_MINUS_COMP = false; // Error, all max/min
-														// comps reduce to
-														// false!
+	public final static boolean USE_CANONICAL_NODES = true; // Store nodes in canonical format?
+	public final static boolean CHECK_MIN_MAX = false; // Will be buggy if min/max of expr
+	public final static boolean USE_MINUS_COMP = false; // Error, all max/min comps reduce to false!
 	public final static int MAX_BRANCH_COUNT = 1000000;
+	public final static double FLUSH_PERCENT_MINIMUM = 0.3d; // Won't flush until < amt
+	public static Runtime RUNTIME = Runtime.getRuntime();
 
 	// Debug
 	public final static boolean CHECK_LOCAL_ORDERING = true;
 	public final static boolean SHOW_DECISION_EVAL = false;
 	public final static boolean DEBUG_EVAL_RANGE = false;
 	public final static boolean DEBUG_CONSTRAINTS = false;
-	public final static boolean ANNOTATE_ENABLE = false;
 	public final static boolean HANDLE_NONLINEAR = false;
-	public static boolean REDUCE_SUB = false;
[…]

### 2012-09-28 trunk/cpomdp/camdp-xadd/src/CPOMDP/discreteObs/DiscreteObsGamma.java

@@ -123,7 +123,7 @@ public class DiscreteObsGamma {
 			for (int j=0;j<newalphas.size();j++)
 			{
 				regressedAlpha[i][j] = _context.apply(newalphas.get(j), obs.getValue(), _context.PROD);
-				regressedAlpha[i][j] = _context.reduceLP(regressedAlpha[i][j] , _pomdp._alContSVars);
+				regressedAlpha[i][j] = _context.reduceLP(regressedAlpha[i][j] );
 			}
 			i++;
 		}

### 2012-09-28 trunk/cpomdp/camdp-xadd/src/camdp/CAMDP.java

@@ -1,84 +1,93 @@
-// Package definition
 package camdp;
 
 // Packages to import
 import graph.Graph;
 
-import java.io.FileOutputStream;
-import java.io.IOException;
-import java.io.PrintStream;
-import java.lang.reflect.AnnotatedElement;
+import java.io.*;
+import java.util.*;
+
 import java.math.BigDecimal;
 import java.text.DecimalFormat;
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.HashMap;
-import java.util.Map;
-
-import java_cup.parser;
 
 import util.IntTriple;
 import xadd.TestXADDDist;
 import xadd.XADD;
-import xadd.XADD.ArithExpr;
-import xadd.XADD.XADDNode;
-import xadd.XADD.XADDTNode;
+import xadd.XADD.*;
 import cmdp.HierarchicalParser;
 
 /**
- * Main CAMDP inference class
- * Class for continuous states and actions
- * Exact solution to the value function considering the maximum points of the Q-function defined by the actions
- * Partitions according to both the state and the action
+ * Main Continuous State and Action MDP (CAMDP) dynamic programming solution class
  * 
  * @version 1.0
- * @author Zahra
+ * @author Zahra Zamani
+ * @author Scott Sanner
  * @language Java (JDK 1.5)
+ * 
+ * TODO: Dump dot file
+ * TODO: Reintroduce policy annotation
+ * TODO: Allow next-state dependent rewards
+ * TODO: Allow alternate initialization of V^0 in input file
+ * TODO: Seems to be a 0 always maxed in as minimum value -- see Rover-nonlinear2, always V > 0
+ * TODO: Believe return XADDs from max_y yields 0's where we don't want them 
+ * TODO: For LB < ROOT < UB, might suppress LB < UB constraints
+ * TODO: Allow conditioning reward on next state
  **/
 public class CAMDP {
 
 	/* Constants */
-	public final static boolean REDUCE_LP = true;
-
-	public final static boolean DISPLAY_Q = false;
+	//public final static String RESULTS_DIR = "results"; // Diagnostic output destination
+	
+	public final static boolean DISPLAY_PREMAX_Q = false;
+	public final static boolean DISPLAY_POSTMAX_Q = false;
 	public final static boolean DISPLAY_V = true;
[…]

### 2012-09-28 trunk/cpomdp/camdp-xadd/src/camdp/ComputeVfunction.java

@@ -1,314 +1,0 @@
-package camdp;
-
-import graph.Graph;
-
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.HashMap;
-import java.util.HashSet;
-import java.util.Map;
-
-import util.IntTriple;
-import xadd.XADD;
-import xadd.XADD.ArithExpr;
-import xadd.XADD.XADDLeafMax;
-
-public class ComputeVfunction {
-	XADD xadd= null;
-	CAMDP camdp = null;
-	public final static boolean PRINTFINALQ = false;
-	public static boolean REDUCE_LP = true;
-	public HashMap<IntTriple,Integer> hashReg;
-
-	public ComputeVfunction(XADD context,CAMDP camdp2)
-	{
-		xadd = context;
-		//need camdp for flush-caches mainly
-		camdp = camdp2;
-		hashReg = camdp.get_hmRegrKey2Node();
-		
-	}
-	
-	/**
-	 * Regress a DD through an action
-	 **/
-	public int computeValueFunction(int vfun, CAction a) {
-      
-		XADD.XADDNode n = xadd.getNode(vfun);
-		HashSet<String> vars = n.collectVars();
-		ArrayList<String> cvar_names = new ArrayList<String>();
-		HashMap<String,Integer> bvar_dds = new HashMap<String,Integer>();
-		HashMap<String,ArithExpr> prime_subs = new HashMap<String,ArithExpr>();
-		ArrayList<XADD.XADDNode> node_list = new ArrayList<XADD.XADDNode>();
-		ArrayList<XADD.ArithExpr> subst = new ArrayList<XADD.ArithExpr>();
-		ArrayList<Integer> node_id_list = new ArrayList<Integer>();
-		
-		//Vars contains both state and action parameters. Take action parameters out
-		//Action parameters can be in the decision tree, but not in the marginalization of the sum and integral 
-		for (int i=0;i< a._actionParam.size();i++)
-			for (int j=0;j<vars.size();j++)
-				if (vars.contains(a._actionParam.get(i)))
-					vars.remove(a._actionParam.get(i));
-		
-		//Also add the variables that do not have transitions but are bounded according to the action:
-		for (int i=0;i< a._contState.size();i++)
-			for (int j=0;j<vars.size();j++)
[…]

### 2012-09-28 trunk/cpomdp/camdp-xadd/src/camdp/ParseCAMDP.java

@@ -186,9 +186,9 @@ public class ParseCAMDP {
 					int ind_false = _camdp._context.getINode(var,  T_ONE,  T_ZERO);
 					//int true_half = _context.applyInt(ind_true, reward_toGoal, _context.PROD,-1); // Note: this enforces canonicity so
 					//int reward_d = _context.apply(reward_2,reward_toGoal, _context.SUM,-1);
-					int false_half = _camdp._context.applyInt(ind_false, reward_2, _camdp._context.PROD,-1); // can use applyInt rather than apply
+					int false_half = _camdp._context.applyInt(ind_false, reward_2, _camdp._context.PROD); // can use applyInt rather than apply
 					//reward_dd = _context.applyInt(true_half, false_half, _context.SUM,-1);
-					reward_dd = _camdp._context.applyInt(reward_2, reward_dd, _camdp._context.SUM,-1);
+					reward_dd = _camdp._context.applyInt(reward_2, reward_dd, _camdp._context.SUM);
 				}
 				o=i.next();
 			}
@@ -326,7 +326,7 @@ public class ParseCAMDP {
 		this._maxCVal = _maxCVal;
 	}
 
-	private int buildObstacleXADD(Obstacle obstacle,int _RS) {
+	/*private int buildObstacleXADD(Obstacle obstacle,int _RS) {
 		
 		//hand coded function for x,y,ax,ay 
 				//according to the points, build C and D for the line intersection 
@@ -337,7 +337,7 @@ public class ParseCAMDP {
 				
 				//first build test for checking slope equality of parallel lines: ay = (y2-y1)/(x2-x1)*ax
 				int obstacleXADD=0;
-				/*String p_lhs = "ay";
+				String p_lhs = "ay";
 				ArithExpr arith_p_lhs = ArithExpr.parse(p_lhs);
 				String p_rhs = "(("+obstacle.getYpoint2()+"-"+obstacle.getYpoint1()+")/("+obstacle.getXpoint2()+"-"+obstacle.getXpoint1()+"))*ax";
 				ArithExpr arith_p_rhs = ArithExpr.parse(p_rhs);
@@ -353,7 +353,7 @@ public class ParseCAMDP {
 				//switch the true and false branch
 				int T_ONE = _context.getTermNode(ONE);
[…]

### 2012-09-28 trunk/cpomdp/camdp-xadd/src/cmdp/TokenStream.java

@@ -1,245 +1,0 @@
-//////////////////////////////////////////////////////////////////////
-//
-// File:     TokenStream.java
-// Author:   Scott Sanner, University of Toronto (ssanner@cs.toronto.edu)
-// Date:     9/1/2003
-//
-//////////////////////////////////////////////////////////////////////
-
-// Package definition
-package cmdp;
-
-// Packages to import
-import java.io.*;
-import java.net.*;
-import java.util.*;
-
-/**
- * Tokenizes a generic text document
- * 
- * @version 1.0
- * @author Scott Sanner
- * @language Java (JDK 1.3)
- **/
-public class TokenStream {
-
-	/** Static Constants **/
-	protected static final int EOF = -1;
-
-	/** Non-static Data Members **/
-	protected String _sFilename;
-	protected BufferedReader _brInput;
-	protected int _nLine;
-	protected int _nLinePos;
-	protected int _nPos;
-	protected boolean _bInMath;
-	protected boolean _bPushBack;
-	protected char _cPushBack;
-
-	/**
-	 * Constructor - default
-	 **/
-	public TokenStream() {
-		_sFilename = null;
-		_brInput = null;
-		_nPos = -1; // No chars read yet
-		_nLinePos = -1; // No chars read yet
-		_nLine = 1; // Start at line 1
-		_bInMath = false;
-		_bPushBack = false;
-	}
-
-	/**
-	 * Process a String for tokenizing
-	 * 
-	 * @param content
-	 *            String content from which to load data
-	 **/
-	public void openFromStringContent(String content) throws TokenStreamException {
-
-		// Initialize components
-		_sFilename = "[local string content]";
-		_nPos = _nLinePos = -1; // No chars read yet
-		_nLine = 1; // Start at line 1
-		if (_brInput != null) {
-			close();
-		}
-
-		_brInput = new BufferedReader(new StringReader(content));
-	}
-	
-	/**
-	 * Open a file for tokenizing
-	 * 
-	 * @param filename
-	 *            File from which to load data
-	 **/
-	public void open(String name) throws TokenStreamException {
-
-		// Initialize components
-		_sFilename = name;
-		_nPos = _nLinePos = -1; // No chars read yet
-		_nLine = 1; // Start at line 1
-		if (_brInput != null) {
-			close();
-		}
-
-		// Open the specified file
-		try {
-
-			if (name.toLowerCase().startsWith("http://")) {
-				URL url = new URL(name);
[…]

### 2012-09-28 trunk/cpomdp/camdp-xadd/src/cmdp/TokenStreamException.java

@@ -1,54 +1,0 @@
-//////////////////////////////////////////////////////////////////////
-//
-// File:     TokenStreamException.java
-// Project:  MIE457F Information Retrieval System
-// Author:   Scott Sanner, University of Toronto (ssanner@cs.toronto.edu)
-// Date:     9/1/2003
-//
-//////////////////////////////////////////////////////////////////////
-
-// Package definition
-package cmdp;
-
-// Packages to import
-import java.lang.*;
-
-/**
- * An exception thrown by TokenStream
- *
- * @version   1.0
- * @author    Scott Sanner
- * @language  Java (JDK 1.3)
- **/
-public class TokenStreamException
-    extends Exception {
-
-    /** Local data members **/
-    public String _sSrc;
-    public int    _nPos;
-    public int    _nLine;
-    public int    _nLinePos;
-
-    /** Constructor
-     *  @param error String specifying the error that occurred
-     *  @param src   Source file/URL where error occurred
-     **/
-    public TokenStreamException(String error, String src) {
-	super(error);
-	_sSrc  = src;
-	_nLine = _nLinePos = -1;
-    }
-
-    /** Constructor
-     *  @param error String specifying the error that occurred
-     *  @param src   Source file/URL where error occurred
-     *  @param line  Line number where error occurred
-     *  @param pos   Char position on line where error occurred
-     **/
-    public TokenStreamException(String error, String src, int line, int line_pos) {
-	super(error);
-	_sSrc     = src;
-	_nLine    = line;
-	_nLinePos = line_pos;
-    }
-}

### 2012-09-28 trunk/cpomdp/camdp-xadd/src/util/WinUNIX.java

@@ -21,6 +21,7 @@ public class WinUNIX {
 	public static String GVIZ_CMD_CLOSE = null;
 	public static String GVIZ2_EXE = null;
 	public static String GVIZ2_CMD = null;
+	public static String SAT_EXE = null;
 	public static String GVIZ2_CMD_CLOSE = null;
 	
 	public static String USER_DIR = System.getProperty("user.dir");
@@ -36,21 +37,25 @@ public class WinUNIX {
 		
 		if (SYSTEM == WINDOWS) {
 			
+			System.out.println("WINDOWS");
 			GVIZ_EXE = "dot.exe -Tdot";
 			GVIZ_CMD = "CMD /C dot.exe -Tdot";
 			GVIZ_CMD_CLOSE = "";
 			GVIZ2_EXE = "neato.exe -Tdot";
 			GVIZ2_CMD = "CMD /C neato.exe -Tdot";
 			GVIZ2_CMD_CLOSE = "";
+			SAT_EXE = "minisat.exe";
 
 		} else if (SYSTEM == LINUX_LIKE) {
 			
+			System.out.println("LINUX");
 			GVIZ_EXE = "dot -Tdot";
 			GVIZ_CMD = "/bin/sh 'dot -Tdot";
 			GVIZ_CMD_CLOSE = "'";
 			GVIZ2_EXE = "neato -Tdot";
 			GVIZ2_CMD = "/bin/sh 'neato -Tdot";
 			GVIZ2_CMD_CLOSE = "'";
+			SAT_EXE = "MiniSat_v1.14_linux";
 
 		} else {
 			System.out.println("util.WinUNIX: Unrecognized OS.");

### 2012-09-28 trunk/cpomdp/camdp-xadd/src/xadd/TestXADD.java

@@ -16,7 +16,7 @@ import xadd.XADD.DoubleExpr;
 import xadd.XADD.ExprDec;
 import xadd.XADD.VarExpr;
 import xadd.XADD.XADDINode;
-import xadd.XADD.XADDLeafMax;
+import xadd.XADD.XADDLeafMaxMin;
 import xadd.XADD.XADDTNode;
 import cmdp.HierarchicalParser;
 
@@ -45,12 +45,13 @@ public class TestXADD {
 		XADD xadd_context = new XADD();
 		
 		//////////////////////
-		int ixadd = TestBuild(xadd_context, "./src/xadd/test7.xadd");
+		int ixadd = TestBuild(xadd_context, "./src/xadd/test12indirect.xadd");
 		
 		Graph g1 = xadd_context.getGraph(ixadd);
 		g1.launchViewer();
 		
-		int reduce = xadd_context.reduceLP(ixadd,Arrays.asList("x"));
+		int reduce = xadd_context.reduceLP(ixadd);
+		reduce = xadd_context.reduceLP(reduce,true);
 		g1 = xadd_context.getGraph(reduce);
 		g1.launchViewer();
 		//ixadd  = xadd_context.reduceProcessXADDLeaf(ixadd, max, false);
@@ -227,24 +228,24 @@ public class TestXADD {
 		System.out.println(xadd_context.getString(xadd_circle));
 		System.out.println("-------------\n" + xadd_context.getString(y));
 
-		System.exit(1);
+		//System.exit(1);
 		xadd_context.getGraph(y).launchViewer();
 		System.out.println(xadd_context.getString(xadd_circle));
 
 		// Pause
-		System.in.read();
+		//System.in.read();
 
 		// *****************TESTING MAX***********
 		int xadd4 = TestBuild(xadd_context, "./src/xadd/test4.xadd");
 		int xadd5 = TestBuild(xadd_context, "./src/xadd/test5.xadd");
 		int xaddrRes = xadd_context.apply(xadd4, xadd5, XADD.MAX);
-		Graph gRes = xadd_context.getGraph(xadd_context.reduceLP(xaddrRes, Arrays.asList("x1", "x2", "r1")));
+		Graph gRes = xadd_context.getGraph(xadd_context.reduceLP(xaddrRes));
 		gRes.launchViewer();
 
 		int xadd_implied = TestBuild(xadd_context, "./src/xadd/implied.xadd");
[…]

### 2012-09-28 cpomdp/camdp-xadd/src/doc/NIPS/draft6-10page.tex
Latest NIPS edits

--- 2012-06-01 cpomdp/camdp-xadd/src/doc/NIPS/draft4-10page.tex
@@
+% TODO: 
+%
+% (1) GenRelObs and propagate fixes to text
+%- i's where they shouldn't be, 
+%- integrating over discrete vars, 
+%- P(z_k|...) needs to condition on belief state, 
+%- line 13 generates does not get rid of xd_s'... did you mean to integrate out alpha(xd_o,xd_s) instead?
+% ... \delta_j(x_{d_o}) should integrate out to 1 right?
+%     then use this?
+%
+% (2) Notation for observation set
+%
+% (3) Improve intuitions
+%
+% (4) Revise conclusion
+%
+% (5) More formal claims, where this works, where PBVI exact
+
 \documentclass{article} % For LaTeX2e
@@
 to the case of continuous observations (while restricting the
-transition and reward functions to piecewise polynomials) and extend
+transition to piecewise linear functions and the reward functions to piecewise constants) and extend
 the work of~\cite{pascal_ijcai05} to \emph{derive} relevant observation
@@
 $P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ are represented by
-\emph{piecewise functions} (PWFs) that are first-order Markov
+\emph{piecewise linear functions} (PLFs) that are first-order Markov
 and deterministic, i.e. the probabilities are encoded using the Dirac
 $\delta[\cdot]$ function. The reward function $R(\vec{d},\vec{x},a)$
-is set to be any general piecewise function.  We next provide an intuitive
+is set to a piecewise constant function. We next provide an intuitive
 example to make this DC-POMDP framework more concrete.
@@
 \begin{align}
+\label{eq:trans}
 P(t_s'|\vec{t_s},a)= \delta\left[ t_s' - 
@@
 a discrete binary observation space with possible
-observations $o \in mathcal{O} = \{\high,\low\}$:
+observations $o \in \mathcal{O} = \{\high,\low\}$:
 {\footnotesize
@@
 \begin{cases}
-  t_s <= 15 &: 0.9 \\
- \neg (t_s <= 15)&: 0.1 \\
+  t_s' <= 15 &: 0.9 \\
+ \neg (t_s' <= 15)&: 0.1 \\
 \end{cases}
@@
 \begin{cases}
- t_s <= 15 &: 0.1 \\
- \neg (t_s <= 15) &: 0.9 \\
+ t_s' <= 15 &: 0.1 \\
+ \neg (t_s' <= 15) &: 0.9 \\
 \end{cases}
@@
 We define the observation probability over $t_o$ as uniform on
-an interval 10 units wide, centered at $t_s$.
+an interval 10 units wide, centered at $t_s'$.
 {\footnotesize
@@
 \begin{cases}
- (t_o>t_s-5) \wedge (t_o<t_s+5) &: 0.1 \\
- \neg((t_o>t_s-5) \wedge (t_o<t_s+5)) &: 0 \\
-\end{cases}
-\end{align}
-}
-
+ (t_o>t_s'-5) \wedge (t_o<t_s'+5) &: 0.1 \\
+ \neg((t_o>t_s'-5) \wedge (t_o<t_s'+5)) &: 0 \\
+\end{cases}
+\end{align}
+}
+While the restrictions may seem too tight in our problems due to the continuous nature of the state and observation space, there has been no exact solution to any of the two set of problems above. 
 \section{DC-POMDP solution}
@@
 iteration} (VI)~\cite{monahan82}.  
-Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (2):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as
+Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (2) and considering $r_a = r(x_s,d_s,a)$ :\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as
 $\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
@@
 \begin{align} 
-g^h_{a,o,j} &=  \int_{x_{s'}} \sum_{d_{s'}} p(o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \label{eq:backup} \\
-\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \label{eq:cross_prod} 
-\end{align}
-\begin{align}
+g^h_{a,o,j} &=  \int_{x_{s'}} \sum_{d_{s'}} p(o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^{h-1}_j(xd'_s) ; \hspace{2mm}  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1} \label{eq:backup} \\
+\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j  \label{eq:cross_prod}\\ 
 \Gamma^h  &= \bigcup_a \Gamma^h_a 
@@
 }
-where $r_a = r(x_s,d_s,a)$.
-
+\vspace{-3mm}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
        			{
-   	 		  		$\alpha_j^{h'},:=\,$ \remapWithPrimes{$\alpha_j^{h}$} 
+   	 		  		$\alpha_j^{h'-1},:=\,$ \remapWithPrimes{$\alpha_j^{h-1}$} 
    	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
@@
         	%monahan's pruning first generates all vectors, then prunes
-              $\Gamma^h \,:=\, $\prune{$\Gamma^h$} \;
+              $\Gamma^h \,:=\, $\prune{$\Gamma^h$} \emph{//optional step}\; 
       }
@@
 to guarantee that the integral on line 14 can be done in closed-form
-and we guarantee this for piecewise functions in
-Section~\ref{sec:disc_obs}.  Because in general we will deal with
+and we guarantee this for the piecewise functions in
+Section~\ref{sec:disc_obs}. We also note that the beliefs are piecewise constant functions. Because in general we will deal with
 $\alpha$-functions represented as piecewise functions, we only assume
@@
 closed-form representation.  For SDP, this closed-form will be a
-piecewise case statement and it will stored compactly as a decision
+piecewise case statement and it will be stored compactly as a decision
 diagram~\cite{sanner_uai11}.  We first review this form and then
@@
 \subsection{Case Representation and Extended ADDs}
+\label{sec:case}
+
 % operations, max, restrict, substitute
@@
 that satisfy some formula $\phi$ as defined in \cite{sanner_uai11}.
-\item Substitution $f\sigma$ : Takes a set $\sigma$ of variables and their substitutions, where the LHS of the substitution operator $/$ represents the substitution variable and the RHS of the $/$ represents the expression that should be substituted in its place as in \cite{sanner_uai11}.
-\item Integration $\int_{x_1} f dx_1$:  Eliminates continuous variables by taking the integral of case statements $f$ restricted to linear constraints and polynomial values~\cite{sanner_aaai12}.
+\item Substitution $f\sigma$ : Takes a set $\sigma$ of variables and their substitutions (which may be case statements), and carries out all variable substitutions in sequence~\cite{sanner_uai11}.
+\item Integration $\int_{x_1} f dx_1$:  There are two forms: If $x_1$ is involved in a $\delta$-function (cf.
+the transition in Eq.~\eqref{eq:trans}) then the integral is equivalent to a symbolic substitution and can be applied to any case statement~\cite{sanner_uai11}.
+Otherwise, if $f$ is restricted to linear constraints and polynomial values, then the approach of~\cite{sanner_aaai12} can be applied to yield a result in the same form.
 \end{itemize}
@@
 For discrete and continuous states and discrete observations, we use the symbolic version of Monahan's algorithm as the DC-POMDP solution represented in \texttt{PBVI}.
-The same approach is required for continuous observations only based on point-based dynamic programming backups.
-
-Given an initial set of beliefs, we can compute the relevant observation partitions explained in the next section and then continue in a similar fashion of having discrete observations.
-%To visualize we explain of the value iteration algorithm as defined in \texttt{VI} for DC-POMDPs.
-Each step of the algorithm is explained using the \textsc{\bf power plant} example for the first three horizons.
-Note that if the input to this algorithm is continuous observations $ContObs=true$ then line 8 finds the relevant observation partitions, else discrete observations are assumed and their probabilities are given in the problem specifications such as section 2.
-Also a set of beliefs are given as the input as often several beliefs are required to cover the belief space and thus the algorithm is executed for each one of $b_i \in B$.
-For each iteration, lines 11--12 product the following equation: 
-{\footnotesize
-\vspace{-2mm}
-\begin{align}
-g^h_{a,o,j} =  \int_{x_{s'}} \large \oplus_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \nonumber
-\end{align}
-}
-where the first probability is from equation (2) and the next probability is from equation (1) of the example.
-The $\alpha$-vectors are case functions primed using substitution in \texttt{Prime}.
-\texttt{Backup} performs regression on the $\alpha$-vectors where \textit{Boolean regression} is performing a symbolic restriction and \textit{Continuous regression} is a symbolic substitution, both SDP operations.
-Lines 14--16 cover closed form operations on the $\Gamma$-set such as cross-sum and union.
-Finally lines 18--20 find the maximum $\alpha$-vector within the $\Gamma^h$ set and add this to the final $\Gamma_{PBVI}^h$ if it is not already in the set.
-%Line 13 takes the cross-sum of the two sets, followed by adding the reward to the discounted result.
-For all actions the  $\Gamma$-set is added to the $\Gamma^h$ of the current horizon.
-The result is pruned as explained earlier.%which takes every possible combination of observations according to the $\alpha$-vectors.
-%All vectors are then multiplied by the discount $\gamma$ and added with the reward in line 14 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
-%The same operation is performed for all actions and then in 
-%Line 15 computes the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors for the next horizon. 
-%The $\alpha$-vector that maximizes the value-function with respect to this belief and save it as the new $\Gamma$-set for the next horizon.
-The action from which this $\alpha$-vector was chosen from is the maximum policy.
-%This operation is performed for each of the belief points $b_i \in B$ and for each belief point one dominant $\alpha$-vector is stored in the $\Gamma$-set. The algorithm iterates until convergence. 
-In the next section we discuss the main part of our algorithm which is generating relevant observation partitions.
+In brief we note that all operations as defined in PBVI apply directly to functions represented as case statements.
+For example, we can rewrite
+Eq~\eqref{eq:backup} in the following case operation form:
+{\footnotesize
+\begin{align}
+g^h_{a,o,j} =  \int_{x_{s'}} \large \bigoplus_{d_{s'}} p(o|xd'_s,a) \otimes p(xd'_s|xd_s,a) \otimes \alpha^h_j(xd'_s) \nonumber
+\end{align}
+}
+Crucially we note that since $p(xd'_s|xd_s,a)$ is always defined using $\delta$-functions (possibly conditioned on stochastic discrete variables) for deterministic transitions as defined in our DC-POMDPs, then the integral can always be computed in closed case form as discussed in Section~\ref{sec:case}.
+In short, nothing additional is required for PBVI on DC-POMDPs
+with continuous state and discrete observations --- the crucial insight
+to understand is simply that $\alpha$-vectors are now represented by
+case statements and can ``grow'' with the horizon as they partition the
+state space more and more finely.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 {
-	$z:=0$;\	
+	$Z:=0$ \;	
 		\For {all $\alpha_j$ in  $\Gamma^{h-1}$}    
@@
 		\emph{// Point-based backup operation}\\
-    	$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) \cdot P(xd'_{s_i}| xd_{s_i},a)\cdot \alpha_j $\;
+    	$\alpha_j(xd_s,xd_o) := \int_{x_s'} \bigoplus_{d_s'} P(xd_{o}|xd'_{s},a) \cdot P(xd'_{s}| xd_{s},a)\cdot \alpha_j \hspace{1mm}dx_s'$\;
 		}  
@@
 		\emph{// Generate value of $\alpha$-vector for belief $b_i$ as a function of observations}\\
-		$\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i \cdot \alpha_j(xd_s,xd_o)$\}\ \;
+		$\delta_{j}(xd_o) := $ $\int_{x_{s}} \bigoplus_{d_s} b_i \cdot \alpha_j(xd_s,xd_o)\hspace{1mm}dx_s$\\ \;
 		}
@@
 		{
-		\emph{// Finding partition points on the observation space}\\
-		$z := max(\delta_j(xd_o),z)$\;
-    	}
-    	\emph{// $\phi_{z_k}$ logical term on each partition of $z$}\\
+		\emph{// Generate partition points on the observation space}\\
+		$Z := max(\delta_j(xd_o),z)$\;
+    	}
+    	\emph{// $\phi_{z_k}$ logical term on each partition of $Z$}\\
     	\emph{// Generate probabilities and partitions of observations that distinguish optimal $\alpha$-vector (conditional plan)}\\
-   		$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$ \;
+   		$P(Z=z_k|b_i) := \int_{x_s'} \int_{x_o} \int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}d_{xd_s'}$ \;
     \Return{$P(z)$} \;
@@
 \caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a,b_i$) $\longrightarrow$ $P(z)$ }
+\label{alg:genrelobs}
 \vspace{-1mm}
@@
 
-The main operations required for value iteration in DC-POMDP is generating relevant observations to perform the bellman backup operation. 
-The \texttt{GenRelObs} algorithm defines the steps required to generate these partitions. 
-We use the second iteration of the \textsc{\bf Power Plant} example to demonstrate partitioning of the entire continuous observation space.
-The main assumption is to consider some belief state and obtaining the relevant partitions according to that belief.
-We choose these beliefs such that they cover the entire probable state space; e.g.
-$b_1: U[t_s;2,6]$ and $b_2: U[t_s;6,11]$.
-
-Starting with an empty set of relevant observation partitions $z$, the algorithm first eliminates the next state variable $xd'_s$.
-It multiplies the observation model and the transition model in the set of $\alpha$-vectors of the previous horizon:
-$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $. 
-This operation of line 3 in \texttt{GenRelObs} is similar to that of \texttt{Regress} multiplied by the observation regressed model.
-For continuous variables the continuous regression and for discrete variables boolean restriction is used as defined in the previous section.
-Assuming that after $h=1$ the set of $\alpha$-vectors are defined equal to the reward function of equation (1).
+The additional operation required for \texttt{PBVI} in DC-POMDPs in
+Algorithm~\ref{alg:vi} over the discrete observation case is simply
+\emph{generating relevant observations} as shown on line 10.
+\texttt{GenRelObs} in Algorithm~\ref{alg:genrelobs} defines the steps
+required to generate these partitions.
+
+We use the second iteration of the \textsc{\bf Power Plant} example to
+demonstrate partitioning of the entire continuous observation
+space. The main assumption is to consider some belief state $b_i$ and
+obtaining the relevant partitions according to that belief. For
+example, we can choose two beliefs as uniform distributions,
+e.g. $b_1: U[t_s;2,6]$ and $b_2: U[t_s;6,11]$ as shown in
+Figure~\ref{fig:beliefs} (left).  Starting with an empty set of
+relevant observation partitions $\mathcal{O}^h$, the algorithm first
+eliminates the next state variable $xd'_s$. It regresses the
+the set of
+$\alpha$-vectors of the previous horizon back one step via
+the transition and observation model: $\alpha_j(xd_s,xd_o) :=
+\int_{x_s'} \bigoplus_{x_d'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j\hspace{1mm}dx_s'
+$.  This operation of line 5 in \texttt{GenRelObs} is similar to 
+the point-based backup in~\eqref{eq:backup}. For
+continuous variables the continuous integral and for discrete
+variables the restriction operation is used as defined in
+Section~\ref{sec:case}.  
+Assuming that after $h=1$ the set of $\alpha$-vectors are
+defined equal to the reward function of equation (1).
 %{\footnotesize
@@
 } 
-We need the $\alpha$-vectors to partition on the observation space, thus next we eliminate the current state variable in line 4--5.
-During this integration the current state variable is factored out by multiplying the belief in each of the current $\alpha$-vectors.
-The resulting $\delta$-functions only depend on the observation: $\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$.
-For $b_1: U[t_s;2,6]$ assume the following simple $\delta$-functions:
-%the two vectors above the resulting $\delta$-functions are defined below:
-{\footnotesize
-\vspace{-2mm}
+We now need the $\alpha$-vectors as a function of the observation
+space for a particular belief state, thus next we eliminate the
+current state variable in line 6--8. The resulting $\delta$-functions
+now only depend on the observation: $\delta_{j}(xd_o) := $
+$\int_{x_s'} \bigoplus_{x_d'} b_i * \alpha_j(xd_s,xd_o) \hspace{1mm}dx_s'$.  For $b_2: U[t_s;6,11]$
+%assume the following simple $\delta$-functions:
+the resulting $\delta$-functions are defined below:
+{\footnotesize
+%\vspace{-2mm}
 \begin{align}
@@
 \begin{cases}
- (-\infty < t_o< \infty) &: -0.4*t_o +6.8\\
+ (14 < t_o< 18) &: 0.025*t_o - 0.45\\
+ (8 < t_o< 14) &:  - 0.1\\
+ (4 < t_o< 8) &: - 0.025*t_o -0.1\\
 \end{cases}
@@
 \delta_{2}(t_o) &= \begin{cases}
- (-\infty < t_o< \infty) &: 1.3*t_o +3.4\\
+ (15 < t_o< 18) &: 25*t_o - 450\\
+ (14 < t_o< 15) &: - 2.5*t_o - 37.5\\
+ (8 < t_o< 14) &:  -72.5\\
+ (5 < t_o< 8) &: - 25*t_o + 127.5\\
+ (4 < t_o< 5) &:  2.5*t_o - 10\\
 \end{cases}
@@
 \end{align}
-\vspace{-5mm}
+%\vspace{-5mm}
 }
@@
 \begin{figure*}[tbp!]
-\vspace{-6mm}
+\vspace{-3mm}
 \centering
+\hspace{-20mm}
+\includegraphics[height=40mm]{pics/beliefs.pdf}
+\hspace{-10mm}
+\includegraphics[width=0.33\textwidth]{pics/delta_b1.pdf}
+\hspace{-2mm}
+\includegraphics[width=0.42\textwidth]{pics/delta_b2.pdf}
 \hspace{-17mm}
-\includegraphics[width=0.41\textwidth]{pics/beliefs.pdf}
-\hspace{-12mm}
-\includegraphics[width=0.42\textwidth]{pics/delta1.pdf}
-\hspace{-12mm}
-\includegraphics[width=0.42\textwidth]{pics/delta2.pdf}
-\hspace{-17mm}
-\vspace{-10mm}
+\vspace{-2mm}
 \caption{\footnotesize 
 {\it (left)} Beliefs $b_1,b_2$ for 1D \textsc{\bf Power Plant} example; 
-{\it (middle)} Observation dependent function for $b_1$ that partition the observation space into 2 regions with different probabilities $P(z_i)$ ; 
-{\it (right)} 3 Relevant observation partitions and their probabilities for $b_2$.
-}
-\label{fig:timeSpace}
-\vspace{-4mm}
+{\it (center)} Observation dependent function for $b_2$ that partition the observation space into 5 regions with different probabilities for $P(z_1),P(z_2)$ ; 
+{\it (right)} 3 Relevant observation partitions (which is only one partition) and their probabilities for $b_2$.
+}
+\label{fig:beliefs}
+%\vspace{-4mm}
 \end{figure*}
@@
 In their work, numerical solutions are proposed to find the roots of any two observation dependent function.
-Here we use the symbolic power of the max-operator to find all the maximum boundary regions of two or more $\delta$-functions at the same time.
-For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximum in line 6--7:
+Instead, here we leverage the symbolic power of the max-operator defined in Section~\ref{sec:case} to find all the boundary regions that define for what partitionings of the observation each
+$\delta$-function is optimal. For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximum in lines 9--11: 
 {\footnotesize
@@
 \begin{cases}
- z_1:(t_o<2 : -0.4*t_o +6.8) \\
-z_2:(t_o>2 : 1.3*t_o +3.4)
+z_1: (14 < t_o< 18) &: 0.025*t_o - 0.45\\
+z_1: (8 < t_o< 14) &:  -0.1\\
+z_1: (5.1 < t_o< 8) &: - 0.025*t_o -0.1\\
+z_2: (5 < t_o< 5.1) &: - 25*t_o + 127.5\\
+z_2: (4 < t_o< 5) &:  2.5*t_o - 10\\
 \end{cases}
@@
 }
-The two partitions define observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the two distinct observation partitions is found.
-If we integrate out the state and observation from the observation model and the belief state regardless of a particular $\alpha$-vector, it integrates to one:
-$\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*P(s) = 1$
-Thus we only need to multiply the indicators of each case partition as as in line 9 in this formula to obtain the probability of each partition such that it sums to one: 
-$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$
-For our example the probabilities are defined as $P(z_1)=0.2$ and $P(z_2)=0.8$.
+Note here that each partitioned is labeled according to the original $\delta$-function where $z_1$ is the label of the partition coming from $\delta_{1}(t_o)$. 
+These partitions define observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the two distinct observation partitions is found.
+This is demonstrated visually
+in Figure~\ref{fig:beliefs} (center) for $b_2$ and for $b_1$
+in Figure~\ref{fig:beliefs} (right). Note here that the maximum of the $\delta$-functions for $b_1$ is equal to $\delta_{2}(t_o)$.
+
+Now with the observation partitions derived, all that remains is to
+calculate probabilities for these relevant observations conditioned on
+the belief state. 
+% This does not check out!
+% 
+%As a reality check, we note that if we integrate out the
+%state and observation from the observation model and the belief state
+%regardless of a particular $\alpha$-vector, it integrates to one:
+%$\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*P(s) = 1$.
+Thus we only need to multiply the indicators of each observation partition 
+in this formula to obtain the probability mass lying in each
+partition:
+$$P(z_k|b_i) := \int_{xd_s'}\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}d_{xd_s'}$$
+For our example the probabilities can be calculated as below:
+{\footnotesize
+\vspace{-2mm}
+\begin{align}
+\mathrm{max} \Bigg(\delta_{1}(t_o),\delta_{2}(t_o)\Bigg) &= 
+\begin{cases}
+(\phi_1 : 14 < t_o< 18) &: 0.2\\
+(\phi_2 : 8 < t_o< 14) &:  0.6\\
+(\phi_3 : 5.1 < t_o< 8) &: 0.183\\
+(\phi_4 : 5 < t_o< 5.1) &: 0.002\\
+(\phi_5 : 4 < t_o< 5) &:  0.0125\\
+\end{cases}
+\nonumber
+\end{align}
+}
+where each constraint is defined as $\phi_i$.
+For any number of actions and observations the total number of observation partitions depends on the number of belief points.
+In our example we must have two observation probabilities which is defined as the sum of the probabilities for $z_1,z_2$ according to the constraints: \\
+\begin{align*}
+z_1&: \phi_1 \vee \phi_2 \vee \phi_3 \longrightarrow P(z_1|b_2)=0.983 \\
+z_2&: \phi_4 \vee \phi_5  \longrightarrow P(z_2|b_2)=0.0127
+\end{align*}
+
 %{\footnotesize
@@
 
-Hence we can now use the algorithms and methods of a discrete observation setting using the probabilities of the partitioned observation space in Monahan's algorithm. 
-%Figure ~\ref{fig:obsPartition} shows the observation partitions obtained in the 1D problem instance for different horizons and their related probabilities. In a lyx file, needed?
-Next we present some results for 2-Dimensional continuous observation spaces.
+Hence we can now use the algorithms and methods of a discrete observation setting using the probabilities of the partitioned observation space in PBVI! 
+We note here that our method applies to a restricted class of piecewise functions which is piecewise linear transitions and piecewise constant reward and belief.
+The main reason is that the integration operation over continuous states and observations only allows constant or linear constraints (upper or lower bounds) over these variables \cite{sanner_aaai12}.
+Thus although in theory we can apply this approach to any piecewise polynomial function, in practice it is limited by the integration bounds.
+%Figure ~\ref{fig:obsPartition} shows the observation partitions obtained in the 1D problem instance for different horizons and their related probabilities.
+In a lyx file, needed?
+Next we present some results for 2-dimensional continuous observation spaces.
 
@@
 \section{Empirical Results}
-%We evaluated our continuous POMDP solution using XADDs on the one-dimensional
-%\textsc{\bf Power Plant} example and another variant of this problem with two variables. \footnote{
-%Full problem specifications and Java code to reproduce these experiments are currently 
-%in Google Code; anonymity restrictions prevent us from providing this link in the submitted paper version.}
-The 2-dimensional model description is provided below:
+We evaluated our continuous POMDP solution using XADDs on the
+\textsc{\bf 1D-Power Plant} example and another variant of this
+problem with two variables, described below.\footnote{ Full problem
+specifications and Java code to reproduce these experiments are available online in
+Google Code: \textit{http://code.google.com/p/cpomdp} .}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure*}[tbp!]
-\vspace{-2mm}
+\vspace{2mm}
 \centering
-\includegraphics[width=0.3\textwidth]{pics/time2.pdf}
-\includegraphics[width=0.3\textwidth]{pics/nodes2.pdf}
-\includegraphics[width=0.3\textwidth]{pics/alpha-vectors2.pdf}
-\vspace{-3mm}
+\includegraphics[width=0.32\textwidth]{pics/time2.pdf}
+\includegraphics[width=0.32\textwidth]{pics/nodes2.pdf}
+\includegraphics[width=0.32\textwidth]{pics/alpha-vectors2.pdf}
+\vspace{-1mm}
 \caption{\footnotesize 
 {\it (left)} Space vs Horizon; 
-{\it (middle)} Time vs Horizon; 
+{\it (center)} Time vs Horizon; 
 {\it (right)} Number of $\alpha$-vectors vs Horizon.
@@
 \label{fig:timeSpace}
-\vspace{-4mm}
+%\vspace{-4mm}
 \end{figure*}
@@
 {\footnotesize
-\vspace{-4mm}
+\vspace{-3mm}
 \begin{align}
@@
 In the fifth iteration, the value function has partitioned into more pieces, showing how higher temperatures can increase the value without considering the effect of the pressure.
-In the last plot, horizon 6 has better tuned the value function so that higher temperatures and pressures increase the value of the maximum $\alpha$-vector and also within the reward range, finer grain partitions have been formed.
+In the last plot, horizon $h=6$ has better tuned the value function so that higher temperatures and pressures increase the value of the maximum $\alpha$-vector and also within the reward range, finer grain partitions have been formed.
 %Also note that the policy depends on the initial belief and for these plots they were based on $b_1$ which was lower than the reward range.
@@
 \begin{figure*}[tbp!]
-\vspace{-2mm}
+%\vspace{-2mm}
 \centering
@@
 {\it (left)}Maximum $\alpha$-vector for $b_1$ in first iteration; 
-{\it (middle)} $\alpha_{max}^5(b_1)$; 
+{\it (center)} $\alpha_{max}^5(b_1)$; 
 {\it (right)} $\alpha_{max}^5(b_1)$.%V^6(p,t)
@@
 \label{fig:3D}
-\vspace{-4mm}
+%\vspace{-4mm}
 \end{figure*}
@@
 %%\section{Conclusion}
-We presented the first exact solution to continuous states and observations in a DC-POMDP framework.
-Transitions and reward functions are defined as piecewise linear functions of the continuous states.
-Continuous observations require definition of explicit partitions based on the observation model and the belief state of the agent.
-The key contribution is to define the partitions on the observation space and derive the related probabilities using symbolic integration and maximization.
-We extend the application to more than one state-observation setting and present the results.
+We presented the first exact symbolic operations for \texttt{PBVI} in
+expressive DC-POMDPs with continuous state \emph{and} observations.
+Unlike related work that has extended to the continuous state and
+observation setting~\cite{Perseus_cont}, we do not approach the
+problem by sampling.  Rather, following~\cite{pascal_ijcai05}, the key
+contribution of this work was to define a discrete set of observation
+partitions on the multivariate continuous observation space via
+symbolic maximization techniques and derive the related probabilities
+using symbolic integration.  An important avenue for future work is to
+determine whether similar techniques can be applied to the difficult
+case of continuous state, observation, \emph{and} action DC-POMDPs.
 

### 2012-10-07 cpomdp/camdp-xadd/doc/NIPS/draft4-10page.tex
Moved NIPS 2012 latex source

--- 2012-09-28 cpomdp/camdp-xadd/src/doc/NIPS/draft6-10page.tex
@@
 to the case of continuous observations (while restricting the
-transition to piecewise linear functions and the reward functions to piecewise constants) and extend
+transition and reward functions to piecewise polynomials) and extend
 the work of~\cite{pascal_ijcai05} to \emph{derive} relevant observation
@@
 $P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ are represented by
-\emph{piecewise linear functions} (PLFs) that are first-order Markov
+\emph{piecewise functions} (PWFs) that are first-order Markov
 and deterministic, i.e. the probabilities are encoded using the Dirac
 $\delta[\cdot]$ function. The reward function $R(\vec{d},\vec{x},a)$
-is set to a piecewise constant function. We next provide an intuitive
+is set to be any general piecewise function.  We next provide an intuitive
 example to make this DC-POMDP framework more concrete.
@@
 }
-While the restrictions may seem too tight in our problems due to the continuous nature of the state and observation space, there has been no exact solution to any of the two set of problems above. 
+
 \section{DC-POMDP solution}
@@
 iteration} (VI)~\cite{monahan82}.  
-Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (2) and considering $r_a = r(x_s,d_s,a)$ :\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as
+Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (2):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as
 $\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
@@
 \begin{align} 
-g^h_{a,o,j} &=  \int_{x_{s'}} \sum_{d_{s'}} p(o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^{h-1}_j(xd'_s) ; \hspace{2mm}  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1} \label{eq:backup} \\
-\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j  \label{eq:cross_prod}\\ 
+g^h_{a,o,j} &=  \int_{x_{s'}} \sum_{d_{s'}} p(o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \label{eq:backup} \\
+\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \label{eq:cross_prod} 
+\end{align}
+\begin{align}
 \Gamma^h  &= \bigcup_a \Gamma^h_a 
@@
 \vspace{-3mm}
+where $r_a = r(x_s,d_s,a)$.
+\vspace{2mm}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
        			{
-   	 		  		$\alpha_j^{h'-1},:=\,$ \remapWithPrimes{$\alpha_j^{h-1}$} 
+   	 		  		$\alpha_j^{h'},:=\,$ \remapWithPrimes{$\alpha_j^{h}$} 
    	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
@@
         	%monahan's pruning first generates all vectors, then prunes
-              $\Gamma^h \,:=\, $\prune{$\Gamma^h$} \emph{//optional step}\; 
+              $\Gamma^h \,:=\, $\prune{$\Gamma^h$} \;
       }
@@
 to guarantee that the integral on line 14 can be done in closed-form
-and we guarantee this for the piecewise functions in
-Section~\ref{sec:disc_obs}. We also note that the beliefs are piecewise constant functions. Because in general we will deal with
+and we guarantee this for piecewise functions in
+Section~\ref{sec:disc_obs}.  Because in general we will deal with
 $\alpha$-functions represented as piecewise functions, we only assume
@@
 state space more and more finely.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 {
-	$Z:=0$ \;	
+	$Z:=0$;\	
 		\For {all $\alpha_j$ in  $\Gamma^{h-1}$}    
@@
 		\emph{// Point-based backup operation}\\
-    	$\alpha_j(xd_s,xd_o) := \int_{x_s'} \bigoplus_{d_s'} P(xd_{o}|xd'_{s},a) \cdot P(xd'_{s}| xd_{s},a)\cdot \alpha_j \hspace{1mm}dx_s'$\;
+    	$\alpha_j(xd_s,xd_o) := \int_{x_s'} \bigoplus_{d_s'} P(xd_{o}|xd'_{s},a) \cdot P(xd'_{s}| xd_{s},a)\cdot \alpha_j $\;
 		}  
@@
 		\emph{// Generate value of $\alpha$-vector for belief $b_i$ as a function of observations}\\
-		$\delta_{j}(xd_o) := $ $\int_{x_{s}} \bigoplus_{d_s} b_i \cdot \alpha_j(xd_s,xd_o)\hspace{1mm}dx_s$\\ \;
+		$\delta_{j}(xd_o) := $ $\int_{x_{s}} \bigoplus_{d_s} b_i \cdot \alpha_j(xd_s,xd_o)$\}\ \;
 		}
@@
     	\emph{// Generate probabilities and partitions of observations that distinguish optimal $\alpha$-vector (conditional plan)}\\
-   		$P(Z=z_k|b_i) := \int_{x_s'} \int_{x_o} \int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}d_{xd_s'}$ \;
+   		$P(Z=z_k|b_i) := \int_{x_o} \int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$ \;
     \Return{$P(z)$} \;
@@
 the transition and observation model: $\alpha_j(xd_s,xd_o) :=
-\int_{x_s'} \bigoplus_{x_d'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j\hspace{1mm}dx_s'
-$.  This operation of line 5 in \texttt{GenRelObs} is similar to 
+\int_{x_s'} \bigoplus_{x_d'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j
+$.  This operation of line 3 in \texttt{GenRelObs} is similar to 
 the point-based backup in~\eqref{eq:backup}. For
@@
 space for a particular belief state, thus next we eliminate the
-current state variable in line 6--8. The resulting $\delta$-functions
+current state variable in line 4--5. The resulting $\delta$-functions
 now only depend on the observation: $\delta_{j}(xd_o) := $
-$\int_{x_s'} \bigoplus_{x_d'} b_i * \alpha_j(xd_s,xd_o) \hspace{1mm}dx_s'$.  For $b_2: U[t_s;6,11]$
-%assume the following simple $\delta$-functions:
-the resulting $\delta$-functions are defined below:
+$\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$.  For $b_1: U[t_s;2,6]$
+assume the following simple $\delta$-functions:
+%the two vectors above the resulting $\delta$-functions are defined below:
 {\footnotesize
@@
 \begin{cases}
- (14 < t_o< 18) &: 0.025*t_o - 0.45\\
- (8 < t_o< 14) &:  - 0.1\\
- (4 < t_o< 8) &: - 0.025*t_o -0.1\\
+ (-\infty < t_o< \infty) &: -0.4*t_o +6.8\\
 \end{cases}
@@
 \delta_{2}(t_o) &= \begin{cases}
- (15 < t_o< 18) &: 25*t_o - 450\\
- (14 < t_o< 15) &: - 2.5*t_o - 37.5\\
- (8 < t_o< 14) &:  -72.5\\
- (5 < t_o< 8) &: - 25*t_o + 127.5\\
- (4 < t_o< 5) &:  2.5*t_o - 10\\
+ (-\infty < t_o< \infty) &: 1.3*t_o +3.4\\
 \end{cases}
@@
 \begin{figure*}[tbp!]
-\vspace{-3mm}
+%\vspace{-3mm}
 \centering
-\hspace{-20mm}
-\includegraphics[height=40mm]{pics/beliefs.pdf}
-\hspace{-10mm}
-\includegraphics[width=0.33\textwidth]{pics/delta_b1.pdf}
-\hspace{-2mm}
-\includegraphics[width=0.42\textwidth]{pics/delta_b2.pdf}
 \hspace{-17mm}
-\vspace{-2mm}
+\includegraphics[width=0.41\textwidth]{pics/beliefs.pdf}
+\hspace{-12mm}
+\includegraphics[width=0.42\textwidth]{pics/delta1.pdf}
+\hspace{-12mm}
+\includegraphics[width=0.42\textwidth]{pics/delta2.pdf}
+\hspace{-17mm}
+\vspace{-10mm}
 \caption{\footnotesize 
 {\it (left)} Beliefs $b_1,b_2$ for 1D \textsc{\bf Power Plant} example; 
-{\it (center)} Observation dependent function for $b_2$ that partition the observation space into 5 regions with different probabilities for $P(z_1),P(z_2)$ ; 
-{\it (right)} 3 Relevant observation partitions (which is only one partition) and their probabilities for $b_2$.
+{\it (center)} Observation dependent function for $b_1$ that partition the observation space into 2 regions with different probabilities $P(z_i)$ ; 
+{\it (right)} 3 Relevant observation partitions and their probabilities for $b_2$.
 }
@@
 Instead, here we leverage the symbolic power of the max-operator defined in Section~\ref{sec:case} to find all the boundary regions that define for what partitionings of the observation each
-$\delta$-function is optimal. For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximum in lines 9--11: 
+$\delta$-function is optimal. For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximum in line 6--7: 
 {\footnotesize
@@
 \begin{cases}
-z_1: (14 < t_o< 18) &: 0.025*t_o - 0.45\\
-z_1: (8 < t_o< 14) &:  -0.1\\
-z_1: (5.1 < t_o< 8) &: - 0.025*t_o -0.1\\
-z_2: (5 < t_o< 5.1) &: - 25*t_o + 127.5\\
-z_2: (4 < t_o< 5) &:  2.5*t_o - 10\\
+ z_1:(t_o<2 : -0.4*t_o +6.8) \\
+z_2:(t_o>2 : 1.3*t_o +3.4)
 \end{cases}
@@
 }
-Note here that each partitioned is labeled according to the original $\delta$-function where $z_1$ is the label of the partition coming from $\delta_{1}(t_o)$. 
-These partitions define observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the two distinct observation partitions is found.
+The two partitions define observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the two distinct observation partitions is found.
 This is demonstrated visually
-in Figure~\ref{fig:beliefs} (center) for $b_2$ and for $b_1$
-in Figure~\ref{fig:beliefs} (right). Note here that the maximum of the $\delta$-functions for $b_1$ is equal to $\delta_{2}(t_o)$.
+in Figure~\ref{fig:beliefs} (center) for this example and for another
+example (and belief state) in Figure~\ref{fig:beliefs} (right).
 
@@
 partition:
-$$P(z_k|b_i) := \int_{xd_s'}\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}d_{xd_s'}$$
-For our example the probabilities can be calculated as below:
-{\footnotesize
-\vspace{-2mm}
-\begin{align}
-\mathrm{max} \Bigg(\delta_{1}(t_o),\delta_{2}(t_o)\Bigg) &= 
-\begin{cases}
-(\phi_1 : 14 < t_o< 18) &: 0.2\\
-(\phi_2 : 8 < t_o< 14) &:  0.6\\
-(\phi_3 : 5.1 < t_o< 8) &: 0.183\\
-(\phi_4 : 5 < t_o< 5.1) &: 0.002\\
-(\phi_5 : 4 < t_o< 5) &:  0.0125\\
-\end{cases}
-\nonumber
-\end{align}
-}
-where each constraint is defined as $\phi_i$.
-For any number of actions and observations the total number of observation partitions depends on the number of belief points.
-In our example we must have two observation probabilities which is defined as the sum of the probabilities for $z_1,z_2$ according to the constraints: \\
-\begin{align*}
-z_1&: \phi_1 \vee \phi_2 \vee \phi_3 \longrightarrow P(z_1|b_2)=0.983 \\
-z_2&: \phi_4 \vee \phi_5  \longrightarrow P(z_2|b_2)=0.0127
-\end{align*}
-
+$$P(z_k|b_i) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$$
+For our example the probabilities can be calculated as $P(z_1|b_1)=0.2$ and $P(z_2|b_1)=0.8$.
 %{\footnotesize
@@
 Hence we can now use the algorithms and methods of a discrete observation setting using the probabilities of the partitioned observation space in PBVI!
-We note here that our method applies to a restricted class of piecewise functions which is piecewise linear transitions and piecewise constant reward and belief.
-The main reason is that the integration operation over continuous states and observations only allows constant or linear constraints (upper or lower bounds) over these variables \cite{sanner_aaai12}.
-Thus although in theory we can apply this approach to any piecewise polynomial function, in practice it is limited by the integration bounds.
-%Figure ~\ref{fig:obsPartition} shows the observation partitions obtained in the 1D problem instance for different horizons and their related probabilities.
-In a lyx file, needed?
+%Figure ~\ref{fig:obsPartition} shows the observation partitions obtained in the 1D problem instance for different horizons and their related probabilities. In a lyx file, needed?
 Next we present some results for 2-dimensional continuous observation spaces.
@@
 problem with two variables, described below.\footnote{ Full problem
-specifications and Java code to reproduce these experiments are available online in
-Google Code: \textit{http://code.google.com/p/cpomdp} .}
+specifications and Java code to reproduce these experiments are in
+Google Code (link suppressed) and also in a fully
+anonymized supplementary archive with instructions, attached to this 
+submission.}
 

### 2012-10-07 cpomdp/camdp-xadd/doc/draft6-after first review3.tex
Moved NIPS 2012 latex source

--- 2012-03-14 cpomdp/camdp-xadd/src/doc/draft6-after first review3.tex
@@
    	 		  		$\alpha_j^{h'}=$ \remapWithPrimes{$\alpha_j^{h}$} \;
-   	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
-   	 		    	$\Gamma_{a,o_i,j}^h \,:=\, o_i \otimes$ \backup{$\alpha_j^{h'},a$}\;
+   	 		  		\emph{// $\forall$ $d_{s_i} \to d'_{s_i}$ and $\forall x_{s_i} \to x'_{s_i}$} \; 
+   	 		    	$\Gamma_{a,o_i,j}^h \,:=\, P(o_i|xd'_s) \otimes$ \backup{$\alpha_j^{h'},a$}\;
        	      	}
@@
 For each observation, two vectors are formed for each corresponding $\alpha$-vector.
-For example line 12 for observations $O_2,O_3$ and $\alpha_1^1,\alpha_2^1$ is defined as:
+For example line 12 for observations $O_1,O_2,O_3$ and $\alpha_1^1,\alpha_2^1$ is defined as:
 
 \begin{align*}
-%\begin{cases}
-%\Gamma_{close,O_1,1}^2= 
-%\begin{cases}
-%(200<t_o<300) \wedge (t_s'>235) &: -900 \\
-%(200<t_o<300)\wedge (200<t_s'<235) &: 90 \\
-%(200<t_o<300) \wedge (t_s'<200) &: -10 \\
-%\end{cases}
-%\\
-%\Gamma_{close,O_1,2}^2= 
-%\begin{cases}
-%(200<t_o<300) \wedge (t_s'>200) &: -0.9 \\
-%(200<t_o<300) \wedge (t_s'<200) &: -0.1 \\
-%\end{cases}
-%\end{cases}
-%\\
+\begin{cases}
+\Gamma_{close,O_1,1}^2= 
+\begin{cases}
+(t_s'>235) &: -900 \\
+(200<t_s'<235) &: 90 \\
+(t_s'<200) &: -10 \\
+\end{cases}
+\\
+\Gamma_{close,O_1,2}^2= 
+\begin{cases}
+(t_s'>200) &: -0.9 \\ 
+(t_s'<200) &: -0.1 \\
+\end{cases}
+\end{cases}
+\\
 \begin{cases}
@@
 \begin{cases}
-(t_o>300) \wedge (t_s'>235) &: -100 \\
-(t_o>300)\wedge (200<t_s'<235) &: 10 \\
-(t_o>300) \wedge (t_s'<200) &: 0 \\
+ (t_s'>235) &: -100 \\
+(200<t_s'<235) &: 10 \\
+(t_s'<200) &: 0 \\
 \end{cases}
@@
 \begin{cases}
-(t_o>300)\wedge (t_s'>200) &: -0.1 \\
-(t_o>300) \wedge (t_s'<200) &: 0 \\
+ (t_s'>200) &: -0.1 \\
+ (t_s'<200) &: 0 \\
 \end{cases}
@@
 \begin{cases}
-(t_o<200) \wedge (t_s'>235) &: -900 \\
-(t_o<200)\wedge (200<t_s'<235) &: 90 \\
-(t_o<200) \wedge (t_s'<200) &: 0 \\
+ (t_s'>235) &: -900 \\
+ (200<t_s'<235) &: 90 \\
+ (t_s'<200) &: 0 \\
 \end{cases}
@@
 \begin{cases}
-(t_o<200) \wedge (t_s'>200) &: -0.9 \\
-(t_o<200)\wedge (t_s'<200) &: 0 \\
-\end{cases}
-\end{cases}
-\end{align*}
-
-Next in line 13 we have to take the cross-sum of the two sets which takes every possible combination of observations according to the $\alpha$-vectors.
-This means for $O_2 \textrm{\large $\boxplus$} O_3$ we have the following vectors:
-
-\begin{align*}
-\begin{cases}
-\begin{cases}
-(t_s'>235) &: -1000 \\
-(200<t_s'<235) &: 100 \\
-(t_s'<200) &: 0 \\
-\end{cases}
-\\
-\begin{cases}
-(t_s'>235) &: -100.9 \\
-(<200<t_s'<235) &: 9.1 \\
-(t_s'<200) &: 0 \\
-\end{cases}
-\\
-\begin{cases}
- (t_s'>235) &: -900 \\
- (200<t_s'<235) &: 89.9 \\
- (t_s'<200) &: 0 \\
-\end{cases}
-\\
-\begin{cases}
-(t_s'>200) &: -1 \\
+ (t_s'>200) &: -0.9 \\
 (t_s'<200) &: 0 \\
@@
 
-Where the observation partitions ($t_o>300 , t_o<200$) are inconsistent and therefore omitted from the result.
-These 4 vectors are added with the two vectors from $\Gamma_{close,O_1}^2$ to get $2^3$ different vectors.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=0.4\textwidth]{policy2.pdf}
+\end{center}
+\vspace{-3mm}
+\caption{%\footnotesize 
+An example of a policy tree in \textsc{Power Plant}
+for the $close$ action, $O_1,O_3$ based on $\alpha_1^1$ and $O_2$ based on $\alpha_2^1$.} 
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Next in line 13 we have to take the cross-sum of the two sets which takes every possible combination of observations according to the $\alpha$-vectors.
+This means for $O_1 \textrm{\large $\boxplus$} O_2 \textrm{\large $\boxplus$} O_3$ we have 8 vectors where the first 4 combination  have been shown below:
+
+\begin{align*}
+\begin{cases}
+\begin{cases}
+(t_s'>235) &: -1900 \\
+(200<t_s'<235) &: 190 \\
+(t_s'<200) &: -10 \\
+\end{cases}
+\\
+\begin{cases}
+(t_s'>235) &: -1000.9 \\
+(<200<t_s'<235) &: 99.1 \\
+(t_s'<200) &: -10 \\
+\end{cases}
+\\
+\begin{cases}
+ (t_s'>235) &: -1800.1 \\
+ (200<t_s'<235) &: 179.9 \\
+ (t_s'<200) &: -10 \\
+\end{cases}
+\\
+\begin{cases}
+ (t_s'>235) &: -901 \\
+ (200<t_s'<235) &: 89 \\
+ (t_s'<200) &: -10 \\
+\end{cases}
+\end{cases}
+\end{align*}
+
+We have shown the policy tree for the third vector in figure 3.
+The weights here are distributed according to each of the observations while the first and last partition choose to take a backup from the $\alpha_1^1$ and the middle partition uses $\alpha_2^1$.
 All 8 vectors are multiplied by the discount $\gamma$ and added with the reward in line 14 where the addition is an augmented $\oplus$ for adding each vector with the reward function.
-% need to add this?
+
+\begin{align*}
+\begin{cases}
+\begin{cases}
+(t_s'>235) &: -2900 \\
+(200<t_s'<235) &: -810 \\
+(t_s'<200) &: 90 \\
+\end{cases}
+\\
+\begin{cases}
+(t_s'>235) &: -2000.9 \\
+(<200<t_s'<235) &: -900.9 \\
+(t_s'<200) &: 90 \\
+\end{cases}
+\\
+\begin{cases}
+ (t_s'>235) &: -2800.1 \\
+ (200<t_s'<235) &: -820.1 \\
+ (t_s'<200) &: 90 \\
+\end{cases}
+\\
+\begin{cases}
+ (t_s'>235) &: -1901 \\
+ (200<t_s'<235) &: -911 \\
+ (t_s'<200) &: 90 \\
+\end{cases}
+\end{cases}
+\end{align*}
+
 The same operation is performed for the $open$ action and 8 different vectors are formed here too.

### 2012-10-07 cpomdp/camdp-xadd/doc/MISC/draft1.tex
Reorganizing NIPS12 files

--- 2012-05-18 trunk/cpomdp/camdp-xadd/src/doc/NIPS/draft2-long version.tex
@@
 % Write intro here
-In many complex systems complete information of the state is not available due to noisy or poor sensors and readings of them.
-Partially observable Markov Decision Processes (POMDPs)\cite{POMDPfirst} can be used for planning under such uncertainty in abstract problems.
-In many real-world problems, resources and sensor inputs are naturally modeled using continuous variables.
 
-Unfortunately most POMDP techniques can not be applied to such settings without assuming prior knowledge about the problem domain.
-Most methods use some approximation to discretize the continuous state, observation or action space which increases computational complexity \cite{Thrun99h}.
-Exact solutions to a multi-variate continuous state and observation setting has not been approached for the curse of dimensionality problem.
-In assuming a continuous observation setting, the agent must search for all possible observations that are read from the environment.
-This can quickly become infeasible even for small problem sizes.
-
-In this paper we provide an exact solution to a subset of POMDPs with continuous states and observations.
-Our symbolic dynamic programming (SDP) approach assumes discrete noise and piecewise linear dynamics and functions.
-We formulize our problems using a Discrete and Continuous POMDP (DC-POMDP) model and apply our SDP algorithm.
-The main problem that rises in any problem modeled using DC-POMDP is the definition of different observation partitions and their probability.
-We apply symbolic solutions such as integration and maximization to provide an optimal closed-form solution to our problem domains.
-As an example we consider the following power plant example which has mainly been solved in the literature using discrete sets of states and observations~\cite{steam2}.
-We use the example throughout the paper to help better understand the definitions.
+The following example is used throughout the paper to help better understand the definitions.
 
@@
 
-We need to extend the basic SDP framework from \cite{sanner_uai11} by adding continuous observations to the model.
-The solution to a closed-form value function is similar to the algorithm from \cite{monahan82} apart from the fact that no explicit observation is defined in the continuous case.
-We redefine the value-iteration algorithm using case representations that are implemented using Extended Algebric decision diagrams (XADD) and provide an efficient automatic way to partition the entire observation space into piecewise linear functions where each partition has a relative probability.
-We use the univariate \textsc{Power Plant} example to define the main operations and algorithms and also provide empirical results for a multi-variate version of the problem.
+Most work in the literature have assumed discretization of the state and observation space ~\cite{steam2}. Here we take a continuous approach using the DC-POMDP framework.
 
@@
 \includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
-\includegraphics[width=0.45\textwidth]{pics/b1-4.pdf}
+\includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
 \end{center}
@@
 Calculating all possible ways of building the value function independent of a certain belief state can define the optimal value function.
-To derive the $\Gamma$-set of all possible $\alpha$-vectors for each horizon, we use the value iteration algorithm from ~\cite{monahan82}.
+To derive the $\Gamma$-set of all possible $\alpha$-vectors for each horizon, we use the value iteration algorithm from (Monahan)~\cite{monahan82}.
 Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (4):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as
@@
 Hence we can now use the algorithms and methods of a discrete observation setting using the probabilities of the partitioned observation space.
-%Figure ~\ref{fig:obsPartition} shows the observation partitions obtained in the 1D problem instance for different horizons and their related probabilities. In a lyx file, needed?
 Next we present some results for 2-Dimensional continuous observation spaces.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{figure}[tbp!]
-%\vspace{-2mm}
-%\centering
-%\includegraphics[width=0.42\textwidth]{pics/nodes.pdf}\\
-%\vspace{-2mm}
-%\includegraphics[width=0.42\textwidth]{pics/time.pdf}
-%\vspace{-2mm}
-%\caption{\footnotesize space 
-%and time vs. horizon.
-%}
-%\label{fig:timeSpace}
-%\vspace{-4mm}
-%\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Empirical Results}
-We evaluated our continuous POMDP solution using XADDs on the one-dimensional
-\textsc{Power Plant} example and another variant of this problem with two variables. \footnote{
-Full problem specifications and Java code to reproduce these experiments are currently 
-in Google Code; anonymity restrictions prevent us from providing this link in the submitted paper version.}
-We start by defining the 2-dimensional description of this problem:
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[tbp!]
-\vspace{-2mm}
-\centering
-\includegraphics[width=0.3\textwidth]{pics/time2.pdf}
-\includegraphics[width=0.3\textwidth]{pics/nodes2.pdf}
-\includegraphics[width=0.3\textwidth]{pics/alpha-vectors2.pdf}
-\vspace{-3mm}
-\caption{\footnotesize 
-{\it (left)} Space vs Horizon; 
-{\it (middle)} Time vs Horizon; 
-{\it (right)} Number of $\alpha$-vectors vs Horizon.
-}
-\label{fig:timeSpace}
-\vspace{-4mm}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-{\bf \textsc{2D-Power Plant}:} We consider the more complex model of the power plant similar to \cite{steam2} where the pressure inside the water tank must be controlled to avoid water mixing into the steam or explosion of the tank.
-We model the pressure variable $p$ as a partially observable variable from the observation readings of the pressure valve $po$.
-The two actions of increase and decrease are defined based on the change in both the temperature and the pressure.
-For the increase action the transition functions and the reward function is defined as below:
-\begin{align}
-P(p_s'|\vec{p_s},a)&= \delta\left[ p_s' - 
-\begin{cases}
- (p+10> 20) &: 20 \\ 
-\neq (p+10> 20) &: p_s + 10 \\
-\end{cases}
-\right]\nonumber
-\\
-P(t_s'|\vec{t_s},a)&= \delta\left[ t_s' -  t_s +10 \right]\nonumber
-\\
-R(t_s,p_s,a) &= 
-\begin{cases}
-(5 \leq p_s \leq 15)\wedge (95 \leq t_s \leq 105)&:50\\
-(5 \leq p_s \leq 15)\wedge (t_s \leq 95)&: -1\\
-(5 \leq p_s \leq 15)\wedge (t_s \geq 95)&: -3\\				
-(p_s \leq 5) &: -3\\						
-(p_s \geq 15) &: -5\\ 
-\end{cases}\nonumber
-\end{align}
-There is a high reward for staying withing the safe temperature and pressure range since it produces power, else depending on how safe it is to have values higher or lower than the safe range, penalty is defined.
-As for the decrease action, the transition functions reduce the temperature by 5 units and the pressure by 10 units as long as the pressure stays above zero.
-For the reward function, we assume that there is always a small penalty for decreasing the values because power can not be generated.
-
-For the observation model we consider two continuous uniform distributions such as the following:  
-\begin{align}
-P(t_o|t_s') &= 
-\begin{cases}
- (t_o>t_s + 80) \wedge (t_o<t_s+ 105) &: 0.4 \\
- \neg( (t_o>t_s + 80) \wedge (t_o<t_s+ 105)) &: 0 \\
-\end{cases}\nonumber
-\end{align}
-\begin{align}
-P(p_o|p_s') &= 
-\begin{cases}
- (p_o>p_s) \wedge (p_o<p_s+10) &: 0.1 \\
- \neg( (p_o>p_s) \wedge (p_o<p_s+10)) &: 0 \\
-\end{cases}\nonumber
-\end{align}
-
-We define two rectangular uniform beliefs around the regions of rewarding, so that one needs to increase the values while the other should decrease them: 
-\begin{align}
-b_1 = 
-\begin{cases}
-(p_o>p_s) \wedge (p_o<p_s+10) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 100)&: 0.01 \\
- \neg( (p_o>p_s) \wedge (p_o<p_s+10) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 100)) &: 0 \\
-\end{cases}\nonumber
-\end{align}
-\begin{align}
-b_2 = 
-\begin{cases}
-(p_o>p_s +10) \wedge (p_o<p_s+30) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 130)&: 0.00125 \\
- \neg((p_o>p_s +10) \wedge (p_o<p_s+30) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 130)) &: 0 
-\end{cases}\nonumber
-\end{align}
-% I can draw this and the 1D belief
-In Figure ~\ref{fig:timeSpace}, a time and space analysis of
-the two versions of the problem have been performed for up to 6 horizons.
-Comparing the two problem sizes demonstrates the effect of the number of state-observation variables in our algorithm.
-As the algorithm progresses, the time required to compute the probability of the partitions and finding the maximum value function increases for both problem sizes and significantly more for the 2D version.
-The stability in time is due to the fact that each stage almost takes the same time since it has converged quickly.%???
-Increase in the problem size, increases the partition numbers on the observation space and this produces more $\alpha$-vectors which also effecting the space required to perform the algorithm.
-The number of vectors stays the same for most horizons and they drop after convergence in the 2D problem instance.
-The number of nodes in each iteration increases as more space is required to perform higher iterations.
-
-In Figure ~\ref{fig:3D}  we show plots of the value function for different iterations of the 2D problem instance.
-Each plot demonstrates the value as a function of the pressure and temperature.
-Starting with the first iteration, the value is highest for the reward range ($5<p<15 \wedge 95<t<105$) and -1 or less for other places.
-In the fifth iteration, the value function has partitioned into more pieces showing how higher temperatures can increase the value without considering the effect of the pressure.
-In the last plot, horizon 6 has better tuned the value function so that higher temperatures and pressures increase the value function and also within the reward range, finer grain partitions have been formed.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% policy2, v2plot.pdf, v9plot
-% annotation?
-\begin{figure*}[tbp!]
-\vspace{-2mm}
-\centering
-\includegraphics[width=0.31\textwidth]{pics/2d1.pdf}
-\includegraphics[width=0.31\textwidth]{pics/2d9.pdf}
-\includegraphics[width=0.31\textwidth]{pics/2d111.pdf}
-\vspace{-3mm}
-\caption{\footnotesize 
-{\it (left)} Value function for first iteration; 
-{\it (middle)} $V^5(p,t)$; 
-{\it (right)} $V^6(p,t)$.
-}
-\label{fig:3D}
-\vspace{-4mm}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \section{Related Work}
-This work has used the concept of continuous states and the SDP solution to MDPs from ~\cite{sanner_uai11} and continuous observations from the work of \cite{pascalPomdp}.
-It is exclusive in the sense that it brings together the continuous state and observation POMDPs using a symbolic approach.
-In many applications of the POMDP framework, the state and observation space have been considered as discrete values such as ~\cite{steam2} here we avoid this general simplification and work with the real values from sensors.
-%sampling, monte carlo, particle filter, gaussians, perseus, pergeus
-
-There has been prior work on approximate solutions to POMDPs with large or continuous state spaces ~\cite{Thrun99h}, ~\cite{Perseus} and some has been extended to the continuous observation setting using methods such as sampling ~\cite{PerseusObs}.
-In most approximate continuous state POMDP work, continuous or large discrete actions has been used.
-Thus we can extend the current work using \cite{zamaniAAAI} to contain continuous actions and provide exact or approximate solutions.
 
 \section{Conclusion}
-We presented the first exact solution to continuous states and observations in a DC-POMDP framework.
-Transitions and reward functions are defined as piecewise linear functions of the continuous states.
-Continuous observations require definition of explicit partitions based on the observation model and the belief state of the agent.
-The key contribution is to define the partitions on the observation space and derive the related probabilities using symbolic integration and maximization.
-We extend the application to more than one state-observation setting and present the results.
 
-%\subsubsection*{References} 
+\subsubsection*{References} 
 \bibliography{dcpomdp}

### 2012-10-11 cpomdp/camdp-xadd/doc/NIPS12/draft6.tex
File reorganization
Edits made on camera-ready

--- 2012-09-28 cpomdp/camdp-xadd/src/doc/NIPS/draft6-10page.tex
@@
 
-\author{
-%Zahra Zamani \\
-%ANU & NICTA\\
-%Canberra, Australia \\
-%\texttt{zahra.zamani@anu.edu.au} \\
-%\And
-%Coauthor \\
-%Affiliation \\
-%Address \\
-%\texttt{email} \\
-%\AND
-%Coauthor \\
-%Affiliation \\
-%Address \\
-%\texttt{email} \\
-%\And
-%Coauthor \\
-%Affiliation \\
-%Address \\
-%\texttt{email} \\
-%\And
-%Coauthor \\
-%Affiliation \\
-%Address \\
-%\texttt{email} \\
-%(if needed)\\
+\author
+{
+Zahra Zamani \\
+NICTA & ANU\\
+Canberra, Australia \\
+\texttt{zahra.zamani@anu.edu.au} \\
+\And
+Scott Sanner \\
+NICTA & ANU\\
+Canberra, Australia \\
+\texttt{ssanner@nicta.com.au} \\
+\And
+Pascal Poupart \\
+University of Waterloo \\
+Waterloo, Canada \\
+\texttt{ppoupart@uwaterloo.ca} \\
+\And
+Kristian Kersting \\
+Fraunhofer IAIS, Sankt Augustin, Germany \\
+IGG, University of Bonn, Germany \\
+\texttt{kristian.kersting@iais.fraunhofer.de} \\
 }
@@
 \hspace{5mm}
-\includegraphics[width=0.45\textwidth]{pics/b1-4.pdf}
+\includegraphics[width=0.45\textwidth]{pics/dag2.pdf}
 \end{center}
 \vspace{-2mm}
-\caption{\footnotesize (left) Example conditional plan $\beta^h$ for discrete observations; (right) an example piecewise $\alpha$-vector for continuous state $t$ represented
+\caption{\footnotesize (left) Example conditional plan $\beta^h$ for discrete observations; (right) an example piecewise $\alpha$-vector for continuous state $x$ represented
 as a 
@@
 {
-	$Z:=0$ \;	
+	$O:=0$ \;	
 		\For {all $\alpha_j$ in  $\Gamma^{h-1}$}    
@@
 		\emph{// Generate partition points on the observation space}\\
-		$Z := max(\delta_j(xd_o),z)$\;
-    	}
-    	\emph{// $\phi_{z_k}$ logical term on each partition of $Z$}\\
+		$O := max(\delta_j(xd_o),\delta_{j+1}(xd_o))$\;
+    	}
+    	\emph{// $\phi_{o_k}$ logical term on each partition of $O$}\\
     	\emph{// Generate probabilities and partitions of observations that distinguish optimal $\alpha$-vector (conditional plan)}\\
-   		$P(Z=z_k|b_i) := \int_{x_s'} \int_{x_o} \int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}d_{xd_s'}$ \;
-    \Return{$P(z)$} \;
+   		$P(O=o_k|b_i) := \int_{x_s'} \int_{x_o} \int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$ \;
+    \Return{$P(o)$} \;
     %do this for each belief in B
 }
-\caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a,b_i$) $\longrightarrow$ $P(z)$ }
+\caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a,b_i$) $\longrightarrow$ $P(o)$ }
 \label{alg:genrelobs}
@@
 \begin{align}
-\alpha_1^2(t_s,t_o) &= 
+\alpha_{close}^2(t_s,t_o) &= 
 \begin{cases}
@@
 \hspace{2mm} 
-\alpha_2^2(t_s,t_o) &= \begin{cases}
+\alpha_{open}^2(t_s,t_o) &= \begin{cases}
 (t_s - 10 < t_o<t_s) &: -0.1 \\
@@
 current state variable in line 6--8. The resulting $\delta$-functions
-now only depend on the observation: $\delta_{j}(xd_o) := $
+now only depend on the observation: $\delta_{j}^{b_i}(xd_o) := $
 $\int_{x_s'} \bigoplus_{x_d'} b_i * \alpha_j(xd_s,xd_o) \hspace{1mm}dx_s'$.  For $b_2: U[t_s;6,11]$
@@
 \begin{align}
-\delta_{1}(t_o) &= 
+\delta_{close}^{b_2}(t_o) &= 
 \begin{cases}
@@
 \hspace{5mm} 
-\delta_{2}(t_o) &= \begin{cases}
+\delta_{open}^{b_2}(t_o) &= \begin{cases}
  (15 < t_o< 18) &: 25*t_o - 450\\
@@
 %}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \hspace{-20mm}
-\includegraphics[height=40mm]{pics/beliefs.pdf}
-\hspace{-10mm}
-\includegraphics[width=0.33\textwidth]{pics/delta_b1.pdf}
-\hspace{-2mm}
-\includegraphics[width=0.42\textwidth]{pics/delta_b2.pdf}
+\includegraphics[width=0.42\textwidth]{pics/beliefs_2.pdf}
+\hspace{10mm}
+%\includegraphics[width=0.33\textwidth]{pics/delta_b1.pdf}
+%\hspace{-2mm}
+\includegraphics[width=0.42\textwidth]{pics/delta_b2_2.pdf}
 \hspace{-17mm}
@@
 {\it (left)} Beliefs $b_1,b_2$ for 1D \textsc{\bf Power Plant} example; 
-{\it (center)} Observation dependent function for $b_2$ that partition the observation space into 5 regions with different probabilities for $P(z_1),P(z_2)$ ; 
-{\it (right)} 3 Relevant observation partitions (which is only one partition) and their probabilities for $b_2$.
+%{\it (center)} Observation dependent function for $b_2$ that partition the observation space into 5 regions with different probabilities for $P(o_1),P(o_2)$ ; 
+{\it (right)} Relevant observation partitions and their probabilities for $b_2$.
+This diagrams shows that for temperatures $t_o < 5.1$ the policy is to open the valve and for higher temperatures, it is unsafe to open it and the policy should follow $\delta_{close}(t_o)$ which is to close the valve.
 }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_1$.
+The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_2$.
 According to continuous observation space defined in \cite{pascal_ijcai05}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space.
@@
 \begin{align}
-\mathrm{max} \Bigg(\delta_{1}(t_o),\delta_{2}(t_o)\Bigg) &= 
-\begin{cases}
-z_1: (14 < t_o< 18) &: 0.025*t_o - 0.45\\
-z_1: (8 < t_o< 14) &:  -0.1\\
-z_1: (5.1 < t_o< 8) &: - 0.025*t_o -0.1\\
-z_2: (5 < t_o< 5.1) &: - 25*t_o + 127.5\\
-z_2: (4 < t_o< 5) &:  2.5*t_o - 10\\
+\mathrm{max} \Bigg(\delta_{close}^{b_2}(t_o),\delta_{open}^{b_2}(t_o)\Bigg) &= 
+\begin{cases}
+o_1: (14 < t_o< 18) &: 0.025*t_o - 0.45\\
+o_1: (8 < t_o< 14) &:  -0.1\\
+o_1: (5.1 < t_o< 8) &: - 0.025*t_o -0.1\\
+o_2: (5 < t_o< 5.1) &: - 25*t_o + 127.5\\
+o_2: (4 < t_o< 5) &:  2.5*t_o - 10\\
 \end{cases}
@@
 }
-Note here that each partitioned is labeled according to the original $\delta$-function where $z_1$ is the label of the partition coming from $\delta_{1}(t_o)$. 
+Note here that each partitioned is labeled according to the original $\delta$-function where $o_1$ is the label of the partition coming from $\delta_{close}^{b_2}(t_o)$. 
 These partitions define observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the two distinct observation partitions is found.
 This is demonstrated visually
-in Figure~\ref{fig:beliefs} (center) for $b_2$ and for $b_1$
-in Figure~\ref{fig:beliefs} (right). Note here that the maximum of the $\delta$-functions for $b_1$ is equal to $\delta_{2}(t_o)$.
+in Figure~\ref{fig:beliefs} (right) for $b_2$. %Note here that the maximum of the $\delta$-functions for $b_2$ is equal to $\delta_{open}^{b_2}(t_o)$.
 
@@
 partition:
-$$P(z_k|b_i) := \int_{xd_s'}\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}d_{xd_s'}$$
-For our example the probabilities can be calculated as below:
+$$P(o_k|b_i) := \int_{x_s'}\int_{x_o}\int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$$
+For our example the non-zero probabilities occur in the following partitions as below:
 {\footnotesize
@@
 \begin{align}
-\mathrm{max} \Bigg(\delta_{1}(t_o),\delta_{2}(t_o)\Bigg) &= 
-\begin{cases}
-(\phi_1 : 14 < t_o< 18) &: 0.2\\
-(\phi_2 : 8 < t_o< 14) &:  0.6\\
-(\phi_3 : 5.1 < t_o< 8) &: 0.183\\
-(\phi_4 : 5 < t_o< 5.1) &: 0.002\\
-(\phi_5 : 4 < t_o< 5) &:  0.0125\\
+P(o_k|b_2) &= 
+\begin{cases}
+(\phi_1 : 14 < t_o< 18) & : 0.2\\
+(\phi_2 : 8 < t_o< 14) & :  0.6\\
+(\phi_3 : 5.1 < t_o< 8) & : 0.183\\
+(\phi_4 : 5 < t_o< 5.1) & : 0.0002\\
+(\phi_5 : 4 < t_o< 5) & :  0.0125\\
 \end{cases}
@@
 For any number of actions and observations the total number of observation partitions depends on the number of belief points.
-In our example we must have two observation probabilities which is defined as the sum of the probabilities for $z_1,z_2$ according to the constraints: \\
+In our example we must have two observation probabilities which is defined as the sum of the probabilities for $o_1,o_2$ according to the constraints:
 \begin{align*}
-z_1&: \phi_1 \vee \phi_2 \vee \phi_3 \longrightarrow P(z_1|b_2)=0.983 \\
-z_2&: \phi_4 \vee \phi_5  \longrightarrow P(z_2|b_2)=0.0127
+o_1&: \phi_1 \vee \phi_2 \vee \phi_3 \longrightarrow P(o_1|b_2)=0.983 \\
+o_2&: \phi_4 \vee \phi_5  \longrightarrow P(o_2|b_2)=0.0127
 \end{align*}
-
 %{\footnotesize
@@
 
-In Figure ~\ref{fig:3D}  we present plots of the maximum $\alpha$-vectors of belief $b_1$ for different iterations of the 2D problem instance.
+In Figure ~\ref{fig:3D}  we present plots of the maximum $\delta$-vectors of belief $b_1$ for different iterations of the 2D problem instance.
 %Each plot demonstrates the value of each vector as a function of the pressure and temperature.
@@
 In the fifth iteration, the value function has partitioned into more pieces, showing how higher temperatures can increase the value without considering the effect of the pressure.
-In the last plot, horizon $h=6$ has better tuned the value function so that higher temperatures and pressures increase the value of the maximum $\alpha$-vector and also within the reward range, finer grain partitions have been formed.
+In the last plot, horizon $h=6$ has better tuned the value function so that higher temperatures and pressures increase the value of the maximum $\delta$-vector and also within the reward range, finer grain partitions have been formed.
 %Also note that the policy depends on the initial belief and for these plots they were based on $b_1$ which was lower than the reward range.
@@
 \centering
-\includegraphics[width=0.31\textwidth]{pics/2d1.pdf}
-\includegraphics[width=0.31\textwidth]{pics/2d9.pdf}
-\includegraphics[width=0.31\textwidth]{pics/2d111.pdf}
+\includegraphics[width=0.31\textwidth]{pics/2d1-4.pdf}
+\includegraphics[width=0.31\textwidth]{pics/2d9-4.pdf}
+\includegraphics[width=0.31\textwidth]{pics/2d111-4.pdf}
 \vspace{-3mm}
 \caption{\footnotesize 
-{\it (left)}Maximum $\alpha$-vector for $b_1$ in first iteration; 
-{\it (center)} $\alpha_{max}^5(b_1)$; 
-{\it (right)} $\alpha_{max}^5(b_1)$.%V^6(p,t)
+{\it (left)}Maximum $\delta$-vector for $b_1$ and action of $close$ in first iteration; 
+{\it (center)} $\delta_{close}^5(b_1)$; 
+{\it (right)} $\delta_{close}^6(b_1)$. Finer grain partitions show that closing (or opening) the valve can occur at more exact temperatures at higher horizons. 
 }
@@
 
+\subsubsection*{Acknowledgments}
+
+NICTA is funded by the Australian Government as represented by the Department of Broadband, Communications and the Digital Economy and the Australian Research Council through the ICT Centre of Excellence program.
+This work was supported by the Fraunhofer ATTRACT fellowship STREAM and by the EC, FP7-248258-First-MM.
 %\subsubsection*{References} 

### 2012-11-01 cpomdp/camdp-xadd/doc/NIPS12/draft6.tex
Intermediate checkin

@@
-% TODO: 
-%
-% (1) GenRelObs and propagate fixes to text
-%- i's where they shouldn't be, 
-%- integrating over discrete vars, 
-%- P(z_k|...) needs to condition on belief state, 
-%- line 13 generates does not get rid of xd_s'... did you mean to integrate out alpha(xd_o,xd_s) instead?
-% ... \delta_j(x_{d_o}) should integrate out to 1 right?
-%     then use this?
-%
-% (2) Notation for observation set
-%
-% (3) Improve intuitions
-%
-% (4) Revise conclusion
-%
-% (5) More formal claims, where this works, where PBVI exact
-
 \documentclass{article} % For LaTeX2e
@@
 
+\renewcommand{\vec}[1]{\mathbf{#1}} % Bold vectors
 \def\argmax{\operatornamewithlimits{arg\max}}
@@
 {
-Zahra Zamani \\
-NICTA & ANU\\
-Canberra, Australia \\
-\texttt{zahra.zamani@anu.edu.au} \\
+Zahra Zamani\\ %\& Scott Sanner\\
+ANU \& NICTA\\
+Canberra, Australia\\
+{\small \texttt{zahra.zamani@anu.edu.au}}\\
+%{\small \texttt{first.last@anu.edu.au}}\\
 \And
-Scott Sanner \\
-NICTA & ANU\\
-Canberra, Australia \\
-\texttt{ssanner@nicta.com.au} \\
+Scott Sanner\\
+NICTA \& ANU\\
+Canberra, Australia\\
+{\small \texttt{scott.sanner@nicta.com.au}}\\
+\AND
+Pascal Poupart\\
+U. of Waterloo\\
+Waterloo, Canada\\
+{\small \texttt{ppoupart@uwaterloo.ca}}\\
 \And 
-Pascal Poupart \\
-University of Waterloo \\
-Waterloo, Canada \\
-\texttt{ppoupart@uwaterloo.ca} \\
-\And
-Kristian Kersting \\
-Fraunhofer IAIS, Sankt Augustin, Germany \\
-IGG, University of Bonn, Germany \\
-\texttt{kristian.kersting@iais.fraunhofer.de} \\
+Kristian Kersting\\
+Fraunhofer IAIS \& U. of Bonn\\
+Bonn, Germany\\
+%{\small \texttt{first.last@ubonn.de???}}\\
+%{\small \texttt{first.last@iais.fraunhofer.de}}\\
+{\small \texttt{kristian.kersting@iais.fraunhofer.de}}\\
 }
@@
 
+\newcommand{\xds}{\mathbf{dx}_s}
+\newcommand{\xdsp}{\mathbf{dx}_s'}
+\newcommand{\xdo}{\mathbf{dx}_o}
 \newcommand{\open}{\mathit{open}}
@@
 \newcommand{\new}{\marginpar{NEW}}
-
-%\nipsfinalcopy % Uncomment for camera-ready version
+\renewcommand{\l}{\langle}
+\renewcommand{\r}{\rangle}
+
+\nipsfinalcopy % Uncomment for camera-ready version
 
@@
 \begin{abstract}
-Partially-observable Markov decision processes (POMDPs) provide a
-powerful model for real-world sequential decision-making problems.  In
-recent years, point-based value iteration methods have proven to be
-extremely effective techniques for finding (approximately) optimal
-dynamic programming solutions to POMDPs when an initial set of belief
-states is known.  However, no point-based work has provided exact
-point-based backups for both continuous state and observation spaces,
-which we tackle in this paper.  Our key insight is that while there
-may be an infinite number of possible observations, there are only a
-finite number of observation partitionings that are relevant for
-optimal decision-making when a finite, fixed set of reachable belief
-states is known.  To this end, we make two important contributions:
-(1) we show how previous exact symbolic dynamic programming solutions
-for continuous state MDPs can be generalized to continuous state
-POMDPs with discrete observations, and (2) we show how this solution
-can be further extended via recently developed symbolic methods to
-continuous state and observations to \emph{derive} the minimal
-relevant observation partitioning for potentially correlated,
-multivariate observation spaces.
-%for domais with piecewise
-%polynomial transition and observation functions.  
-We demonstrate our algorithm on a power plant
-control problem requiring continuous state 
-\emph{and} observations.
-%empirically evaluate on real-world inspired power 
-%proof-of-concept results on uni- and multi-variate state and
-%observation power plant control.
+%Partially-observable Markov decision processes (POMDPs) provide a
+%powerful model for real-world sequential decision-making problems.  
+Point-based value iteration (PBVI) methods have
+proven extremely effective for finding
+(approximately) optimal dynamic programming solutions to
+partially-observable Markov decision processes (POMDPs) when a 
+set of initial belief states is known.  However, no PBVI work has
+provided \emph{exact point-based backups for both continuous state and
+observation spaces}, which we tackle in this paper.  Our key insight is
+that while there may be an infinite number of observations,
+there are only a finite number of continuous observation partitionings
+that are relevant for optimal decision-making when a finite, fixed set
+of reachable belief states is considered.  To this end, we make two
+important contributions: (1) we show how previous exact symbolic
+dynamic programming solutions for continuous state MDPs can be
+generalized to \emph{continuous state POMDPs with discrete observations}, and
+(2) we show how recently developed symbolic integration methods
+allow this solution to be extended 
+to PBVI for \emph{continuous state and observation POMDPs} with
+potentially correlated, multivariate continuous observation spaces.
+%We demonstrate a proof-of-concept implementation on power plant regulation.
 \end{abstract}
@@
 problems~\cite{kaebling}.  In recent years, point-based value
-iteration methods~\cite{pbvi_jair06,hsvi2,Perseus,gapmin} have proved
+iteration methods (PBVI)~\cite{pbvi_jair06,hsvi2,Perseus,gapmin} have proved
 extremely successful at scaling (approximately) optimal POMDP
@@
 
-More recent work has extended point-based methods to both continuous
-state and continuous observation spaces, but no work has tackled both
-jointly without sampling.  \cite{Perseus_cont} provides exact
-point-based backups for continuous state and discrete observation
-problems (with approximate sample-based extensions to continuous
-actions and observations), while~\cite{pascal_ijcai05} provides exact
-point-based backups for discrete state and continuous observation
-problems (where multivariate observations must be conditionally
-independent).  While
+While PBVI has been extended to both continuous state and continuous
+observation spaces, no prior work has tackled both jointly without sampling.
+\cite{Perseus_cont} provides exact point-based backups for continuous
+state and discrete observation problems (with approximate sample-based
+extensions to continuous actions and observations),
+while~\cite{pascal_ijcai05} provides exact point-based backups (PBBs)
+for discrete state and continuous observation problems (where
+multivariate observations must be conditionally independent).  While
 restricted to discrete states, \cite{pascal_ijcai05} provides an
@@
 
-Our contributions in this work are two-fold.  First, we extend
+We propose two major contributions:  First, we extend
 symbolic dynamic programming for continuous state
-MDPs~\cite{sanner_uai11} to the case of continuous state and discrete
-observation POMDPs.  This provides an expressive and concrete
-instantiation of the framework in~\cite{Perseus_cont} (which only
-abstractly required that integrals were computable) in that it ensures
-the closed-form of all $\alpha$-functions for \emph{all horizons} is a
-symbolic piecewise case statement, even for POMDPs with
-\emph{arbitrary} continuous rewards and transitions with discrete
-noise (i.e., a finite mixture of deterministic transitions).  
-Second, we extend this symbolic dynamic programming algorithm
-to the case of continuous observations (while restricting the
-transition to piecewise linear functions and the reward functions to piecewise constants) and extend
-the work of~\cite{pascal_ijcai05} to \emph{derive} relevant observation
-partitions for potentially correlated multivariate continuous
-observation spaces by exploiting the piecewise polynomial integration
-operation of~\cite{sanner_aaai12} and the multivariate symbolic maximization
-technique of~\cite{sanner_uai11}.  
-We conclude by demonstrating our algorithm on a power plant
-control problem requiring bivariate continuous state and observations.
-
-%\footnote{All code for these
-%experiments is freely available online (link suppressed) and
-%attached as anonymized supplementary material to this submission.}
+MDPs~\cite{sanner_uai11} to POMDPs with discrete 
+observations, 
+%This provides an expressive and concrete
+%instantiation of the framework in~\cite{Perseus_cont} (which only
+%abstractly required that integrals were computable) in that it ensures
+%the closed-form of all $\alpha$-functions for \emph{all horizons} is a
+%symbolic piecewise case statement, even for POMDPs 
+\emph{arbitrary} continuous reward and transitions with discrete noise
+(i.e., a finite mixture of deterministic transitions).  Second, we
+extend this symbolic dynamic programming algorithm to PBVI and the case of
+continuous observations (while restricting 
+transition dynamics to be piecewise linear with discrete noise, rewards to be
+piecewise constant, and observation probabilities and 
+beliefs to be uniform) by building 
+on~\cite{pascal_ijcai05} to \emph{derive} relevant observation
+partitions for potentially correlated, multivariate continuous
+observations. % spaces.
+
+%by exploiting the piecewise polynomial integration
+%operation of~\cite{sanner_aaai12} and the multivariate symbolic
+%maximization technique of~\cite{sanner_uai11}.  We conclude by
+%demonstrating our algorithm on a power plant control problem requiring
+%bivariate continuous state and observations.
 
@@
 
-We assume familiarity with MDPs and introduce discrete and continuous
-partially observable MDPs (DC-POMDPs) as an extension to their DC-MDP
-subset~\cite{sanner_uai11}.  A DC-POMDP is a tuple $\langle
+\label{sec:model}
+
+%% TODO: further restrictions on observations!  also discrete problems!
+%We assume familiarity with MDPs and
+%We introduce discrete and continuous
+%partially observable MDPs (DC-POMDPs) as an extension to 
+%DC-MDPs~\cite{sanner_uai11}.  
+A discrete and continuous
+partially observable MDP (DC-POMDP) is a tuple $\langle
 \mathcal{S},\mathcal{A},\mathcal{O},\mathcal{T},\mathcal{R},\mathcal{Z},\gamma,h
-\rangle$.  States $\mathcal{S}$ are represented by vectors of
-$(\vec{d_s},\vec{x_s}) = (
+\rangle$.  States $\mathcal{S}$ are given by vector 
+$\xds = (\vec{d}_s,\vec{x}_s) = (
 d_{s_1},\ldots,d_{s_n},x_{s_1},\ldots,x_{s_m} )$ where each $d_{s_i}
-\in \{ 0,1 \}$ ($1 \leq i \leq n$) is a discrete boolean and each
-$x_{s_j} \in \mathbb{R}$ ($1 \leq j \leq m$) is a continuous variable.
+\in \{ 0,1 \}$ ($1 \leq i \leq n$) is boolean and each
+$x_{s_j} \in \mathbb{R}$ ($1 \leq j \leq m$) is continuous.
 We assume a finite, discrete action space $\mathcal{A} = \{ a_1,
-\ldots, a_p \}$. Observations
-$\mathcal{O}$ are given by the vector $(\vec{d_o},\vec{x_o}) = (
+\ldots, a_r \}$. Observations
+$\mathcal{O}$ are given by the vector $\xdo = (\vec{d}_o,\vec{x}_o) = (
 d_{o_1},\ldots,d_{o_p},x_{o_1},\ldots,x_{o_q} )$ where each $d_{o_i}
@@
 
-Three functions are required for modeling DC-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state given the action and previous state%$P(\vec{d}',\vec{x}'|\cdots,a)$
+Three functions are required for modeling DC-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state %$(\vec{d}_s',\vec{x}_s')$
+given the action and previous state%
+%$(\vec{d}_s,\vec{x}_s)$ 
+%$p(\vec{d}',\vec{x}'|\cdots,a)$
 ; (2)  $\mathcal{R}:\mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R}$ a reward function which returns the immediate reward of taking an action in some state; and (3) an observation function defined as $\mathcal{Z} : \mathcal{S} \times \mathcal{A} \times \mathcal{O} \rightarrow [ 0, 1 ]$  which gives the probability of an observation given the outcome of a state after executing an action.
@@
 
-We use a dynamic Bayesian network (DBN) to compactly represent the transition model over the factored state variables and we likewise use a Bayesian network to represent the observation model:
-{\footnotesize
-\begin{align*}
+We use a dynamic Bayes net (DBN)\footnote{We disallow general 
+  synchronic arcs for simplicity of exposition but
+  note their inclusion only places restrictions on the variable
+  elimination ordering used during the dynamic programming backup
+  operation.} to compactly represent the transition model $\mathcal{T}$ over the
+factored state variables and we use a two-layer Bayes net to
+represent the observation model $\mathcal{Z}$: {\footnotesize
+\begin{align}
 \mathcal{T}: \;\; &
-P(\vec{d_s}',\vec{x_s}'|\vec{d_s},\vec{x_s},a) = 
-\prod_{i=1}^n P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \prod_{j=1}^m P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a). \nonumber \\
+%p(\vec{d}_s',\vec{x}_s'|\vec{d}_s,\vec{x}_s,a) = 
+p(\xdsp|\xds,a) = 
+\prod_{i=1}^n p(d_{s_i}'|\xds,a) \prod_{j=1}^m p(x_{s_j}'|\xds, \vec{d}_s',a). \label{eq:trans_model} \\
 \mathcal{Z}: \;\; & 
-P(\vec{d_o},\vec{x_o}|\vec{d_s},\vec{x_s},a) = 
-\prod_{i=1}^n P(d_{o_i}|d_{s_i},a) \prod_{j=1}^m P(x_{o_j}|x_{s_j},a). \nonumber 
-\end{align*}}
-For simplicity of algorithmic exposition, we do not allow synchronic
-arcs in the transition representation, but note that to incorporate
-such arcs only requires restrictions on the variable elimination
-ordering during dynamic programming.  In these transition and
-observation Bayes nets, 
-probabilities over 
-\emph{discrete} variables are represented using conditional probability
-functions (CPF) in the form of
-$P(d_i'|\vec{d},\vec{x},a)$. Probabilities for \emph{continuous} variables
-$P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ are represented by
-\emph{piecewise linear functions} (PLFs) that are first-order Markov
-and deterministic, i.e. the probabilities are encoded using the Dirac
-$\delta[\cdot]$ function. The reward function $R(\vec{d},\vec{x},a)$
-is set to a piecewise constant function. We next provide an intuitive
-example to make this DC-POMDP framework more concrete.
+%p(\vec{d}_o,\vec{x}_o|\vec{d}_s,\vec{x}_s,a) = 
+p(\xdo|\xdsp,a) = 
+\prod_{i=1}^p p(d_{o_i}|\xdsp,a) \prod_{j=1}^q p(x_{o_j}|\xdsp,a). \label{eq:obs_model}
+\end{align}}
+%% TRANSITION LINEAR, BELIEF, REWARD, OBSERVATION PWC???
+Probabilities over \emph{discrete} variables $p(d_{s_i}'|\xds,a)$ and
+$p(d_{o_i}|\xdsp,a)$ may condition on both discrete variables and
+(nonlinear) inequalities of continuous variables; this is further
+restricted to linear inequalities in the case of continuous
+observations.  Transitions over \emph{continuous}
+variables $p(x_{s_j}'|\xds, \vec{d}_s',a)$ must be deterministic (but
+arbitrary nonlinear) piecewise functions;
+%encoded using the Dirac $\delta$ function; 
+in the case of continuous observations they are further restricted to
+be piecewise linear; this permits discrete noise in the continuous 
+transitions since they may condition on stochastically sampled
+discrete next-state variables $\vec{d}_s'$.
+%(hence allowing discrete noise, but not, e.g., Gaussian noise).
+Observation probabilities over continuous variables $p(x_{o_j}|\xdsp,a)$ 
+only occur in the case of continuous observation and are required to be
+piecewise constant (a mixture of uniform distributions); the same
+restriction holds for belief state representations.
+The reward $R(\vec{d},\vec{x},a)$ may be 
+an arbitrary (nonlinear) piecewise function in the case of
+deterministic observations and a piecewise constant function in the
+case of continuous observations.  
+We now provide concrete examples.
+%To make this concrete, we now provide examples of both discrete and
+%continuous observation DC-POMDPs.
+
+\textbf{Example} \textsc{\bf (Power Plant)~\cite{steam2}} \emph{The steam
+generation system of a power plant evaporates feed-water under restricted 
+pressure and temperature conditions to turn a stream turbine.
+A reward is obtained when electricity is generated from the turbine 
+and the steam pressure and temperature are within safe ranges.
+Mixing water and steam makes the
+respective pressure and temperature observations $p_o \in \mathbb{R}$
+and $t_o \in \mathbb{R}$ on the underlying state $p_s \in \mathbb{R}$
+and $t_s \in \mathbb{R}$ highly uncertain.  Actions $A = \{ \open, \close \}$
+control temperature and pressure by means of a pressure valve.}
+
+We initially present two DC-POMDP variants labeled \textsc{\bf 1D-Power
+  Plant} using a single temperature state variable $t_s$.  
+The transition and reward are common to both ---
+temperature increments (decrements) with a closed (opened) valve, a
+large negative reward is given for a closed valve with $t_s$ exceeding
+critical threshold $15$, and positive reward is given for a safe, 
+electricity-producing state:
+% For the transition probability, we use the Dirac function to model the deterministic equations.
+We can also define stochasticity in the transition model using boolean random variables that are sampled stochastically.
+The reward function can be any linear function of the state or action.
+Going above a threshold temperature (e.g.
+$T=10$) will cause an explosion in the plant when trying to close the valve which results in the negative reward of $-1000$.
+Staying below this temperature is safe and will produce electricity an gain the reward of $100$.
+The reward of an open valve is $-1$.
+{\footnotesize
+\vspace{-1mm}
+\begin{align}
+\label{eq:trans}
+p(t_s'|t_s,a)= \delta\left[ t_s' - 
+\begin{cases}
+ (a=\open) &: t_s - 5 \\ 
+(a = \close) &: t_s + 7 \\
+\end{cases}
+\right]
+\hspace{2mm}
+R(t_s,a) = 
+\begin{cases}
+ (a=\open) &: -1 \\
+(a = \close)\wedge (t_s>15) &: -1000 \\
+(a = \close)\wedge \neg(t_s>15) &: 100 \\
+\end{cases} 
+\end{align}
+\vspace{-4mm}
+}
+
+%contribution? 
+Next we introduce the \textsc{\bf Discrete Obs. 1D-Power Plant} variant where
+we define an \emph{observation space with a single discrete binary 
+variable} $o \in \mathcal{O} = \{\high,\low\}$:
+{\footnotesize
+\vspace{-1mm} 
+\begin{align}
+\hspace{-2.5mm} p(o=\high|t_s',a=\open) = 
+\begin{cases}
+  t_s' \leq 15 &: 0.9 \\
+  t_s' > 15    &: 0.1 \\
+\end{cases}
+\;\;\;\;
+p(o=\high|t_s',a=\close) = 
+\begin{cases}
+ t_s' \leq 15 &: 0.7 \\
+ t_s' > 15    &: 0.3 \\
+\end{cases} \label{eq:ex_disc_obs}
+\end{align}
+\vspace{-4mm}
+}
+%\\
+%p(t_{o_1}|t_s',\close) = 
+%\begin{cases}
+% \delta\left[ t_s \leq 15 \right] &: 0.1 \\
+% \delta\left[ \neg (t_s \leq 15)\right] &: 0.8 \\
+%\end{cases}\nonumber
+%,
+%p(t_{o_2}|t_s',\close) = 
+%\begin{cases}
+% \delta\left[ t_s \leq 15 \right] &: 0.9 \\
+% \delta\left[ \neg (t_s \leq 15)\right] &: 0.2 \\
+%\end{cases} \nonumber
+%\end{align}
+
+Finally we introduce the \textsc{\bf Cont. Obs. 1D-Power Plant}
+variant where we define an \emph{observation space with
+a single continuous variable} $t_o$ uniformly distributed on
+an interval of 10 units centered at $t_s'$.
+{\footnotesize
+\vspace{-1mm}
+\begin{align}
+p(t_o|t_s',a=\open) = U(t_o;t_s' - 5, t_s' + 5) = 
+\begin{cases}
+ (t_o>t_s'-5) \wedge (t_o<t_s'+5)         &: 0.1 \\
+ (t_o \leq t_s'-5) \vee (t_o \geq t_s'+5) &: 0 \\
+\end{cases} \label{eq:ex_cont_obs}
+\end{align}
+\vspace{-4mm}
+}
+While simple, we note no prior method could perform exact point-based backups for either problem.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
-\hspace{5mm}
+\hspace{10mm}
 \includegraphics[width=0.45\textwidth]{pics/dag2.pdf}
@@
 \vspace{-2mm}
-\caption{\footnotesize (left) Example conditional plan $\beta^h$ for discrete observations; (right) an example piecewise $\alpha$-vector for continuous state $x$ represented
-as a 
-%The maximum $\alpha$ vector for a belief state at iteration 4 for the \textsc{\bf Power Plant} problem in Section~\ref{}
-decision diagram: 
-the \emph{true} branch is solid, the \emph{false}
-branch is dashed.}
+\caption{\footnotesize (left) Example conditional plan $\beta^h$ for
+discrete observations; (right) example $\alpha$-function for $\beta^h$
+over state $b \in \{0,1\}, x \in \mathbb{R}$ in decision
+diagram form: the \emph{true} (1) branch is solid, the \emph{false} (0) branch is
+dashed.}
 \label{fig:cond_plan}
-\vspace{-1mm}
 \end{figure}
@@
 
-\textbf{Example} \textsc{\bf (Power Plant)~\cite{steam2}} \emph{The steam
-generation system of a power plant aims to provide hot steam to the
-steam turbine which in turn produces electricity. Feed-water is
-exposed to heat in the tubes leading to the water tank, where water
-evaporation is performed under specific pressure and temperature.
-A reward is obtained when electricity is generated through steam
-production and the pressure and temperature are within safe ranges.
-Mixing water and steam makes the
-respective pressure and temperature observations $p_o \in \mathbb{R}$
-and $t_o \in \mathbb{R}$ on the underlying state $p_s \in \mathbb{R}$
-and $t_s \in \mathbb{R}$ highly uncertain.  Action choices are either
-to open or close a pressure valve $a \in \{ \open, \close \}$.}
-
-We initially present two 
-simple examples labeled \textsc{\bf 1D-Power Plant} 
-using only the temperature state $t_s$.
-First, the transition and reward functions common to both problems
-are defined as below where simply, the temperate increments or
-decrements according to the action:
-% For the transition probability, we use the Dirac function to model the deterministic equations.
-We can also define stochasticity in the transition model using boolean random variables that are sampled stochastically.
-The reward function can be any linear function of the state or action.
-Going above a threshold temperature (e.g.
-$T=10$) will cause an explosion in the plant when trying to close the valve which results in the negative reward of $-1000$.
-Staying below this temperature is safe and will produce electricity an gain the reward of $100$.
-The reward of an open valve is $-1$.
-{\footnotesize
-\vspace{-1mm}
-\begin{align}
-\label{eq:trans}
-P(t_s'|\vec{t_s},a)= \delta\left[ t_s' - 
-\begin{cases}
- (a=\open) &: t_s - 5 \\ 
-(a \neq \open) &: t_s + 7 \\
-\end{cases}
-\right]
-\hspace{2mm}
-R(t_s,a) = 
-\begin{cases}
- (a=\open) &: -1 \\
-(a \neq \open)\wedge (t_s>T) &: -1000 \\
-(a \neq \open)\wedge \neg(t_s>T) &: 100 \\
-\end{cases}
-\end{align}
-\vspace{-4mm}
-}
-
-%contribution? 
-Next we introduce the \textsc{\bf Discrete Obs. 1D-Power Plant} variant using 
-a discrete binary observation space with possible
-observations $o \in \mathcal{O} = \{\high,\low\}$:
-{\footnotesize
-\vspace{-1mm}
-\begin{align}
-P(o=\high|t_s',\open) = 
-\begin{cases}
-  t_s' <= 15 &: 0.9 \\
- \neg (t_s' <= 15)&: 0.1 \\
-\end{cases}
-\hspace{5mm}
-P(o=\low|t_s',\open) = 
-\begin{cases}
- t_s' <= 15 &: 0.1 \\
- \neg (t_s' <= 15) &: 0.9 \\
-\end{cases}
-\end{align}
-\vspace{-3mm}
-}
-%\\
-%P(t_{o_1}|t_s',\close) = 
-%\begin{cases}
-% \delta\left[ t_s <= 15 \right] &: 0.1 \\
-% \delta\left[ \neg (t_s <= 15)\right] &: 0.8 \\
-%\end{cases}\nonumber
-%,
-%P(t_{o_2}|t_s',\close) = 
-%\begin{cases}
-% \delta\left[ t_s <= 15 \right] &: 0.9 \\
-% \delta\left[ \neg (t_s <= 15)\right] &: 0.2 \\
-%\end{cases} \nonumber
-%\end{align}
-
-Finally we introduce the \textsc{\bf Cont. Obs. 1D-Power Plant}
-variant with continuous observation $t_o$.  
-We define the observation probability over $t_o$ as uniform on
-an interval 10 units wide, centered at $t_s'$.
-{\footnotesize
-\begin{align}
-P(t_o|t_s',\open) = 
-\begin{cases}
- (t_o>t_s'-5) \wedge (t_o<t_s'+5) &: 0.1 \\
- \neg((t_o>t_s'-5) \wedge (t_o<t_s'+5)) &: 0 \\
-\end{cases}
-\end{align}
-}
-While the restrictions may seem too tight in our problems due to the continuous nature of the state and observation space, there has been no exact solution to any of the two set of problems above. 
-\section{DC-POMDP solution}
+\section{Value Iteration for DC-POMDPs}
+
+\label{sec:vi}
 
 In a DC-POMDP, the agent does not directly observe the states and thus
-must maintain a belief state $b(xd_s) = P(xd_s)$.\footnote{For
-compactness we define $xd_s := (x_s,d_s)$ and $xd_o:=(x_o,d_o)$.}  For a
-given belief state $\vec{b} = b(\cdot)$, a POMDP policy $\pi$ can be
+must maintain a belief state $b(\xds) = p(\xds)$.  For a
+given belief state $\vec{b} = b(\xds)$, a POMDP policy $\pi$ can be
 represented by a tree corresponding to a conditional plan $\beta$.  An
@@
 of $(h-1)$-step conditional plans as shown in
-Fig.~\ref{fig:cond_plan}.
-
+Fig.~\ref{fig:cond_plan} (left).
 Our goal is to find a policy $\pi$ that maximizes the value function,
@@
 starting from initial belief state $\vec{b}$:
+{\footnotesize
+\vspace{-1mm}
 \begin{equation}
@@
 \end{equation}
+\vspace{-4mm}
+}
+
 where $r_t$ is the reward obtained at time $t$ and $\vec{b}_0$ is the
 belief state at $t=0$.  For finite $h$ and belief state $\vec{b}$, the
-optimal policy $\pi$ takes the form of an $h$-step conditional plan
+optimal policy $\pi$ is given by an $h$-step conditional plan
 $\beta^h$.  For $h = \infty$, the optimal discounted ($\gamma < 1$)
-value function can be approximated arbitrarily closely by using
-a sufficiently large, but finite $h$~\cite{kaebling}.
+value can be approximated arbitrarily closely by 
+a sufficiently large $h$~\cite{kaebling}.
 
@@
 are discrete), 
-the optimal POMDP value function for finite horizon $h$ is a piecewise linear
-and convex function of the belief state $\vec{b}$~\cite{Perseus_cont} hence 
+the optimal POMDP value function for finite horizon $h$ is a piecewise linear and
+convex function of the belief state $\vec{b}$~\cite{Perseus_cont}, hence 
 $V^h$ is given by a maximum over a finite set of
-``$\alpha$-vectors'' $\alpha^h_i$:
+``$\alpha$-functions'' $\alpha^h_i$:
+{\footnotesize 
+\vspace{-1mm}
 \begin{equation}
-V^h(\vec{b}) = \max\nolimits_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot \vec{b}
+V^h(\vec{b}) = \max_{\alpha^h_i \in \Gamma^h} \l \alpha^h_i, \vec{b} \r = \max_{\alpha^h_i \in \Gamma^h} \int_{\vec{x}_s} \sum_{\vec{d}_s} \alpha^h_i(\xds) \cdot \vec{b}(\xds) \; d\vec{x}_s 
 \end{equation}
+\vspace{-4mm}
+}
+
 Later on when we tackle continuous state \emph{and} observations,
@@
 
-The $\Gamma^h$ in this optimal $h$-stage-to-go value function can be
-computed via Monahan's dynamic programming approach to \emph{value
-iteration} (VI)~\cite{monahan82}.  
-Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (2) and considering $r_a = r(x_s,d_s,a)$ :\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as
-$\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
-\textrm{\large $\boxplus$} Q = \{ \vec{p} + \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}
-{\footnotesize
-\begin{align} 
-g^h_{a,o,j} &=  \int_{x_{s'}} \sum_{d_{s'}} p(o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^{h-1}_j(xd'_s) ; \hspace{2mm}  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1} \label{eq:backup} \\
-\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j  \label{eq:cross_prod}\\ 
+Given discrete observations $o \in \mathcal{O}^h$, the $\Gamma^h$ in
+this optimal $h$-stage-to-go value function can be computed via
+Monahan's dynamic programming approach to \emph{value iteration}
+(VI)~\cite{monahan82}.  Initializing $\alpha^0_1 = \vec{0}$ and
+$\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from
+$\Gamma^{h-1}$ using the backup operation defined in (2):\footnote{The
+  $\textrm{\large $\boxplus$}$ of sets is defined as $\textrm{\large
+    $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large
+    $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the
+  pairwise cross-sum $P \textrm{\large $\boxplus$} Q = \{ \vec{p} +
+  \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}  
+{\footnotesize
+\vspace{-1mm}
+\begin{align} 
+g^h_{a,o,j}(\xds) &=  \int_{\vec{x}_{s'}} \sum_{\vec{d_{s'}}} p(o|\xds',a)p(\xds'|\xds,a) \alpha^{h-1}_j(\xds') d\vec{x}_{s'}; \hspace{2mm}  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1} \label{eq:backup} \\
+\Gamma^{h}_a   &= R(\xds,a) + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\xds) \right\}_j  \label{eq:cross_prod}\\ 
 \Gamma^h  &= \bigcup_a \Gamma^h_a 
 \end{align}
-}
-\vspace{-3mm}
+\vspace{-4mm}
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{algorithm}[t!]
+\footnotesize
 \vspace{-.5mm}
@@
        $h:=h+1, \Gamma^h :=\emptyset, \Gamma_{PBVI}^h :=\emptyset$\;
-       \ForEach {$b_i \in B$}
+       \ForEach {$\vec{b}_i \in B$}
        {
@@
 			$\Gamma_{a}^h :=\emptyset$ \;       		
-       		\If {$ContObs =true$}
-       			{\emph{// Continuous observation so derive observation set $O_i^h$ for belief $b_i$} \;
-			  $(O_i^h,P(O_i^h|xd_{s})) \,:=\,$ \genObs{$\Gamma_{PBVI}^{h-1},a,b_i$}\;}
+       		\If {(continuous observations: $q > 0$)}
+       			{\emph{// Derive relevant observation partitions $O_i^h$ for belief $\vec{b}_i$} \;
+			  $\l O_i^h,p(O_i^h|\vec{b}_i,a) \r \,:=\,$ \genObs{$\Gamma_{PBVI}^{h-1},a,\vec{b}_i$}\;}
        		 \ForEach {$o \in \mathcal{O}_i^{h}$}
@@
        			{
-   	 		  		$\alpha_j^{h'-1},:=\,$ \remapWithPrimes{$\alpha_j^{h-1}$} 
-   	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
-%   	 		    	$\Gamma_{a,xd_{o_i},j}^h \,:=\, P(o|xd_{s}) \cdot$ \backup{$\alpha_j^{h'},a$}\;
+   	 		  		$\alpha_j^{h-1} :=\,$ \remapWithPrimes{$\alpha_j^{h-1}$} 
+   	 		  		\emph{// $\forall d_i$: $d_i \to d_i'$ and $\forall x_i$: $x_i \to x_i'$} \; 
+%   	 		    	$\Gamma_{a,xd_{o_i},j}^h \,:=\, p(o|xd_{s}) \cdot$ \backup{$\alpha_j^{h},a$}\;
    	 		    	{$g_{a,o,j}^h \,:=\, $ see Eq~\eqref{eq:backup}}
        	      	}
-%       	      	$\Gamma_{a,xd_o}^h\,:=\,\textrm{\large $\boxplus$} \Gamma_{a,xd_{o_i}}^h$\;
-%       	      	$\Gamma_{a,xd_o}^h\,:=\, $ see Eq~\eqref{eq:cross_prod}\;
-       	     }
-%           $\Gamma_a^{h} \,:=\,R_a \oplus \gamma \cdot \Gamma_{a,xd_o}^h$\;
+%       	      	$\Gamma_{a,\xdo}^h\,:=\,\textrm{\large $\boxplus$} \Gamma_{a,xd_{o_i}}^h$\;
+%       	      	$\Gamma_{a,\xdo}^h\,:=\, $ see Eq~\eqref{eq:cross_prod}\;
+       	     }
+%           $\Gamma_a^{h} \,:=\,R_a \oplus \gamma \cdot \Gamma_{a,\xdo}^h$\;
             $\Gamma_a^{h} \,:=\, $ see Eq~\eqref{eq:cross_prod}\;
@@
         	%monahan's pruning first generates all vectors, then prunes
-              $\Gamma^h \,:=\, $\prune{$\Gamma^h$} \emph{//optional step}\; 
-      }
-             % $V^h \,:=\, \mathrm{max}_{\alpha_j \in \Gamma^h} b_i \cdot \alpha_j$\;
+                %but not for PBVI since we only retain alpha-functions provably dominant at a belief point
+%              $\Gamma^h \,:=\, $\prune{$\Gamma^h$} \emph{// optional strict dominance testing of $\alpha$-functions}\; 
+      }
+             % $V^h \,:=\, \mathrm{max}_{\alpha_j \in \Gamma^h} \vec{b}_i \cdot \alpha_j$\;
              % $\pi^{*,h} \,:=\, \argmax_{a} \, \Gamma_a^{h}$\;
-      \ForEach {$b_i \in B$}
+      \ForEach {$\vec{b}_i \in B$}
       {
-     	$\alpha_{b_i}^h :=\ \argmax_{\alpha_j \in \Gamma^h} \alpha_j \cdot b_i$\;
-     	$\Gamma_{PBVI}^h :=\ \Gamma_{PBVI}^h \cup \alpha_{b_i}^h$\;
+     	$\alpha_{\vec{b}_i}^h :=\ \argmax_{\alpha_j \in \Gamma^h} \alpha_j \cdot \vec{b}_i$\;
+     	$\Gamma_{PBVI}^h :=\ \Gamma_{PBVI}^h \cup \alpha_{\vec{b}_i}^h$\;
       }
@@
 }
-\caption{\footnotesize \texttt{PBVI}(DC-POMDP, $H$,$ContObs$, $B=\left\{b_i \right\}$) $\longrightarrow$ $(V^h,\pi^{*,h})$ \label{alg:vi}}
+\caption{\footnotesize \texttt{PBVI}(DC-POMDP, $H$, $B=\left\{\vec{b}_i \right\}$) $\longrightarrow$ $\l V^h \r$ \label{alg:vi}}
 \vspace{-1mm}
@@
 
-\textbf{Point-based value iteration (PBVI)} computes the optimal value function only for a set of belief states $\{ b_i \}$ where $b_i := P(xd_s)$.
+\textbf{Point-based value iteration (PBVI)} computes the value function only for a set of belief states $\{ \vec{b}_i \}$ where $\vec{b}_i := p(\xds)$.
 The idea is straightforward and the main modification needed to Monahan's algorithm in Algorithm~\ref{alg:vi} is the loop from lines 19--21 where only $\alpha$-vectors optimal at some belief state are retained for subsequent iterations.
 We note that \emph{if} all beliefs reachable in $h$ steps from some initial set can be finitely enumerated and PBVI is executed for this set of beliefs, then optimality is guaranteed for the PBVI policy over these initial beliefs up to horizon $h$.
+
 %In actuality in the DC-POMDP setting, the generation of continuous observations will most often lead to an infinite number of reachable belief states even in one step, so this optimality guarantee of PBVI will only hold for DC-POMDPs in very special cases when the set of reachable beliefs is finite.
@@
 We pause for a moment to make a few notes about the \texttt{PBVI}
-algorithm in the context of DC-POMDPs.  If we have a fixed discrete
-observation set, i.e., $\forall h \, \mathcal{O}^h = \mathcal{O}$, then
-the algorithm proceeds as usual, but if we have continuous observation
-variables then we will need to derive a relevant set of observations
-on line 10, which we describe in Section~\ref{sec:cont_obs}.  We need
-to guarantee that the integral on line 14 can be done in closed-form
-and we guarantee this for the piecewise functions in
-Section~\ref{sec:disc_obs}. We also note that the beliefs are piecewise constant functions. Because in general we will deal with
-$\alpha$-functions represented as piecewise functions, we only assume
-that strict dominance testing is implemented for \texttt{Prune} in
-line 18.
+algorithm in the context of DC-POMDPs.  If we have no continuous
+observations ($q=0$), the the discrete observation set is
+$\mathcal{O}^h = \{ \vec{d}_o \}, \forall h$ with $p(\vec{d}_o)$ as
+defined by~\eqref{eq:obs_model}.  However, if we have continuous
+observation variables ($q > 0$) then we will need to derive a relevant
+set of observations on line 10, as described in
+Section~\ref{sec:cont_obs}.
 
@@
 In this section we take a symbolic dynamic programming (SDP) approach
-to performing \texttt{PBVI} in Algorithm~\ref{alg:vi}.  This requires
-us to define all of the operations used in \texttt{PBVI} on a
-closed-form representation.  For SDP, this closed-form will be a
-piecewise case statement and it will be stored compactly as a decision
-diagram~\cite{sanner_uai11}.  We first review this form and then
-proceed to define SDP operations specific to the case of DC-POMDPs.
+to implementing VI and PBVI as defined in the last section.  To do this,
+we need only show that all required operations can be computed efficiently
+and in closed-form, which we do next, building on SDP for 
+MDPs~\cite{sanner_uai11}.
 
@@
 %overview + example plant
-The \textsc{\bf Power Plant} example represented all functions using
-the case structure which can be generally defined as {\footnotesize
-\vspace{-3mm}
+The previous \textsc{\bf Power Plant} examples represented all functions in case form,
+generally defined as {\footnotesize
+\vspace{-1mm}
 \begin{align}
@@
 \end{align}
-}
+\vspace{-4mm}
+}
+
 and this is the form we use to represent all functions in a DC-POMDP.
-Here $\phi_i$ are disjoint logical formulae defined over the state $(\vec{d},\vec{x})$ with logical ($\land,\lor,\neg$) combinations of boolean variables and inequalities ($\geq,>,\leq,<$) over continuous variables.
-The $f_i$ are general functions, we will restrict them later for the
-case of continuous observation POMDPs.
+The $\phi_i$ are disjoint logical formulae defined over $\xds$ and/or $\xdo$ with logical ($\land,\lor,\neg$) combinations of boolean variables and inequalities ($\geq,>,\leq,<$) over continuous variables.
+For discrete observation DC-POMDPs, the $f_i$ and inequalities may use any function (e.g., $\sin(x_1) > \log(x_2)\cdot x_3)$; for continuous observations, they are restricted to linear inequalities and linear or piecewise constant $f_i$ as described in Section~\ref{sec:model}.
 
@@
 {\footnotesize 
-\vspace{-3mm}
+\vspace{-4mm}
 \begin{center}
@@
 }
+
 The following SDP operations on case statements require more detail than can be provided here, hence we refer the reader to the relevant literature:
-\begin{itemize}
-\item Restriction $f|_{\phi}$:  Takes a function $f$ to restrict only in cases
+%\begin{itemize}
+%\item 
+{\it Restriction $f|_{\phi}$:}  Takes a function $f$ to restrict only in cases
 that satisfy some formula $\phi$ as defined in \cite{sanner_uai11}.
-\item Substitution $f\sigma$ : Takes a set $\sigma$ of variables and their substitutions (which may be case statements), and carries out all variable substitutions in sequence~\cite{sanner_uai11}.
-\item Integration $\int_{x_1} f dx_1$:  There are two forms: If $x_1$ is involved in a $\delta$-function (cf.
-the transition in Eq.~\eqref{eq:trans}) then the integral is equivalent to a symbolic substitution and can be applied to any case statement~\cite{sanner_uai11}.
-Otherwise, if $f$ is restricted to linear constraints and polynomial values, then the approach of~\cite{sanner_aaai12} can be applied to yield a result in the same form.
-\end{itemize}
+%\item 
+{\it Substitution $f\sigma$:} Takes a set $\sigma$ of variables and their substitutions (which may be case statements), and carries out all variable substitutions in sequence~\cite{sanner_uai11}.
+%\item 
+{\it Integration $\int_{x_1} f dx_1$:}  There are two forms: If $x_1$ is involved in a $\delta$-function (cf.
+the transition in Eq~\eqref{eq:trans}) then the integral is equivalent to a symbolic substitution and can be applied to any case statement~\cite{sanner_uai11}.
+Otherwise, if $f$ is restricted to linear constraints and constant values, then the approach of~\cite{sanner_aaai12} can be applied to yield a linearly constained piecewise linear result.
+%\end{itemize}
+
 %\vspace{-2mm}
@@
 %\hspace{15mm}
-%$\int_{x_j'} P(x_j'|\cdots) V'^{h} dx_j' = \begin{cases}
+%$\int_{x_j'} p(x_j'|\cdots) V'^{h} dx_j' = \begin{cases}
 %    \phi_1: & V'^{h} \{ x_j' = f_1 \} \\ 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{PBVI for DC State and Discrete Observations} 
+\subsection{VI for DC State and Discrete Observations} 
 \label{sec:disc_obs}
 
-For discrete and continuous states and discrete observations, we use the symbolic version of Monahan's algorithm as the DC-POMDP solution represented in \texttt{PBVI}.
-In brief we note that all operations as defined in PBVI apply directly to functions represented as case statements.
-For example, we can rewrite
-Eq~\eqref{eq:backup} in the following case operation form:
-{\footnotesize
-\begin{align}
-g^h_{a,o,j} =  \int_{x_{s'}} \large \bigoplus_{d_{s'}} p(o|xd'_s,a) \otimes p(xd'_s|xd_s,a) \otimes \alpha^h_j(xd'_s) \nonumber
-\end{align}
-}
-Crucially we note that since $p(xd'_s|xd_s,a)$ is always defined using $\delta$-functions (possibly conditioned on stochastic discrete variables) for deterministic transitions as defined in our DC-POMDPs, then the integral can always be computed in closed case form as discussed in Section~\ref{sec:case}.
+For DC-POMDPs with only discrete observations $o
+\in \mathcal{O}$ and observation function $p(o|\xds',a)$ (e.g., 
+Eq~\eqref{eq:ex_disc_obs}), we introduce a symbolic version of
+Monahan's VI algorithm.  In brief, we note that all VI operations
+needed in Section~\ref{sec:vi} apply \emph{directly} to 
+DC-POMDPs, e.g., we can rewrite
+Eq~\eqref{eq:backup}: {\footnotesize
+\vspace{-1mm}
+\begin{equation}
+g^h_{a,o,j}(\xds) \!
+=  \!\!
+\int_{\vec{x}_{s'}} \!\!
+\bigoplus_{\vec{d_{s'}}} \!
+\left[ p(o|\xds',\!a) \!
+\otimes \!\!
+\left( \!
+\bigotimes_{i=1}^n p(d_{s_i}'|\xds,\!a) \!\!
+\right) \!\!
+\otimes \!\!
+\left( \!
+\bigotimes_{j=1}^m p(x_{s_j}'|\xds, \vec{d}_s',\!a) \!\!
+\right) \!\!
+\otimes \!
+\alpha^{h-1}_j(\xds') \!
+\right] \!\!
+d\vec{x}_{s'} \label{eq:backup_sdp}
+\end{equation}
+\vspace{-4mm}
+}
+
+Crucially we note since the continuous transition cpfs $p(x_{s_j}'|\xds, \vec{d}_s',a)$ are deterministic and hence defined with Dirac $\delta$'s (e.g., Eq~\ref{eq:trans}) as described in Section~\ref{sec:model}, the integral $\int_{\vec{x}_{s'}}\!$ can always be computed in closed case form as discussed in Section~\ref{sec:case}.
 In short, nothing additional is required for PBVI on DC-POMDPs
-with continuous state and discrete observations --- the crucial insight
-to understand is simply that $\alpha$-vectors are now represented by
+in this case --- the key insight
+is simply that $\alpha$-functions are now represented by
 case statements and can ``grow'' with the horizon as they partition the
@@
 \begin{algorithm}[t!]
+\footnotesize 
 \vspace{-.5mm}
@@
 {
-	$O:=0$ \;	
-		\For {all $\alpha_j$ in  $\Gamma^{h-1}$}    
+		\For {$\alpha_j(\xdsp) \in \Gamma^{h-1}$}    
 		{
-		\emph{// Point-based backup operation}\\
-    	$\alpha_j(xd_s,xd_o) := \int_{x_s'} \bigoplus_{d_s'} P(xd_{o}|xd'_{s},a) \cdot P(xd'_{s}| xd_{s},a)\cdot \alpha_j \hspace{1mm}dx_s'$\;
-		}  
-		\For {all $j$ in $\alpha_j(xd_s,xd_o)$}    
+		\emph{// Point-based backup (PBB) operation}\\
+    	$\alpha_j(\xds,\xdo) := \int_{\vec{x}_s'} \bigoplus_{\vec{d}_s'} p(\xdo|\xdsp,a) \otimes p(\xdsp| \xds,a) \otimes \alpha_j(\xdsp) \; d\vec{x}_s'$\;
+		}  
+		\ForEach {$\alpha_j(\xds,\xdo)$}    
 		{
-		\emph{// Generate value of $\alpha$-vector for belief $b_i$ as a function of observations}\\
-		$\delta_{j}(xd_o) := $ $\int_{x_{s}} \bigoplus_{d_s} b_i \cdot \alpha_j(xd_s,xd_o)\hspace{1mm}dx_s$\\ \;
-		}
-		\For {all $j$ in $\delta_{j}(xd_o)$}    
-		{
-		\emph{// Generate partition points on the observation space}\\
-		$O := max(\delta_j(xd_o),\delta_{j+1}(xd_o))$\;
-    	}
-    	\emph{// $\phi_{o_k}$ logical term on each partition of $O$}\\
-    	\emph{// Generate probabilities and partitions of observations that distinguish optimal $\alpha$-vector (conditional plan)}\\
-   		$P(O=o_k|b_i) := \int_{x_s'} \int_{x_o} \int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$ \;
-    \Return{$P(o)$} \;
+		\emph{// Generate value of $\alpha$-vector for belief $\vec{b}_i(\xds)$ as a function of observations}\\
+		$\delta_{j}(\xdo) := \int_{\vec{x}_{s}} \bigoplus_{\vec{d}_s} \vec{b}_i(\xds) \otimes \alpha_j(\xds,\xdo) \; d\vec{x}_s$\\ \;
+		}
+		\emph{// Generate observation partitions -- see text for details}\\
+		$\mathcal{O}^h := \max(\delta_1(\xdo),\ldots,\delta_{j}(\xdo))$\;
+
+	\ForEach {$o_k \in \mathcal{O}^h$}{
+    	  \emph{// Let $\phi_{o_k}$ be the partition constraints for observation $o_k \in \mathcal{O}^h$}\\
+            $p(\mathcal{O}^h = o_k|\vec{b}_i) := \int_{x_s'} \int_{x_o} \int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} p(\xdo|\xds',a) \otimes p(\xds'|\xds,a) \otimes \vec{b}_i \otimes  \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$ \;
+        }
+    \Return{$\l \mathcal{O}^h, p(\mathcal{O}^h|\vec{b}_i,a) \r$} \;
     %do this for each belief in B
 }
-\caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a,b_i$) $\longrightarrow$ $P(o)$ }
+\caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a,\vec{b}_i$) $\longrightarrow$ $\l \mathcal{O}^h, p(\mathcal{O}^h|\vec{b}_i,a) \r$ }
 \label{alg:genrelobs}
@@
 
-The additional operation required for \texttt{PBVI} in DC-POMDPs in
-Algorithm~\ref{alg:vi} over the discrete observation case is simply
-\emph{generating relevant observations} as shown on line 10.
-\texttt{GenRelObs} in Algorithm~\ref{alg:genrelobs} defines the steps
-required to generate these partitions.
-
-We use the second iteration of the \textsc{\bf Power Plant} example to
-demonstrate partitioning of the entire continuous observation
-space. The main assumption is to consider some belief state $b_i$ and
-obtaining the relevant partitions according to that belief. For
-example, we can choose two beliefs as uniform distributions,
-e.g. $b_1: U[t_s;2,6]$ and $b_2: U[t_s;6,11]$ as shown in
-Figure~\ref{fig:beliefs} (left).  Starting with an empty set of
+In general, it would be impossible to apply standard VI to 
+DC-POMDPs with continuous observations since the number of observations
+is infinite.  However, building on ideas in~\cite{pascal_ijcai05},
+in the case of PBVI, it is possible to \emph{derive} a finite set of
+continuous observation partitions that permit exact point-based backups.
+This additional operation (\texttt{GenRelObs}) appears on line 10 of PBVI in 
+Algorithm~\ref{alg:vi} in the case of continuous observations and is
+formally defined in Algorithm~\ref{alg:genrelobs}.
+
+To demonstrate the generation of relevant continuous observation partitions, 
+we use the second iteration of the \textsc{\bf Cont. Obs. 1D-Power
+  Plant} along with two beliefs points represented as 
+uniform distributions: $b_1: U(t_s;2,6)$ and $b_2: U(t_s;6,11)$ as
+shown in Figure~\ref{fig:beliefs} (left).
+
+Starting with an empty set of
 relevant observation partitions $\mathcal{O}^h$, the algorithm first
-eliminates the next state variable $xd'_s$. It regresses the
+eliminates the next state variable $\xds'$. It regresses the
 the set of
 $\alpha$-vectors of the previous horizon back one step via
-the transition and observation model: $\alpha_j(xd_s,xd_o) :=
-\int_{x_s'} \bigoplus_{x_d'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j\hspace{1mm}dx_s'
+the transition and observation model: $\alpha_j(\xds,\xdo) :=
+\int_{x_s'} \bigoplus_{x_d'} p(xd_{o_i}|xd'_{s_i},a)  \otimes  p(xd'_{s_i}| xd_{s_i},a) \otimes  \alpha_j\hspace{1mm}dx_s'
 $.  This operation of line 5 in \texttt{GenRelObs} is similar to 
@@
 current state variable in line 6--8. The resulting $\delta$-functions
-now only depend on the observation: $\delta_{j}^{b_i}(xd_o) := $
-$\int_{x_s'} \bigoplus_{x_d'} b_i * \alpha_j(xd_s,xd_o) \hspace{1mm}dx_s'$.  For $b_2: U[t_s;6,11]$
+now only depend on the observation: $\delta_{j}^{\vec{b}_i}(\xdo) := $
+$\int_{x_s'} \bigoplus_{x_d'} \vec{b}_i  \otimes  \alpha_j(\xds,\xdo) \hspace{1mm}dx_s'$.  For $b_2: U[t_s;6,11]$
 %assume the following simple $\delta$-functions:
@@
 {\it (left)} Beliefs $b_1,b_2$ for 1D \textsc{\bf Power Plant} example; 
-%{\it (center)} Observation dependent function for $b_2$ that partition the observation space into 5 regions with different probabilities for $P(o_1),P(o_2)$ ; 
+%{\it (center)} Observation dependent function for $b_2$ that partition the observation space into 5 regions with different probabilities for $p(o_1),p(o_2)$ ; 
 {\it (right)} Relevant observation partitions and their probabilities for $b_2$.
@@
 %regardless of a particular $\alpha$-vector, it integrates to one:
-%$\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*P(s) = 1$.
+%$\int_{\xdo}\int_{\xds} p(\xdo|\xds',a)*p(\xds'|\xds,a)*p(s) = 1$.
 Thus we only need to multiply the indicators of each observation partition 
@@
 partition:
-$$P(o_k|b_i) := \int_{x_s'}\int_{x_o}\int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$$
+$$p(o_k|\vec{b}_i) := \int_{x_s'}\int_{x_o}\int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} p(\xdo|\xds',a)*p(\xds'|\xds,a)*\vec{b}_i* \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$$
 For our example the non-zero probabilities occur in the following partitions as below:
@@
 \begin{align}
-P(o_k|b_2) &= 
+p(o_k|b_2) &= 
 \begin{cases}
@@
 \begin{align*}
-o_1&: \phi_1 \vee \phi_2 \vee \phi_3 \longrightarrow P(o_1|b_2)=0.983 \\
-o_2&: \phi_4 \vee \phi_5  \longrightarrow P(o_2|b_2)=0.0127
+o_1&: \phi_1 \vee \phi_2 \vee \phi_3 \longrightarrow p(o_1|b_2)=0.983 \\
+o_2&: \phi_4 \vee \phi_5  \longrightarrow p(o_2|b_2)=0.0127
 \end{align*}
@@
 %\begin{align}
-%P(z_k)=
+%p(z_k)=
 %\begin{cases}
@@
 \begin{align}
-P(p_s'|\vec{p_s},inc)&= \delta\left[ p_s' - 
+p(p_s'|\vec{p_s},inc)&= \delta\left[ p_s' - 
 \begin{cases}
@@
 \hspace{5mm} 
-P(t_s'|\vec{t_s},inc)= \delta\left[ t_s' - (t_s +10) \right]\nonumber
+p(t_s'|\vec{t_s},inc)= \delta\left[ t_s' - (t_s +10) \right]\nonumber
 \end{align}
@@
 \begin{align}
-P(t_o|t_s') = 
+p(t_o|t_s') = 
 \begin{cases}
@@
 \hspace{5mm} 
-P(p_o|p_s') = 
+p(p_o|p_s') = 
 \begin{cases}

### 2012-11-02 cpomdp/camdp-xadd/doc/NIPS12/draft6.tex
Intermediate checkin

@@
 \renewcommand{\r}{\rangle}
+\newcommand{\casemax}{\mathrm{casemax}}
 
@@
 
-Given discrete observations $o \in \mathcal{O}^h$, the $\Gamma^h$ in
+The $\Gamma^h$ in
 this optimal $h$-stage-to-go value function can be computed via
 Monahan's dynamic programming approach to \emph{value iteration}
-(VI)~\cite{monahan82}.  Initializing $\alpha^0_1 = \vec{0}$ and
-$\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from
-$\Gamma^{h-1}$ using the backup operation defined in (2):\footnote{The
+(VI)~\cite{monahan82}.  Initializing $\alpha^0_1 = \vec{0}$, 
+$\Gamma^0 = \{ \alpha^0_1 \}$, and assuming discrete 
+observations $o \in \mathcal{O}^h$,
+$\Gamma^h$ is obtained from
+$\Gamma^{h-1}$ as follows:\footnote{The
   $\textrm{\large $\boxplus$}$ of sets is defined as $\textrm{\large
@@
 			$\Gamma_{a}^h :=\emptyset$ \;       		
-       		\If {(continuous observations: $q > 0$)}
-       			{\emph{// Derive relevant observation partitions $O_i^h$ for belief $\vec{b}_i$} \;
-			  $\l O_i^h,p(O_i^h|\vec{b}_i,a) \r \,:=\,$ \genObs{$\Gamma_{PBVI}^{h-1},a,\vec{b}_i$}\;}
+       		\eIf {(continuous observations: $q > 0$)}
+       	         {\emph{// Derive relevant observation partitions $\mathcal{O}_i^h$ for belief $\vec{b}_i$} \;
+			  $\l \mathcal{O}_i^h,p(\mathcal{O}_i^h|\xdsp,a) \r \,:=\,$ \genObs{$\Gamma_{PBVI}^{h-1},a,\vec{b}_i$}\;}
+                 {\emph{// Discrete observations and model already known}\;
+                          $\mathcal{O}_i^h := \{ \vec{d}_o \}$; $p(\mathcal{O}_i^h|\xdsp,a) := $ see Eq~\eqref{eq:obs_model}\;}
        		 \ForEach {$o \in \mathcal{O}_i^{h}$}
@@
 
-\textbf{Point-based value iteration (PBVI)} computes the value function only for a set of belief states $\{ \vec{b}_i \}$ where $\vec{b}_i := p(\xds)$.
-The idea is straightforward and the main modification needed to Monahan's algorithm in Algorithm~\ref{alg:vi} is the loop from lines 19--21 where only $\alpha$-vectors optimal at some belief state are retained for subsequent iterations.
-We note that \emph{if} all beliefs reachable in $h$ steps from some initial set can be finitely enumerated and PBVI is executed for this set of beliefs, then optimality is guaranteed for the PBVI policy over these initial beliefs up to horizon $h$.
-
-%In actuality in the DC-POMDP setting, the generation of continuous observations will most often lead to an infinite number of reachable belief states even in one step, so this optimality guarantee of PBVI will only hold for DC-POMDPs in very special cases when the set of reachable beliefs is finite.
-Nonetheless, PBVI has been quite successful in practice without exhaustive enumeration of all reachable beliefs~\cite{pbvi_jair06,hsvi2,Perseus,gapmin}, which motivates our use of PBVI in this work.
-
-We pause for a moment to make a few notes about the \texttt{PBVI}
-algorithm in the context of DC-POMDPs.  If we have no continuous
-observations ($q=0$), the the discrete observation set is
-$\mathcal{O}^h = \{ \vec{d}_o \}, \forall h$ with $p(\vec{d}_o)$ as
-defined by~\eqref{eq:obs_model}.  However, if we have continuous
-observation variables ($q > 0$) then we will need to derive a relevant
-set of observations on line 10, as described in
-Section~\ref{sec:cont_obs}.
+\textbf{Point-based value iteration (PBVI)} computes the value
+function only for a set of belief states $\{ \vec{b}_i \}$ where
+$\vec{b}_i := p(\xds)$.  The idea is straightforward and the main
+modification needed to Monahan's VI approach in Algorithm~\ref{alg:vi}
+(following~\cite{Perseus_cont}) is the loop from lines 22--24 where
+only $\alpha$-vectors optimal at some belief state are retained for
+subsequent iterations.  In the case of continuous observation
+variables ($q > 0$), we will need to derive a relevant set of
+observations on line 10, a key contribution of this work as described
+in Section~\ref{sec:cont_obs}.  Whereas PBVI is optimal if all
+reachable belief states within horizon $H$ are enumerated in $B$, in the DC-POMDP
+setting, the generation of continuous observations will most often
+lead to an infinite number of reachable belief states, even with
+finite horizon.  Nonetheless, PBVI has been quite successful in
+practice without exhaustive enumeration of all reachable
+beliefs~\cite{pbvi_jair06,hsvi2,Perseus,gapmin}, which motivates our
+use of PBVI in this work.
 
@@
 &
-\hspace{-7mm} $\mathrm{max} \Bigg(
+\hspace{-7mm} $\casemax \Bigg(
   \begin{cases}
@@
 %\item 
-{\it Integration $\int_{x_1} f dx_1$:}  There are two forms: If $x_1$ is involved in a $\delta$-function (cf.
+{\it Integration $\int_{x_1} f \; dx_1$:}  There are two forms: If $x_1$ is involved in a $\delta$-function (cf.
 the transition in Eq~\eqref{eq:trans}) then the integral is equivalent to a symbolic substitution and can be applied to any case statement~\cite{sanner_uai11}.
@@
 {
-		\For {$\alpha_j(\xdsp) \in \Gamma^{h-1}$}    
+		\ForEach {$\alpha_j(\xdsp) \in \Gamma^{h-1}$ and $a \in A$}    
 		{
-		\emph{// Point-based backup (PBB) operation}\\
-    	$\alpha_j(\xds,\xdo) := \int_{\vec{x}_s'} \bigoplus_{\vec{d}_s'} p(\xdo|\xdsp,a) \otimes p(\xdsp| \xds,a) \otimes \alpha_j(\xdsp) \; d\vec{x}_s'$\;
-		}  
-		\ForEach {$\alpha_j(\xds,\xdo)$}    
+		\emph{// Perform exact 1-step DP backup of $\alpha$-functions at horizon $h-1$}\\
+    	$\alpha^a_j(\xds,\xdo) := \int_{\vec{x}_s'} \bigoplus_{\vec{d}_s'} p(\xdo|\xdsp,a) \otimes p(\xdsp| \xds,a) \otimes \alpha_j(\xdsp) \; d\vec{x}_s'$\;
+		}  
+		\ForEach {$\alpha^a_j(\xds,\xdo)$}    
 		{
-		\emph{// Generate value of $\alpha$-vector for belief $\vec{b}_i(\xds)$ as a function of observations}\\
-		$\delta_{j}(\xdo) := \int_{\vec{x}_{s}} \bigoplus_{\vec{d}_s} \vec{b}_i(\xds) \otimes \alpha_j(\xds,\xdo) \; d\vec{x}_s$\\ \;
-		}
-		\emph{// Generate observation partitions -- see text for details}\\
-		$\mathcal{O}^h := \max(\delta_1(\xdo),\ldots,\delta_{j}(\xdo))$\;
+		\emph{// Generate value of each $\alpha$-vector at belief point $\vec{b}_i(\xds)$ as a function of observations}\\
+		$\delta^a_{j}(\xdo) := \int_{\vec{x}_{s}} \bigoplus_{\vec{d}_s} \vec{b}_i(\xds) \otimes \alpha^a_j(\xds,\xdo) \; d\vec{x}_s$\\ \;
+		}
+		\emph{// Using $\casemax$, generate observation partitions relevant to each policy -- see text for details}\\
+		$\mathcal{O}^h := \mathrm{extract\text{-}partition\text{-}constraints}[ \casemax(\delta^{a_1}_1(\xdo),\delta^{a_2}_1(\xdo),\ldots,\delta^{a_r}_{j}(\xdo))]$\;
 
@@
     	  \emph{// Let $\phi_{o_k}$ be the partition constraints for observation $o_k \in \mathcal{O}^h$}\\
-            $p(\mathcal{O}^h = o_k|\vec{b}_i) := \int_{x_s'} \int_{x_o} \int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} p(\xdo|\xds',a) \otimes p(\xds'|\xds,a) \otimes \vec{b}_i \otimes  \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$ \;
-        }
-    \Return{$\l \mathcal{O}^h, p(\mathcal{O}^h|\vec{b}_i,a) \r$} \;
+            $p(\mathcal{O}^h = o_k|\xdsp,a) := \int_{\vec{x}_o} \bigoplus_{\vec{d}_o} p(\xdo|\xds',a) \mathbb{I}[\phi_{o_k}] d_{\vec{x}_o}$ \;
+%            $p(\mathcal{O}^h = o_k|\xdsp) := \int_{\vec{x}_s'} \int_{\vec{x}_o} \int_{\vec{x}_s} \bigoplus_{\vec{d}_o} \bigoplus_{\vec{d}_s} \bigoplus_{\vec{d}_s'} p(\xdo|\xds',a) \otimes p(\xds'|\xds,a) \otimes \vec{b}_i \otimes  \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$ \;
+        }
+    \Return{$\l \mathcal{O}^h, p(\mathcal{O}^h|\xdsp,a) \r$} \;
     %do this for each belief in B
 }
-\caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a,\vec{b}_i$) $\longrightarrow$ $\l \mathcal{O}^h, p(\mathcal{O}^h|\vec{b}_i,a) \r$ }
+\caption{\footnotesize \texttt{GenRelObs}($\Gamma^{h-1},a,\vec{b}_i$) $\longrightarrow$ $\l \mathcal{O}^h, p(\mathcal{O}^h|\xdsp,a) \r$ }
 \label{alg:genrelobs}
@@
 in the case of PBVI, it is possible to \emph{derive} a finite set of
-continuous observation partitions that permit exact point-based backups.
+continuous observation partitions that permit exact point-based backups
+\emph{at a belief point}.
 This additional operation (\texttt{GenRelObs}) appears on line 10 of PBVI in 
@@
 we use the second iteration of the \textsc{\bf Cont. Obs. 1D-Power
-  Plant} along with two beliefs points represented as 
+  Plant} along with two belief points represented as 
 uniform distributions: $b_1: U(t_s;2,6)$ and $b_2: U(t_s;6,11)$ as
 shown in Figure~\ref{fig:beliefs} (left).
-
-Starting with an empty set of
-relevant observation partitions $\mathcal{O}^h$, the algorithm first
-eliminates the next state variable $\xds'$. It regresses the
-the set of
-$\alpha$-vectors of the previous horizon back one step via
-the transition and observation model: $\alpha_j(\xds,\xdo) :=
-\int_{x_s'} \bigoplus_{x_d'} p(xd_{o_i}|xd'_{s_i},a)  \otimes  p(xd'_{s_i}| xd_{s_i},a) \otimes  \alpha_j\hspace{1mm}dx_s'
-$.  This operation of line 5 in \texttt{GenRelObs} is similar to 
-the point-based backup in~\eqref{eq:backup}. For
-continuous variables the continuous integral and for discrete
-variables the restriction operation is used as defined in
-Section~\ref{sec:case}.  
-Assuming that after $h=1$ the set of $\alpha$-vectors are
-defined equal to the reward function of equation (1).
-%{\footnotesize
-%\vspace{-3mm}
-%\begin{align}
-%\alpha_1^1(t_s) &= 
-%\begin{cases}
-% (t_s>200) &: -1000 \\
-%\neg(t_s>200) &: 100 \\
-%\end{cases}
-%\hspace{10mm} 
-%\alpha_2^1(t_s) = \top: -1 \nonumber
-%\end{align}
-%}
-For $h=2$ the resulting $\alpha$-vectors after the regression step are functions of the observation and current state: 
-{\footnotesize
-\vspace{-2mm}
-\begin{align}
-\alpha_{close}^2(t_s,t_o) &= 
-\begin{cases}
- (t_s<15)\wedge (t_s - 10 < t_o<t_s) &: 10 \\
-(t_s>15)\wedge (t_s - 10 < t_o<t_s) &: -100  \\
-\neg(t_s - 10 < t_o<t_s) &: 0
+Letting $h=2$, we assume that $\Gamma^{1}$ contains only the 
+reward as defined in~\eqref{eq:trans}.
+Then in lines 2--4 of Algorithm~\ref{alg:genrelobs}, we compute the 
+following two $\alpha$-functions for $a \in \{ \open, \close \}$ 
+($j=1$ since $|\Gamma^{1}| = 1$):
+{\footnotesize
+\vspace{-1mm}
+\begin{align}
+\alpha_1^{\close}(t_s,t_o) &= 
+\begin{cases}
+ (t_s<15)\wedge (t_s \! - \! 10 < t_o<t_s) &\!\!\!: 10 \\
+(t_s\geq15)\wedge (t_s \! - \! 10 < t_o<t_s) &\!\!\!: -100  \\
+\neg(t_s \! - \! 10 < t_o<t_s) &\!\!\! : 0
 \end{cases}
-\hspace{2mm} 
-\alpha_{open}^2(t_s,t_o) &= \begin{cases}
-(t_s - 10 < t_o<t_s) &: -0.1 \\
-\neg(t_s - 10 < t_o<t_s) &: 0
+\;\;
+\alpha_1^{\open}(t_s,t_o) = \begin{cases}
+(t_s \! - \! 10 < t_o<t_s) &\!\!\!: -0.1 \\
+\neg(t_s \! - \! 10 < t_o<t_s) &\!\!\!: 0
 \end{cases}
@@
 \end{align}
-} 
+\vspace{-4mm}
+}
+
 We now need the $\alpha$-vectors as a function of the observation
-space for a particular belief state, thus next we eliminate the
-current state variable in line 6--8. The resulting $\delta$-functions
-now only depend on the observation: $\delta_{j}^{\vec{b}_i}(\xdo) := $
-$\int_{x_s'} \bigoplus_{x_d'} \vec{b}_i  \otimes  \alpha_j(\xds,\xdo) \hspace{1mm}dx_s'$.  For $b_2: U[t_s;6,11]$
-%assume the following simple $\delta$-functions:
-the resulting $\delta$-functions are defined below:
-{\footnotesize
-%\vspace{-2mm}
-\begin{align}
-\delta_{close}^{b_2}(t_o) &= 
-\begin{cases}
- (14 < t_o< 18) &: 0.025*t_o - 0.45\\
+space for a particular belief state, thus next we marginalize out $\xds$
+in lines 5--7. The resulting $\delta$-functions are as follows:
+{\footnotesize
+\vspace{-1mm}
+\begin{align}
+\delta^{\close}_1(t_o) &= 
+\begin{cases}
+ (14 < t_o< 18) &: 0.025 t_o - 0.45\\
  (8 < t_o< 14) &:  - 0.1\\
- (4 < t_o< 8) &: - 0.025*t_o -0.1\\
+ (4 < t_o< 8) &: - 0.025 t_o -0.1\\
 \end{cases}
 \hspace{5mm} 
-\delta_{open}^{b_2}(t_o) &= \begin{cases}
- (15 < t_o< 18) &: 25*t_o - 450\\
- (14 < t_o< 15) &: - 2.5*t_o - 37.5\\
+\delta^{\open}_1(t_o) = \begin{cases}
+ (15 < t_o< 18) &: 25 t_o - 450\\
+ (14 < t_o< 15) &: - 2.5 t_o - 37.5\\
  (8 < t_o< 14) &:  -72.5\\
- (5 < t_o< 8) &: - 25*t_o + 127.5\\
- (4 < t_o< 5) &:  2.5*t_o - 10\\
+ (5 < t_o< 8) &: - 25 t_o + 127.5\\
+ (4 < t_o< 5) &:  2.5 t_o - 10\\
 \end{cases}
@@
 \end{align}
-%\vspace{-5mm}
-}
-%{\footnotesize
-%\begin{align}
-%\delta_{1}(t_o) &= 
-%\begin{cases}
-% (2<t_o<6) &: 15 - 2.5*t_o \\
-%(-4<t_o<2) &: 10 \\
-%(-8<t_o<-4) &: 20 + 2.5*t_o 
-%\end{cases}
-%\hspace{5mm} 
-%\delta_{2}(t_o) &= \begin{cases}
-% (2<t_o<6) &: -0.15 + 0.025*t_o \\
-%(-4<t_o<2) &: -0.1 \\
-%(-8<t_o<-4) &: -0.2 - 0.025*t_o 
-%\end{cases}
-%\nonumber
-%\end{align}
-%}
+\vspace{-4mm}
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \caption{\footnotesize 
-{\it (left)} Beliefs $b_1,b_2$ for 1D \textsc{\bf Power Plant} example; 
+{\it (left)} Beliefs $b_1,b_2$ for \textsc{\bf Cont. 1D-Power Plant}; 
 %{\it (center)} Observation dependent function for $b_2$ that partition the observation space into 5 regions with different probabilities for $p(o_1),p(o_2)$ ; 
-{\it (right)} Relevant observation partitions and their probabilities for $b_2$.
+{\it (right)} derived observations partitions for $b_2$.
+%, and optimal conditional policies.
 This diagrams shows that for temperatures $t_o < 5.1$ the policy is to open the valve and for higher temperatures, it is unsafe to open it and the policy should follow $\delta_{close}(t_o)$ which is to close the valve.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_2$.
-According to continuous observation space defined in \cite{pascal_ijcai05}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space.
-In their work, numerical solutions are proposed to find the roots of any two observation dependent function.
-Instead, here we leverage the symbolic power of the max-operator defined in Section~\ref{sec:case} to find all the boundary regions that define for what partitionings of the observation each
-$\delta$-function is optimal. For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximum in lines 9--11: 
-{\footnotesize
+
+Both $\delta^{\close}_1(t_o)$ and $\delta^{\open}_1(t_o)$ are drawn graphically
+in Figure~\ref{fig:beliefs} (right).  Here we see that the only relevant
+observation is whether $t_o < 5.1$ or not -- this is the only distinction
+that determines whether the policy associated with 
+$\delta^{\close}_1(t_o)$ is higher than $\delta^{\open}_1(t_o)$.
+
+% *** Only relevant observations, need a way to automatically derive...
+
+% *** Later show extract-relevant-partitions
+
+The observation dependent $\delta$-functions divide the observation
+space into regions which can yield the optimal policy according to the
+belief state $b_2$. According to continuous observation space defined
+in \cite{pascal_ijcai05}, for a 1-dimensional observation space with
+two or more functions of the observation, we need to find the optimal
+boundaries or partitions of the space. In their work, numerical
+solutions are proposed to find the roots of any two observation
+dependent function. Instead, here we leverage the symbolic power of
+the max-operator defined in Section~\ref{sec:case} to find all the
+boundary regions that define for what partitionings of the observation
+each $\delta$-function is optimal. For the two $\delta$-functions
+above the following partitions of the observation space is derived
+after taking their maximum in lines 9--11: {\footnotesize
 \vspace{-2mm}
@@
 \begin{cases}
-o_1: (14 < t_o< 18) &: 0.025*t_o - 0.45\\
+o_1: (14 < t_o< 18) &: 0.025 t_o - 0.45\\
 o_1: (8 < t_o< 14) &:  -0.1\\
-o_1: (5.1 < t_o< 8) &: - 0.025*t_o -0.1\\
-o_2: (5 < t_o< 5.1) &: - 25*t_o + 127.5\\
-o_2: (4 < t_o< 5) &:  2.5*t_o - 10\\
+o_1: (5.1 < t_o< 8) &: - 0.025 t_o -0.1\\
+o_2: (5 < t_o< 5.1) &: - 25 t_o + 127.5\\
+o_2: (4 < t_o< 5) &:  2.5 t_o - 10\\
 \end{cases}
@@
 partition:
-$$p(o_k|\vec{b}_i) := \int_{x_s'}\int_{x_o}\int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} p(\xdo|\xds',a)*p(\xds'|\xds,a)*\vec{b}_i* \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$$
+$$p(o_k|\vec{b}_i) := \int_{x_s'}\int_{x_o}\int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} p(\xdo|\xds',a) p(\xds'|\xds,a) \vec{b}_i  \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$$
 For our example the non-zero probabilities occur in the following partitions as below:

### 2012-11-02 cpomdp/camdp-xadd/doc/NIPS12/draft6.tex
Intermediate checkin

@@
 %\item 
-{\it Restriction $f|_{\phi}$:}  Takes a function $f$ to restrict only in cases
-that satisfy some formula $\phi$ as defined in \cite{sanner_uai11}.
+%%{\it Restriction $f|_{\phi}$:}  Takes a function $f$ to restrict only in cases
+%%that satisfy some formula $\phi$ as defined in \cite{sanner_uai11}.
 %\item 
@@
 
-In general, it would be impossible to apply standard VI to 
-DC-POMDPs with continuous observations since the number of observations
-is infinite.  However, building on ideas in~\cite{pascal_ijcai05},
-in the case of PBVI, it is possible to \emph{derive} a finite set of
-continuous observation partitions that permit exact point-based backups
-\emph{at a belief point}.
-This additional operation (\texttt{GenRelObs}) appears on line 10 of PBVI in 
-Algorithm~\ref{alg:vi} in the case of continuous observations and is
-formally defined in Algorithm~\ref{alg:genrelobs}.
-
-To demonstrate the generation of relevant continuous observation partitions, 
-we use the second iteration of the \textsc{\bf Cont. Obs. 1D-Power
-  Plant} along with two belief points represented as 
-uniform distributions: $b_1: U(t_s;2,6)$ and $b_2: U(t_s;6,11)$ as
-shown in Figure~\ref{fig:beliefs} (left).
-Letting $h=2$, we assume that $\Gamma^{1}$ contains only the 
-reward as defined in~\eqref{eq:trans}.
-Then in lines 2--4 of Algorithm~\ref{alg:genrelobs}, we compute the 
-following two $\alpha$-functions for $a \in \{ \open, \close \}$ 
-($j=1$ since $|\Gamma^{1}| = 1$):
-{\footnotesize
-\vspace{-1mm}
-\begin{align}
-\alpha_1^{\close}(t_s,t_o) &= 
-\begin{cases}
- (t_s<15)\wedge (t_s \! - \! 10 < t_o<t_s) &\!\!\!: 10 \\
-(t_s\geq15)\wedge (t_s \! - \! 10 < t_o<t_s) &\!\!\!: -100  \\
-\neg(t_s \! - \! 10 < t_o<t_s) &\!\!\! : 0
-\end{cases}
-\;\;
-\alpha_1^{\open}(t_s,t_o) = \begin{cases}
-(t_s \! - \! 10 < t_o<t_s) &\!\!\!: -0.1 \\
-\neg(t_s \! - \! 10 < t_o<t_s) &\!\!\!: 0
-\end{cases}
-\nonumber
-\end{align}
-\vspace{-4mm}
-}
-
-We now need the $\alpha$-vectors as a function of the observation
-space for a particular belief state, thus next we marginalize out $\xds$
-in lines 5--7. The resulting $\delta$-functions are as follows:
-{\footnotesize
-\vspace{-1mm}
-\begin{align}
-\delta^{\close}_1(t_o) &= 
-\begin{cases}
- (14 < t_o< 18) &: 0.025 t_o - 0.45\\
- (8 < t_o< 14) &:  - 0.1\\
- (4 < t_o< 8) &: - 0.025 t_o -0.1\\
-\end{cases}
-\hspace{5mm} 
-\delta^{\open}_1(t_o) = \begin{cases}
- (15 < t_o< 18) &: 25 t_o - 450\\
- (14 < t_o< 15) &: - 2.5 t_o - 37.5\\
- (8 < t_o< 14) &:  -72.5\\
- (5 < t_o< 8) &: - 25 t_o + 127.5\\
- (4 < t_o< 5) &:  2.5 t_o - 10\\
-\end{cases}
-\nonumber
-\end{align}
-\vspace{-4mm}
-}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Both $\delta^{\close}_1(t_o)$ and $\delta^{\open}_1(t_o)$ are drawn graphically
-in Figure~\ref{fig:beliefs} (right).  Here we see that the only relevant
-observation is whether $t_o < 5.1$ or not -- this is the only distinction
-that determines whether the policy associated with 
-$\delta^{\close}_1(t_o)$ is higher than $\delta^{\open}_1(t_o)$.
-
-% *** Only relevant observations, need a way to automatically derive...
-
-% *** Later show extract-relevant-partitions
-
-The observation dependent $\delta$-functions divide the observation
-space into regions which can yield the optimal policy according to the
-belief state $b_2$. According to continuous observation space defined
-in \cite{pascal_ijcai05}, for a 1-dimensional observation space with
-two or more functions of the observation, we need to find the optimal
-boundaries or partitions of the space. In their work, numerical
-solutions are proposed to find the roots of any two observation
-dependent function. Instead, here we leverage the symbolic power of
-the max-operator defined in Section~\ref{sec:case} to find all the
-boundary regions that define for what partitionings of the observation
-each $\delta$-function is optimal. For the two $\delta$-functions
-above the following partitions of the observation space is derived
-after taking their maximum in lines 9--11: {\footnotesize
-\vspace{-2mm}
+In general, it would be impossible to apply standard VI to 
+DC-POMDPs with continuous observations since the number of observations
+is infinite.  However, building on ideas in~\cite{pascal_ijcai05},
+in the case of PBVI, it is possible to \emph{derive} a finite set of
+continuous observation partitions that permit exact point-based backups
+\emph{at a belief point}.
+This additional operation (\texttt{GenRelObs}) appears on line 10 of PBVI in 
+Algorithm~\ref{alg:vi} in the case of continuous observations and is
+formally defined in Algorithm~\ref{alg:genrelobs}.
+
+To demonstrate the generation of relevant continuous observation partitions, 
+we use the second iteration of the \textsc{\bf Cont. Obs. 1D-Power
+  Plant} along with two belief points represented as 
+uniform distributions: $b_1: U(t_s;2,6)$ and $b_2: U(t_s;6,11)$ as
+shown in Figure~\ref{fig:beliefs} (left).
+Letting $h=2$, we assume that $\Gamma^{1}$ contains only the 
+reward as defined in~\eqref{eq:trans}.
+Then in lines 2--4 of Algorithm~\ref{alg:genrelobs}, we compute the 
+following two $\alpha$-functions for $a \in \{ \open, \close \}$ 
+($j=1$ since $|\Gamma^{1}| = 1$):\footnote{For brevity, 0 partitions
+are suppressed in all case forms in this section.}
+{\footnotesize
+\vspace{-1mm}
 \begin{align}
-\mathrm{max} \Bigg(\delta_{close}^{b_2}(t_o),\delta_{open}^{b_2}(t_o)\Bigg) &= 
-\begin{cases}
-o_1: (14 < t_o< 18) &: 0.025 t_o - 0.45\\
-o_1: (8 < t_o< 14) &:  -0.1\\
-o_1: (5.1 < t_o< 8) &: - 0.025 t_o -0.1\\
-o_2: (5 < t_o< 5.1) &: - 25 t_o + 127.5\\
-o_2: (4 < t_o< 5) &:  2.5 t_o - 10\\
+\alpha_1^{\close}(t_s,t_o) &= 
+\begin{cases}
+ (t_s<15)\wedge (t_s \! - \! 10 < t_o<t_s) &\!\!\!: 10 \\
+(t_s\geq15)\wedge (t_s \! - \! 10 < t_o<t_s) &\!\!\!: -100  \\
+\neg(t_s \! - \! 10 < t_o<t_s) &\!\!\! : 0
+\end{cases}
+\;\;
+\alpha_1^{\open}(t_s,t_o) = \begin{cases}
+(t_s \! - \! 10 < t_o<t_s) &\!\!\!: -0.1 \\
+\neg(t_s \! - \! 10 < t_o<t_s) &\!\!\!: 0
 \end{cases}
@@
 \end{align}
-}
-Note here that each partitioned is labeled according to the original $\delta$-function where $o_1$ is the label of the partition coming from $\delta_{close}^{b_2}(t_o)$. 
-These partitions define observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the two distinct observation partitions is found.
-This is demonstrated visually
-in Figure~\ref{fig:beliefs} (right) for $b_2$. %Note here that the maximum of the $\delta$-functions for $b_2$ is equal to $\delta_{open}^{b_2}(t_o)$.
-
-Now with the observation partitions derived, all that remains is to
-calculate probabilities for these relevant observations conditioned on
-the belief state. 
-% This does not check out!
-% 
-%As a reality check, we note that if we integrate out the
-%state and observation from the observation model and the belief state
-%regardless of a particular $\alpha$-vector, it integrates to one:
-%$\int_{\xdo}\int_{\xds} p(\xdo|\xds',a)*p(\xds'|\xds,a)*p(s) = 1$.
-Thus we only need to multiply the indicators of each observation partition 
-in this formula to obtain the probability mass lying in each
-partition:
-$$p(o_k|\vec{b}_i) := \int_{x_s'}\int_{x_o}\int_{x_s} \bigoplus_{d_o} \bigoplus_{d_s} \bigoplus_{d_s'} p(\xdo|\xds',a) p(\xds'|\xds,a) \vec{b}_i  \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$$
-For our example the non-zero probabilities occur in the following partitions as below:
-{\footnotesize
-\vspace{-2mm}
+\vspace{-4mm}
+}
+
+We now need the $\alpha$-vectors as a function of the observation
+space for a particular belief state, thus next we marginalize out $\xds$
+in lines 5--7. The resulting $\delta$-functions are as follows:
+{\footnotesize
+\vspace{-1mm}
 \begin{align}
-p(o_k|b_2) &= 
-\begin{cases}
-(\phi_1 : 14 < t_o< 18) & : 0.2\\
-(\phi_2 : 8 < t_o< 14) & :  0.6\\
-(\phi_3 : 5.1 < t_o< 8) & : 0.183\\
-(\phi_4 : 5 < t_o< 5.1) & : 0.0002\\
-(\phi_5 : 4 < t_o< 5) & :  0.0125\\
+\delta^{\close}_1(t_o) &= 
+\begin{cases}
+ (14 < t_o< 18) &: 0.025 t_o - 0.45\\
+ (8 < t_o< 14) &:  - 0.1\\
+ (4 < t_o< 8) &: - 0.025 t_o -0.1\\
+\end{cases}
+\hspace{5mm} 
+\delta^{\open}_1(t_o) = \begin{cases}
+ (15 < t_o< 18) &: 25 t_o - 450\\
+ (14 < t_o< 15) &: - 2.5 t_o - 37.5\\
+ (8 < t_o< 14) &:  -72.5\\
+ (5 < t_o< 8) &: - 25 t_o + 127.5\\
+ (4 < t_o< 5) &:  2.5 t_o - 10\\
 \end{cases}
@@
 \end{align}
-}
-where each constraint is defined as $\phi_i$.
-For any number of actions and observations the total number of observation partitions depends on the number of belief points.
-In our example we must have two observation probabilities which is defined as the sum of the probabilities for $o_1,o_2$ according to the constraints:
-\begin{align*}
-o_1&: \phi_1 \vee \phi_2 \vee \phi_3 \longrightarrow p(o_1|b_2)=0.983 \\
-o_2&: \phi_4 \vee \phi_5  \longrightarrow p(o_2|b_2)=0.0127
-\end{align*}
-%{\footnotesize
-%\begin{align}
-%p(z_k)=
-%\begin{cases}
-% (t_o<2) &: 0.2 \\
-%(2<t_o) &: 0.8\\
-%\end{cases} 
-%\nonumber
-%\end{align}
-%}
-
-Hence we can now use the algorithms and methods of a discrete observation setting using the probabilities of the partitioned observation space in PBVI! 
-We note here that our method applies to a restricted class of piecewise functions which is piecewise linear transitions and piecewise constant reward and belief.
-The main reason is that the integration operation over continuous states and observations only allows constant or linear constraints (upper or lower bounds) over these variables \cite{sanner_aaai12}.
-Thus although in theory we can apply this approach to any piecewise polynomial function, in practice it is limited by the integration bounds.
-%Figure ~\ref{fig:obsPartition} shows the observation partitions obtained in the 1D problem instance for different horizons and their related probabilities.
-In a lyx file, needed?
-Next we present some results for 2-dimensional continuous observation spaces.
+\vspace{-4mm}
+}
+
+Both $\delta^{\close}_1(t_o)$ and $\delta^{\open}_1(t_o)$ are drawn
+graphically in Figure~\ref{fig:beliefs} (right).  These observation-dependent 
+$\delta$'s divide the observation space into regions
+which can yield the optimal policy according to the belief state
+$b_2$.  Following \cite{pascal_ijcai05}, we need to find the optimal
+boundaries or partitions of the observation space; in their work, numerical
+solutions are proposed to find these boundaries in \emph{one dimension}
+(multiple observations are handled through an independence assumption).
+Instead, here we leverage the symbolic power of
+the $\casemax$ operator defined in Section~\ref{sec:case} to find all the
+partitions where each \emph{potentially correlated, multivariate} observation
+$\delta$ is optimal. For the two $\delta$'s 
+above, the following partitions of the observation space are derived
+by the $\casemax$ operator in line 9:
+{\footnotesize
+\vspace{-1mm}
+\begin{align}
+\casemax \left( \delta^{\close}_1(t_o),\delta^{\open}_1(t_o) \right) &= 
+\begin{cases}
+o_1: (14 < t_o \leq 18) &: 0.025 t_o - 0.45\\
+o_1: (8 < t_o \leq 14) &:  -0.1\\
+o_1: (5.1 < t_o \leq  8) &: - 0.025 t_o -0.1\\
+o_2: (5 < t_o \leq 5.1) &: - 25 t_o + 127.5\\
+o_2: (4 < t_o \leq  5) &:  2.5 t_o - 10\\
+\end{cases}
+\nonumber
+\end{align}
+\vspace{-4mm}
+}
+
+Here we have labeled with $o_1$ the observations where 
+$\delta^\close_1$ is maximal and with $o_2$ the observations where
+$\delta^\open_1$ is maximal.
+What we really care about though are just the constraints 
+identifying $o_1$ and $o_2$ and this is the
+task of $\mathrm{extract\text{-}partition\text{-}constraints}$ in line 9.
+This would associate with $o_1$ the partition constraint
+$\phi_{o_1} \equiv (5.1 < t_o \leq 8) \lor (8 < t_o \leq 14) \lor (14 < t_o \leq 18)$
+and with $o_2$ the 
+partition constraint $\phi_{o_2} \equiv (4 < t_o \leq 5) \lor (5 < t_o \leq 5.1)$
+--- taking into account the 0 partitions and the 1D nature of this example, 
+we can further simplify
+$\phi_{o_1} \equiv (t_o > 5.1)$ and $\phi_{o_2} \equiv (t_o \leq 5.1)$.
+
+Given these relevant observation partitons, our final task in lines
+10-12 is to compute the probabilities of each observation partition
+$\phi_{o_k}$.  This is simply done by marginalizing over the
+observation function $p(\mathcal{O}^h|\xdsp,a)$ within each region
+defined by $\phi_{o_k}$ (achieved by multiplying by an indicator
+function $\mathbb{I}[\phi_{o_k}]$ over these constraints).  To better understand 
+what is computed here, we can compute
+the probability $p(o_k|\vec{b}_i,a)$ of each observation for a belief,
+calculated as follows:
+{\footnotesize
+\vspace{-1mm}
+\begin{equation}
+p(o_k|\vec{b}_i,a) := \int_{\vec{x}_{s}} \int_{\vec{x}_s'} \bigoplus_{\vec{d}_s} \bigoplus_{\vec{d}_s'} p(o_k|\xdsp,a) \otimes p(\xdsp| \xds,a) \otimes \alpha_j(\xdsp) \otimes \vec{b}_i(\xds)\; d\vec{x}_s' d\vec{x}_s
+\end{equation}
+\vspace{-4mm}
+}
+
+Specifically, for $\vec{b}_2$, 
+we obtain $p(o_1|\vec{b}_2,a=\close) = 0.0127$ and $p(o_2|\vec{b}_2,a=\close) = 0.933$
+as shown in Figure~\ref{fig:beliefs} (right).  Incidentally we note that
+$p(o_k|\vec{b}_i,a)$ can be computed and used in other variants of PBVI
+that directly enumerate and work with reachable beliefs.
+
+In summary, in this section we have shown how we can now use the algorithms and methods of a discrete observation POMDP setting to compute exact 1-step point-based backups.
+Next we present some empirical results for 2-dimensional continuous observation spaces.
 
@@
 
-{\bf \textsc{\bf 2D-Power Plant}:} We consider the more complex model of the power plant similar to \cite{steam2} where the pressure inside the water tank must be controlled to avoid mixing water into the steam or explosion of the tank.
-We model the pressure variable $p$ as a partially observable variable from the observation readings of the pressure $po$.
-The two actions of increase and decrease are defined based on the change in both the temperature and the pressure.
-For the increase action we define: %the transition functions and the reward function is defined as below:
+{\bf \textsc{\bf 2D-Power Plant}:} We consider the more complex model
+of the power plant similar to \cite{steam2} where the pressure inside
+the water tank must be controlled to avoid mixing water into the steam
+(leading to explosion of the tank).  We model an observable pressure
+reading $p_o$ as a function of the underlying pressure state $p_s$.
+Again we have two actions for opening and closing a pressure valve.
+The $\close$ action has transition 
 {\footnotesize
 \begin{align}
-p(p_s'|\vec{p_s},inc)&= \delta\left[ p_s' - 
+p(p_s'|p_s,a=\close)&= \delta\left[ p_s' - 
 \begin{cases}
@@
 \hspace{5mm} 
-p(t_s'|\vec{t_s},inc)= \delta\left[ t_s' - (t_s +10) \right]\nonumber
+p(t_s'|t_s,a=\close)= \delta\left[ t_s' - (t_s +10) \right]\nonumber
 \end{align}
 }
-There is a high reward for staying withing the safe temperature and pressure range since it produces power, else depending on how safe it is to have values higher or lower than the safe range, penalty is defined.
-{\footnotesize
-\vspace{-3mm}
+and yields high reward for staying withing the safe temperature and pressure range:
+{\footnotesize
+\vspace{-1mm}
 \begin{align}
-R(t_s,p_s,inc) &= 
+R(t_s,p_s,a=\close) &= 
 \begin{cases}
@@
 (5 \leq p_s \leq 15)\wedge (t_s \leq 95)&: -1\\
-%(5 \leq p_s \leq 15)\wedge (t_s \geq 95)&: -3\\				
-%(p_s \leq 5) &: -3\\						
 (p_s \geq 15) &: -5\\ 
@@
 \end{align}
-}
-As for the decrease action, the transition functions reduce the temperature by 5 units and the pressure by 10 units as long as the pressure stays above zero.
-For the reward function, we assume that there is always a small penalty for decreasing the values because power can not be generated.
-For the observation model we consider two continuous uniform distributions such as the following:  
+\vspace{-4mm}
+}
+
+Alternately, for the $\open$ action, the transition functions reduce
+the temperature by 5 units and the pressure by 10 units as long as the
+pressure stays above zero. For the $\open$ reward function, we assume
+that there is always a small constant penalty (-1) since no electricity
+is produced.
+
+Observations are distributed uniformly within a region depending on
+their underlying state:
 {\footnotesize
@@
 }
-We define two rectangular uniform beliefs around the regions of rewarding, so that one needs to increase the values while the other should decrease them: $b_1: U[t_s;90,100]*U[p_s;0,10]$ and $b_2: U[t_s;90,130]*U[p_s;10,30]$
-%\begin{align}
-%b_1 = 
-%\begin{cases}
-%(p_o>p_s) \wedge (p_o<p_s+10) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 100)&: 0.01 \\
-% \neg( (p_o>p_s) \wedge (p_o<p_s+10) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 100)) &: 0 \\
-%\end{cases}\nonumber
-%\end{align}
-%\begin{align}
-%b_2 = 
-%\begin{cases}
-%(p_o>p_s +10) \wedge (p_o<p_s+30) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 130)&: 0.00125 \\
-% \neg((p_o>p_s +10) \wedge (p_o<p_s+30) \wedge (t_o>t_s+ 90) \wedge (t_o<t_s+ 130)) &: 0 
-%\end{cases}\nonumber
-%\end{align}
-% I can draw this and the 1D belief
-In Figure ~\ref{fig:timeSpace}, a time and space analysis of
-the two versions of \textsc{\bf Power Plant} have been performed for up to 6 horizons.
-%Comparing the two problem sizes demonstrates the effect of the number of state-observation variables in our algorithm.
-As the algorithm progresses, the time required to compute the probability of the partitions and finding the maximum $\alpha$-vector with respect to beliefs increases for both problem sizes and significantly more for the 2D version.
-%The stability in time is due to the fact that each stage almost takes the same time since it has converged quickly.%???
-Increase in the problem size, increases the partition numbers on the observation space and this produces more $\alpha$-vectors which also effects the space required to perform the algorithm.
-The number of vectors stays the same for most horizons and they drop after convergence in the 2D problem instance.
-%The number of nodes in each iteration increases as more space is required to perform higher iterations.
-This shows that although the 2D instance takes more time and space than the 1D instance, it still converges within reasonable resources.
-
-In Figure ~\ref{fig:3D}  we present plots of the maximum $\delta$-vectors of belief $b_1$ for different iterations of the 2D problem instance.
-%Each plot demonstrates the value of each vector as a function of the pressure and temperature.
-Starting with the first iteration, the value is highest for the reward range ($5<p<15 \wedge 95<t<105$) and -1 or less for other places.
+Finally for PBVI, we define two uniform beliefs as follows: 
+$\vec{b}_1: U[t_s;90,100]*U[p_s;0,10]$ and $\vec{b}_2: U[t_s;90,130]*U[p_s;10,30]$
+
+In Figure~\ref{fig:timeSpace}, a time and space analysis of the two
+versions of \textsc{\bf Power Plant} have been performed for up to 6
+horizons.\footnote{These experiments as shown rely on one additional
+  approximation in that they substitute $p(\mathcal{O}^h|\vec{b},a)$
+  in place of $p(\mathcal{O}^h|\xdsp,a)$.  While this yields correct
+  observation probabilities for a point-based backup at a particular
+  belief state $\vec{b}$, the resulting $\alpha$-functions represent
+  an approximation for other belief states.  In general, the PBVI
+  framework in this paper does \emph{not} require this approximation,
+  although when appropriate, using it should increase computational
+  efficiency.}  As the algorithm progresses, the time required to
+compute the probability of the partitions and finding the maximum
+$\alpha$-vector with respect to beliefs increases for both problem
+sizes and significantly more for the 2D version.  As the problem size
+increases, this increases the number of observation partitions and
+this produces more $\alpha$-functions which also affects the space
+required to perform the algorithm.  Later as the control policy
+simplifies in both cases, we see a drop in the number of
+$\alpha$-functions.
+
+%In Figure ~\ref{fig:3D}  we present plots of the maximum $\delta$-vectors of belief $b_1$ for different iterations of the 2D problem instance. 
+%Starting with the first iteration, the value is highest for the reward range ($5<p<15 \wedge 95<t<105$) and -1 or less for other places.
 In the fifth iteration, the value function has partitioned into more pieces, showing how higher temperatures can increase the value without considering the effect of the pressure.
 In the last plot, horizon $h=6$ has better tuned the value function so that higher temperatures and pressures increase the value of the maximum $\delta$-vector and also within the reward range, finer grain partitions have been formed.
-%Also note that the policy depends on the initial belief and for these plots they were based on $b_1$ which was lower than the reward range.
-The first action would then mean it has performed an increase while the other iterations decrease and increase respectively.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% policy2, v2plot.pdf, v9plot
-% annotation?
-\begin{figure*}[tbp!]
-%\vspace{-2mm}
-\centering
-\includegraphics[width=0.31\textwidth]{pics/2d1-4.pdf}
-\includegraphics[width=0.31\textwidth]{pics/2d9-4.pdf}
-\includegraphics[width=0.31\textwidth]{pics/2d111-4.pdf}
-\vspace{-3mm}
-\caption{\footnotesize 
-{\it (left)}Maximum $\delta$-vector for $b_1$ and action of $close$ in first iteration; 
-{\it (center)} $\delta_{close}^5(b_1)$; 
-{\it (right)} $\delta_{close}^6(b_1)$. Finer grain partitions show that closing (or opening) the valve can occur at more exact temperatures at higher horizons. 
-}
-\label{fig:3D}
-%\vspace{-4mm}
-\end{figure*}
+%\begin{figure*}[tbp!]
+%\centering
+%\includegraphics[width=0.31\textwidth]{pics/2d1-4.pdf}
+%\includegraphics[width=0.31\textwidth]{pics/2d9-4.pdf}
+%\includegraphics[width=0.31\textwidth]{pics/2d111-4.pdf}
+%\vspace{-3mm}
+%\caption{\footnotesize 
+%{\it (left)}Maximum $\delta$-vector for $b_1$ and action of $close$ in first iteration; 
+%{\it (center)} $\delta_{close}^5(b_1)$; 
+%{\it (right)} $\delta_{close}^6(b_1)$. Finer grain partitions show that closing (or opening) the valve can occur at more exact temperatures at higher horizons. 
+%}
+%\label{fig:3D}
+%\end{figure*}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-11-08 trunk/cpomdp/camdp-xadd/doc/NIPS12/draft6.tex
NIPS Camera-ready

--- 2012-11-02 cpomdp/camdp-xadd/doc/NIPS12/draft6.tex
@@
 
-\newcommand{\xds}{\mathbf{dx}_s}
-\newcommand{\xdsp}{\mathbf{dx}_s'}
-\newcommand{\xdo}{\mathbf{dx}_o}
+\newcommand{\xds}{\mathbf{x}_s,\!\mathbf{d}_s}
+\newcommand{\xdsp}{\mathbf{x}_s',\!\mathbf{d}_s'}
+%\newcommand{\xds}{\mathbf{dx}_s}
+%\newcommand{\xdsp}{\mathbf{dx}_s'}
+\newcommand{\xdo}{\mathbf{x}_o,\!\mathbf{d}_o}
+%\newcommand{\xdo}{\mathbf{dx}_o}
 \newcommand{\open}{\mathit{open}}
@@
 
-\section{DC-POMDP Model}
+\section{Hybrid POMDP Model}
 
@@
 %We introduce discrete and continuous
-%partially observable MDPs (DC-POMDPs) as an extension to 
+%partially observable MDPs (H-POMDPs) as an extension to 
 %DC-MDPs~\cite{sanner_uai11}.  
-A discrete and continuous
-partially observable MDP (DC-POMDP) is a tuple $\langle
+A \emph{hybrid} (discrete and continuous)
+\emph{partially observable MDP} (H-POMDP) is a tuple $\langle
 \mathcal{S},\mathcal{A},\mathcal{O},\mathcal{T},\mathcal{R},\mathcal{Z},\gamma,h
 \rangle$.  States $\mathcal{S}$ are given by vector 
-$\xds = (\vec{d}_s,\vec{x}_s) = (
+$(\vec{d}_s,\vec{x}_s) = (
 d_{s_1},\ldots,d_{s_n},x_{s_1},\ldots,x_{s_m} )$ where each $d_{s_i}
@@
 \ldots, a_r \}$. Observations
-$\mathcal{O}$ are given by the vector $\xdo = (\vec{d}_o,\vec{x}_o) = (
+$\mathcal{O}$ are given by the vector $(\vec{d}_o,\vec{x}_o) = (
 d_{o_1},\ldots,d_{o_p},x_{o_1},\ldots,x_{o_q} )$ where each $d_{o_i}
@@
 
-Three functions are required for modeling DC-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state %$(\vec{d}_s',\vec{x}_s')$
+Three functions are required for modeling H-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state %$(\vec{d}_s',\vec{x}_s')$
 given the action and previous state%
@@
 %% TRANSITION LINEAR, BELIEF, REWARD, OBSERVATION PWC???
-Probabilities over \emph{discrete} variables $p(d_{s_i}'|\xds,a)$ and
-$p(d_{o_i}|\xdsp,a)$ may condition on both discrete variables and
+Probabilities over \emph{discrete} variables $p(d_{s_i}'\!|\xds,\!a)$ and
+$p(d_{o_i}\!|\xdsp,\!a)$ may condition on both discrete variables and
 (nonlinear) inequalities of continuous variables; this is further
@@
 observations.  Transitions over \emph{continuous}
-variables $p(x_{s_j}'|\xds, \vec{d}_s',a)$ must be deterministic (but
+variables $p(x_{s_j}'\!|\xds,\!\vec{d}_s',\!a)$ must be deterministic (but
 arbitrary nonlinear) piecewise functions;
@@
 %(hence allowing discrete noise, but not, e.g., Gaussian noise).
-Observation probabilities over continuous variables $p(x_{o_j}|\xdsp,a)$ 
-only occur in the case of continuous observation and are required to be
+Observation probabilities over continuous variables $p(x_{o_j}\!|\xdsp,\!a)$ 
+only occur in the case of continuous observations and are required to be
 piecewise constant (a mixture of uniform distributions); the same
@@
 %To make this concrete, we now provide examples of both discrete and
-%continuous observation DC-POMDPs.
+%continuous observation H-POMDPs.
 
@@
 generation system of a power plant evaporates feed-water under restricted 
-pressure and temperature conditions to turn a stream turbine.
+pressure and temperature conditions to turn a steam turbine.
 A reward is obtained when electricity is generated from the turbine 
@@
 
-We initially present two DC-POMDP variants labeled \textsc{\bf 1D-Power
+We initially present two H-POMDP variants labeled \textsc{\bf 1D-Power
   Plant} using a single temperature state variable $t_s$.  
@@
 
-\section{Value Iteration for DC-POMDPs}
+\section{Value Iteration for Hybrid POMDPs}
 
@@
 
-In a DC-POMDP, the agent does not directly observe the states and thus
+In an H-POMDP, the agent does not directly observe the states and thus
 must maintain a belief state $b(\xds) = p(\xds)$.  For a
@@
 \begin{align} 
-g^h_{a,o,j}(\xds) &=  \int_{\vec{x}_{s'}} \sum_{\vec{d_{s'}}} p(o|\xds',a)p(\xds'|\xds,a) \alpha^{h-1}_j(\xds') d\vec{x}_{s'}; \hspace{2mm}  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1} \label{eq:backup} \\
+g^h_{a,o,j}(\xds) &=  \int_{\vec{x}_{s'}} \sum_{\vec{d_{s'}}} p(o|\xdsp,a)p(\xdsp|\xds,a) \alpha^{h-1}_j(\xdsp) d\vec{x}_{s'}; \hspace{2mm}  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1} \label{eq:backup} \\
 \Gamma^{h}_a   &= R(\xds,a) + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\xds) \right\}_j  \label{eq:cross_prod}\\ 
@@
              % $\pi^{*,h} \,:=\, \argmax_{a} \, \Gamma_a^{h}$\;
+      \emph{// Retain only $\alpha$-functions optimal at each belief point}\;
       \ForEach {$\vec{b}_i \in B$}
       {
-     	$\alpha_{\vec{b}_i}^h :=\ \argmax_{\alpha_j \in \Gamma^h} \alpha_j \cdot \vec{b}_i$\;
+     	$\alpha_{\vec{b}_i}^h :=\ \argmax_{\alpha_j \in \Gamma^h} \alpha_j \cdot \vec{b}_i$ $\,$\;
      	$\Gamma_{PBVI}^h :=\ \Gamma_{PBVI}^h \cup \alpha_{\vec{b}_i}^h$\;
@@
 
+        \emph{// Terminate if early convergence}\;
        \If{$\Gamma_{PBVI}^h = \Gamma_{PBVI}^{h-1}$}
-           {break $\,$ \emph{// Terminate if early convergence}\;}
+           {break $\,$\;}
    }
@@
 }
-\caption{\footnotesize \texttt{PBVI}(DC-POMDP, $H$, $B=\left\{\vec{b}_i \right\}$) $\longrightarrow$ $\l V^h \r$ \label{alg:vi}}
+\caption{\footnotesize \texttt{PBVI}(H-POMDP, $H$, $B=\left\{\vec{b}_i \right\}$) $\longrightarrow$ $\l V^h \r$ \label{alg:vi}}
 \vspace{-1mm}
@@
 
-\textbf{Point-based value iteration (PBVI)} computes the value
+\textbf{Point-based value iteration (PBVI)}~\cite{pbvi_jair06,Perseus} 
+computes the value
 function only for a set of belief states $\{ \vec{b}_i \}$ where
@@
 modification needed to Monahan's VI approach in Algorithm~\ref{alg:vi}
-(following~\cite{Perseus_cont}) is the loop from lines 22--24 where
-only $\alpha$-vectors optimal at some belief state are retained for
-subsequent iterations.  In the case of continuous observation
+is the loop from lines 23--25 where
+only $\alpha$-functions optimal at some belief state are retained for
+subsequent iterations.
+In the case of continuous observation
 variables ($q > 0$), we will need to derive a relevant set of
 observations on line 10, a key contribution of this work as described
-in Section~\ref{sec:cont_obs}.  Whereas PBVI is optimal if all
-reachable belief states within horizon $H$ are enumerated in $B$, in the DC-POMDP
-setting, the generation of continuous observations will most often
-lead to an infinite number of reachable belief states, even with
-finite horizon.  Nonetheless, PBVI has been quite successful in
-practice without exhaustive enumeration of all reachable
+in Section~\ref{sec:cont_obs}.  Otherwise if the observations are only 
+discrete ($q=0$), then a finite set of observations is already known and
+the observation function as given in Eq~\eqref{eq:obs_model}.
+
+We remark that Algorithm~\ref{alg:vi} is a generic framework that 
+can be used for both PBVI and other variants of 
+approximate VI.  If used for PBVI, an 
+efficient direct backup computation of the optimal $\alpha$-function for belief 
+state $\vec{b}_i$ should be used in line 17 
+that is linear in the number of observations~\cite{pbvi_jair06,Perseus}
+and which obviates the need for lines 23--25.  
+However, for an alternate version of approximate value iteration that will often
+produce more accurate values for belief states other than those in $B$, 
+one may instead retain the full cross-sum backup
+of line 17, but omit lines 23--25 --- 
+this yields an approximate VI approach 
+(using discretized observations 
+relevant only to a chosen set of belief states $B$ if continuous observations
+are present) that is not restricted to alpha-functions only optimal at $B$, 
+hence allowing greater flexibility
+in approximating the value function over all belief states.%}
+
+Whereas PBVI is optimal if all reachable belief states within horizon
+$H$ are enumerated in $B$, in the H-POMDP setting, the generation of
+continuous observations will most often lead to an infinite number of
+reachable belief states, even with finite horizon --- this makes
+it quite difficult to provide optimality guarantees in the general case of PBVI
+for continuous observation settings.  Nonetheless, PBVI has been quite
+successful in practice without exhaustive enumeration of all reachable
 beliefs~\cite{pbvi_jair06,hsvi2,Perseus,gapmin}, which motivates our
@@
 
-and this is the form we use to represent all functions in a DC-POMDP.
+and this is the form we use to represent all functions in an H-POMDP.
 The $\phi_i$ are disjoint logical formulae defined over $\xds$ and/or $\xdo$ with logical ($\land,\lor,\neg$) combinations of boolean variables and inequalities ($\geq,>,\leq,<$) over continuous variables.
-For discrete observation DC-POMDPs, the $f_i$ and inequalities may use any function (e.g., $\sin(x_1) > \log(x_2)\cdot x_3)$; for continuous observations, they are restricted to linear inequalities and linear or piecewise constant $f_i$ as described in Section~\ref{sec:model}.
+For discrete observation H-POMDPs, the $f_i$ and inequalities may use any function (e.g., $\sin(x_1) > \log(x_2)\cdot x_3)$; for continuous observations, they are restricted to linear inequalities and linear or piecewise constant $f_i$ as described in Section~\ref{sec:model}.
 
@@
 The following SDP operations on case statements require more detail than can be provided here, hence we refer the reader to the relevant literature:
-%\begin{itemize}
+\begin{itemize}
 %\item 
@@
 %%that satisfy some formula $\phi$ as defined in \cite{sanner_uai11}.
-%\item 
-{\it Substitution $f\sigma$:} Takes a set $\sigma$ of variables and their substitutions (which may be case statements), and carries out all variable substitutions in sequence~\cite{sanner_uai11}.
-%\item 
-{\it Integration $\int_{x_1} f \; dx_1$:}  There are two forms: If $x_1$ is involved in a $\delta$-function (cf.
-the transition in Eq~\eqref{eq:trans}) then the integral is equivalent to a symbolic substitution and can be applied to any case statement~\cite{sanner_uai11}.
-Otherwise, if $f$ is restricted to linear constraints and constant values, then the approach of~\cite{sanner_aaai12} can be applied to yield a linearly constained piecewise linear result.
-%\end{itemize}
+\item 
+{\it Substitution $f\sigma$:} Takes a set $\sigma$ of variables and their substitutions (which may be case statements themselves), and carries out all variable substitutions~\cite{sanner_uai11}.
+\item 
+{\it Integration $\int_{x_1} f \; dx_1$:}  There are two forms: If $x_1$ is involved in a $\delta$-function ({\it cf.} the transition in Eq~\eqref{eq:trans}) then the integral is equivalent to a symbolic substitution and can be applied to \emph{any} case statement ({\it cf.} ~\cite{sanner_uai11}).
+Otherwise, if $f$ is in linearly constrained polynomial case form, then the approach of~\cite{sanner_aaai12} can be applied to yield a result in the same form.
+\end{itemize}
 
@@
 %xadd representation
-The data structure of the \emph{extended algebraic decision diagram}
-(XADD)~\cite{sanner_uai11} is used to support case statements and the
-required operations.  Figure~\ref{fig:cond_plan} (right) 
-is an example of an XADD representation.
+Case operations yield a combinatorial explosion in size if
+na\"{i}vely implemented, hence 
+we use the data structure of the \emph{extended algebraic
+  decision diagram} (XADD)~\cite{sanner_uai11} as shown in
+Figure~\ref{fig:cond_plan} (right) to \emph{compactly} 
+represent case statements and \emph{efficiently} support the above
+case operations with them.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{VI for DC State and Discrete Observations} 
+\subsection{VI for Hybrid State and Discrete Observations} 
 \label{sec:disc_obs}
 
-For DC-POMDPs with only discrete observations $o
-\in \mathcal{O}$ and observation function $p(o|\xds',a)$ (e.g., 
-Eq~\eqref{eq:ex_disc_obs}), we introduce a symbolic version of
+For H-POMDPs with only discrete observations $o
+\in \mathcal{O}$ and observation function $p(o|\xdsp,a)$ 
+as in the form of Eq~\eqref{eq:ex_disc_obs}, we introduce a symbolic version of
 Monahan's VI algorithm.  In brief, we note that all VI operations
 needed in Section~\ref{sec:vi} apply \emph{directly} to 
-DC-POMDPs, e.g., we can rewrite
+H-POMDPs, e.g., rewriting 
 Eq~\eqref{eq:backup}: {\footnotesize
@@
 \bigoplus_{\vec{d_{s'}}} \!
-\left[ p(o|\xds',\!a) \!
+\left[ p(o|\xdsp,\!a) \!
 \otimes \!\!
 \left( \!
-\bigotimes_{i=1}^n p(d_{s_i}'|\xds,\!a) \!\!
+\bigotimes_{i=1}^n p(d_{s_i}'\!|\xds,\!a) \!\!
 \right) \!\!
@@
 \left( \!
-\bigotimes_{j=1}^m p(x_{s_j}'|\xds, \vec{d}_s',\!a) \!\!
+\bigotimes_{j=1}^m p(x_{s_j}'\!|\xds, \vec{d}_s',\!a) \!\!
 \right) \!\!
 \otimes \!
-\alpha^{h-1}_j(\xds') \!
+\alpha^{h\!-\!1}_j(\xdsp) \!
 \right] \!\!
@@
 
-Crucially we note since the continuous transition cpfs $p(x_{s_j}'|\xds, \vec{d}_s',a)$ are deterministic and hence defined with Dirac $\delta$'s (e.g., Eq~\ref{eq:trans}) as described in Section~\ref{sec:model}, the integral $\int_{\vec{x}_{s'}}\!$ can always be computed in closed case form as discussed in Section~\ref{sec:case}.
-In short, nothing additional is required for PBVI on DC-POMDPs
+Crucially we note since the continuous transition cpfs $p(x_{s_j}' \!
+|\xds, \!
+\vec{d}_s',\!a)$ are deterministic and hence defined with Dirac $\delta$'s (e.g., Eq~\ref{eq:trans}) as described in Section~\ref{sec:model}, the integral $\int_{\vec{x}_{s'}}\!$ can always be computed in closed case form as discussed in Section~\ref{sec:case}.
+In short, nothing additional is required for PBVI on H-POMDPs
 in this case --- the key insight
@@
     	  \emph{// Let $\phi_{o_k}$ be the partition constraints for observation $o_k \in \mathcal{O}^h$}\\
-            $p(\mathcal{O}^h = o_k|\xdsp,a) := \int_{\vec{x}_o} \bigoplus_{\vec{d}_o} p(\xdo|\xds',a) \mathbb{I}[\phi_{o_k}] d_{\vec{x}_o}$ \;
-%            $p(\mathcal{O}^h = o_k|\xdsp) := \int_{\vec{x}_s'} \int_{\vec{x}_o} \int_{\vec{x}_s} \bigoplus_{\vec{d}_o} \bigoplus_{\vec{d}_s} \bigoplus_{\vec{d}_s'} p(\xdo|\xds',a) \otimes p(\xds'|\xds,a) \otimes \vec{b}_i \otimes  \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$ \;
+            $p(\mathcal{O}^h = o_k|\xdsp,a) := \int_{\vec{x}_o} \bigoplus_{\vec{d}_o} p(\xdo|\xdsp,a) \mathbb{I}[\phi_{o_k}] d_{\vec{x}_o}$ \;
+%            $p(\mathcal{O}^h = o_k|\xdsp) := \int_{\vec{x}_s'} \int_{\vec{x}_o} \int_{\vec{x}_s} \bigoplus_{\vec{d}_o} \bigoplus_{\vec{d}_s} \bigoplus_{\vec{d}_s'} p(\xdo|\xdsp,a) \otimes p(\xdsp|\xds,a) \otimes \vec{b}_i \otimes  \mathbb{I}[\phi_{o_k}] d_{x_o} d_{x_s}d_{x_s'}$ \;
         }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{PBVI for DC State and DC Observations} 
+\subsection{PBVI for Hybrid State and Hybrid Observations} 
 \label{sec:cont_obs}
@@
 %\hspace{-2mm}
-\includegraphics[width=0.42\textwidth]{pics/delta_b2_2.pdf}
+\includegraphics[width=0.42\textwidth]{pics/delta_b2_3.pdf}
 \hspace{-17mm}
@@
 %{\it (center)} Observation dependent function for $b_2$ that partition the observation space into 5 regions with different probabilities for $p(o_1),p(o_2)$ ; 
-{\it (right)} derived observations partitions for $b_2$.
+{\it (right)} derived observation partitions for $b_2$ at $h=2$.
 %, and optimal conditional policies.
@@
 In general, it would be impossible to apply standard VI to 
-DC-POMDPs with continuous observations since the number of observations
+H-POMDPs with continuous observations since the number of observations
 is infinite.  However, building on ideas in~\cite{pascal_ijcai05},
@@
 shown in Figure~\ref{fig:beliefs} (left).
-Letting $h=2$, we assume that $\Gamma^{1}$ contains only the 
-reward as defined in~\eqref{eq:trans}.
-Then in lines 2--4 of Algorithm~\ref{alg:genrelobs}, we compute the 
-following two $\alpha$-functions for $a \in \{ \open, \close \}$ 
-($j=1$ since $|\Gamma^{1}| = 1$):\footnote{For brevity, 0 partitions
-are suppressed in all case forms in this section.}
-{\footnotesize
-\vspace{-1mm}
+Letting $h=2$, we will assume simply for expository purposes that 
+$|\Gamma^{1}| = 1$ (i.e., it contains only one $\alpha$-function) and
+that in lines 2--4 of Algorithm~\ref{alg:genrelobs} we have computed the 
+following two $\alpha$-functions for $a \in \{ \open, \close \}$:
+{\footnotesize
+\vspace{-2mm}
 \begin{align}
@@
 \alpha_1^{\open}(t_s,t_o) = \begin{cases}
-(t_s \! - \! 10 < t_o<t_s) &\!\!\!: -0.1 \\
+(t_s \! - \! 10 < t_o<t_s) &\!\!\!: 0.1 \\
 \neg(t_s \! - \! 10 < t_o<t_s) &\!\!\!: 0
@@
 \end{align}
-\vspace{-4mm}
+\vspace{-3mm}
 }
@@
 space for a particular belief state, thus next we marginalize out $\xds$
-in lines 5--7. The resulting $\delta$-functions are as follows:
+in lines 5--7. The resulting $\delta$-functions are shown as follows
+where for brevity from this point forward, 0 partitions
+are suppressed in the cases:
 {\footnotesize
@@
 what is computed here, we can compute
-the probability $p(o_k|\vec{b}_i,a)$ of each observation for a belief,
-calculated as follows:
+the probability $p(o_k|\vec{b}_i,a)$ of each observation for a 
+particular belief, calculated as follows:
 {\footnotesize
@@
 
-Specifically, for $\vec{b}_2$, 
-we obtain $p(o_1|\vec{b}_2,a=\close) = 0.0127$ and $p(o_2|\vec{b}_2,a=\close) = 0.933$
-as shown in Figure~\ref{fig:beliefs} (right).  Incidentally we note that
-$p(o_k|\vec{b}_i,a)$ can be computed and used in other variants of PBVI
-that directly enumerate and work with reachable beliefs.
-
-In summary, in this section we have shown how we can now use the algorithms and methods of a discrete observation POMDP setting to compute exact 1-step point-based backups.
-Next we present some empirical results for 2-dimensional continuous observation spaces.
+Specifically, for $\vec{b}_2$, we obtain $p(o_1|\vec{b}_2,a=\close) =
+0.0127$ and $p(o_2|\vec{b}_2,a=\close) = 0.933$ as shown in
+Figure~\ref{fig:beliefs} (right).
+
+In summary, in this section we have shown how we can extend the exact
+dynamic programming algorithm for the continuous state, discrete
+observation POMDP setting from Section~\ref{sec:disc_obs} to compute
+exact 1-step point-based backups in the continuous observation setting;
+this was accomplished through the crucial insight that despite the
+infinite number of observations, using Algorithm~\ref{alg:genrelobs}
+we can symbolically 
+derive a set of relevant observations for each belief point that
+distinguish the optimal policy and hence value as graphically
+illustrated in Figure~\ref{fig:beliefs} (right).  Next we present some
+empirical results for 1- and 2-dimensional continuous state and 
+observation spaces.
 
@@
 specifications and Java code to reproduce these experiments are available online in
-Google Code: \textit{http://code.google.com/p/cpomdp} .}
+Google Code: \texttt{http://code.google.com/p/cpomdp} .}
 
@@
 \centering
-\includegraphics[width=0.32\textwidth]{pics/time2.pdf}
-\includegraphics[width=0.32\textwidth]{pics/nodes2.pdf}
-\includegraphics[width=0.32\textwidth]{pics/alpha-vectors2.pdf}
+\includegraphics[width=0.35\textwidth]{pics/time3.pdf} \hspace{15mm}
+\includegraphics[width=0.35\textwidth]{pics/nodes3.pdf}
+%\includegraphics[width=0.32\textwidth]{pics/alpha-vectors2.pdf}
 \vspace{-1mm}
 \caption{\footnotesize 
-{\it (left)} Space vs Horizon; 
-{\it (center)} Time vs Horizon; 
-{\it (right)} Number of $\alpha$-vectors vs Horizon.
+{\it (left)} time vs. horizon, and 
+{\it (right)} space (total \# XADD nodes in $\alpha$-functions) vs. horizon.
+%{\it (right)} number of $\alpha$-functions that would be generated before
+%pruning for each horizon, belief state, and problem if the full 
+%exponential cross-sum backup were used.
+%{\it (right)} Number of $\alpha$-vectors vs Horizon.
+%{\bf TODO:} *** \# of alpha-vectors that would be computed in an
+%approximate value iteration approach.  COVER in text as well.
 }
@@
 }
-and yields high reward for staying withing the safe temperature and pressure range:
+and yields high reward for staying within the 
+safe temperature and pressure range:
 {\footnotesize
@@
 \begin{cases}
-(t_s + 80<t_o<t_s+ 105) &: 0.4 \\
+(t_s + 80<t_o<t_s+ 105) &: 0.04 \\
  \neg (t_s + 80<t_o<t_s+ 105) &: 0 \\
@@
 In Figure~\ref{fig:timeSpace}, a time and space analysis of the two
-versions of \textsc{\bf Power Plant} have been performed for up to 6
-horizons.\footnote{These experiments as shown rely on one additional
-  approximation in that they substitute $p(\mathcal{O}^h|\vec{b},a)$
-  in place of $p(\mathcal{O}^h|\xdsp,a)$.  While this yields correct
+versions of \textsc{\bf Power Plant} have been performed for up to 
+horizon $h=6$. This experimental evaluation relies on one additional
+  approximation over the PBVI approach of Algorithm~\ref{alg:vi} 
+  in that it substitutes $p(\mathcal{O}^h|\vec{b},a)$
+  in place of $p(\mathcal{O}^h|\xdsp,a)$ --- while this yields correct
   observation probabilities for a point-based backup at a particular
@@
   although when appropriate, using it should increase computational
-  efficiency.}  As the algorithm progresses, the time required to
-compute the probability of the partitions and finding the maximum
-$\alpha$-vector with respect to beliefs increases for both problem
-sizes and significantly more for the 2D version.  As the problem size
-increases, this increases the number of observation partitions and
-this produces more $\alpha$-functions which also affects the space
-required to perform the algorithm.  Later as the control policy
-simplifies in both cases, we see a drop in the number of
-$\alpha$-functions.
+  efficiency.
+
+Figure~\ref{fig:timeSpace} shows that the computation time required
+per iteration generally increases since more complex $\alpha$-functions lead to
+a larger number of observation partitions and thus a more expensive
+backup operation.  While an order of magnitude more time is required
+to double the number of state and observation variables, one can
+see that the PBVI approach leads to a fairly constant amount of computation
+time per horizon, which indicates that long horizons should be computable
+for any problem for which at least one horizon can be computed in an
+acceptable amount of time.
 
@@
 We presented the first exact symbolic operations for \texttt{PBVI} in
-expressive DC-POMDPs with continuous state \emph{and} observations.
+an expressive subset of H-POMDPs with continuous state \emph{and} observations.
 Unlike related work that has extended to the continuous state and
@@
 using symbolic integration.  An important avenue for future work is to
-determine whether similar techniques can be applied to the difficult
-case of continuous state, observation, \emph{and} action DC-POMDPs.
+extend these techniques to the case of
+continuous state, observation, \emph{and} action H-POMDPs.
+
+{\small
 
@@
 
-NICTA is funded by the Australian Government as represented by the Department of Broadband, Communications and the Digital Economy and the Australian Research Council through the ICT Centre of Excellence program.
+NICTA is funded by the Australian Government as represented by the Department of Broadband, Communications and the Digital Economy and the ARC through the ICT Centre of Excellence program.
 This work was supported by the Fraunhofer ATTRACT fellowship STREAM and by the EC, FP7-248258-First-MM.
+}
+
 %\subsubsection*{References} 
